### 2016-08-11 article/unrave.tex
Paper frame

%% This file is part of the UNRAVE Project 
%% Copyright 2016 the authors.  All rights reserved.

\documentclass[preprint2,trackchanges]{aastex}

\usepackage{amsmath,amssymb}
\usepackage{bm}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\logg}{\log g}
\newcommand{\Teff}{T_{\mathrm{eff}}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}

%\AuthorCallLimit=10

\begin{document}

\title{The \project{UNRAVE} catalog}

\author{Andrew R. Casey}
\affil{Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}

\author{Some combination of: Hawkins, Ness, Hogg, HWR, Gilmore, Matijevic, RAVE collaboration}
%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}

\begin{abstract}
\end{abstract}

\keywords{}

\section{Introduction} \label{sec:intro}

\acknowledgements

\begin{thebibliography}{dummy}
\end{thebibliography}

\clearpage

%\begin{figure}[p]
%\caption{Choose two wavelengths (one continuum and one interesting)
%  and plot some scatter plots of flux vs various parameters, and also
%  cross-validation results for the
%  hyper-parameters.\label{fig:onewavelength}}
%\end{figure}

\end{document}

### 2016-08-11 article/unrave.tex
foo

@@
 
-\title{The \project{UNRAVE} catalog}
+\title{The \project{UNRAVE} catalog of stellar astrophysical parameters and chemical abundances}
 
@@
 \begin{abstract}
+% Gaia will release stuff
+% Much we want to do will involve chemistry + kinematics
+% RAVE has the largest overlap with Gaia DR1, but chemistry is unavailable for most stars
+% We perform an independent analysis of RAVE spectra w/ The Cannon, censoring, + upgrades, to derive astrophysical parameters and detailed chemical abundances.
+% We produce a catalog of up to N abundances for X stars, constituting the largest
+% catalog of chemical abundances that overlap with Gaia by a factor of X.
+
 \end{abstract}
@@
 
-\section{Introduction} \label{sec:intro}
+\section{Introduction} 
+\label{sec:introduction}
+
+\section{Methods}
+\label{sec:methods}
+
+\section{Validation}
+\label{sec:validation}
+
+\section{Conclusion}
+\label{sec:conclusion}
 
 \acknowledgements
+% We thank,......
+
+This research made use of: Astropy, a community-developed core Python package for
+Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
+ and \project{TOPCAT} \citep{Taylor2005}.
+This work was partly supported by the European Union FP7 programme through ERC 
+grant number 320360.
+Funding for RAVE has been provided by: the Australian Astronomical Observatory; the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National University; the Australian Research Council; the French National Research Agency; the German Research Foundation (SPP 1177 and SFB 881); the European Research Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at Padova; The Johns Hopkins University; the National Science Foundation of the USA (AST-0908326); the W.
+M.
+Keck foundation; the Macquarie University; the Netherlands Research School for Astronomy; the Natural Sciences and Engineering Research Council of Canada; the Slovenian Research Agency; the Swiss National Science Foundation; the Science \& Technology Facilities Council of the UK; Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg and Sydney.
+The RAVE web site is at https://www.rave-survey.org.
 
 \begin{thebibliography}{dummy}
+\bibitem[Astropy Collaboration et 
+al.(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, Astronomy \& Astrophysics, 558, AA33
+
+\bibitem[Taylor(2005)]{Taylor2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+
 \end{thebibliography}

### 2016-08-12 article/unrave.tex
Added a bad abstract

@@
 \newcommand{\logg}{\log g}
-\newcommand{\Teff}{T_{\mathrm{eff}}}
+\newcommand{\teff}{T_{\mathrm{eff}}}
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}
@@
 
-\title{The \project{UNRAVE} catalog of stellar astrophysical parameters and chemical abundances}
+\title{The \project{UNRAVE} catalog}
 
@@
 
-\author{Some combination of: Hawkins, Ness, Hogg, HWR, Gilmore, Matijevic, RAVE collaboration}
+\author{Some combination of: Gerry Gilmore, Keith Hawkins, David W. Hogg, Gal Matijevic, Melissa Ness, Hans Walter-Rix, and the \project{RAVE} collaboration}
 %\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
@@
 \begin{abstract}
-% Gaia will release stuff
-% Much we want to do will involve chemistry + kinematics
-% RAVE has the largest overlap with Gaia DR1, but chemistry is unavailable for most stars
-% We perform an independent analysis of RAVE spectra w/ The Cannon, censoring, + upgrades, to derive astrophysical parameters and detailed chemical abundances.
-% We produce a catalog of up to N abundances for X stars, constituting the largest
-% catalog of chemical abundances that overlap with Gaia by a factor of X.
-
+On 14 September 2016 the \project{Gaia} mission will release positions, proper motions, and parallaxes for about two million stars.
+Most studies intending to make use of those data would benefit enormously from stellar chemical abundances.
+  The RAdial Velocity Experiment (\project{RAVE}) survey has acquired spectra for approximately 280,000 stars in common with the first \project{Gaia} data release, constituting the largest overlap of any stellar spectroscopic sample.
+Here we perform an independent analysis of the \project{RAVE} spectra using a new non-parametric version of \thecannon.
+We deliver stellar astrophysical parameters ($\teff$, $\logg$) and chemical abundances of up to 8 elements (Fe, Na, Mg, Al, C, N, Ti, Ca) for X stars.
+The typical precision in abundances is $X.XX$~dex.
+We validate our results internally from repeat visits, and externally against existing high-resolution spectroscopic surveys.
 \end{abstract}

### 2016-08-12 article/unrave.tex
Added some points for discussion

@@
 
+\newcommand{\Nstars}{483,330}
+
 %\AuthorCallLimit=10
@@
 
-\author{Some combination of: Gerry Gilmore, Keith Hawkins, David W. Hogg, Gal Matijevic, Melissa Ness, Hans Walter-Rix, and the \project{RAVE} collaboration}
+\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W.
+Hogg, Gal Matijevic, Melissa Ness, Jason Sanders, Hans Walter-Rix, and any other  \project{RAVE} people who have contributed to the data analysis}
 %\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
+
+\author{and the \project{RAVE} collaboration}
 
@@
 The RAdial Velocity Experiment (\project{RAVE}) survey has acquired spectra for approximately 280,000 stars in common with the first \project{Gaia} data release, constituting the largest overlap of any stellar spectroscopic sample.
-Here we perform an independent analysis of the \project{RAVE} spectra using a new non-parametric version of \thecannon.
-We deliver stellar astrophysical parameters ($\teff$, $\logg$) and chemical abundances of up to 8 elements (Fe, Na, Mg, Al, C, N, Ti, Ca) for X stars.
-The typical precision in abundances is $X.XX$~dex.
+Here we perform an independent analysis of the \project{RAVE} spectra using a new non-parametric implementation of \thecannon\ that employs prior probabilities on atomic line formation and stellar astrophysical parameters.
+We deliver effective temperature $\teff$, surface gravity $\logg$, and chemical abundances of up to 8 elements (Fe, Na, Mg, Al, C, N, Ti, Ca) for \Nstars\ stars.
 We validate our results internally from repeat visits, and externally against existing high-resolution spectroscopic surveys.
+The typical precision in chemical abundances is $X.XX$~dex.
 \end{abstract}
@@
 
+% Chemistry and kinematics are essential for unfolding the Galaxy.
+
+% Gaia will do this, and until Gaia releases abundances for stars from RVS spectra, we rely on ground-based surveys to provide chemical abundances.
+% This includes GES, APOGEE, etc.
+
+% We examined the overlap between different surveys and found RAVE to have the largest overlap with Gaia DR1.
+% RAVE observed from X to Y, given blah blah blah resolution, etc
+% RAVE was a radial velocity survey, so the S/N peaks at ~40
+% Typically lower than the requisite S/N ratio required for chemical abundances.
+
+\section{Data}
+\label{sec:data}
+
+% How many spectra, acquired over what time
+% Data reduction
+% Normalization
+% Flux errors.
+
 \section{Methods}
 \label{sec:methods}
+
+% Describe The Cannon briefly
+% Initial tests using a single model for all RAVE spectra
+% Complex structure led us to ivnestigate other models.
+% Non-parametric form
+% Censoring masks
+% Priors on astrophysical parameters from isochrones
 
@@
 
+% Chi-sq of the sample.
+% Identify outliers.
+% Stars with morphological classification
+
+% Internal validation:
+% --> repeat visits: are our formal errors reasonable?
+% --> bootstrap resampling of the training set?
+
+% External validation: 
+% --> RAVE Data Release 4
+% --> Gaia-ESO Survey
+% --> Asteroseismic nu_max and delta_nu available anywhere?
+% --> Furhman and other studies in Kordopatis+ 2014
+
+% Astrophysical validation:
+% --> Identify open clusters and look at their metallicity spreads
+% --> Identify globular clusters and look to see whether we capture "expected" abundance trends/spreads that are seen elsewhere in the literature
+% --> Residuals as a function of galactic position? (DIBS)
+
 \section{Conclusion}
@@
 
+% We have introduced a non-parametric version of The Cannon that uses strict priors on line formation (censoring masks), and prior probability distributions on astrophysical parameters, using isochrones.
+% We ran it on the entire RAVE spectra.
+% We deliver effective temperature, surface gravity, and abundances of up to 8 elements for \Nstars.
+Our internal and external validation tests suggest that we achieve a typical precision of X~K in $\teff$, 0.0X~dex in $\logg$, and $\approx{}X.XX$ in individual chemical abundances.
+This catalog constitutes the largest collection of stellar abundances for stars in the first \project{Gaia} data release.
+When combined with positions and 3D velocities from \project{Gaia}, the \project{UNRAVE} catalog will likely be crucial for understanding our local place in the Milky Way.
+
 \acknowledgements
-% We thank,......
-
+We thank Sven Buder (MPIA), others??.
 This research made use of: Astropy, a community-developed core Python package for
@@
 Keck foundation; the Macquarie University; the Netherlands Research School for Astronomy; the Natural Sciences and Engineering Research Council of Canada; the Slovenian Research Agency; the Swiss National Science Foundation; the Science \& Technology Facilities Council of the UK; Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg and Sydney.
-The RAVE web site is at https://www.rave-survey.org.
+The RAVE web site is https://www.rave-survey.org.
 
@@
 %\begin{figure}[p]
-%\caption{Choose two wavelengths (one continuum and one interesting)
-%  and plot some scatter plots of flux vs various parameters, and also
-%  cross-validation results for the
-%  hyper-parameters.\label{fig:onewavelength}}
+%\caption{A H-R diagram showing the training set labels.\label{fig:training-set-hrd}}
+%\end{figure}
+
+%\begin{figure}[p]
+%\caption{A H-R diagram showing the test set labels.\label{fig:test-set-hrd}}
+%\end{figure}
+
+%\begin{figure}[p]
+%\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
+%\end{figure}
+
+%\begin{figure}[p]
+%\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
 %\end{figure}

### 2016-08-12 article/unrave.tex
some shit

@@
 Most studies intending to make use of those data would benefit enormously from stellar chemical abundances.
-The RAdial Velocity Experiment (\project{RAVE}) survey has acquired spectra for approximately 280,000 stars in common with the first \project{Gaia} data release, constituting the largest overlap of any stellar spectroscopic sample.
+The RAdial Velocity Experiment (\project{RAVE}) survey has acquired spectra for approximately 290,000 stars in common with the first \project{Gaia} data release, constituting the largest overlap of any stellar spectroscopic sample.
 Here we perform an independent analysis of the \project{RAVE} spectra using a new non-parametric implementation of \thecannon\ that employs prior probabilities on atomic line formation and stellar astrophysical parameters.
-We deliver effective temperature $\teff$, surface gravity $\logg$, and chemical abundances of up to 8 elements (Fe, Na, Mg, Al, C, N, Ti, Ca) for \Nstars\ stars.
+We deliver effective temperature $\teff$, surface gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) for \Nstars\ stars.
 We validate our results internally from repeat visits, and externally against existing high-resolution spectroscopic surveys.
@@
 
+\section{Data}
+\label{sec:data}
+
+The first \project{Gaia} data release includes positions, proper motions, and parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.
+After cross-matching all major stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, \project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST} \citep{lamost}, and \project{RAVE} \citep{rave}}, we found the \project{RAVE} survey to have the largest overlap with the first \project{Gaia} data release: 292,036 stars.
+We then used the \project{Gaia} universe model to estimate the precision in parallax and proper motions available in \project{Gaia} DR1 for every star in these cross-matches.
+Comparing the expected precision to what is currently available, we further found that the \project{RAVE} survey would benefit most from \project{Gaia} DR1.
+The distances of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to improve with the first \project{Gaia} data release, and 47\% are likely to have better proper motions (137,211 stars).
+
+This motivated us to examine what chemical abundance information is currently available for \project{RAVE}, and to evaluate whether we could contribute additional abundances.
+
 % We examined the overlap between different surveys and found RAVE to have the largest overlap with Gaia DR1.
@@
 % Typically lower than the requisite S/N ratio required for chemical abundances.
-
-\section{Data}
-\label{sec:data}
 
@@
 % Priors on astrophysical parameters from isochrones
+
+% Include velocities at test time?
+% Train on fibre number? Fit for spectral resolution at test time?
 

### 2016-08-12 article/figures/rave_io.py
Updated DR4 comparison figure and RAVE IO

@@ -1,13 +1,9 @@
 
+import os
 import numpy as np
 from astropy.table import Table
 
-rave_kordopatis_dr4 = Table.read("../../RAVE-DR4.fits")
-rave_cannon_dr1 = Table.read("../../results/rave-tgas-v3a+b-xm-dr4.fits")
+DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
 
-w1 = np.exp(-0.5 * rave_cannon_dr1["chi-sq_1"])
-w2 = np.exp(-0.5 * rave_cannon_dr1["chi-sq_2"])
-for label in ("TEFF", "LOGG", "FE_H"):
-    rave_cannon_dr1[label] \
-        = (rave_cannon_dr1["{}_1".format(label)] * w1 \
-            + rave_cannon_dr1["{}_2".format(label)] * w2)/(w1 + w2)
+rave_kordopatis_dr4 = Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
+rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))

### 2016-08-12 article/unrave.tex

@@
 
-%\begin{figure}[p]
-%\caption{A H-R diagram showing the training set labels.\label{fig:training-set-hrd}}
-%\end{figure}
+\begin{figure}[p]
+\caption{A H-R diagram showing the training set labels.\label{fig:training-set-hrd}}
+\end{figure}
 
-%\begin{figure}[p]
-%\caption{A H-R diagram showing the test set labels.\label{fig:test-set-hrd}}
-%\end{figure}
+\begin{figure}[p]
+\caption{A H-R diagram showing the test set labels.\label{fig:test-set-hrd}}
+\end{figure}
 
-%\begin{figure}[p]
-%\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
-%\end{figure}
+\begin{figure}[p]
+\caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:test-set-repeats}}
+\end{figure}
 
-%\begin{figure}[p]
-%\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
-%\end{figure}
+\begin{figure}[p]
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
+\end{figure}
+
+\begin{figure}[p]
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:dr4-comparison}}
+\end{figure}
+
+\begin{figure}[p]
+\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{kordopatis_2014}}
+\end{figure}
 

### 2016-08-12 article/figures/plot_dr4_comparison.py
Finish script that makes comparisons to DR4 stellar parameters #1

@@ -25,8 +25,24 @@ else:
 QC = combined_table["OK"] * (combined_table["QK"] == 0)
 combined_table = combined_table[QC]
 
+N_bins = 50
 
-K, factor = (3, 3.5)
+all_columns = [
+    # RAVE label, DR4 label
+    ("TEFF", "TeffK"),
+    ("LOGG", "loggK"),
+    ("FE_H", "__M_H_K"),
+]
+label_limits = {
+    "TEFF": (3000, 8000),
+    "LOGG": (0, 5),
+    "FE_H": (-2, 0.75)
+}
+latex_labels = (r"$T_{\rm eff}$", r"$\log{g}$", r"$[{\rm Fe/H}]$")
+
+
+
+K, factor = (len(all_columns), 3.5)
 lbdim = 0.2 * factor
 trdim = 0.1 * factor
 whspace = 0.05
@@ -40,26 +56,9 @@ fig.subplots_adjust(
     left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
     top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
 
-N_bins = 50
+for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labels)):
 
-labels = [
-    # RAVE label, DR4 label
-    ("TEFF", "TeffK"),
-    ("LOGG", "loggK"),
-    ("FE_H", "__M_H_K"),
-]
-label_limits = {
-    "TEFF": (3000, 8000),
-    "LOGG": (0, 5),
-    "FE_H": (-2, 0.75)
-}
-latex_labels = (r"$T_{\rm eff}$", r"$\log{g}$", r"$[{\rm Fe/H}]$")
-
-for i, ((cannon_label, dr4_label), latex_label) \
-in enumerate(zip(labels, latex_labels)):
-
-
-    ax = axes[i]
+    cannon_label, dr4_label = columns
 
     x = combined_table[cannon_label]._data
     y = combined_table[dr4_label]._data
@@ -76,8 +75,11 @@ in enumerate(zip(labels, latex_labels)):
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_xlabel("{} (UNRAVE)".format(latex_label))
[…]

### 2016-08-15 article/unrave.tex
and MS

@@
 \author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W.
-Hogg, Gal Matijevic, Melissa Ness, Jason Sanders, Hans Walter-Rix, and any other  \project{RAVE} people who have contributed to the data analysis}
+Hogg, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any other \project{RAVE} people who have contributed to the data analysis}
 %\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}

### 2016-08-15 article/figures/plot_literature_comparisons.py
Typo; Added script to plot results from LOOCV #1

@@ -6,7 +6,7 @@ from matplotlib.colors import LogNorm
 from matplotlib.ticker import MaxNLocator
 
 try:
-    combined_tabledata_table
+    data_table
 
 except NameError:
     from rave_io import (rave_cannon_dr1, literature_pastel)

### 2016-08-15 article/figures/plot_loocv_results.py

@@ -1,0 +1,74 @@
+
+
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+
+import AnniesLasso as tc
+
+try:
+    cannon_model_path
+
+except NameError:
+    from rave_io import (cannon_loocv_results, cannon_model_path, )
+
+else:
+    print("Warning: Using pre-loaded data!")
+
+
+label_limits = {
+    "TEFF": (3000, 8000),
+    "LOGG": (0, 5),
+    "FE_H": (-2, 0.75)
+}
+
+latex_labels = {
+    "TEFF": r"$T_{\rm eff}$",
+    "LOGG": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$"
+}
+
+
+
+model = tc.load_model(cannon_model_path)
+
+K = len(model.vectorizer.label_names)
+
+N = (1, 5)
+
+
+assert K <= (N[0] * N[1])
+
+fig, axes = plt.subplots(N[0], N[1])
+
+c = model.labelled_set["SNRK"]
+
+for i, (ax, label_name) in enumerate(zip(axes, model.vectorizer.label_names)):
+
+
+    x = model.labelled_set[label_name]
+    #xerr = model.labelled_set["E_{}".format(label_name)]
+    y = cannon_loocv_results[label_name]
+    #yerr = cannon_loocv_results["E_{}".format(label_name)]
+
+    ax.scatter(x, y, c=c, cmap="viridis", s=50)
+
+    _ = limits[label_name]
+    ax.plot(_, _, c="#666666", zorder=-1)
+    ax.set_xlim(_)
+    ax.set_ylim(_)
+    ax.xaxis.set_major_locator(MaxNLocator(6))
+    ax.yaxis.set_major_locator(MaxNLocator(6))
+    ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm Labelled})$"]))
+    ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm LOOCV})$"]))
+
+    [_.set_rotation(30) for _ in ax.get_xticklabels()]
+    [_.set_rotation(30) for _ in ax.get_yticklabels()]
+
+
+
+fig.tight_layout()
+
+fig.savefig("loocv-results.pdf", dpi=300)
+fig.savefig("loocv-results.png", dpi=300)

### 2016-08-15 article/figures/plot_ges_comparison.py
Added script to plot GES comparison #1

@@ -1,0 +1,88 @@
+
+
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+
+try:
+    data_table
+
+except NameError:
+    from rave_io import (rave_cannon_dr1, ges_idr4)
+
+    from astropy.table import join
+
+    data_table = join(rave_cannon_dr1, ges_idr4, keys=("Name", ))
+
+else:
+    print("Warning: Using pre-loaded data!")
+
+
+# Plot labels against literature.
+
+label_names = ("TEFF_1", "LOGG_2", "FE_H")
+ges_label_names = {
+    "TEFF_1": "TEFF_2",
+    "LOGG_1": "LOGG_2",
+    "FE_H": "FEH",
+}
+
+label_limits = {
+    "TEFF_1": (3000, 8000),
+    "LOGG_1": (0, 5),
+    "FE_H": (-2, 0.75)
+}
+
+latex_labels = {
+    "TEFF_1": r"$T_{\rm eff}$",
+    "LOGG_1": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$"
+}
+
+
+# Exclude ones we think are bad.
+ok = data_table["OK"]
+
+data_table = data_table[ok]
+
+K = len(label_names)
+factor = 3.5
+lbdim = 0.2 * factor
+trdim = 0.1 * factor
+whspace = 0.05
+yspace = factor
+xspace = factor * K + factor * (K - 1) * whspace + lbdim * (K - 1)
+xdim = lbdim + xspace + trdim
+ydim = lbdim + yspace + trdim
+
+fig, axes = plt.subplots(1, K, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+for i, (ax, label_name) in enumerate(zip(axes, label_names)):
+
+    x = data_table[label_name]
+    #xerr
+    y = data_table[ges_label_names[label_name]]
+    # yerr
+
+    ax.scatter(x, y, c=data_table["SNRK"], alpha=0.75, s=50, 
+        linewidths=0.5, edgecolors="#000000")
+
+    ax.set_xlim(limits[label_name])
[…]

### 2016-08-15 article/figures/rave_io.py

@@ -7,5 +7,6 @@ DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
 
 rave_kordopatis_dr4 = Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
 rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))
+ges_idr4 = Table.read(os.path.join(DATA_PATH, "literature-ges-idr4-xm.fits"))
 
 literature_pastel = Table.read(os.path.join(DATA_PATH, "literature-pastel-database-xm.fits"))

### 2016-08-15 article/unrave.tex
Some stuff about data, etc

@@
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}
-
+\newcommand{\stub}[1]{\textbf{#1}}
 \newcommand{\Nstars}{483,330}
@@
 
-% Chemistry and kinematics are essential for unfolding the Galaxy.
+\stub{Chemistry and kinematics are essential for unfolding the Galaxy, etc.}
 
-% Gaia will do this, and until Gaia releases abundances for stars from RVS spectra, we rely on ground-based surveys to provide chemical abundances.
-% This includes GES, APOGEE, etc.
+\stub{
+Gaia will do this, and until Gaia releases abundances for stars from RVS 
+spectra, we rely on ground-based surveys to provide chemical abundances.
+This includes GES, APOGEE, etc.
+}
+
+The first \project{Gaia} data release includes positions, proper motions, and 
+parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.  
+After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
+we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, 
+\project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST}
+\citep{lamost}, and \project{RAVE} \citep{rave}}, we found the RAdial Velocity
+Experiment (hereafter \rave) 
+survey to have the largest overlap with the first \project{Gaia} data release: 
+292,036 stars.  We then used the \project{Gaia} universe model \citep{GUMS} to 
+estimate the precision in parallax and proper motions that will be available in 
+\project{Gaia} DR1 for every star in these cross-matches.  Comparing the 
+expected precision to what is currently available, we further found that the 
+\project{RAVE} survey will benefit most from \project{Gaia} DR1.  The distances
+of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars)
+are expected to improve with the first \project{Gaia} data release, and 47\% of
+stars likely to have better proper motions (137,211 stars).  This motivated us to
+examine what chemical abundance information is currently available for \rave, and
+to evaluate whether we could contribute to this existing set.
+
+We briefly describe the \rave\ data in Section \ref{sec:data}, before explaining
+the model we employ in Section \ref{sec:model}.  In Section \ref{sec:validation}
+we outline a number of validation experiments, including: internal sanity checks,
+comparisons with literature samples, and investigations to ensure our results
+are consistent with expectations from astrophysics.  We provide concluding
+remarks in Section \ref{sec:conclusion}, as well as a description of how to
+access the data.
 
@@
 
-The first \project{Gaia} data release includes positions, proper motions, and parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.
-After cross-matching all major stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, \project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST} \citep{lamost}, and \project{RAVE} \citep{rave}}, we found the \project{RAVE} survey to have the largest overlap with the first \project{Gaia} data release: 292,036 stars.
-We then used the \project{Gaia} universe model to estimate the precision in parallax and proper motions available in \project{Gaia} DR1 for every star in these cross-matches.
-Comparing the expected precision to what is currently available, we further found that the \project{RAVE} survey would benefit most from \project{Gaia} DR1.
-The distances of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to improve with the first \project{Gaia} data release, and 47\% are likely to have better proper motions (137,211 stars).
+The \rave\ survey started observations using the 6dF system on the 1.2~m UK 
+Schmidt telescope at the Australian Astronomical Observatory\footnote{Formerly 
+the Anglo-Australian Observatory}.  The large 5.7~degree field-of-view and 
+robotic fibre positioner made for very efficient observing: up to 150 targets 
+could be simultaneously obtained.  When observations concluded in April 2013, a 
+total of 574,630 spectra had been collected of 483,330 unique objects.
 
-This motivated us to examine what chemical abundance information is currently available for \project{RAVE}, and to evaluate whether we could contribute additional abundances.
+RAVE is a magnitude-limited survey, targeting stars with apparent $I$-band
+magnitudes between $9 < I < 13$.  The $I$ band was used as the selection
+magnitude because it overlaps with the wavelength range that \rave\ operates in:
+8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared triplet,
+strong stellar absorption lines that are visible even in very low S/N data 
+(i.e., they are very useful for radial velocity determination).
 
-% We examined the overlap between different surveys and found RAVE to have the largest overlap with Gaia DR1.
-% RAVE observed from X to Y, given blah blah blah resolution, etc
-% RAVE was a radial velocity survey, so the S/N peaks at ~40
-% Typically lower than the requisite S/N ratio required for chemical abundances.
+% When did RAVE observe?
+% Where from
 
-% How many spectra, acquired over what time
+% target selection.
+% wavelength range
+
+% How many data releases
+% What was in previous data releases?
+
 % Data reduction
 % Normalization
-% Flux errors.
+% Flux errors
 
-\section{Methods}
-\label{sec:methods}
+As the name implies, the \rave\ survey was principally designed to obtain radial
+velocity measurements for up to $10^6$ stars.  Detailed chemical abundances were
+a secondary science goal of the survey.  For this reason the distribution
+of S/N ratios for \rave\ spectra is considerably lower than stellar 
+spectroscopic surveys where chemical abundances are the defining objective.  The 
+\rave\ spectra cover $\approx38$~nm in the infrared with an effective resolution
+$\mathcal{R} \approx 7500$, and the S/N peaks at $\approx$50~per~pixel$^{-1}$.
+For comparison, the \project{GALAH} survey \citep{DaSilva_2015} --- which was 
+specifically constructed for detailed chemical abundance analyses --- includes a
+wavelength range about 2.5 times larger at resolution $\mathcal{R} \approx 28,000$,
+and target for S/N $\gtrsim100$.
+
+Despite the relatively low resolution and S/N of the spectra, the \rave\ data
+releases have demonstrated the team's enormous success in deriving radial 
+velocities, stellar astrophysical parameters ($\teff$, $\logg$), as well as 
+individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,
+Kordopatis_2014, Kunder_2016}.  The data used for this analysis are the result
+of a thorough reprocessing of all survey data.  This includes a detailed
+re-reduction of the original data frames, an updated continuum-normalization
+procedure, as well as an improvement on the quantification of flux errors in
+each pixel.  At the end of this processing for each survey observation we are 
+provided with: wavelengths, continuum-normalized fluxes, as well as 
+$1\sigma$ uncertainties in flux values.  We refer the reader to the official 
+fifth data release paper of the \rave\ survey, as presented by
+\citet{Kunder_2016}, for more details of this reprocessing.
+
+We chose not to re-normalize the spectra, as our tests proved to us that the
+normalization procedure outlined in \citet{Kunder_2016} is sufficient for our
+analysis.  We re-sampled all spectra onto a common rest-wavelength map.
+Depending on the fibre used and stellar radial velocity, the rest-frame wavelength
+range varied for each star.  Given that fluxes were unavailable in the edge
+pixels for most stars, we excluded pixels outside of the rest wavelength range
+$8423.2{\rm \AA} < \lambda < 8777.6{\rm \AA}$.
+
+\section{Model}
+\label{sec:model}
+
+Given the relatively low S/N ratios in the \rave\ spectra, we chose to adopt a 
+data-driven model (rather than a physics-driven model) for this analysis.
 
 % Describe The Cannon briefly
+% training set from APOGEE
+
 % Initial tests using a single model for all RAVE spectra
-% Complex structure led us to ivnestigate other models.
-% Non-parametric form
 % Censoring masks
@@
 % Identify outliers.
+% Convex hull?
+
 % Stars with morphological classification
@@
 % --> Asteroseismic nu_max and delta_nu available anywhere?
-% --> Furhman and other studies in Kordopatis+ 2014
+% --> Literature studies presented in Kordopatis+ 2014
 

### 2016-08-15 article/unrave.tex
Formatting, etc

@@
 %% Copyright 2016 the authors.  All rights reserved.
+
+% Look for \stub{}s. If you can expand it, do so!
 
@@
 \begin{abstract}
-On 14 September 2016 the \project{Gaia} mission will release positions, proper motions, and parallaxes for about two million stars.
-Most studies intending to make use of those data would benefit enormously from stellar chemical abundances.
-The RAdial Velocity Experiment (\project{RAVE}) survey has acquired spectra for approximately 290,000 stars in common with the first \project{Gaia} data release, constituting the largest overlap of any stellar spectroscopic sample.
-Here we perform an independent analysis of the \project{RAVE} spectra using a new non-parametric implementation of \thecannon\ that employs prior probabilities on atomic line formation and stellar astrophysical parameters.
-We deliver effective temperature $\teff$, surface gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) for \Nstars\ stars.
-We validate our results internally from repeat visits, and externally against existing high-resolution spectroscopic surveys.
-The typical precision in chemical abundances is $X.XX$~dex.
+On 14 September 2016 the \project{Gaia} mission will release positions, proper 
+motions, and parallaxes for about two million stars.  Most studies intending to
+make use of those data would benefit enormously by coupling that astrometry with
+stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
+survey has acquired spectra for approximately 290,000 stars in common with the 
+first \project{Gaia} data release, constituting the largest overlap of any 
+stellar spectroscopic sample.  Here we perform an independent analysis of the 
+\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
+that employs prior probabilities on atomic line formation and stellar 
+astrophysical parameters.  We deliver effective temperature $\teff$, surface 
+gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
+Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
+repeat visits, and externally against existing high-resolution spectroscopic 
+surveys.  The typical precision in chemical abundances is $X.XX$~dex.  
 \end{abstract}
@@
 (i.e., they are very useful for radial velocity determination).
-
-% When did RAVE observe?
-% Where from
-
-% target selection.
-% wavelength range
-
-% How many data releases
-% What was in previous data releases?
-
-% Data reduction
-% Normalization
-% Flux errors
 

### 2016-08-15 article/unrave.tex
Add citations

@@
 The first \project{Gaia} data release includes positions, proper motions, and 
-parallaxes for approximately two million stars in the Tycho-2 catalog \citep{tycho2}.  
+parallaxes for approximately two million stars in the Tycho-2 catalog \citep{Hog_2000}.  
 After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
-we cross-matched the Tycho-2 catalog against {APOGEE} \citep{apogee}, 
-\project{GALAH} \citep{galah}, \project{Gaia-ESO} \citep{ges}, \project{LAMOST}
-\citep{lamost}, and \project{RAVE} \citep{rave}}, we found the RAdial Velocity
+we cross-matched the Tycho-2 catalog against {APOGEE} \citep{Zasowski_2013}, 
+\project{GALAH} \citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, \project{LAMOST}
+\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006}}, we found the RAdial Velocity
 Experiment (hereafter \rave) 
 survey to have the largest overlap with the first \project{Gaia} data release: 
-292,036 stars.  We then used the \project{Gaia} universe model \citep{GUMS} to 
+292,036 stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
 estimate the precision in parallax and proper motions that will be available in 
@@
 $\mathcal{R} \approx 7500$, and the S/N peaks at $\approx$50~per~pixel$^{-1}$.
-For comparison, the \project{GALAH} survey \citep{DaSilva_2015} --- which was 
+For comparison, the \project{GALAH} survey \citep{DeSilva_2015} --- which was 
 specifically constructed for detailed chemical abundance analyses --- includes a
@@
 individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,
-Kordopatis_2014, Kunder_2016}.  The data used for this analysis are the result
+Kordopatis_2013, Kunder_2016}.  The data used for this analysis are the result
 of a thorough reprocessing of all survey data.  This includes a detailed
@@
 pixels for most stars, we excluded pixels outside of the rest wavelength range
-$8423.2{\rm \AA} < \lambda < 8777.6{\rm \AA}$.
+$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.
 
@@
 
+\bibitem[Cui et al.(2012)]{Cui_2012} Cui, X.-Q., Zhao, Y.-H., Chu, Y.-Q., et al.\ 2012, Research in Astronomy and Astrophysics, 12, 1197
+
+\bibitem[De Silva et al.(2015)]{DeSilva_2015} De Silva, G.~M., Freeman, K.~C., Bland-Hawthorn, J., et al.\ 2015, \mnras, 449, 2604
+
+\bibitem[Gilmore et al.(2012)]{Gilmore_2012} Gilmore, G., Randich, S., Asplund, M., et al.\ 2012, The Messenger, 147, 25
+
+\bibitem[H{\o}g et al.(2000)]{Hog_2000} H{\o}g, E., Fabricius, C., Makarov, V.~V., et al.\ 2000, \aap, 355, L27
+
+\bibitem[Kordopatis et al.(2013)]{Kordopatis_2013} Kordopatis, G., Gilmore, G., Steinmetz, M., et al.\ 2013, \aj, 146, 134
+
+\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, in preparation
+
+\bibitem[Ness et al.(2015)]{Ness_2015} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16
+
+\bibitem[Randich et al.(2013)]{Randich_2013} Randich, S., Gilmore, G., \& Gaia-ESO Consortium 2013, The Messenger, 154, 47
+
+\bibitem[Robin et al.(2012)]{Robin_2012} Robin, A.~C., Luri, X., Reyl{\'e}, C., et al.\ 2012, \aap, 543, A100
+
+\bibitem[Siebert et al.(2011)]{Siebert_2011} Siebert, A., Williams, M.~E.~K., Siviero, A., et al.\ 2011, \aj, 141, 187
+
+\bibitem[Steinmetz et al.(2006)]{Steinmetz_2006} Steinmetz, M., Zwitter, T., Siebert, A., et al.\ 2006, \aj, 132, 1645
+
 \bibitem[Taylor(2005)]{Taylor2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+
+\bibitem[Zasowski et al.(2013)]{Zasowski_2013} Zasowski, G., Johnson, J.~A., Frinchaboy, P.~M., et al.\ 2013, \aj, 146, 81
+
+\bibitem[Zwitter et al.(2008)]{Zwitter_2008} Zwitter, T., Siebert, A., Munari, U., et al.\ 2008, \aj, 136, 421
 
@@
 \begin{figure}[p]
-\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{kordopatis_2014}}
+\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2013}}
 \end{figure}

### 2016-08-15 article/unrave.tex
Minor things

@@
 
-\title{The \project{UNRAVE} catalog}
+\title{The \project{unRAVE} catalog}
 
@@
 repeat visits, and externally against existing high-resolution spectroscopic 
-surveys.  The typical precision in chemical abundances is $X.XX$~dex.  
+surveys.  The typical precision in chemical abundances is X.XX~dex.  
 \end{abstract}
@@
 As the name implies, the \rave\ survey was principally designed to obtain radial
-velocity measurements for up to $10^6$ stars.  Detailed chemical abundances were
-a secondary science goal of the survey.  For this reason the distribution
-of S/N ratios for \rave\ spectra is considerably lower than stellar 
-spectroscopic surveys where chemical abundances are the defining objective.  The 
+velocity measurements.  Detailed chemical abundances were a secondary science 
+goal of the survey.  For this reason the distribution of S/N ratios for \rave\ 
+spectra is considerably lower than stellar spectroscopic surveys where chemical
+abundances are the defining objective.  The 
 \rave\ spectra cover $\approx38$~nm in the infrared with an effective resolution
@@
 
+% Inserting 2MASS photometry as fake pixels
+
 % Initial tests using a single model for all RAVE spectra
@@
 % Identify outliers.
-% Convex hull?
+% Convex hull? Distance to plane?
 
@@
 Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
- and \project{TOPCAT} \citep{Taylor2005}.
+and \project{TOPCAT} \citep{Taylor_2005}.
 This work was partly supported by the European Union FP7 programme through ERC 
 grant number 320360.
-Funding for RAVE has been provided by: the Australian Astronomical Observatory; the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National University; the Australian Research Council; the French National Research Agency; the German Research Foundation (SPP 1177 and SFB 881); the European Research Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at Padova; The Johns Hopkins University; the National Science Foundation of the USA (AST-0908326); the W.
-M.
-Keck foundation; the Macquarie University; the Netherlands Research School for Astronomy; the Natural Sciences and Engineering Research Council of Canada; the Slovenian Research Agency; the Swiss National Science Foundation; the Science \& Technology Facilities Council of the UK; Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg and Sydney.
-The RAVE web site is https://www.rave-survey.org.
+Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
+the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
+University; the Australian Research Council; the French National Research Agency;
+the German Research Foundation (SPP 1177 and SFB 881); the European Research 
+Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
+Padova; The Johns Hopkins University; the National Science Foundation of the USA
+(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
+Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
+Research Council of Canada; the Slovenian Research Agency; the Swiss National 
+Science Foundation; the Science \& Technology Facilities Council of the UK; 
+Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
+and Sydney. The RAVE web site is https://www.rave-survey.org.
 
@@
 
-\bibitem[Taylor(2005)]{Taylor2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+\bibitem[Taylor(2005)]{Taylor_2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
 
@@
 
-\begin{figure}[p]
+\begin{figure*}[p]
 \caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:test-set-repeats}}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[p]
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:dr4-comparison}}
-\end{figure}
+\begin{figure*}[p]
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:rave-dr4-comparison}}
+\end{figure*}
 
-\begin{figure}[p]
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:dr4-comparison}}
-\end{figure}
+\begin{figure*}[p]
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:ges-dr4-comparison}}
+\end{figure*}
 
-\begin{figure}[p]
+\begin{figure*}[p]
 \caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2013}}
-\end{figure}
+\end{figure*}
 

### 2016-08-16 article/unrave.tex
I should really spent *all* of that time writing..

@@
 
-% Look for \stub{}s. If you can expand it, do so!
+%% TODO:
+%% Co-authors: Look for \stub{}s. If you can expand it, do so!
 
-\documentclass[preprint2,trackchanges]{aastex}
+\documentclass[preprint,trackchanges]{aastex}
 
@@
 \usepackage{bm}
+
+% This section can be removed at submission.
+% ------------------------------------------
+\usepackage{color}
+\usepackage{datenumber}
+
+\newcounter{dateone}
+\newcounter{datetwo}
+
+\newcommand{\difftoday}[3]{%
+  \setmydatenumber{dateone}{\the\year}{\the\month}{\the\day}%
+  \setmydatenumber{datetwo}{#1}{#2}{#3}%
+  \addtocounter{datetwo}{-\thedateone}%
+  \the\numexpr(\thedatetwo)\relax\space days
+}
+% ------------------------------------------
 
@@
 \begin{document}
+% Remove at submission:
+\slugcomment{{\color{red} \textbf{To appear on arXiv on 9 September 2016 (\difftoday{2016}{9}{7}away)}}}
 
-\title{The \project{unRAVE} catalog}
+\title{The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies}
 
@@
 \author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W.
-Hogg, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any other \project{RAVE} people who have contributed to the data analysis}
+Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any \project{RAVE} members who have contributed to the data analysis in this study}
 %\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
@@
 surveys.  The typical precision in chemical abundances is X.XX~dex.  
+% For stars not in TGAS, we calculate distances
 \end{abstract}
@@
 \label{sec:introduction}
+
+The Milky Way is considered to be the best laboratory for understanding galaxy formation and evolution.
+This premise depends on the ability to precisely measure the position, motion, and chemistry of (many) individual stars, and to infer the structure, kinematics, and chemical enrichment of the Galaxy from those data \citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,Casey_2014a,Casey_2014b}.
+% include references on: disk chemistry, halo structures
+However, these quantities are not known for even 1\% of stars in the Milky Way.
+Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,Madler_2016}, proper motions are frequently plagued by unquantified systematics from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and stellar spectroscopists frequently report significantly different chemical abundance patterns \citep{Smiljanic_2014}.
+If these measurements were known with high fidelity then \stub{xyz}.
+
+% These scientific questions were only some of them that prompted Gaia.
+% What will Gaia provide eventually.
+% In the first data release what do we get.
+% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
+% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry
+
+% Gaia will produce a very clear snapshot, producing astrometry space motions
+
+% but if we want to infer the past then we need additional information about how the galaxy has been enriched. so we need chemical abundances.
+% the number of stars that overlap with Gaia DR1 between stellar surveys is very different, due to targeting, etc
+% and those stars are not on the same metallicity scale
+% therefore a large-scale catalog of homogenous abundances that overlap with Gaia DR1 is currently lacking
 
@@
 % --> Gaia-ESO Survey
-% --> Asteroseismic nu_max and delta_nu available anywhere?
+% --> APOGEE / EPIC (whichever was not used)
 % --> Literature studies presented in Kordopatis+ 2014
@@
 \acknowledgements
-We thank Sven Buder (MPIA), others??.
+We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), others??.
 This research made use of: Astropy, a community-developed core Python package for
@@
 
+\bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88
+
+\bibitem[Casey et al.(2013)]{Casey_2013} Casey, A.~R., Da Costa, G., Keller, S.~C., \& Maunder, E.\ 2013, \apj, 764, 39
+
+\bibitem[Casey et al.(2014a)]{Casey_2014a} Casey, A.~R., Keller, S.~C., Da Costa, G., Frebel, A., \& Maunder, E.\ 2014, \apj, 784, 19
+
+\bibitem[Casey et al.(2014b)]{Casey_2014b} Casey, A.~R., Keller, S.~C., Alves-Brito, A., et al.\ 2014, \mnras, 443, 828
+
+\bibitem[Casey \& Schlaufman(2015)]{Casey_Schlaufman_2015} Casey, A.~R., \& Schlaufman, K.~C.\ 2015, \apj, 809, 110
+
+\bibitem[Casey et al.(2016)]{Casey_2016} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
+
 \bibitem[Cui et al.(2012)]{Cui_2012} Cui, X.-Q., Zhao, Y.-H., Chu, Y.-Q., et al.\ 2012, Research in Astronomy and Astrophysics, 12, 1197
@@
 \bibitem[De Silva et al.(2015)]{DeSilva_2015} De Silva, G.~M., Freeman, K.~C., Bland-Hawthorn, J., et al.\ 2015, \mnras, 449, 2604
+
+\bibitem[Deason et al.(2011)]{Deason_2011} Deason, A.~J., Belokurov, V., \& Evans, N.~W.\ 2011, \mnras, 416, 2903
 
@@
 
+\bibitem[Jofr{\'e} et al.(2015)]{Jofre_2015} Jofr{\'e}, P., M{\"a}dler, T., Gilmore, G., et al.\ 2015, \mnras, 453, 1428
+
 \bibitem[Kordopatis et al.(2013)]{Kordopatis_2013} Kordopatis, G., Gilmore, G., Steinmetz, M., et al.\ 2013, \aj, 146, 134
@@
 \bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, in preparation
+
+\bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, arXiv:1606.03015
+
+\bibitem[Ness et al.(2012)]{Ness_2012} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2012, \apj, 756, 22
+
+\bibitem[Ness et al.(2013a)]{Ness_2013a} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 430, 836
+
+\bibitem[Ness et al.(2013b)]{Ness_2013b} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 432, 2092
 
@@
 
+\bibitem[Smiljanic et al.(2014)]{Smiljanic_2014} Smiljanic, R., Korn, A.~J., Bergemann, M., et al.\ 2014, \aap, 570, A122
+
 \bibitem[Steinmetz et al.(2006)]{Steinmetz_2006} Steinmetz, M., Zwitter, T., Siebert, A., et al.\ 2006, \aj, 132, 1645
@@
 \bibitem[Taylor(2005)]{Taylor_2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+
+\bibitem[van Leeuwen(2007)]{van_Leeuwen_2007} van Leeuwen, F.\ 2007, \aap, 474, 653
 

### 2016-08-16 article/unrave.tex
Minor edits

@@
 stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
-survey has acquired spectra for approximately 290,000 stars in common with the 
+survey has acquired spectra for 292,036 stars in common with the 
 first \project{Gaia} data release, constituting the largest overlap of any 
@@
 surveys.  The typical precision in chemical abundances is X.XX~dex.  
-% For stars not in TGAS, we calculate distances
+We provide improved distances by isochrone-fitting for \project{RAVE} stars
+where astrometric solutions are not available in the first \project{Gaia}
+data release.
 \end{abstract}
@@
 
-% These scientific questions were only some of them that prompted Gaia.
+% Introducing Gaia
+% Gaia will produce a very clear snapshot, producing astrometry space motions
 % What will Gaia provide eventually.
@@
 
-% Gaia will produce a very clear snapshot, producing astrometry space motions
-
-% but if we want to infer the past then we need additional information about how the galaxy has been enriched. so we need chemical abundances.
-% the number of stars that overlap with Gaia DR1 between stellar surveys is very different, due to targeting, etc
-% and those stars are not on the same metallicity scale
-% therefore a large-scale catalog of homogenous abundances that overlap with Gaia DR1 is currently lacking
-
-\stub{Chemistry and kinematics are essential for unfolding the Galaxy, etc.}
-
-\stub{
-Gaia will do this, and until Gaia releases abundances for stars from RVS 
-spectra, we rely on ground-based surveys to provide chemical abundances.
-This includes GES, APOGEE, etc.
-}
-
-The first \project{Gaia} data release includes positions, proper motions, and 
-parallaxes for approximately two million stars in the Tycho-2 catalog \citep{Hog_2000}.  
+The first \project{Gaia} data release will include positions, proper motions, and 
+parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} catalog \citep{Michalik_2015a,Michalik_2015b}.  
 After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
-we cross-matched the Tycho-2 catalog against {APOGEE} \citep{Zasowski_2013}, 
+we cross-matched the Tycho-2 catalog against the \project{APOGEE} \citep{Zasowski_2013}, 
 \project{GALAH} \citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, \project{LAMOST}
-\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006}}, we found the RAdial Velocity
+\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} surveys}, we found the RAdial Velocity
 Experiment (hereafter \rave) 
@@
 
+\bibitem[Michalik et al.(2015a)]{Michalik_2015a} Michalik, D., Lindegren, L., \& Hobbs, D.\ 2015, \aap, 574, A115
+
+\bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
+
 \bibitem[Ness et al.(2012)]{Ness_2012} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2012, \apj, 756, 22

### 2016-08-16 article/unrave.tex
Wrap wrap wrap

@@
 stellar spectroscopic sample.  Here we perform an independent analysis of the 
-\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
-that employs prior probabilities on atomic line formation and stellar 
+\project{RAVE} spectra using an implementation of \thecannon\ 
+that uses prior probabilities on atomic line formation and stellar 
 astrophysical parameters.  We deliver effective temperature $\teff$, surface 
@@
 
-The Milky Way is considered to be the best laboratory for understanding galaxy formation and evolution.
-This premise depends on the ability to precisely measure the position, motion, and chemistry of (many) individual stars, and to infer the structure, kinematics, and chemical enrichment of the Galaxy from those data \citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,Casey_2014a,Casey_2014b}.
-% include references on: disk chemistry, halo structures
+The Milky Way is considered to be the best laboratory for understanding galaxy
+formation and evolution.  This premise hinges on the ability to precisely measure 
+the position, motion, and chemistry of (many) individual stars, and to infer the 
+structure, kinematics, and chemical enrichment of the Galaxy from those data 
+\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
+Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
 However, these quantities are not known for even 1\% of stars in the Milky Way.  
-Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,Madler_2016}, proper motions are frequently plagued by unquantified systematics from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and stellar spectroscopists frequently report significantly different chemical abundance patterns \citep{Smiljanic_2014}.
-If these measurements were known with high fidelity then \stub{xyz}.
+Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
+Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
+from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
+stellar spectroscopists frequently report significantly different chemical 
+abundance patterns \citep{Smiljanic_2014}.  Countless long-standing scientific
+questions could be definitively answered if these measurements were known with
+high fidelity: \stub{x? \citep{someone}; y; and z, to list just a few}.
 
@@
 The first \project{Gaia} data release will include positions, proper motions, and 
-parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} catalog \citep{Michalik_2015a,Michalik_2015b}.  
-After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
-we cross-matched the Tycho-2 catalog against the \project{APOGEE} \citep{Zasowski_2013}, 
-\project{GALAH} \citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, \project{LAMOST}
-\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} surveys}, we found the RAdial Velocity
-Experiment (hereafter \rave) 
-survey to have the largest overlap with the first \project{Gaia} data release: 
-292,036 stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
+parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
+catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
+stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
+catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{GALAH} 
+\citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, 
+\project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
+surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
+have the largest overlap with the first \project{Gaia} data release: 292,036 
+stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
 estimate the precision in parallax and proper motions that will be available in 
-\project{Gaia} DR1 for every star in these cross-matches.  Comparing the 
-expected precision to what is currently available, we further found that the 
-\project{RAVE} survey will benefit most from \project{Gaia} DR1.  The distances
-of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars)
-are expected to improve with the first \project{Gaia} data release, and 47\% of
-stars likely to have better proper motions (137,211 stars).  This motivated us to
-examine what chemical abundance information is currently available for \rave, and
-to evaluate whether we could contribute to this existing set.
+\project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected
+precision to what is currently available, we further found that the \project{RAVE}
+survey will benefit most from \project{Gaia} DR1.  The distances of 63\% of the 
+\project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to
+improve with the first \project{Gaia} data release, and 47\% of stars likely to
+have better proper motions (137,211 stars).  This motivated us to examine what 
+chemical abundance information is currently available for \rave, and to evaluate
+whether we could contribute to this existing set.
 

### 2016-08-16 article/unrave.tex
Minor things

@@
 formation and evolution.  This premise hinges on the ability to precisely measure 
-the position, motion, and chemistry of (many) individual stars, and to infer the 
-structure, kinematics, and chemical enrichment of the Galaxy from those data 
+the position, motion, and chemistry of (many) individual stars, and use those data 
+to infer the structure, kinematics, and chemical enrichment of the Galaxy 
 \citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
@@
 stellar spectroscopists frequently report significantly different chemical 
-abundance patterns \citep{Smiljanic_2014}.  Countless long-standing scientific
-questions could be definitively answered if these measurements were known with
-high fidelity: \stub{x? \citep{someone}; y; and z, to list just a few}.
+abundance patterns \citep{Smiljanic_2014}.  Countless long-standing astrophysics
+problems could be resolved if those measurements were known with high fidelity: 
+\stub{x? \citep{someone}; y; and z, to list just a few}.
 
+\stub{Introducing Gaia; rely on ground-based spectroscopic surveys for chemistry}
 % Introducing Gaia
@@
 have better proper motions (137,211 stars).  This motivated us to examine what 
-chemical abundance information is currently available for \rave, and to evaluate
-whether we could contribute to this existing set.
+chemical abundance information was available for \rave, and to evaluate whether 
+we could enable new chemodynamic studies by contributing to this existing set of
+chemical abundances.
 
-We briefly describe the \rave\ data in Section \ref{sec:data}, before explaining
-the model we employ in Section \ref{sec:model}.  In Section \ref{sec:validation}
+We briefly describe the \rave\ data in section \ref{sec:data}, before explaining
+the model we employ in section \ref{sec:model}.  In section \ref{sec:validation}
 we outline a number of validation experiments, including: internal sanity checks,
@@
 are consistent with expectations from astrophysics.  We provide concluding
-remarks in Section \ref{sec:conclusion}, as well as a description of how to
-access the data.
+remarks in section \ref{sec:conclusion}, as well as a description of how to
+access our results electronically.
 
@@
 
-The \rave\ survey started observations using the 6dF system on the 1.2~m UK 
-Schmidt telescope at the Australian Astronomical Observatory\footnote{Formerly 
-the Anglo-Australian Observatory}.  The large 5.7~degree field-of-view and 
-robotic fibre positioner made for very efficient observing: up to 150 targets 
-could be simultaneously obtained.  When observations concluded in April 2013, a 
-total of 574,630 spectra had been collected of 483,330 unique objects.
+\rave\ is a magnitude-limited stellar spectroscopic survey of the (nearby) Milky Way,
+principally designed to measure radial velocities for up to $10^6$ stars.
+Observations were conducted on the 1.2~m UK Schmidt telescope at the Australian 
+Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
+2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
+made for very efficient observing: spectra for up to 150 targets could be 
+simultaneously acquired.  When observations concluded in April 2013, a total of 
+574,630 spectra had been collected of 483,330 unique objects.
 
-RAVE is a magnitude-limited survey, targeting stars with apparent $I$-band
-magnitudes between $9 < I < 13$.  The $I$ band was used as the selection
-magnitude because it overlaps with the wavelength range that \rave\ operates in:
-8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared triplet,
-strong stellar absorption lines that are visible even in very low S/N data 
-(i.e., they are very useful for radial velocity determination).
+The target selection for \rave\ is solely based on the $I$-band apparent magnitude: 
+$9 < I < 13$.  The $I$ band was used because it overlaps with the wavelength range 
+that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
+near-infrared triplet.  These stellar absorption lines are dominated by pressure 
+broadening, which makes them strong and therefore visible even in very low S/N data
+or metal-poor stars.  This spectral region is also well-studied.  It is one of the
+key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
+\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
+coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
+space telescope \citep{Recio-Blanco_2016}.
 
-As the name implies, the \rave\ survey was principally designed to obtain radial
-velocity measurements.  Detailed chemical abundances were a secondary science 
-goal of the survey.  For this reason the distribution of S/N ratios for \rave\ 
-spectra is considerably lower than stellar spectroscopic surveys where chemical
-abundances are the defining objective.  The 
-\rave\ spectra cover $\approx38$~nm in the infrared with an effective resolution
-$\mathcal{R} \approx 7500$, and the S/N peaks at $\approx$50~per~pixel$^{-1}$.
-For comparison, the \project{GALAH} survey \citep{DeSilva_2015} --- which was 
-specifically constructed for detailed chemical abundance analyses --- includes a
-wavelength range about 2.5 times larger at resolution $\mathcal{R} \approx 28,000$,
-and target for S/N $\gtrsim100$.
+The exposure times (and resultant S/N ratios) for \rave\ were optimised to obtain
+radial velocities for as many stars as possible.  Detailed chemical abundances were
+always an important science goal of the survey, but this was a secondary objective.  
+For this reason the distribution of S/N ratios of \rave\ spectra is considerably 
+lower than other stellar spectroscopic surveys where chemical abundances are the 
+primary motivation.  The \rave\ spectra have an effective resolution 
+$\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
+$\approx$50~per~pixel$^{-1}$.  For comparison, the \project{GALAH} survey 
+\citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
+abundance analyses --- includes a wavelength range about 2.5 times larger at 
+resolution $\mathcal{R} \approx 28,000$, and yet the \project{GALAH} project still 
+targets for S/N $\gtrsim100$.
 
-Despite the relatively low resolution and S/N of the spectra, the \rave\ data
-releases have demonstrated the team's enormous success in deriving radial 
-velocities, stellar astrophysical parameters ($\teff$, $\logg$), as well as 
-individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,
-Kordopatis_2013, Kunder_2016}.  The data used for this analysis are the result
-of a thorough reprocessing of all survey data.  This includes a detailed
-re-reduction of the original data frames, an updated continuum-normalization
-procedure, as well as an improvement on the quantification of flux errors in
-each pixel.  At the end of this processing for each survey observation we are 
-provided with: wavelengths, continuum-normalized fluxes, as well as 
-$1\sigma$ uncertainties in flux values.  We refer the reader to the official 
-fifth data release paper of the \rave\ survey, as presented by
-\citet{Kunder_2016}, for more details of this reprocessing.
+Despite the relatively low resolution and S/N of the spectra compared to other
+surveys, the \rave\ data releases have demonstrated the team's enormous success 
+in deriving radial velocities, stellar astrophysical parameters ($\teff$, $\logg$),
+as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
+Siebert_2011,Kordopatis_2013, Kunder_2016}.  In this work we make use of spectra
+that has been reprocessed for the fifth \rave\ data release.  These re-processing
+steps include: a detailed re-reduction of all the original data frames, with flux
+variances propagated at every step; an updated continuum-normalization procedure;
+as well as revised determinations of stellar radial velocities and morphological
+classifications. At the end of this processing for each survey observation we are 
+provided with: rest-frame wavelengths, continuum-normalized fluxes, fractional 
+uncertainties in continuum-normalized flux values, as well as relevant metadata.  
+We refer the reader to the official fifth data release paper of the \rave\ survey, 
+as presented by \citet{Kunder_2016}, for more details of this re-processing.
 
-We chose not to re-normalize the spectra, as our tests proved to us that the
-normalization procedure outlined in \citet{Kunder_2016} is sufficient for our
-analysis.  We re-sampled all spectra onto a common rest-wavelength map.
-Depending on the fibre used and stellar radial velocity, the rest-frame wavelength
-range varied for each star.  Given that fluxes were unavailable in the edge
-pixels for most stars, we excluded pixels outside of the rest wavelength range
+Given the high-quality of the normalization performed by the \rave\ team, we chose
+not to re-normalize the spectra.  Our tests demonstrated that the procedure 
+outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
+there is a limited number of pre-processing steps that we conducted before our
+analysis could begin.  Firstly, we calculated inverse variance arrays from the
+fractional uncertainties provided, then re-sampled the flux and inverse variance
+arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
+used and the stellar radial velocity, the range of rest-frame wavelength values
+varied for each star.  Given that fluxes were unavailable in the edge pixels for 
+most stars, we excluded pixels outside of the rest wavelength range 
 $8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.
@@
 
-% Describe The Cannon briefly
-% training set from APOGEE
+\stub{data-driven goodness}
+% Describe The Cannon briefly, refer to Ness.
+
+\subsection{The training set}
+\label{sec:the-training-set}
+
+% training set from APOGEE or K2/EPIC?!!
 
@@
 
-% Initial tests using a single model for all RAVE spectra
-% Censoring masks
+% Censoring masks on abundance labels
 % Priors on astrophysical parameters from isochrones
@@
 % Include velocities at test time?
-% Train on fibre number? Fit for spectral resolution at test time?
 
@@
 \acknowledgements
-We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), others??.
+A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
 This research made use of: Astropy, a community-developed core Python package for
@@
 
+\bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
+
 \bibitem[Robin et al.(2012)]{Robin_2012} Robin, A.~C., Luri, X., Reyl{\'e}, C., et al.\ 2012, \aap, 543, A100

### 2016-08-16 article/unrave.tex
Add conditions

@@
 
-The exposure times (and resultant S/N ratios) for \rave\ were optimised to obtain
-radial velocities for as many stars as possible.  Detailed chemical abundances were
+The exposure times for \rave\ observations were optimised to obtain radial 
+velocities for as many stars as possible.  Detailed chemical abundances were
 always an important science goal of the survey, but this was a secondary objective.  
-For this reason the distribution of S/N ratios of \rave\ spectra is considerably 
+For this reason the distribution of S/N ratios in \rave\ spectra is considerably 
 lower than other stellar spectroscopic surveys where chemical abundances are the 
@@
 most stars, we excluded pixels outside of the rest wavelength range 
-$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.
+$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
+30~pixels excluded on either side of the common wavelength array.
 
@@
 When combined with positions and 3D velocities from \project{Gaia}, the \project{UNRAVE} catalog will likely be crucial for understanding our local place in the Milky Way.
+
+\subsection*{Access the results electronically}
+
+% TODO: Replace ZENODO-URL with a data url
+Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  This document was compiled from revision hash \githash\
+in that repository.
+Inferred labels and covariance matrices are available with associated metadata electronically at \texttt{\url{ZENODO-URL}}.
+Please note that it is a condition of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must (also) be cited, as the work presented here would not have been possible without the tireless efforts of the entire \rave\ collaboration, past and present.
 

### 2016-08-16 article/unrave.tex
githash ftw

@@
 % ------------------------------------------
+
+\input{vc.tex}
 
@@
 % TODO: Replace ZENODO-URL with a data url
-Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  This document was compiled from revision hash \githash\
-in that repository.
-Inferred labels and covariance matrices are available with associated metadata electronically at \texttt{\url{ZENODO-URL}}.
-Please note that it is a condition of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must (also) be cited, as the work presented here would not have been possible without the tireless efforts of the entire \rave\ collaboration, past and present.
+Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  
+This document was compiled from revision hash \texttt{\githash} in that repository.
+Inferred labels and covariance matrices are available electronically with 
+associated metadata at \texttt{\url{ZENODO-URL}}.  Please note that it is a condition
+of using these results that the fifth \rave\ data release by \citet{Kunder_2016} 
+must (also) be cited, as the work presented here would not have been possible 
+without the tireless efforts of the entire \rave\ collaboration, past and present.
 

### 2016-08-17 article/unrave.tex
minor

@@
 
-\title{The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies}
+\title{The \project{unRAVE} catalog}
 
@@
 stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
-catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{GALAH} 
-\citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, 
+catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{Gaia-ESO} 
+\citep{Gilmore_2012,Randich_2013}, \project{GALAH} \citep{DeSilva_2015},
 \project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
@@
 % --> Residuals as a function of galactic position? (DIBS)
+
+\section{Discussion}
+\label{sec:discussion}
 

### 2016-08-18 article/unrave.tex
Add core RAVE DR5 members as prospective co-authors (list provided by Steinmetz)

@@
 
-\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W.
-Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any \project{RAVE} members who have contributed to the data analysis in this study}
-%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
+% Specific contributors to this work
+\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W. Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
+
+% RAVE DR5 core team
+\author{Luca Casagrande, Cristina Chiappini, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
 
 \author{and the \project{RAVE} collaboration}
+
+%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
 
@@
 
-The target selection for \rave\ is solely based on the $I$-band apparent magnitude: 
-$9 < I < 13$.  The $I$ band was used because it overlaps with the wavelength range 
+The target selection for \rave\ is based on the $I$-band apparent magnitude,
+$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge.
+The $I$ band was used because it overlaps with the wavelength range 
 that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 

### 2016-08-18 article/unrave.tex
Some filler words for internal validation tests

@@
 The target selection for \rave\ is based on the $I$-band apparent magnitude,
-$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge.
+$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge\footnote{
+However, see the works of \citet{Kunder_2016} and \citet{Kordopatis_2013} for 
+subtleties of the selection function}.
 The $I$ band was used because it overlaps with the wavelength range 
@@
 
-\section{Validation}
+\section{Validation experiments}
 \label{sec:validation}
 
-% Chi-sq of the sample.
-% Identify outliers.
-% Convex hull? Distance to plane?
+\subsection{Internal validation}
+We first conducted leave-one-out cross-validation to validate our results.
+For each training set star we removed that star from the training set and
+then re-trained the model.  Using that re-trained model, we then treated
+the excluded star as a test star and compared our inferred labels to the
+test set labels.  We show the results of this experiment in Figure
+\ref{fig:loocv-comparison}.  The bias and root mean square (RMS) error in 
+$\teff$ is XX~K and XX~K, and in $\logg$ is X.XX~dex and X.XX~dex,
+respectively.  The bias in abundance labels is negligible --- between
+X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).
 
-% Stars with morphological classification
+The \rave\ survey performed repeat observations for 23,288 stars  with time 
+intervals ranging from a few hours to up to four years.  This timing was 
+constructed to be quasi-logarithmic such that spectroscopic binaries could
+be optimally identified. Most of the stars that were observed multiple times
+were only observed twice, with thirteen visits being the maximum number 
+of observations for any target.  These repeat observations allow us to 
+quantify the level of (in)correctness in our formal errors.  For every star
+with multiple visits we constructed a high S/N stacked spectrum for that
+star, weighted by the inverse variances in the individual visit spectra,
+which then served as the basis of comparison to all individual visits.  
+Figure \ref{fig:formal-errors-comparison} shows the difference in labels 
+between the comparison spectrum and a repeat visit, normalized by their 
+formal errors summed in quadrature (e.g., $\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
+If our measurements were unbiased by S/N and the errors were representative, 
+these values should be normally distributed with a zero mean and unit variance.
+\stub{Are they? Should we inflate our formal errors?}
 
 % Internal validation:
-% --> repeat visits: are our formal errors reasonable?
 % --> bootstrap resampling of the training set?
+
+% Identify outliers:
+% --> Chi-sq of the sample.
+% --> Convex hull? Distance to plane?
+% --> Stars with morphological classification
+% --> Anything else?
+
+\subsection{External validation}
 
@@
 % --> Literature studies presented in Kordopatis+ 2014
+% --> Gold standards (incl. exoplanet host star studies): Reddy (2003, 2006), Bensby (2014), Valenti & Fischer (2005)
 
+\subsection{Astrophysical validation}
 % Astrophysical validation:
-% --> Identify open clusters and look at their metallicity spreads
 % --> Identify globular clusters and look to see whether we capture "expected" abundance trends/spreads that are seen elsewhere in the literature
@@
 Inferred labels and covariance matrices are available electronically with 
-associated metadata at \texttt{\url{ZENODO-URL}}.  Please note that it is a condition
-of using these results that the fifth \rave\ data release by \citet{Kunder_2016} 
-must (also) be cited, as the work presented here would not have been possible 
-without the tireless efforts of the entire \rave\ collaboration, past and present.
+associated metadata at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  
+Please note that it is a condition of using these results that the fifth \rave\ 
+data release by \citet{Kunder_2016} must (also) be cited, as the work presented 
+here would not have been possible without the tireless efforts of the entire 
+\rave\ collaboration, past and present.
 

### 2016-08-18 article/unrave.tex
woe is us

@@
 
+We sought to construct a training set consisting of stars across the main-sequence,
+sub-giant, and the red giant branch.  We require stars with precisely measured
+effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
+of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
+range of \rave\ does not overlap substantially with high-resolution spectroscopic
+surveys.  The fourth data release of the \project{Gaia-ESO} survey includes giant
+and main-sequence stars, but only 142 overlap with \rave.  The most recent 
+\project{APOGEE} data release \citep{apogee_dr13} reports labels for stars on the
+giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
+that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
+then `up-turning' main-sequence is present, and the metallicity gradient trends in 
+the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
+stars sit above the isochrones, not below).  If we move to lower-resolution studies,
+there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
+S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
+seen in the main-sequence, the labels are expectedly less precise given the lower
+resolution, no detailed abundance information is available\footnote{Abundance 
+information is available for \project{LAMOST} stars from \citet{Ho_2015}, but that 
+sample contains only giant stars}, and there is only sparse sampling across the 
+Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
+but results from that sample are not yet precise enough for our purposes.
+
 % training set from APOGEE or K2/EPIC?!!

### 2016-08-18 article/unrave.tex
Citations for Ho and SDSS DR13

@@
 and main-sequence stars, but only 142 overlap with \rave.  The most recent 
-\project{APOGEE} data release \citep{apogee_dr13} reports labels for stars on the
+\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
@@
 resolution, no detailed abundance information is available\footnote{Abundance 
-information is available for \project{LAMOST} stars from \citet{Ho_2015}, but that 
+information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
 sample contains only giant stars}, and there is only sparse sampling across the 
@@
 
+\bibitem[Ho et al.(2016)]{Ho_2016} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., et al.\ 2016, arXiv:1602.00303
+
 \bibitem[H{\o}g et al.(2000)]{Hog_2000} H{\o}g, E., Fabricius, C., Makarov, V.~V., et al.\ 2000, \aap, 355, L27
@@
 \bibitem[Siebert et al.(2011)]{Siebert_2011} Siebert, A., Williams, M.~E.~K., Siviero, A., et al.\ 2011, \aj, 141, 187
+
+\bibitem[SDSS Collaboration et al.(2016)]{sloan_dr13} SDSS Collaboration, Albareti, F.~D., Allende Prieto, C., et al.\ 2016, arXiv:1608.02013
 

### 2016-08-18 article/unrave.tex
Data driven goodness

@@
 $8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
-30~pixels excluded on either side of the common wavelength array.
+30~pixels excluded on either side of the common wavelength array, leaving us with
+945~pixels for science.
 
@@
 
-Given the relatively low S/N ratios in the \rave\ spectra, we chose to adopt a 
-data-driven model (rather than a physics-driven model) for this analysis.
+We chose to adopt a data-driven model --- rather than a physics-driven model ---
+for this analysis.  Specifically we will use an implementation of \thecannon\
+\citep{Ness_2015}.  Although this choice complicated the construction of our model
+(e.g., see Section \ref{sec:the-training-set}), a data-driven model makes
+use of all available information in the spectrum, and lowers the S/N ratio at
+which systematic effects begin to dominate.  In other words, a well-constructed
+data-driven model will yield more precise labels (e.g., stellar parameters and
+chemical abundances) at low S/N ratios than most physics-driven models\footnote{
+However, see \citet{Casey_2016}}.  This is particularly relevant for the 
+low-resolution \rave\ data analysed here, because about half of the spectra have
+S/N $\lesssim 50$.
 
-\stub{data-driven goodness}
-% Describe The Cannon briefly, refer to Ness.
+Throughout this work we adopt the assumptions and nomenclature as set out in
+\citet{Casey_2016}.  One distinction here is that the \emph{labelled} set and
+the \emph{training} set in this work will be the same set.  In \citet{Casey_2016}
+the labelled set contained two sub-sets: the training set and a distinct
+validation set, allowing us to train on the \emph{training set} and then test
+on the \emph{validation set}, a subset of the labelled set.  However in this work
+the labelled set is relatively small, about an order of magnitude smaller than 
+what was available in \citet{Casey_2016}, and therefore we will use all labelled 
+set stars in the training step and supplement our validation experiments with 
+literature comparisons.  Thus, throughout this work we will use the terms 
+\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
+how the training set was constructed before outlining the model we adopt.
 
@@
 range of \rave\ does not overlap substantially with high-resolution spectroscopic
-surveys.  The fourth data release of the \project{Gaia-ESO} survey includes giant
-and main-sequence stars, but only 142 overlap with \rave.  The most recent 
+surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
+giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
 \project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
@@
 S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
-seen in the main-sequence, the labels are expectedly less precise given the lower
+seen on the main-sequence: the labels are expectedly less precise given the lower
 resolution, no detailed abundance information is available\footnote{Abundance 
@@
 but results from that sample are not yet precise enough for our purposes.
+
+% These restrictions forced us to use a heterogeneous training set.
+%
+
+\stub{We used the APOGEE DR13 cross-over for training set stars on the giant branch with cuts XYZ}
 
@@
 
-% Include velocities at test time?
+% Include velocities and rotational velocities at test time?
 

### 2016-08-18 article/unrave.tex
Minor

@@
 for this analysis.  Specifically we will use an implementation of \thecannon\
-\citep{Ness_2015}.  Although this choice complicated the construction of our model
-(e.g., see Section \ref{sec:the-training-set}), a data-driven model makes
-use of all available information in the spectrum, and lowers the S/N ratio at
-which systematic effects begin to dominate.  In other words, a well-constructed
+\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
+of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven model 
+makes use of all available information in the spectrum, and lowers the S/N ratio 
+at which systematic effects begin to dominate.  In other words, a well-constructed
 data-driven model will yield more precise labels (e.g., stellar parameters and
 chemical abundances) at low S/N ratios than most physics-driven models\footnote{
-However, see \citet{Casey_2016}}.  This is particularly relevant for the 
+However, see \citet{Casey_2016a}}.  This is particularly relevant for the 
 low-resolution \rave\ data analysed here, because about half of the spectra have
@@
 Throughout this work we adopt the assumptions and nomenclature as set out in
-\citet{Casey_2016}.  One distinction here is that the \emph{labelled} set and
-the \emph{training} set in this work will be the same set.  In \citet{Casey_2016}
-the labelled set contained two sub-sets: the training set and a distinct
+\citet{Casey_2016b}.  One distinction here is that the \emph{labelled} set and
+the \emph{training} set in this work will be the same set.  In \citet{Casey_2016b}
+the labelled set contained two subsets: the training set and a distinct
 validation set, allowing us to train on the \emph{training set} and then test
@@
 the labelled set is relatively small, about an order of magnitude smaller than 
-what was available in \citet{Casey_2016}, and therefore we will use all labelled 
+what was available in \citet{Casey_2016b}, and therefore we will use all labelled 
 set stars in the training step and supplement our validation experiments with 
@@
 \emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
-how the training set was constructed before outlining the model we adopt.
+how we constructed the training set, before we outline the model that we adopt.
 
@@
 
-\bibitem[Casey et al.(2016)]{Casey_2016} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
+\bibitem[Casey(2016)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
+
+\bibitem[Casey et al.(2016)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
 
@@
 \bibitem[Ness et al.(2015)]{Ness_2015} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16
+
+\bibitem[Ness et al.(2016)]{Ness_2016} Ness, M., Hogg, D.~W., Rix, H.-W., et al.\ 2016, \apj, 823, 114
 

### 2016-08-18 article/unrave.tex
Paragraph about APOGEE giant training set

@@
 
-% These restrictions forced us to use a heterogeneous training set.
-%
+These constraints forced us to construct a heterogeneous training set.  Given previous
+successes in transferring high S/N ratio labels from \project{APOGEE} \citep{Ness_2015,
+Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \project{APOGEE}
+---\rave\ overlap sample for giant star labels in the training set.  Of these, about
+900 are giants according to \project{APOGEE}.  From this sample we selected stars to
+have: metallicity determinations ([Fe/H] $> -5$); high S/N ratios in both the 
+\project{APOGEE} and \rave\ spectra ($>200$ in \project{APOGEE}; $>50$ in \rave); and
+we further required that the \project{ASPCAP} pipeline did not report any peculiar
+flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
+giant branch, with metallicities ranging from [Fe/H] $= -1.62$ to 0.32.
 
-\stub{We used the APOGEE DR13 cross-over for training set stars on the giant branch with cuts XYZ}
+Assembling a training set for the main-sequence and sub-giant branch was not as
+straightforward.
+% training set from K2/EPIC?!!
 
-% training set from APOGEE or K2/EPIC?!!
+% Identifying obviously bad stars that have been mis-classified by K2/EPIC
 

### 2016-08-18 article/unrave.tex
Some framework text about globular clusters

@@
 
-Assembling a training set for the main-sequence and sub-giant branch was not as
-straightforward.
+Assembling a training set for the main-sequence and sub-giant branch was more complex.
 % training set from K2/EPIC?!!
+% Identifying obviously bad stars that have been mis-classified by K2/EPIC
 
-% Identifying obviously bad stars that have been mis-classified by K2/EPIC
+% Magic wizardry for missing and imperfect labels.
 
@@
 \subsection{External validation}
+\label{sec:external-validation}
 
@@
 \subsection{Astrophysical validation}
+
+After verifying that our stellar labels are comparable with other
+high-resolution studies, here we verify that the abundance labels we find
+are consistent with expectations from astrophysics.  One example is stars in
+globular clusters, which show anti-correlations in light element abundances 
+\citep[e.g.,][and references therein]{Da_Costa,Carretta_2009}.  In the \rave\
+survey we found 70 stars with positions and radial velocities that are
+consistent with being members of globular clusters.  Specifically we identified
+49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
+globular cluster NGC~3201, and 10 members of NGC~362.  
+% TODO: Were any of these stars unintentionally in the training set?
+All of these systems are well-studied and exhibit abundance anti-correlations 
+in Na--O and Mg--Al \citep{people}.  Other correlations are also present 
+(e.g., C--N), but those relationships are less relevant for this work because 
+we do not measure C and N abundances here.
+
+We show the detailed abundances for these globular cluster members in
+Figure \ref{fig:globular-cluster-abundances}.  For comparison purposes we
+have included abundances from the high-resolution studies of \citep{people}.
+\stub{What do we find?}
+
+% Assuming we recover these relationships...
+Indeed, this demonstrates that our abundance labels can be used to identify
+globular cluster members that are now tidally disrupted \citep{Anguiano_2016,Kuzma_2016,Navin_2016},
+even for stars with low S/N ratios.
+
+% Some literature notes:
+% See http://iopscience.iop.org/article/10.1088/0004-637X/731/1/64/pdf
+% NGC 3201 has Na-O anti-correlation: Carretta et al. (2009)
+% NGC 362 has "two discrete groups along the Na-O anti-correlation": https://arxiv.org/abs/1307.4085
+% NGC 362 has Mg-Al relationship (low Mg scatter, large Al scatter) --> Carretta (2013)
+% NGC 3201 has Mg-Al relationship (http://arxiv.org/pdf/1305.3645.pdf and references therein)
+
 % Astrophysical validation:
-% --> Identify globular clusters and look to see whether we capture "expected" abundance trends/spreads that are seen elsewhere in the literature
 % --> Residuals as a function of galactic position? (DIBS)
@@
 \begin{thebibliography}{dummy}
+
+\bibitem[Anguiano et al.(2016)]{Anguiano_2016} Anguiano, B., De Silva, G.~M., Freeman, K., et al.\ 2016, \mnras, 457, 2078
+
 \bibitem[Astropy Collaboration et 
@@
 
+\bibitem[Kuzma et al.(2016)]{Kuzma_2016} Kuzma, P.~B., Da Costa, G.~S., Mackey, A.~D., \& Roderick, T.~A.\ 2016, \mnras, 461, 3639
+
 \bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, arXiv:1606.03015
@@
 \bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
+
+\bibitem[Navin et al.(2016)]{Navin_2016} Navin, C.~A., Martell, S.~L., \& Zucker, D.~B.\ 2016, arXiv:1606.06430
 

### 2016-08-18 article/unrave.tex
Add paragraph stem about rave dr4/5 comparisons

@@
 
+Here we cross-match our results against literature samples to verify that
+our labels are accurate.  Because this work is an independent analysis, our
+first point of reference is against the official \rave\ data releases.  These
+comparisons are shown in Figures \ref{fig:rave-dr4-comparison} and 
+\ref{fig:rave-dr5-comparison}, where we show our labels with respect to the
+results in the fourth and fifth \rave\ data releases, respectively.  In order
+to provide a fair comparison we only show stars that meet a number of quality
+flags in both samples. 
+\stub{What selections qualify as a fair comparison in both samples?}
+There is very good agreement in $\teff$, with a bias
+and RMS of just XX~K and XXX~K, respectively.  We find our $\logg$ values are
+closest to the `calibrated' \rave\ surface gravities, which make use of 
+asteroseismic information to correct (increase) the $\logg$ values along the
+giant branch.  The calibrated \rave\ surface gravities also correct the 
+location of the red clump (by $\sim0.5$~dex) to be in the same position in our
+results.  However, while our $\logg$ values generally agree with the calibrated
+\rave\ values along the giant branch, our $\logg$ values do differ slightly 
+along the main-sequence.  This is principally because the main-sequence in 
+this work tends to taper down towards higher $\logg$ values at cooler 
+temperatures, whereas the comparison sample tend to have a slightly flatter
+main-sequence.  This difference is not likely to have a very significant 
+effect on the detailed abundance or spectrophotometric distance determinations
+between these studies.
+
+\stub{What abundances are available?}
+\stub{Any differences in RAVE abundances?}
+
 % External validation: 
-% --> RAVE Data Release 4
 % --> Gaia-ESO Survey
@@
 globular cluster NGC~3201, and 10 members of NGC~362.  
-% TODO: Were any of these stars unintentionally in the training set?
+\stub{Were any of these stars unintentionally in the training set?}
 All of these systems are well-studied and exhibit abundance anti-correlations 
@@
 \begin{figure*}[p]
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release.\label{fig:rave-dr4-comparison}}
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release \citep{Kordopatis_2014}.\label{fig:rave-dr4-comparison}}
+\end{figure*}
+
+\begin{figure*}[p]
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fifth data release \citep{Kunder_2016}.\label{fig:rave-dr5-comparison}}
 \end{figure*}
@@
 \begin{figure*}[p]
-\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2013}}
+\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2014}}
 \end{figure*}

### 2016-08-19 article/unrave.tex
Update abstract

@@
 
-\input{vc.tex}
+\newcommand{\githash}{UNKNOWN}
+%\input{vc.tex}
 
@@
 \newcommand{\Nstars}{483,330}
+\newcommand{\gaia}{\project{Gaia}}
 
@@
 \begin{abstract}
-On 14 September 2016 the \project{Gaia} mission will release positions, proper 
-motions, and parallaxes for about two million stars.  Most studies intending to
-make use of those data would benefit enormously by coupling that astrometry with
-stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
-survey has acquired spectra for 292,036 stars in common with the 
-first \project{Gaia} data release, constituting the largest overlap of any 
-stellar spectroscopic sample.  Here we perform an independent analysis of the 
-\project{RAVE} spectra using an implementation of \thecannon\ 
-that uses prior probabilities on atomic line formation and stellar 
-astrophysical parameters.  We deliver effective temperature $\teff$, surface 
-gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
-Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
-repeat visits, and externally against existing high-resolution spectroscopic 
-surveys.  The typical precision in chemical abundances is X.XX~dex.  
-We provide improved distances by isochrone-fitting for \project{RAVE} stars
-where astrometric solutions are not available in the first \project{Gaia}
-data release.
+
+The Milky Way may be our greatest laboratory for understanding galaxy formation.
+This premise relies on the ability to measure [1] 3D positions and motions, and
+[2] detailed chemical abundances, for large numbers of individual stars.  On 14
+September 2016 the \gaia\ mission will release the former for $\approx2\times10^6$
+stars.  However, a catalog of precise chemical abundances to complement that data
+release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
+acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
+constituting the largest overlap of any stellar spectroscopic sample.  Here we 
+perform an independent analysis of \rave\ spectra using an implementation of 
+\thecannon\ that uses prior probabilities on atomic line formation and stellar
+astrophysical parameters.  Our approach yields precise stellar labels at lower
+S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
+improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
+and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
+for \Nstars\ stars.  We validate our results internally from repeat visits, 
+and externally against existing high-resolution spectroscopic surveys.  The 
+typical precision in chemical abundances is X.XX~dex.  We provide 
+spectrophotometric distances based on our labels, improving upon the parallax
+distances available for the fainter targets in the first \gaia\ data release.
+% A one-line science result based on RAVE/Hipparcos overlap.
 \end{abstract}

### 2016-08-22 article/figures/plot_ges_comparison.py
Update RAVE io file and plots

@@ -21,7 +21,7 @@ else:
 
 # Plot labels against literature.
 
-label_names = ("TEFF_1", "LOGG_2", "FE_H")
+label_names = ("TEFF_1", "LOGG_1", "FE_H")
 ges_label_names = {
     "TEFF_1": "TEFF_2",
     "LOGG_1": "LOGG_2",
@@ -29,7 +29,7 @@ ges_label_names = {
 }
 
 label_limits = {
-    "TEFF_1": (3000, 8000),
+    "TEFF_1": (3000, 6500),
     "LOGG_1": (0, 5),
     "FE_H": (-2, 0.75)
 }
@@ -42,7 +42,8 @@ latex_labels = {
 
 
 # Exclude ones we think are bad.
-ok = data_table["OK"]
+#ok = data_table["OK"]
+ok = data_table["R_CHI_SQ"] < 3
 
 data_table = data_table[ok]
 
@@ -69,11 +70,15 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     y = data_table[ges_label_names[label_name]]
     # yerr
 
-    ax.scatter(x, y, c=data_table["SNRK"], alpha=0.75, s=50, 
+    scat = ax.scatter(x, y, c=[[0, 1][each.startswith("U5")] for each in data_table["REC_SETUP"]], 
+        alpha=0.75, s=50, 
         linewidths=0.5, edgecolors="#000000")
 
-    ax.set_xlim(limits[label_name])
-    ax.set_ylim(limits[label_name])
+    limits = label_limits[label_name]
+    ax.plot(limits, limits, c="#666666", linestyle=":", zorder=-1)
+
+    ax.set_xlim(limits)
+    ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
     ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm UNRAVE})$"]))
@@ -81,6 +86,10 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
 
     [_.set_rotation(30) for _ in ax.get_xticklabels()]
     [_.set_rotation(30) for _ in ax.get_yticklabels()]
+
+    diff = y - x
+    print(label_name, np.nanmean(diff), np.nanstd(diff))
+
 
 fig.tight_layout()
 

### 2016-08-22 article/figures/plot_literature_comparisons.py

@@ -42,7 +42,7 @@ pastel_label_names = {
 label_limits = {
     "TEFF": (3000, 8000),
     "LOGG": (0, 5),
-    "FE_H": (-2, 0.75)
+    "FE_H": (-3, 0.75)
 }
 
 latex_labels = {
@@ -105,8 +105,8 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
             c=colors[j], alpha=0.75, marker=(3, 1, angles[j]), s=50, 
             linewidths=0.5, edgecolors="#000000")
 
-    ax.set_xlim(limits[label_name])
-    ax.set_ylim(limits[label_name])
+    ax.set_xlim(label_limits[label_name])
+    ax.set_ylim(label_limits[label_name])
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
     ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm UNRAVE})$"]))

### 2016-08-22 article/figures/rave_io.py

@@ -7,6 +7,16 @@ DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
 
 rave_kordopatis_dr4 = Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
 rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))
+
+if "OK" not in rave_cannon_dr1.dtype.names:
+    rave_cannon_dr1["OK"] = (rave_cannon_dr1["SNRK"] > 50) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
+    print("Setting OK as stars with S/N > 50 and R_CHI_SQ < 3")
+
+else:
+    print("Using pre-defined subset of 'OK'")
+
+
 ges_idr4 = Table.read(os.path.join(DATA_PATH, "literature-ges-idr4-xm.fits"))
+ges_idr4["Name"] = [each.strip() for each in ges_idr4["Name"]]
 
 literature_pastel = Table.read(os.path.join(DATA_PATH, "literature-pastel-database-xm.fits"))

### 2016-08-22 article/figures/make_plots.sh
Include comparison figures (as PNGs, temporarily)

@@ -1,0 +1,6 @@
+python  plot_clusters.py
+python  plot_dr4_comparison.py
+python  plot_ges_comparison.py
+#python  plot_label_precision.py
+python  plot_literature_comparisons.py
+#python  plot_loocv_results.py

### 2016-08-22 article/figures/plot_dr4_comparison.py

@@ -22,24 +22,23 @@ except NameError:
 else:
     print("Warning: Using pre-loaded data!")
 
-QC = combined_table["OK"] * (combined_table["QK"] == 0)
+QC = combined_table["OK"] * (combined_table["QK_1"] == 0)
 combined_table = combined_table[QC]
 
 N_bins = 50
 
 all_columns = [
     # RAVE label, DR4 label
-    ("TEFF", "TeffK"),
-    ("LOGG", "loggK"),
-    ("FE_H", "__M_H_K"),
+    ("TEFF", "TeffK_1"),
+    ("LOGG", "loggK_1"),
+    ("FE_H", "__M_H_K_1"),
 ]
 label_limits = {
     "TEFF": (3000, 8000),
     "LOGG": (0, 5),
-    "FE_H": (-2, 0.75)
+    "FE_H": (-2.5, 0.75)
 }
 latex_labels = (r"$T_{\rm eff}$", r"$\log{g}$", r"$[{\rm Fe/H}]$")
-
 
 
 K, factor = (len(all_columns), 3.5)
@@ -81,7 +80,10 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
     [_.set_rotation(30) for _ in ax.get_xticklabels()]
     [_.set_rotation(30) for _ in ax.get_yticklabels()]
 
+    diff = y - x
+    print(latex_labels, np.nanmean(diff), np.nanstd(diff))
+
 fig.tight_layout()
 
 fig.savefig("dr4-comparison.pdf", dpi=300)
-fig.savefig("dr4-comparison.png", dpi=300)
+fig.savefig("dr4-comparison.png")

### 2016-08-22 article/figures/plot_ges_comparison.py

@@ -70,8 +70,7 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     y = data_table[ges_label_names[label_name]]
     # yerr
 
-    scat = ax.scatter(x, y, c=[[0, 1][each.startswith("U5")] for each in data_table["REC_SETUP"]], 
-        alpha=0.75, s=50, 
+    scat = ax.scatter(x, y, c=data_table["SNRK"], alpha=0.75, s=50, 
         linewidths=0.5, edgecolors="#000000")
 
     limits = label_limits[label_name]
@@ -90,8 +89,9 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     diff = y - x
     print(label_name, np.nanmean(diff), np.nanstd(diff))
 
-
+cbar = plt.colorbar(scat)
+cbar.set_label(r"$S/N$")
 fig.tight_layout()
 
 fig.savefig("ges-comparison.pdf", dpi=300)
-fig.savefig("ges-comparison.png", dpi=300)
+fig.savefig("ges-comparison.png")

### 2016-08-22 article/figures/plot_literature_comparisons.py

@@ -51,13 +51,15 @@ latex_labels = {
     "FE_H": r"$[{\rm Fe/H}]$"
 }
 
+"""
 exclude_comparisons = ("2013MNRAS.429..126R", )
 ok = np.ones(len(data_table), dtype=bool)
 for bibcode in exclude_comparisons:
     ok *= (data_table["bibcode"] != bibcode)
+"""
 
 # Exclude ones we think are bad.
-ok *= data_table["OK"]
+ok = data_table["OK"]
 
 data_table = data_table[ok]
 
@@ -118,4 +120,4 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
 fig.tight_layout()
 
 fig.savefig("literature-comparison.pdf", dpi=300)
-fig.savefig("literature-comparison.png", dpi=300)
+fig.savefig("literature-comparison.png")

### 2016-08-22 article/unrave.tex
Triggering dd
Trigger
muh triggers
Nothing
Comment to trigger dd
fuck
fix shit with dd
sick of this
grr
test
Include comparison figures (as PNGs, temporarily)
Fix citation to Kordopatis

@@
 
-\usepackage{amsmath,amssymb}
+\usepackage{amsmath}
 \usepackage{bm}
@@
 \newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\gaia}{\project{Gaia}}
 \newcommand{\thecannon}{\project{The~Cannon}}
+\newcommand{\rave}{\project{\acronym{RAVE}}}
+
 \newcommand{\acronym}[1]{{\small{#1}}}
-\newcommand{\rave}{\project{\acronym{RAVE}}}
+
+\newcommand{\stub}[1]{\textbf{#1}}
+
+\newcommand{\teff}{T_{\mathrm{eff}}}
 \newcommand{\logg}{\log g}
-\newcommand{\teff}{T_{\mathrm{eff}}}
+
+\newcommand{\Nstars}{483,330}
+
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}
-\newcommand{\stub}[1]{\textbf{#1}}
-\newcommand{\Nstars}{483,330}
-\newcommand{\gaia}{\project{Gaia}}
 
@@
 
-% Specific contributors to this work
+% Specific contributors to this work:
+% There is some order already in mind, but final order will be determined by  contribution.
 \author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W. Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
 
-% RAVE DR5 core team
+% RAVE DR5 core team:
 \author{Luca Casagrande, Cristina Chiappini, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
 
-\author{and the \project{RAVE} collaboration}
-
-%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}
+\author{and the \rave\ collaboration}
 
 \begin{abstract}
-
 The Milky Way may be our greatest laboratory for understanding galaxy formation.
@@
 distances available for the fainter targets in the first \gaia\ data release.
-% A one-line science result based on RAVE/Hipparcos overlap.
+\stub{A one-line science result based on RAVE/Hipparcos overlap}
 \end{abstract}
@@
 \begin{figure*}[p]
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release \citep{Kordopatis_2014}.\label{fig:rave-dr4-comparison}}
+\includegraphics[width=\textwidth]{figures/dr4-comparison.png}
+\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release \citep{Kordopatis_2013}.\label{fig:rave-dr4-comparison}}
 \end{figure*}
@@
 \begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/ges-comparison.png}
 \caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:ges-dr4-comparison}}
@@
 \begin{figure*}[p]
-\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2014}}
+\includegraphics[width=\textwidth]{figures/literature-comparison.png}
+\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2013}}
 \end{figure*}

### 2016-08-23 article/figures/plot_hrd_test.py
Added script to plot HRD for the test set

@@ -1,0 +1,84 @@
+
+
+import locale
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+
+locale.setlocale(locale.LC_ALL, 'en_US')
+
+try:
+    rave_cannon_dr1
+
+except NameError:
+    from rave_io import rave_cannon_dr1
+
+else:
+    print("Using pre-loaded data")
+
+
+
+
+t = rave_cannon_dr1
+
+#ok = (t["SNRK"] > 50) * (t["r_chi_sq_ms"] < 3) * (t["r_chi_sq_giant"] < 3) #* (t["WEIGHTED_VSINI"] < 1)
+
+
+snrs = (100, 50, 25, 10)
+M, N = (len(snrs), 50)
+
+
+factor = 3.5
+lbdim = 0.2 * factor
+trdim = 0.1 * factor
+whspace = 0.05
+yspace = factor
+xspace = factor * M + factor * (M - 1) * whspace + lbdim * (M - 1)
+xdim = lbdim + xspace + trdim
+ydim = lbdim + yspace + trdim
+
+fig, axes = plt.subplots(1, M, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+for ax, snr in zip(axes, snrs):
+    
+    ok = (t["SNRK"] > snr) * (t["r_chi_sq_ms"] < 3) * (t["r_chi_sq_giant"] < 3)
+
+    ax.hexbin(t["WEIGHTED_TEFF"][ok], t["WEIGHTED_LOGG"][ok], gridsize=N,
+        extent=(3500, 7000, 0.5, 5),
+        cmap="Blues", norm=LogNorm(), edgecolor="#eeeeee", linewidths=0.1)
+
+
+    K = locale.format("%d", sum(ok), grouping=True).replace(",", "$,$")
+
+    ax.text(0.05, 0.9, r"$S/N > {:.0f}$".format(snr),
+        horizontalalignment="left", verticalalignment="bottom",
+        transform=ax.transAxes, fontsize=14)
+    ax.text(0.05, 0.82, r"${}$".format(K) + r" ${\rm stars}$",
[…]

### 2016-08-23 article/figures/plot_spectral_derivatives.py

@@ -1,0 +1,117 @@
+
+"""
+Plot model coefficients.
+"""
+
+
+import AnniesLasso as tc
+
+
+giant_model = tc.load_model("/data/gaia-eso/arc/rave/rave-tgas-v23giants.model") # giants
+ms_model = tc.load_model("/data/gaia-eso/arc/rave/rave-tgas-v16b.model") # dwarfs
+
+kwds = [
+    dict(color="r", alpha=0.8, label="Giant model - Combined model"),
+    dict(color="g", alpha=0.8, label="Dwarf model - Combined model"),
+    #dict(color="b", alpha=0.8),
+    dict(color="m", alpha=0.8)
+]
+
+# For comparison sake
+ms_model.vectorizer._label_names = ["TEFF", "LOGG", "FE_H", "VSINI"] 
+
+models = (giant_model, ms_model)
+
+comparison_model = tc.load_model("/data/gaia-eso/arc/rave/rave-tgas-v23all.model")   # both
+
+N = comparison_model.theta.shape[1]
+
+
+fig, axes = plt.subplots(4, 1, sharex=True, figsize=(9, 13))
+
+offset = 0
+for i in range(N):
+
+
+    term = comparison_model.vectorizer.get_human_readable_label_vector().split(" + ")[i]
+
+    if "FE_H" not in term:
+        offset += 1
+        continue
+
+    ax = axes[i - offset]
+
+    for k, (model, kwd) in enumerate(zip(models, kwds)):
+
+        # Get the index.
+        
+
+        try:
+            index = model.vectorizer.get_human_readable_label_vector().split(" + ").index(term)
+        
+        except ValueError:
+            print("Term {} not in model".format(term))
+            continue
+
+
+
+        y1 = model.theta[:, index].copy()
+
+        # Scale the coefficients back by their isotropic scales.s
+        
+        if i > 0:
+            for j, z in model.vectorizer.terms[index - 1]:
+                y1 = y1 * model.vectorizer.scales[j]
+                print(term, k, model, j, model.vectorizer.scales[j])
+
+        y2 = comparison_model.theta[:, i].copy()
+        if i > 0:
+            for j, k in comparison_model.vectorizer.terms[i - 1]:
+                y2 = y2 * comparison_model.vectorizer.scales[j]
+        
+
+        """
+        if "EPIC_TEFF" in term:
[…]

### 2016-08-23 article/figures/rave_io.py

@@ -9,8 +9,13 @@ rave_kordopatis_dr4 = Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
 rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))
 
 if "OK" not in rave_cannon_dr1.dtype.names:
-    rave_cannon_dr1["OK"] = (rave_cannon_dr1["SNRK"] > 50) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
-    print("Setting OK as stars with S/N > 50 and R_CHI_SQ < 3")
+    try:
+        rave_cannon_dr1["OK"] = (rave_cannon_dr1["SNRK"] > 50) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
+    except KeyError:
+        print("No 'OK' subset")
+
+    else:
+        print("Setting OK as stars with S/N > 50 and R_CHI_SQ < 3")
 
 else:
     print("Using pre-defined subset of 'OK'")

### 2016-08-23 article/unrave.tex
Show git hash when compiling, or default to unknown if vc.tex doesn't exist

@@
 
-\newcommand{\githash}{UNKNOWN}
-%\input{vc.tex}
+\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}}
 
@@
 effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
-of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
+of Fe, Mg, Si, Ti, Ca, Al, and Ni. We note that the while Ti can be measured, its values in both 
+APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
+This proved to be difficult because the magnitude
 range of \rave\ does not overlap substantially with high-resolution spectroscopic
@@
 results in the fourth and fifth \rave\ data releases, respectively.  In order
-to provide a fair comparison we only show stars that meet a number of quality
+to provide a fair comparison, we only show stars that meet a number of quality
 flags in both samples. 
@@
 \citep[e.g.,][and references therein]{Da_Costa,Carretta_2009}. In the \rave\
-survey we found 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters.  Specifically we identified
-49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
+survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
+consistent with being members of globular clusters.  We have used that sample
+Specifically, this GC sample includes 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
 globular cluster NGC~3201, and 10 members of NGC~362.  
-\stub{Were any of these stars unintentionally in the training set?}
+\stub{Were any of these stars unintentionally in the training set? \\ KH: Not using my training sets.}
 All of these systems are well-studied and exhibit abundance anti-correlations 
@@
 % NGC 362 has Mg-Al relationship (low Mg scatter, large Al scatter) --> Carretta (2013)
+% NGC 362 has a large metallicity spread in RAVE (Anguiano_2015) but not the case in high-res studies (e.g.
+https://arxiv.org/abs/1307.4085 has a metallicity spread of 0.05/0.08 dex) may indicate better cluster selection needed and can be done with cannon results.
 % NGC 3201 has Mg-Al relationship (http://arxiv.org/pdf/1305.3645.pdf and references therein)
@@
 This work was partly supported by the European Union FP7 programme through ERC 
-grant number 320360.
+grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
 Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
@@
 \begin{thebibliography}{dummy}
+\bibitem[Anguiano et al.(2015)]{Anguiano_2015} Anguiano, B., Zucker, D.~B., Scholz, R.-D., et al.\ 2015, \mnras, 451, 1229
 
@@
 \bibitem[Gilmore et al.(2012)]{Gilmore_2012} Gilmore, G., Randich, S., Asplund, M., et al.\ 2012, The Messenger, 147, 25
+
+\bibitem[Hawkins et al.(2016)]{Hawkins_2016} Hawkins, K., Masseron, T., Jofre, P., et al.\ 2016, arXiv:1604.08800
+
+\bibitem[Holtzman et al.(2015)]{Holtzman_2015} Holtzman, J.~A., Shetrone, M., Johnson, J.~A., et al.\ 2015, \aj, 150, 148
 

### 2016-08-23 article/unrave.tex
Expanded the stubs in the introduction

@@
 
+\newcommand{\acronym}[1]{{\small{#1}}}
+
 \newcommand{\project}[1]{\textsl{#1}}
@@
 \newcommand{\rave}{\project{\acronym{RAVE}}}
-
-\newcommand{\acronym}[1]{{\small{#1}}}
+\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
+\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
+\newcommand{\hipparcos}{\project{Hipparcos}}
 
@@
 The Milky Way may be our greatest laboratory for understanding galaxy formation.
-This premise relies on the ability to measure [1] 3D positions and motions, and
-[2] detailed chemical abundances, for large numbers of individual stars.  On 14
-September 2016 the \gaia\ mission will release the former for $\approx2\times10^6$
-stars.  However, a catalog of precise chemical abundances to complement that data
+This premise relies on the ability to measure --- for large numbers of stars ---
+3D positions and motions, and their detailed chemical abundances.  On 14 September
+2016 the \gaia\ mission will release the former for $\approx2\times10^6$ stars.
+However, a catalog of precise chemical abundances to complement that data
 release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
@@
 formation and evolution.  This premise hinges on the ability to precisely measure 
-the position, motion, and chemistry of (many) individual stars, and use those data 
+the astrometry and chemistry for (many) individual stars, and to use those data 
 to infer the structure, kinematics, and chemical enrichment of the Galaxy 
-\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
-Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
-However, these quantities are not known for even 1\% of stars in the Milky Way.  
-Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
-Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
-from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
+\citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
+Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Bovy_2016}.  However, these 
+quantities are not known for even 1\% of stars in the Milky Way.  Stellar 
+distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,
+Madler_2016}, proper motions can be plagued by unquantified systematics from 
+the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
 stellar spectroscopists frequently report significantly different chemical 
-abundance patterns \citep{Smiljanic_2014}.  Countless long-standing astrophysics
-problems could be resolved if those measurements were known with high fidelity: 
-\stub{x? \citep{someone}; y; and z, to list just a few}.
+abundance patterns \citep{Smiljanic_2014}.  The impact this lack of data has on
+scientific inferences cannot be understated.  It limits our understanding on a
+number of sub-fields in astrophysics, including the properties of exoplanet host
+stars, the formation (and destruction) of stars and clusters, as well as the 
+structure of the Galactic disk, to name a few.
 
-\stub{Introducing Gaia; rely on ground-based spectroscopic surveys for chemistry}
-% Introducing Gaia
-% Gaia will produce a very clear snapshot, producing astrometry space motions
-% What will Gaia provide eventually.
-% In the first data release what do we get.
-% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
-% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry
+These formed just some of the scientific drivers for the \gaia\ space telescope.
+\gaia\ is primarily an astrometric mission, providing precise positions,
+parallaxes and proper motions for more than $10^9$ stars in its final data
+release, scheduled for 2022.  While this is a sample size about three orders 
+of magnitude larger than its predecessor \hipparcos, both astrometry and 
+chemistry are required to fully characterize the formation and evolution of our Galaxy.  
+For this reason \gaia\ will also provide radial velocities, stellar parameters 
+and chemical abundances for a subset of brighter stars, primarily from the on-board
+radial velocity spectrometer.  However, those chemical abundances will not be 
+available until later releases.  Until those abundances are available,
+astronomers seeking to use chemistry and kinematic measurements are reliant on 
+ground-based spectroscopic surveys in order to complement the available astrometry
+from \gaia.
 
@@
 
-\section{Model}
-\label{sec:model}
+\section{Method}
+\label{sec:method}
 
@@
 effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
-of Fe, Mg, Si, Ti, Ca, Al, and Ni. We note that the while Ti can be measured, its values in both 
-APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
-This proved to be difficult because the magnitude
+of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
 range of \rave\ does not overlap substantially with high-resolution spectroscopic
@@
 % Identifying obviously bad stars that have been mis-classified by K2/EPIC
+% Refer to the training set figure.
+% Two separate models, with an overlap region.
 
-% Magic wizardry for missing and imperfect labels.
+% Initially we train on three labels.
 
@@
 
+% Magic wizardry for missing and imperfect labels.
 % Censoring masks on abundance labels
@@
 \stub{What abundances are available?}
+
+% This is more relevant for the discussion, not description of the training set.
+We note that the while Ti can be measured, its values in both 
+APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
+
 \stub{Any differences in RAVE abundances?}
@@
 
+\bibitem[Bovy et al.(2016)]{Bovy_2016} Bovy, J., Rix, H.-W., Schlafly, E.~F., et al.\ 2016, \apj, 823, 30
+
 \bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88
@@
 
+\bibitem[Schlaufman et al.(2009)]{Schlaufman_2009} Schlaufman, K.~C., Rockosi, C.~M., Allende Prieto, C., et al.\ 2009, \apj, 703, 2177
+
 \bibitem[Siebert et al.(2011)]{Siebert_2011} Siebert, A., Williams, M.~E.~K., Siviero, A., et al.\ 2011, \aj, 141, 187
@@
 \begin{figure}[p]
-\caption{A H-R diagram showing the test set labels.\label{fig:test-set-hrd}}
+\includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
+\caption{A `spectroscopic' Hertzsprung-Russell diagram showing the effective temperature $\teff$ and surface gravity $\logg$ for stars in the test set.
+Only stars with $\chi_r < 3$ are shown, and each panel shows stars that meet a threshold in $S/N$ ratio.\label{fig:test-set-hrd}}
 \end{figure}

### 2016-08-23 article/unrave.tex
Some edits to the method section

@@
 \newcommand{\hipparcos}{\project{Hipparcos}}
+\newcommand{\epic}{\project{K2/EPIC}}
 
@@
 The target selection for \rave\ is based on the $I$-band apparent magnitude,
-$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge\footnote{
-However, see the works of \citet{Kunder_2016} and \citet{Kordopatis_2013} for 
-subtleties of the selection function}.
-The $I$ band was used because it overlaps with the wavelength range 
-that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
+$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
+The $I$ band was used because it overlaps with the wavelength range that \rave\ 
+operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
 near-infrared triplet.  These stellar absorption lines are dominated by pressure 
@@
 as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
-Siebert_2011,Kordopatis_2013, Kunder_2016}.  In this work we make use of spectra
+Siebert_2011,Kordopatis_2013,Kunder_2016}.  In this work we make use of spectra
 that has been reprocessed for the fifth \rave\ data release.  These re-processing
@@
 of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven model 
-makes use of all available information in the spectrum, and lowers the S/N ratio 
+makes use of all available information in the spectrum and lowers the S/N ratio 
 at which systematic effects begin to dominate.  In other words, a well-constructed
@@
 
-Throughout this work we adopt the assumptions and nomenclature as set out in
-\citet{Casey_2016b}.  One distinction here is that the \emph{labelled} set and
-the \emph{training} set in this work will be the same set.  In \citet{Casey_2016b}
-the labelled set contained two subsets: the training set and a distinct
-validation set, allowing us to train on the \emph{training set} and then test
-on the \emph{validation set}, a subset of the labelled set.  However in this work
-the labelled set is relatively small, about an order of magnitude smaller than 
-what was available in \citet{Casey_2016b}, and therefore we will use all labelled 
-set stars in the training step and supplement our validation experiments with 
-literature comparisons.  Thus, throughout this work we will use the terms 
-\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
-how we constructed the training set, before we outline the model that we adopt.
+Throughout this work we adopt the assumptions and nomenclature first set out in
+\citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
+here with \citet{Caset_2016} is that in this work the \emph{labelled} set and the
+\emph{training} set will be the same set.  In \citet{Casey_2016b} the labelled set
+contained two subsets: the training set and a distinct validation set, allowing us
+to train on the \emph{training set} and then test on the \emph{validation set}, a 
+subset of the labelled set.  However in this work the labelled set is relatively 
+small, about an order of magnitude smaller than what was available in 
+\citet{Casey_2016b}, and therefore we will use all labelled set stars in the 
+training step and supplement our validation experiments with literature comparisons.
+Thus, throughout this work we will use the terms \emph{labelled set} and 
+\emph{training set} interchangeably.  Below we describe how we constructed the 
+training set, before we outline the model that we adopt.
 
@@
 the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
-stars sit above the isochrones, not below).  If we move to lower-resolution studies,
-there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
-S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
-seen on the main-sequence: the labels are expectedly less precise given the lower
-resolution, no detailed abundance information is available\footnote{Abundance 
-information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
-sample contains only giant stars}, and there is only sparse sampling across the 
-Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
-but results from that sample are not yet precise enough for our purposes.
+stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
+If we consider lower-resolution studies, there are 2,369 stars that overlap with 
+\project{LAMOST} --- of which 2,213 have positive S/N ratios in the $g$-band 
+(\texttt{snrg}).  However, similar systematic effects are seen on the main-sequence: 
+the labels are expectedly less precise given the lower resolution, no detailed 
+abundance information is available\footnote{Abundance information is available for 
+\project{LAMOST} stars from \citet{Ho_2016}, but that sample contains only giant stars},
+and there is only sparse sampling across the Hertzsprung-Russell diagram.  The 
+\project{GALAH} survey has a good overlap with \rave, but results from that sample 
+are not yet precise enough for our purposes.
 
@@
 flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
-giant branch, with metallicities ranging from [Fe/H] $= -1.62$ to 0.32.
+giant branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Our tests
+showed that the metallicity range of the training set needed to extend below
+$[{\rm Fe/H}] \lesssim -2$, and thus below we describe how we supplemented this
+sample with giant stars from the \epic\ catalog \citep{Huber_2016}.
 
@@
 % Identifying obviously bad stars that have been mis-classified by K2/EPIC
+
+% Giant stars of low metallicity.
+
 % Refer to the training set figure.
 % Two separate models, with an overlap region.
+% Worry later about label offsets.
 
@@
 % Include velocities and rotational velocities at test time?
+
+\subsection{The model}
+\label{sec:the-model}
 
@@
 
+\bibitem[Huber et al.(2016)]{Huber_2016} Huber, D., Bryson, S.~T., Haas, M.~R., et al.\ 2016, \apjs, 224, 2
+
 \bibitem[Jofr{\'e} et al.(2015)]{Jofre_2015} Jofr{\'e}, P., M{\"a}dler, T., Gilmore, G., et al.\ 2015, \mnras, 453, 1428
@@
 
-\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, in preparation
+\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, submitted
 
@@
 \bibitem[van Leeuwen(2007)]{van_Leeuwen_2007} van Leeuwen, F.\ 2007, \aap, 474, 653
+
+\bibitem[Wonjo et al.(2016)]{Wonjo_2016} Wonjo, J., et al.\ 2016, submitted
 

### 2016-08-23 article/figures/plot_ges_comparison.py
Update figures; include script to plot HRD of training sets

@@ -29,7 +29,7 @@ ges_label_names = {
 }
 
 label_limits = {
-    "TEFF_1": (3000, 6500),
+    "TEFF_1": (3500, 7000),
     "LOGG_1": (0, 5),
     "FE_H": (-2, 0.75)
 }
@@ -43,7 +43,7 @@ latex_labels = {
 
 # Exclude ones we think are bad.
 #ok = data_table["OK"]
-ok = data_table["R_CHI_SQ"] < 3
+ok = (data_table["R_CHI_SQ"] < 3) * (data_table["SNRK"] > 50)
 
 data_table = data_table[ok]
 
@@ -70,7 +70,7 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     y = data_table[ges_label_names[label_name]]
     # yerr
 
-    scat = ax.scatter(x, y, c=data_table["SNRK"], alpha=0.75, s=50, 
+    scat = ax.scatter(x, y, facecolor="#666666", alpha=0.5, s=50, 
         linewidths=0.5, edgecolors="#000000")
 
     limits = label_limits[label_name]
@@ -89,8 +89,8 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     diff = y - x
     print(label_name, np.nanmean(diff), np.nanstd(diff))
 
-cbar = plt.colorbar(scat)
-cbar.set_label(r"$S/N$")
+#cbar = plt.colorbar(scat)
+#cbar.set_label(r"$S/N$")
 fig.tight_layout()
 
 fig.savefig("ges-comparison.pdf", dpi=300)

### 2016-08-23 article/figures/plot_hrd_test.py

@@ -46,11 +46,11 @@ fig.subplots_adjust(
 
 for ax, snr in zip(axes, snrs):
     
-    ok = (t["SNRK"] > snr) * (t["r_chi_sq_ms"] < 3) * (t["r_chi_sq_giant"] < 3)
-
-    ax.hexbin(t["WEIGHTED_TEFF"][ok], t["WEIGHTED_LOGG"][ok], gridsize=N,
+    ok = (t["SNRK"] > snr) * (t["R_CHI_SQ"] < 3)
+    
+    ax.hexbin(t["TEFF"][ok], t["LOGG"][ok], gridsize=N,
         extent=(3500, 7000, 0.5, 5),
-        cmap="Blues", norm=LogNorm(), edgecolor="#eeeeee", linewidths=0.1)
+        cmap="Blues", norm=LogNorm(), edgecolor="#ffffff", linewidths=0.25)
 
 
     K = locale.format("%d", sum(ok), grouping=True).replace(",", "$,$")

### 2016-08-23 article/figures/plot_hrd_train.py

@@ -1,0 +1,94 @@
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+
+
+try:
+    ms_training_set, giant_training_set, joint_training_set
+
+except NameError:
+    from rave_io import get_training_sets
+
+    ms_training_set, giant_training_set, joint_training_set = get_training_sets()    
+
+else:
+    print("Using pre-loaded data")
+
+
+names = (
+    r"${\rm Main-sequence}$ ${\rm star}$ ${\rm model}$", 
+    r"${\rm Giant}$ ${\rm star}$ ${\rm model}$", 
+    r"${\rm Joint}$ ${\rm star}$ ${\rm model}$"
+)
+
+labelled_sets = (ms_training_set, giant_training_set, joint_training_set)
+M = len(labelled_sets)
+
+factor = 3.5
+cbar = 0.15 * factor
+lbdim = 0.2 * factor
+trdim = 0.1 * factor
+whspace = 0.05
+yspace = factor
+xspace = factor * M + factor * (M - 1) * whspace + lbdim * (M - 1) + cbar
+xdim = lbdim + xspace + trdim
+ydim = lbdim + yspace + trdim
+
+fig, axes = plt.subplots(1, M, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+xlim = (7000, 3500)
+ylim = (5.5, 0)
+
+vmin, vmax = (-3, 0.5)
+
+# TODO: Add uncertainties.
+# TODO: Show other training sets in the background of each axes as grey points.
+
+for ax, labelled_set, name in zip(axes, labelled_sets, names):
+    
+    scat = ax.scatter(labelled_set["TEFF"], labelled_set["LOGG"],
+        c=labelled_set["FE_H"], cmap="magma",
+        vmin=vmin, vmax=vmax, alpha=0.75, s=50)
+
+    ax.text(0.05, 0.9, name,
+        horizontalalignment="left", verticalalignment="bottom",
+        transform=ax.transAxes, fontsize=14)
[…]

### 2016-08-23 article/figures/plot_spectral_derivatives.py

@@ -22,7 +22,7 @@ ms_model.vectorizer._label_names = ["TEFF", "LOGG", "FE_H", "VSINI"]
 
 models = (giant_model, ms_model)
 
-comparison_model = tc.load_model("/data/gaia-eso/arc/rave/rave-tgas-v23all.model")   # both
+comparison_model = tc.load_model("/data/gaia-eso/arc/rave/rave-tgas-v26.model")   # both
 
 N = comparison_model.theta.shape[1]
 
@@ -101,6 +101,8 @@ for i in range(N):
 fig.tight_layout()
 fig.subplots_adjust(wspace=0, hspace=0)
 
+
+raise a
 
 def objective_function(xdata, a, b):
     y = xdata * a + b

### 2016-08-23 article/unrave.tex
Update with proper giturl from vc.tex
Minor text changes about labelled/training set

@@
 \citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
-here with \citet{Caset_2016} is that in this work the \emph{labelled} set and the
-\emph{training} set will be the same set.  In \citet{Casey_2016b} the labelled set
-contained two subsets: the training set and a distinct validation set, allowing us
-to train on the \emph{training set} and then test on the \emph{validation set}, a 
-subset of the labelled set.  However in this work the labelled set is relatively 
+here with \citet{Casey_2016b} is that in this work the \emph{labelled} set and the
+\emph{training} set will be the same set.  In \citet{Casey_2016b} two subsets of
+the labelled set were defined: the training set and a distinct validation set.
+This allowing us to train on the \emph{training set} and then test on the 
+\emph{validation set}.  However in this work the labelled set is relatively 
 small, about an order of magnitude smaller than what was available in 
@@
 % Worry later about label offsets.
+% Make sure you mention the lack of abundance information in K2/EPIC
 

### 2016-08-23 article/unrave.tex
Some words about K2/EPIC mis-classifications

@@
 
-\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}}
+\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}\newcommand{\giturl}{UNKNOWN}}
 
@@
 Assembling a training set for the main-sequence and sub-giant branch was more complex.
-% training set from K2/EPIC?!!
+There are no spectroscopic studies that have a large enough overlap with \rave, and
+also cover our required range of stellar parameters (e.g., FGKM-type stars).  Moreover,
+most of the spectroscopic studies we considered exhibit a flat lower main-sequence, a
+consequence of violating the assumption of local thermal equilibrium 
+\citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make
+use of the \epic\ catalog \citep{Huber_2016} for the training set labels.  The \epic\
+catalog follows from the successful \project{Kepler} input catalog \citep{Brown_2011},
+and provides probabilistic stellar classifications for $\gtrsim180{,}000$ stars in 
+the \project{K2} fields based on the astrometric, photometric, and spectroscopic
+information available.  There are 4,611 stars that overlap between \epic\ and \rave.
+
+When the \epic\ posteriors are narrow (i.e., the uncertainties on labels are small),
+\citet{Huber_2016} show that the labels are accurate and trustworthy.  However,
+\epic\ differs from the \project{Kepler} input catalog because \epic\ does not
+benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant
+classification.  For this reason there are a number of stars that are mis-classified
+in the \epic\ catalog.  \citet{Huber_2016} acknowledges this, noting: the fraction 
+of X classified as Y could be as much as Z\%; up to Z2\% of X2 could be marked as Y2, 
+and \stub{other things}.  In other words, the \epic\ labels are usually of high 
+fidelity, and mis-classifications are clearly detectable if spectra are available.
+As we will show, when the mis-classifications can be identified from \rave\ spectra, 
+the \epic\ catalog makes a suitable training set for the main-sequence and sub-giant
+branch in \rave.
+
 % Identifying obviously bad stars that have been mis-classified by K2/EPIC
@@
 % TODO: Replace ZENODO-URL with a data url
-Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  
+Source code for this project is available at \texttt{\giturl}.
 This document was compiled from revision hash \texttt{\githash} in that repository.
@@
 
+\bibitem[Bensby et al.(2014)]{Bensby_2014} Bensby, T., Feltzing, S., \& Oey, M.~S.\ 2014, \aap, 562, A71
+
 \bibitem[Bovy et al.(2016)]{Bovy_2016} Bovy, J., Rix, H.-W., Schlafly, E.~F., et al.\ 2016, \apj, 823, 30
+
+\bibitem[Brown et al.(2011)]{Brown_2011} Brown, T.~M., Latham, D.~W., Everett, M.~E., \& Esquerdo, G.~A.\ 2011, \aj, 142, 112
 

### 2016-08-23 article/unrave.tex
Some stubs for me to write out tomorrow

@@
 
+% Stubs for discussion & analysis.
+% TODO: Inserting 2MASS photometry as fake pixels
+% TODO: Censoring masks on abundance labels
+% TODO: Priors on astrophysical parameters from isochrones
+% TODO: Include velocities and rotational velocities at test time?
+
 \begin{abstract}
@@
 This premise relies on the ability to measure --- for large numbers of stars ---
-3D positions and motions, and their detailed chemical abundances.  On 14 September
-2016 the \gaia\ mission will release the former for $\approx2\times10^6$ stars.
+their detailed chemical abundances, and their 3D positions and motions.  On 14 September
+2016 the \gaia\ mission will release the latter for $\approx2\times10^6$ stars.
 However, a catalog of precise chemical abundances to complement that data
@@
 % Identifying obviously bad stars that have been mis-classified by K2/EPIC
-
-% Giant stars of low metallicity.
-
-% Refer to the training set figure.
-% Two separate models, with an overlap region.
-% Worry later about label offsets.
+% Requiring stars with narrow posteriors.
 % Make sure you mention the lack of abundance information in K2/EPIC
 
-% Initially we train on three labels.
-
-% Inserting 2MASS photometry as fake pixels
-
-% Magic wizardry for missing and imperfect labels.
-% Censoring masks on abundance labels
-% Priors on astrophysical parameters from isochrones
-
-% Include velocities and rotational velocities at test time?
+% Including Giant stars of low metallicity.
+% Refer to the training set figure.
+% Worry later about label offsets.
 
@@
 \label{sec:the-model}
+
+% Describe The Cannon basic model, vectorizer, etc.
+% Fitting time errors, etc.
+
+% Here we will first consider only a three parameter model.
+% APOGEE used only quadratic model, worked for giants
+% We experimented with a quadratic model only. 
+% Useful for identifying mis-classifications but the structure wasn't all there..
+% worked but didn't show the structure in detail.
+% Split into two different models with partial overlap.
+% One for giants and one for dwarfs.
+% The labels for stars in the training sets are shown in Figure X.
+
+% We trained all three models.
+% Note the differences in metallicity scales.
+% The K2/EPIC stuff is accurate but imprecise, whereas the APOGEE stuff is precise, usually precise (except Ti?)
+% Look at spectral derivatives.
+% Show the model internals are relatable.
+% Map spectral derivatives from giant model to main-sequence model.
+% Update the model with those new spectral derivatives.
+
+% We ran all three models on all spectra.
+% How we combined the results into one.
+% Refer to test set HRD.
+
+% Restricted to 3-label model, but the giant set includes abundance labels.
+% The spectral range covered by \rave\ includes atomic transitions of X, Y, Z elements.
+% The strongest lines are indicated in Figure XXX.
+
+% Censoring masks
+% Transfering coefficients, etc.
 
@@
 X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).
+% Fill this in tomorrow.
 
@@
 % --> Stars with morphological classification
-% --> Anything else?
+% --> Some rogue gallery of a few high S/N cases that have clearly gone haywire?
 
@@
 \acknowledgements
-A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
+A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sergey Koposov (Cambridge),
+and Sven Buder (MPIA).
 This research made use of: Astropy, a community-developed core Python package for

### 2016-08-24 article/unrave.tex
Text edits and new text on the k2/epic ms training set

@@
 \newcommand{\rave}{\project{\acronym{RAVE}}}
-\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
-\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
+\newcommand{\galah}{\project{\acronym{GALAH}}}
+\newcommand{\ges}{\project{Gaia-ESO}}
+\newcommand{\apogee}{\project{\acronym{APOGEE}}}
+\newcommand{\aspcap}{\project{\acronym{ASPCAP}}}
+\newcommand{\lamost}{\project{\acronym{LAMOST}}}
 \newcommand{\hipparcos}{\project{Hipparcos}}
@@
 The Milky Way may be our greatest laboratory for understanding galaxy formation.
-This premise relies on the ability to measure --- for large numbers of stars ---
-their detailed chemical abundances, and their 3D positions and motions.  On 14 September
-2016 the \gaia\ mission will release the latter for $\approx2\times10^6$ stars.
-However, a catalog of precise chemical abundances to complement that data
-release is currently lacking.  The RAdial Velocity Experiment (\rave) survey has 
-acquired spectra for 292,036 stars in common with the first \gaia\ data release, 
-constituting the largest overlap of any stellar spectroscopic sample.  Here we 
-perform an independent analysis of \rave\ spectra using an implementation of 
-\thecannon\ that uses prior probabilities on atomic line formation and stellar
+This premise relies on the ability to precisely determine the astrometry and
+chemistry for many individual stars, and to use those data to make inferences
+about the Galaxy's formation and evolution. On 14 September 2016 the \gaia\ 
+mission will release precise astrometry for $\approx2\times10^6$ stars.  
+However, a comprehensive catalog of detailed chemical abundances to complement
+those data is currently lacking.  The RAdial Velocity Experiment (\rave) survey 
+has acquired spectra for 292,036 stars in common with the first \gaia\ data 
+release, constituting the largest overlap of any stellar spectroscopic sample.  
+Here we perform an independent analysis of \rave\ spectra using an implementation
+of \thecannon\ that uses prior probabilities on atomic line formation and stellar
 astrophysical parameters.  Our approach yields precise stellar labels at lower 
-S/N ratios than most \emph{ab initio} approaches, allowing us to deliver
-improved radial velocities, effective temperature $\teff$, surface gravity $\logg$,
-and chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) 
-for \Nstars\ stars.  We validate our results internally from repeat visits, 
-and externally against existing high-resolution spectroscopic surveys.  The 
-typical precision in chemical abundances is X.XX~dex.  We provide 
-spectrophotometric distances based on our labels, improving upon the parallax
-distances available for the fainter targets in the first \gaia\ data release.
-\stub{A one-line science result based on RAVE/Hipparcos overlap}
+S/N ratios than most \emph{ab initio} approaches, allowing us to deliver improved
+radial velocities, effective temperature $\teff$, surface gravity $\logg$, and 
+chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) for 
+\Nstars\ stars.  We validate our results internally from repeat visits, and 
+externally against existing high-resolution spectroscopic surveys.  The typical
+precision in chemical abundances is X.XX~dex.  We provide spectrophotometric 
+distances based on our labels, improving upon the parallax distances available
+for the fainter targets in the first \gaia\ data release. \stub{A one-line 
+science result based on RAVE/Hipparcos overlap}
 \end{abstract}
@@
 stellar spectroscopists frequently report significantly different chemical 
-abundance patterns \citep{Smiljanic_2014}.  The impact this lack of data has on
-scientific inferences cannot be understated.  It limits our understanding on a
+abundance patterns \citep{Smiljanic_2014}.  The impact these issues have on our
+scientific inferences cannot be understated.  They limit our understanding on a
 number of sub-fields in astrophysics, including the properties of exoplanet host
@@
 parallaxes and proper motions for more than $10^9$ stars in its final data
-release, scheduled for 2022.  While this is a sample size about three orders 
-of magnitude larger than its predecessor \hipparcos, both astrometry and 
-chemistry are required to fully characterize the formation and evolution of our Galaxy.  
-For this reason \gaia\ will also provide radial velocities, stellar parameters 
-and chemical abundances for a subset of brighter stars, primarily from the on-board
-radial velocity spectrometer.  However, those chemical abundances will not be 
-available until later releases.  Until those abundances are available,
+release in 2022.  While this is a sample size about three orders of magnitude 
+larger than its predecessor \hipparcos, both astrometry and chemistry are 
+required to fully characterize the formation and evolution of our Galaxy.  For
+this reason \gaia\ will also provide radial velocities, stellar parameters and
+chemical abundances for a subset of brighter stars.  While astrometry for bright
+and nearby stars will be imminently available, chemical abundances from \gaia\ 
+will not be available until later releases.  Until those abundances are available,
 astronomers seeking to use chemistry and kinematic measurements are reliant on 
-ground-based spectroscopic surveys in order to complement the available astrometry
-from \gaia.
+ground-based spectroscopic surveys in order to complement the available astrometry.
 
@@
 stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
-catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{Gaia-ESO} 
-\citep{Gilmore_2012,Randich_2013}, \project{GALAH} \citep{DeSilva_2015},
-\project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
+catalog against the \apogee\ \citep{Zasowski_2013}, \ges\
+\citep{Gilmore_2012,Randich_2013}, \galah\ \citep{DeSilva_2015},
+\lamost\ \citep{Cui_2012}, and \rave\ \citep{Steinmetz_2006} 
 surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
-have the largest overlap with the first \project{Gaia} data release: 292,036 
-stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
+have the largest overlap with the first \gaia\ data release: 292,036 
+stars.  We then used the \gaia\ universe model snapshot \citep{Robin_2012} to 
 estimate the precision in parallax and proper motions that will be available in 
-\project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected
-precision to what is currently available, we further found that the \project{RAVE}
-survey will benefit most from \project{Gaia} DR1.  The distances of 63\% of the 
-\project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to
-improve with the first \project{Gaia} data release, and 47\% of stars likely to
-have better proper motions (137,211 stars).  This motivated us to examine what 
-chemical abundance information was available for \rave, and to evaluate whether 
-we could enable new chemodynamic studies by contributing to this existing set of
-chemical abundances.
+the first \gaia\ data release (DR1) for stars in those overlap samples. Comparing
+the expected precision to what is currently available, we further found that the
+\rave\ survey will benefit most from \gaia\ DR1.  The distances of 63\% of the 
+\rave--\gaia\ DR1 overlap sample (182,862 stars) are expected to improve with 
+the first \gaia\ data release, and 47\% of stars are likely to have better 
+proper motions (137,211 stars).  This motivated us to examine what chemical 
+abundance information was available for \rave, and to evaluate whether we could
+enable new chemodynamic studies by contributing to this existing set of chemical
+abundances.
 
-We briefly describe the \rave\ data in section \ref{sec:data}, before explaining
-the model we employ in section \ref{sec:model}.  In section \ref{sec:validation}
+We briefly describe the \rave\ data in Section~\ref{sec:data}, before explaining
+our method in Section~\ref{sec:method}.  In Section~\ref{sec:validation}
 we outline a number of validation experiments, including: internal sanity checks,
 comparisons with literature samples, and investigations to ensure our results
-are consistent with expectations from astrophysics.  We provide concluding
-remarks in section \ref{sec:conclusion}, as well as a description of how to
-access our results electronically.
+are consistent with expectations from astrophysics.  We discuss these results in
+Section~\ref{sec:discussion}, and in Section~\ref{sec:conclusion} we provide 
+concluding remarks as well as a description of how to access our results electronically.
 
@@
 2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
-made for very efficient observing: spectra for up to 150 targets could be 
-simultaneously acquired.  When observations concluded in April 2013, a total of 
-574,630 spectra had been collected of 483,330 unique objects.
+made for very efficient observing.  Spectra for up to 150 targets were simultaneously
+acquired, and when observations concluded in April 2013, a total of 574,630 spectra
+had been collected of 483,330 unique objects.
 
@@
 $9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
-The $I$ band was used because it overlaps with the wavelength range that \rave\ 
-operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
-near-infrared triplet.  These stellar absorption lines are dominated by pressure 
-broadening, which makes them strong and therefore visible even in very low S/N data
-or metal-poor stars.  This spectral region is also well-studied.  It is one of the
-key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
+The $I$ band was selected because it overlaps with the wavelength range that \rave\ 
+operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared 
+triplet, and a number of atomic transitions of light-, $\alpha$-, and Fe-peak elements.
+The \ion{Ca}{2} near infrared triplet lines are dominated by pressure broadening, 
+which makes them strong and therefore visible even in very low S/N data or 
+metal-poor stars.  This spectral region is also very well-studied.  It is one of the
+key setups used for the ground-based high-resolution \ges\ survey
 \citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
-coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
+coverage is comparable to the Radial Velocity Spectrometer on board the \gaia\
 space telescope \citep{Recio-Blanco_2016}.
@@
 $\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
-$\approx$50~per~pixel$^{-1}$.  For comparison, the \project{GALAH} survey 
+$\approx$50~per~pixel$^{-1}$.  For comparison, the \galah\ survey 
 \citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
 abundance analyses --- includes a wavelength range about 2.5 times larger at 
-resolution $\mathcal{R} \approx 28,000$, and yet the \project{GALAH} project still 
+resolution $\mathcal{R} \approx 28,000$, and yet the \galah\ project still 
 targets for S/N $\gtrsim100$.
@@
 30~pixels excluded on either side of the common wavelength array, leaving us with
-945~pixels for science.
+945~pixels per spectrum for science.
 
@@
 small, about an order of magnitude smaller than what was available in 
-\citet{Casey_2016b}, and therefore we will use all labelled set stars in the 
-training step and supplement our validation experiments with literature comparisons.
-Thus, throughout this work we will use the terms \emph{labelled set} and 
-\emph{training set} interchangeably.  Below we describe how we constructed the 
-training set, before we outline the model that we adopt.
+\citet{Casey_2016b}, and therefore no validation set exists. We will use all 
+labelled set stars in the training step and supplement our validation experiments 
+with literature comparisons.  Thus, throughout this work we will use the terms 
+\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
+how we constructed the training set, before we outline the model that we adopt.
 
@@
 We sought to construct a training set consisting of stars across the main-sequence,
-sub-giant, and the red giant branch.  We require stars with precisely measured
+sub-giant, and the red giant branch.  We required stars with precisely measured
 effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
@@
 range of \rave\ does not overlap substantially with high-resolution spectroscopic
-surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
+surveys.  The fourth internal data release of the \ges\ survey includes 
 giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
-\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
+\apogee\ data release \citep{sloan_dr13} reports labels for stars on the
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
-that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
+that the main-sequence labels suffered from significant systematic effects.  A flat, 
 then `up-turning' main-sequence is present, and the metallicity gradient trends in 
@@
 If we consider lower-resolution studies, there are 2,369 stars that overlap with 
-\project{LAMOST} --- of which 2,213 have positive S/N ratios in the $g$-band 
-(\texttt{snrg}).  However, similar systematic effects are seen on the main-sequence: 
-the labels are expectedly less precise given the lower resolution, no detailed 
-abundance information is available\footnote{Abundance information is available for 
-\project{LAMOST} stars from \citet{Ho_2016}, but that sample contains only giant stars},
-and there is only sparse sampling across the Hertzsprung-Russell diagram.  The 
-\project{GALAH} survey has a good overlap with \rave, but results from that sample 
-are not yet precise enough for our purposes.
+\lamost\ --- of which 2,213 have positive S/N ratios in the $g$-band 
+(\texttt{snrg}).  However, the labels are expectedly less precise given the lower 
+resolution, no detailed abundance information is available for main-sequence stars
+\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
+but that sample contains only giant stars}, and similar systematic trends are apparent
+in the lower main-sequence.  The \galah\ survey has a good overlap with \rave, but 
+results from that sample are not yet precise enough for our purposes.
 
 These constraints forced us to construct a heterogeneous training set.  Given previous
-successes in transferring high S/N ratio labels from \project{APOGEE} \citep{Ness_2015,
-Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \project{APOGEE}
----\rave\ overlap sample for giant star labels in the training set.  Of these, about
-900 are giants according to \project{APOGEE}.  From this sample we selected stars to
-have: metallicity determinations ([Fe/H] $> -5$); high S/N ratios in both the 
-\project{APOGEE} and \rave\ spectra ($>200$ in \project{APOGEE}; $>50$ in \rave); and
-we further required that the \project{ASPCAP} pipeline did not report any peculiar
-flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
-giant branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Our tests
-showed that the metallicity range of the training set needed to extend below
-$[{\rm Fe/H}] \lesssim -2$, and thus below we describe how we supplemented this
-sample with giant stars from the \epic\ catalog \citep{Huber_2016}.
+successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
+Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
+overlap sample for giant star labels in the training set.  Of these, about 900 are 
+giants according to \apogee.  From this sample we selected stars to have: metallicity
+determinations ([Fe/H] $> -5$); high S/N ratios in both the \apogee\ and \rave\ spectra
+($>200$ in \apogee; $>50$ in \rave); and we further required that the \aspcap\ pipeline 
+did not report any peculiar flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left 
+us with 357 stars along the giant branch, with metallicities ranging from 
+$[{\rm Fe/H}] = -1.62$ to 0.32.  Intermediate tests with globular cluster members showed
+that the metallicity range of the training set needed to extend below 
+$[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be practically useful.
+Therefore, as we will describe in more detail below, we supplemented this sample with 
+giant stars from the \epic\ catalog \citep{Huber_2016}.
 
 Assembling a training set for the main-sequence and sub-giant branch was more complex.
-There are no spectroscopic studies that have a large enough overlap with \rave, and
-also cover our required range of stellar parameters (e.g., FGKM-type stars).  Moreover,
-most of the spectroscopic studies we considered exhibit a flat lower main-sequence, a
-consequence of violating the assumption of local thermal equilibrium 
-\citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make
-use of the \epic\ catalog \citep{Huber_2016} for the training set labels.  The \epic\
-catalog follows from the successful \project{Kepler} input catalog \citep{Brown_2011},
-and provides probabilistic stellar classifications for $\gtrsim180{,}000$ stars in 
-the \project{K2} fields based on the astrometric, photometric, and spectroscopic
-information available.  There are 4,611 stars that overlap between \epic\ and \rave.
+There are no spectroscopic studies that have a large enough sample size that overlap 
+with \rave, and which also cover our required range in stellar parameters 
+(e.g., FGKM-type stars).  Moreover, most of the spectroscopic studies we considered 
+exhibit a flat lower main-sequence, a systematic consequence of the analysis method 
+adopted \citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose
+to make use of the \epic\ catalog \citep{Huber_2016} for the training set labels on
+the main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
+\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
+classifications for 138,600 stars in the \project{K2} fields based on the 
+astrometric, asteroseismic, photometric, and spectroscopic information available for
+every star.  There are 4,611 stars that overlap between \epic\ and \rave.
 
-When the \epic\ posteriors are narrow (i.e., the uncertainties on labels are small),
-\citet{Huber_2016} show that the labels are accurate and trustworthy.  However,
 \epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
 benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
-classification.  For this reason there are a number of stars that are mis-classified
-in the \epic\ catalog.  \citet{Huber_2016} acknowledges this, noting: the fraction 
-of X classified as Y could be as much as Z\%; up to Z2\% of X2 could be marked as Y2, 
-and \stub{other things}.  In other words, the \epic\ labels are usually of high 
-fidelity, and mis-classifications are clearly detectable if spectra are available.
-As we will show, when the mis-classifications can be identified from \rave\ spectra, 
-the \epic\ catalog makes a suitable training set for the main-sequence and sub-giant
-branch in \rave.
+classification.  Despite this limitation the labels in the \epic\ catalog have 
+already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
+when the posteriors are wide (i.e., the quoted confidence intervals are large) 
+due to limited information available, it is possible that a star has been 
+mis-classified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
+note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
+probability of misclassification is usually quantified in the uncertainties of 
+each star; most sub-giants that could be dwarfs have large confidence intervals.  
+Therefore, requiring low uncertainties will decrease the total sample size, but 
+will remove most misclassifications.  The situation is far more favourable for 
+dwarfs and giants.  Only 1-4\% of giant stars are misclassified as dwarfs, and 
+about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
+with narrow confidence intervals are usually of high fidelity, and any spurious 
+misclassifications would be clearly distinguishable if spectra were available.
 
-% Identifying obviously bad stars that have been mis-classified by K2/EPIC
-% Requiring stars with narrow posteriors.
+We sought to have a small overlap between our giant and main-sequence star training
+sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
+however there is a sparse sampling of stars reaching to $\logg \approx 4$.  For
+the \epic\ main-sequence/sub-giant star training set we required $\logg > 3.5$,
+allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
+employed the following quality constraints: the upper and lower confidence intervals 
+in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
+must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
+30~per~pixel$^{-1}$; and $\teff \leqslant 6750$.  Unfortunately these strict constraints
+removed most metal-poor stars, which we later found to cause the test labels to have
+under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
+(ignored) those quality constraints for stars in the \epic\ catalog with 
+$[{\rm Fe/H}] < -1$.  Of the stars we forcibly included with $[{\rm Fe/H}] < -1$,
+none were classified as being sub-giants in \epic, therefore they have a low (7\%)
+likelihood of misclassification.  The distilled sample of the \rave--\epic\ overlap
+catalog contains 583 stars (of 4,611).
+
+% We can also identify misclassified stars by cross-validation.
+
 % Make sure you mention the lack of abundance information in K2/EPIC

### 2016-08-24 article/unrave.tex
Finish description of the training sets

@@
 will remove most misclassifications.  The situation is far more favourable for 
-dwarfs and giants.  Only 1-4\% of giant stars are misclassified as dwarfs, and 
+dwarfs and giants.  Only $1-4$\% of giant stars are misclassified as dwarfs, and 
 about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
@@
 must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
-30~per~pixel$^{-1}$; and $\teff \leqslant 6750$.  Unfortunately these strict constraints
+30~per~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
 removed most metal-poor stars, which we later found to cause the test labels to have
@@
 none were classified as being sub-giants in \epic, therefore they have a low (7\%)
-likelihood of misclassification.  The distilled sample of the \rave--\epic\ overlap
-catalog contains 583 stars (of 4,611).
+likelihood of misclassification.  After training a model based on main-sequence and
+giant stars (Section \ref{sec:the-model}), we found we could further identify 
+misclassifications by leave-one-out cross-validation (iteratively, even).  However,
+we chose not to do this because the number of likely misclassifications in the training
+set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
+was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
+stars (of 4,611).
 
-% We can also identify misclassified stars by cross-validation.
+We previously alluded to including giant stars from \epic\ in our training set based 
+on \apogee\ labels.  This was principally to ensure an adequate balance of metal-poor
+giant stars in the training set.  Without this inclusion, we again found that the
+abundances of giant stars would be under-predicted at low metallicity, critically 
+affecting any inferences we wish to make with low metallicity stars (e.g., globular
+clusters).  The metallicities in the \epic\ catalog come from a variety of spectroscopic
+sources, or photometrically-selected when no spectroscopic information is available.
+While these heterogeneous metallicities may introduce scaling differences, on balance we
+later found that it was more important to include reasonably accurate (albeit imprecise)
+metallicities of metal-poor stars in the training set rather than excluding metal-poor
+stars entirely.  In Section~\ref{sec:validation} we show that these potential issues
+have a negligible impact on the test set labels, and we further show that our abundances
+are consistent with high-resolution studies in the literature.
 
-% Make sure you mention the lack of abundance information in K2/EPIC
-
-% Including Giant stars of low metallicity.
-% Refer to the training set figure.
-% Worry later about label offsets.
+We have constructed a justified training set for stars across the main-sequence, sub-giant,
+and red giant branch.  However the lack of overlap between \rave\ and other works have
+resulted in a somewhat peculiar situation. Detailed abundances are available from \apogee\
+for all giant stars in our sample, but only imprecise metallicities are available from
+\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section
+we will construct a simplified model that does not include detailed abundances, before
+demonstrating how we (1) improve upon the imprecise metallicities for main-sequence/sub-giant
+stars, and (2) translate abundances from the giant branch stars to the main-sequence/sub-giant
+branch sample.
 

### 2016-08-26 article/figures/plot_bensby.py
Update figure scripts

@@ -1,0 +1,79 @@
+
+"""
+Make label comparisons with Bensby et al. (2014).
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+try:
+    bensby
+
+except NameError: # Do you know who I am?
+
+    from rave_io import get_cannon_dr1, get_literature_bensby
+
+
+    rave_cannon_dr1 = get_cannon_dr1()
+    #OK = (data["SNRK"] > 10) * (data["R_CHI_SQ"] < 3) * (data["R"] > 25)
+    #rave_cannon_dr1 = rave_cannon_dr1[OK]
+
+    bensby = get_literature_bensby()
+
+    from astropy.table import join
+
+    data = join(rave_cannon_dr1, bensby, keys=("Name", ))
+    ok = (data["SNRK"] > 10) * (data["R_CHI_SQ"] < 3) * (data["R"] > 25)
+    data = data[ok].filled()
+
+else:
+    print("Using pre-loaded data!")
+
+latex_labels = {
+    "TEFF": r"$T_{\rm eff}$",
+    "LOGG": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$"
+}
+
+def scatter_comparison(axis, bensby_label_name, label_name, c=None):
+    """
+    Show a scatter plot on the given axis.
+    """
+
+    x = data[bensby_label_name]
+    y = data[label_name]
+    c = data["Teff"]
+
+    axis.scatter(x, y, c=c)
+
+    limits = np.array([ax.get_xlim(), ax.get_ylim()]).flatten()
+    limits = [np.min(limits), np.max(limits)]    
+    axis.plot(limits, limits, c="#666666", zorder=-1, linestyle=":")
+    axis.set_xlim(limits)
+    axis.set_ylim(limits)
+
+    diff = y - x
+    print(label_name, np.nanmean(diff), np.nanstd(diff))
+
+    axis.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm Bensby}+$ $2014)$"]))
+    axis.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm unRAVE})$"]))
+
+
+# Compare teff, logg, [Fe/H]
+fig, axes = plt.subplots(1, 3)
+
+labels = [
+    ("TEFF", "Teff"),
+    ("LOGG", "logg"),
+    ("FE_H", "Fe_H")
+]
+
[…]

### 2016-08-26 article/figures/plot_ges_comparison.py

@@ -9,11 +9,11 @@ try:
     data_table
 
 except NameError:
-    from rave_io import (rave_cannon_dr1, ges_idr4)
+    from rave_io import (get_cannon_dr1, get_ges_idr4)
 
     from astropy.table import join
 
-    data_table = join(rave_cannon_dr1, ges_idr4, keys=("Name", ))
+    data_table = join(get_cannon_dr1(), get_ges_idr4(), keys=("Name", ))
 
 else:
     print("Warning: Using pre-loaded data!")
@@ -65,12 +65,12 @@ fig.subplots_adjust(
 
 for i, (ax, label_name) in enumerate(zip(axes, label_names)):
 
-    x = data_table[label_name]
+    y = data_table[label_name]
     #xerr
-    y = data_table[ges_label_names[label_name]]
+    x = data_table[ges_label_names[label_name]]
     # yerr
 
-    scat = ax.scatter(x, y, facecolor="#666666", alpha=0.5, s=50, 
+    scat = ax.scatter(x, y, c=data_table[ges_label_names["LOGG_1"]], facecolor="#666666", alpha=0.5, s=50, 
         linewidths=0.5, edgecolors="#000000")
 
     limits = label_limits[label_name]
@@ -80,8 +80,8 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm UNRAVE})$"]))
-    ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm GES})$"]))
+    ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm UNRAVE})$"]))
+    ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm GES})$"]))
 
     [_.set_rotation(30) for _ in ax.get_xticklabels()]
     [_.set_rotation(30) for _ in ax.get_yticklabels()]

### 2016-08-26 article/figures/plot_gold_standard_comparisons.py

@@ -1,0 +1,126 @@
+
+"""
+Compare our results to the gold standard studies of Bensby et al. (2014), 
+Reddy et al. (2003, 2006), and Valenti & Fischer (2005).
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+try:
+    bensby_2014, reddy_2003, reddy_2006, valenti_2005
+
+except NameError: # Do you know who I am? That's Jeff Vader!
+
+    from rave_io import \
+        (get_cannon_dr1, get_literature_bensby, get_literature_reddy_2003,
+            get_literature_reddy_2006, get_literature_valenti_2005)
+
+    from astropy.table import join
+
+    rave_cannon_dr1 = get_cannon_dr1()
+    
+    OK = (rave_cannon_dr1["SNRK"] > 10) * (rave_cannon_dr1["R_CHI_SQ"] < 3) * (rave_cannon_dr1["R"] > 25)
+    rave_cannon_dr1 = rave_cannon_dr1[OK].filled()
+
+    bensby_2014 = join(rave_cannon_dr1, get_literature_bensby(), keys=("Name", ))
+    reddy_2003 = join(rave_cannon_dr1, get_literature_reddy_2003(), keys=("Name", ))
+    reddy_2006 = join(rave_cannon_dr1, get_literature_reddy_2006(), keys=("Name", ))
+    valenti_2005 = join(rave_cannon_dr1, get_literature_valenti_2005(), keys=("Name", ))
+
+else:
+    print("Using pre-loaded data")
+
+
+# Create convenience functions 
+
+def get_data(table, cannon_label, translated_label):
+    x = table[translated_label]
+    #xerr #TODO
+    y = table[cannon_label]
+    # yerr #TODO
+    return (x, y)
+
+
+def bensby_2014_data(cannon_label):
+
+    label_name = {
+        "TEFF": "Teff",
+        "LOGG": "logg",
+        "FE_H": "Fe_H"
+    }[cannon_label]
+
+    return get_data(bensby_2014, cannon_label, label_name)
+
+
+def reddy_2003_data(cannon_label):
+
+    label_name = {
+        "TEFF": "Teff",
+        "LOGG": "logg",
+        "FE_H": "__Fe_H__2"
+    }[cannon_label]
+
+    return get_data(reddy_2003, cannon_label, label_name)
+
+
+def reddy_2006_data(cannon_label):
+
+    label_name = {
[…]

### 2016-08-26 article/figures/plot_hrd_test_color.py

@@ -1,0 +1,93 @@
+
+
+import locale
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+
+locale.setlocale(locale.LC_ALL, 'en_US')
+
+try:
+    rave_cannon_dr1
+
+except NameError:
+    from rave_io import rave_cannon_dr1
+
+else:
+    print("Using pre-loaded data")
+
+
+
+
+t = rave_cannon_dr1
+
+#ok = (t["SNRK"] > 50) * (t["r_chi_sq_ms"] < 3) * (t["r_chi_sq_giant"] < 3) #* (t["WEIGHTED_VSINI"] < 1)
+
+xlim = (7500, 3500)
+ylim = (5.5, 0)
+vmin, vmax = (-3, 0.5)
+
+snrs = (100, 50, 25, 10)
+M, N = (len(snrs), 50)
+
+
+factor = 3.5
+lbdim = 0.2 * factor
+trdim = 0.1 * factor
+whspace = 0.05
+yspace = factor
+xspace = factor * M + factor * (M - 1) * whspace + lbdim * (M - 1)
+xdim = lbdim + xspace + trdim
+ydim = lbdim + yspace + trdim
+
+fig, axes = plt.subplots(1, M, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+for ax, snr in zip(axes, snrs):
+    
+    ok = (t["SNRK"] > snr) * (t["R_CHI_SQ"] < 3) * (t["R"] > 25) #* (t["VSINI"] < 1)
+
+    ax.scatter(t["TEFF"][ok], t["LOGG"][ok], c=t["FE_H"][ok],
+        vmin=vmin, vmax=vmax, alpha=0.25, s=50, edgecolor="none", cmap="plasma",
+        rasterized=True)
+
+    if ax.is_last_col():
+        scat = ax.scatter([0], [0], c=[0], vmin=vmin, vmax=vmax, cmap="plasma")
+
+    K = locale.format("%d", sum(ok), grouping=True).replace(",", "$,$")
+
+    ax.text(0.05, 0.9, r"$S/N > {:.0f}$".format(snr),
[…]

### 2016-08-26 article/figures/plot_reddy.py

@@ -1,0 +1,83 @@
+
+"""
+Make label comparisons with Reddy et al. (2003, 2006).
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+try:
+    reddy_data
+
+except NameError: # Do you know who I am?
+
+    from rave_io import get_cannon_dr1, get_literature_reddy_2003, get_literature_reddy_2006
+
+
+    rave_cannon_dr1 = get_cannon_dr1()
+    #OK = (data["SNRK"] > 10) * (data["R_CHI_SQ"] < 3) * (data["R"] > 25)
+    #rave_cannon_dr1 = rave_cannon_dr1[OK]
+
+    reddy_2003 = get_literature_reddy_2003()
+    reddy_2006 = get_literature_reddy_2006()
+
+    from astropy.table import join, vstack
+
+    reddy = vstack([reddy_2003, reddy_2006])
+
+    reddy_data = join(rave_cannon_dr1, reddy, keys=("Name", ))
+    
+    #ok = (data["SNRK"] > 10) * (data["R_CHI_SQ"] < 3) * (data["R"] > 25)
+    #data = data[ok].filled()
+
+else:
+    print("Using pre-loaded data!")
+
+latex_labels = {
+    "TEFF": r"$T_{\rm eff}$",
+    "LOGG": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$"
+}
+
+def scatter_comparison(axis, reddy_label, label_name, c=None):
+    """
+    Show a scatter plot on the given axis.
+    """
+
+    x = reddy_data[reddy_label]
+    y = reddy_data[label_name]
+    c = reddy_data["__Fe_H__2"]
+
+    axis.scatter(x, y, c=c)
+
+    limits = np.array([ax.get_xlim(), ax.get_ylim()]).flatten()
+    limits = [np.min(limits), np.max(limits)]    
+    axis.plot(limits, limits, c="#666666", zorder=-1, linestyle=":")
+    axis.set_xlim(limits)
+    axis.set_ylim(limits)
+
+    diff = y - x
+    print(label_name, np.nanmean(diff), np.nanstd(diff))
+
+    axis.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm Reddy}+$ $2003$,$2006)$"]))
[…]

### 2016-08-26 article/figures/plot_valenti.py

@@ -1,0 +1,80 @@
+
+"""
+Make label comparisons with Reddy et al. (2003, 2006).
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+try:
+    vf_data
+
+except NameError: # Do you know who I am?
+
+    from rave_io import get_cannon_dr1, get_literature_valenti_2005
+
+
+    rave_cannon_dr1 = get_cannon_dr1()
+    #OK = (data["SNRK"] > 10) * (data["R_CHI_SQ"] < 3) * (data["R"] > 25)
+    #rave_cannon_dr1 = rave_cannon_dr1[OK]
+
+    valenti_2005 = get_literature_valenti_2005()
+    
+    from astropy.table import join
+
+    vf_data = join(rave_cannon_dr1, valenti_2005, keys=("Name", ))
+    
+    ok = (vf_data["SNRK"] > 10) * (vf_data["R_CHI_SQ"] < 3) * (vf_data["R"] > 25)
+    vf_data = vf_data[ok].filled()
+
+else:
+    print("Using pre-loaded data!")
+
+latex_labels = {
+    "TEFF": r"$T_{\rm eff}$",
+    "LOGG": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$"
+}
+
+def scatter_comparison(axis, literature_label, label_name, c=None):
+    """
+    Show a scatter plot on the given axis.
+    """
+
+    x = vf_data[literature_label]
+    y = vf_data[label_name]
+    c = vf_data["__Fe_H__2"]
+
+    axis.scatter(x, y, c=c)
+
+    limits = np.array([ax.get_xlim(), ax.get_ylim()]).flatten()
+    limits = [np.min(limits), np.max(limits)]    
+    axis.plot(limits, limits, c="#666666", zorder=-1, linestyle=":")
+    axis.set_xlim(limits)
+    axis.set_ylim(limits)
+
+    diff = y - x
+    print(label_name, np.nanmean(diff), np.nanstd(diff))
+
+    axis.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm Valenti}+$ $2005)$"]))
+    axis.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm unRAVE})$"]))
+
+
+# Compare teff, logg, [Fe/H]
+fig, axes = plt.subplots(1, 3)
+
+labels = [
+    ("TEFF", "Teff"),
[…]

### 2016-08-26 article/figures/rave_io.py

@@ -5,27 +5,64 @@ from astropy.table import Table
 
 DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
 
-rave_kordopatis_dr4 = Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
-rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))
+
+def get_rave_kordopatis_dr4():
+    return Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
 
 
-if "OK" not in rave_cannon_dr1.dtype.names:
-    try:
-        rave_cannon_dr1["OK"] = (rave_cannon_dr1["SNRK"] > 50) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
-    except KeyError:
-        print("No 'OK' subset")
+
+def get_cannon_dr1():
+    rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, "rave-tgas-v1.fits"))
+
+    if "OK" not in rave_cannon_dr1.dtype.names:
+        try:
+            rave_cannon_dr1["OK"] = (rave_cannon_dr1["SNRK"] > 50) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
+        except KeyError:
+            print("No 'OK' subset")
+
+        else:
+            print("Setting OK as stars with S/N > 50 and R_CHI_SQ < 3")
 
     else:
-        print("Setting OK as stars with S/N > 50 and R_CHI_SQ < 3")
+        print("Using pre-defined subset of 'OK'")
 
-else:
-    print("Using pre-defined subset of 'OK'")
+    return rave_cannon_dr1
 
 
-ges_idr4 = Table.read(os.path.join(DATA_PATH, "literature-ges-idr4-xm.fits"))
-ges_idr4["Name"] = [each.strip() for each in ges_idr4["Name"]]
+def get_ges_idr4():
 
-literature_pastel = Table.read(os.path.join(DATA_PATH, "literature-pastel-database-xm.fits"))
+    ges_idr4 = Table.read(os.path.join(DATA_PATH, "literature-ges-idr4-xm.fits"))
+    ges_idr4["Name"] = [each.strip() for each in ges_idr4["Name"]]
+    return ges_idr4
+
+
+def get_literature_pastel():
+    return Table.read(os.path.join(DATA_PATH, "literature-pastel-database-xm.fits"))
+
+
+def get_literature_bensby():
[…]

### 2016-08-26 article/unrave.tex
new problem..

@@
 
+\newcommand{\Dvector}[1]{\boldsymbol{#1}}
+\newcommand{\vectheta}{\Dvector{\theta}}
+\newcommand{\vecv}{\Dvector{v}}
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\,}
@@
 proper motions (137,211 stars).  This motivated us to examine what chemical 
-abundance information was available for \rave, and to evaluate whether we could
-enable new chemodynamic studies by contributing to this existing set of chemical
-abundances.
+abundances were available for \rave, and to evaluate whether we could enable new 
+chemodynamic studies by contributing to this existing set of chemical abundances.
 
@@
 for all giant stars in our sample, but only imprecise metallicities are available from
-\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section
-we will construct a simplified model that does not include detailed abundances, before
-demonstrating how we (1) improve upon the imprecise metallicities for main-sequence/sub-giant
-stars, and (2) translate abundances from the giant branch stars to the main-sequence/sub-giant
-branch sample.
+\epic\ for stars on the main-sequence and the sub-giant branch.  To our knowledge this 
+problem has not been addressed with data-driven models in the astrophysics literature.
+In the following section we will construct a simplified model that does not include detailed
+abundances, before demonstrating how we (1) improve upon the imprecise metallicities for 
+main-sequence/sub-giant stars, and (2) translate abundances from the giant branch stars 
+to the main-sequence/sub-giant branch sample.
 
@@
 
-% Describe The Cannon basic model, vectorizer, etc.
-% Fitting time errors, etc.
+\noindent{}The model we adopt is
+\begin{eqnarray}\label{eq:model}
+y_{jn} & = & \vecv(\ell_n)\cdot\vectheta_j + e_{jn}\quad ,
+\end{eqnarray}
 
-% Here we will first consider only a three parameter model.
-% APOGEE used only quadratic model, worked for giants
-% We experimented with a quadratic model only. 
-% Useful for identifying mis-classifications but the structure wasn't all there..
+\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
+$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
+$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
+$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
+wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
+vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Section 
+\ref{sec:a-simple-model}).  The noise values $e_{jn}$ can be considered to be drawn 
+from a Gaussian distribution with zero mean and variance $\sigma_{jn}^2 + s_j^2$, where
+$\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and $s_j^2$ describes the excess 
+variance at the $j$-th wavelength pixel.  We direct the reader to \citet{Ness_2015} for
+a more detailed introduction and justification of \thecannon.
+
+At the training step we fix the $K$-lists of labels for all the training set stars $n$.
+At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
+by optimizing the likelihood function:
+\begin{eqnarray}\label{eq:train}
+\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
+    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
+    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2)
+    \right]
+  \quad .
+\end{eqnarray}
+
+In practice this is achieved by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a
+convex optimization problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
+We can also adopt strict priors on $\vectheta$ such that labels can only make non-zero
+contributions to the data in certain wavelength ranges.  That is to say, we can specify
+that certain labels (e.g., [Al/H]) can only contribute to the model at wavelength pixels
+where we know an aluminium transition exists.  This method, which we call \emph{censoring},
+is described in Section~\ref{sec:censoring} and \citet{Ness_2016b}.
+
+The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
+pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
+the likelihood function is:
+\begin{eqnarray}\label{eq:test}
+  \ell_m &\leftarrow& \argmin{\ell}\left[
+    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
+    \right]
+  \quad .
+\end{eqnarray}
+
+At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
+$\ell$, which provides us with the formal errors on $\ell$.  We judge the veracity of these
+formal errors in our validation experiments in Section \ref{sec:validation}.
+
+\subsection{Three-label models ($\teff$, $\logg$, [Fe/H])}
+\label{sec:a-simple-model}
+
+% Given previous successes, we will first adopt a three-parameter model for giants + dwarfs.
+% Polynomial thing --> expand out \vectheta
 % worked but didn't show the structure in detail.
@@
 % Note the differences in metallicity scales.
-% The K2/EPIC stuff is accurate but imprecise, whereas the APOGEE stuff is precise, usually precise (except Ti?)
 % Look at spectral derivatives.
@@
 
-% We ran all three models on all spectra.
 % How we combined the results into one.
@@
 
+\subsection{A 10-label model for detailed abundances of giant stars}
+\label{sec:giant-abundance-model}
 % Restricted to 3-label model, but the giant set includes abundance labels.
@@
 % The strongest lines are indicated in Figure XXX.
+% Censoring masks
 
-% Censoring masks
-% Transfering coefficients, etc.
+% Train
 

### 2016-08-31 article/unrave.tex
regularization essential; start writing up all the hacks!

@@
 \newcommand{\epic}{\project{K2/EPIC}}
+\newcommand{\sdss}{\project{\acronym{SDSS}}}
 
@@
 \newcommand{\logg}{\log g}
+\newcommand{\feh}{[\mathrm{Fe/H}]}
 
@@
 stellar spectroscopists frequently report significantly different chemical 
-abundance patterns \citep{Smiljanic_2014}.  The impact these issues have on our
-scientific inferences cannot be understated.  They limit our understanding on a
-number of sub-fields in astrophysics, including the properties of exoplanet host
-stars, the formation (and destruction) of stars and clusters, as well as the 
-structure of the Galactic disk, to name a few.
+abundance patterns for the same star \citep{Smiljanic_2014}.  The impact 
+these issues have on scientific inferences cannot be understated.  Imperfect 
+astrometry or chemistry limits understanding in a number of sub-fields in
+astrophysics, including the properties of exoplanet host stars, the formation 
+(and destruction) of stars and clusters, as well as the structure of the 
+Galactic disk, to name a few.
 
@@
 larger than its predecessor \hipparcos, both astrometry and chemistry are 
-required to fully characterize the formation and evolution of our Galaxy.  For
-this reason \gaia\ will also provide radial velocities, stellar parameters and
-chemical abundances for a subset of brighter stars.  While astrometry for bright
-and nearby stars will be imminently available, chemical abundances from \gaia\ 
-will not be available until later releases.  Until those abundances are available,
+required to fully characterize the formation and evolution of our Galaxy. 
+\gaia\ will also provide radial velocities, stellar parameters and chemical 
+abundances for a subset of brighter stars, but these measurements will not be 
+available in the first few data releases. Until those abundances are available,
 astronomers seeking to use chemistry and kinematic measurements are reliant on 
@@
 We briefly describe the \rave\ data in Section~\ref{sec:data}, before explaining
-our method in Section~\ref{sec:method}.  In Section~\ref{sec:validation}
+our methods in Section~\ref{sec:method}.  In Section~\ref{sec:validation}
 we outline a number of validation experiments, including: internal sanity checks,
@@
 The $I$ band was selected because it overlaps with the wavelength range that \rave\ 
-operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared 
-triplet, and a number of atomic transitions of light-, $\alpha$-, and Fe-peak elements.
-The \ion{Ca}{2} near infrared triplet lines are dominated by pressure broadening, 
-which makes them strong and therefore visible even in very low S/N data or 
-metal-poor stars.  This spectral region is also very well-studied.  It is one of the
-key setups used for the ground-based high-resolution \ges\ survey
+operates in:  8410--8795~\AA.  This spectral region is very well-studied.  It is one 
+of the key setups used for the ground-based high-resolution \ges\ survey
 \citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
 coverage is comparable to the Radial Velocity Spectrometer on board the \gaia\
-space telescope \citep{Recio-Blanco_2016}.
+space telescope \citep{Recio-Blanco_2016}.  The region includes the \ion{Ca}{2} 
+near infrared triplet lines, strong transitions that are dominated by pressure 
+broadening, which are visible even in very low S/N data or metal-poor stars.  
+Atomic transitions of light-, $\alpha$-, and Fe-peak elements are also present, 
+allowing for detailed chemical abundance studies.
 
@@
 there is a limited number of pre-processing steps that we conducted before our
-analysis could begin.  Firstly, we calculated inverse variance arrays from the
-fractional uncertainties provided, then re-sampled the flux and inverse variance
+analysis could begin.  First, we calculated inverse variance arrays from the
+fractional uncertainties provided, and then re-sampled the flux and inverse variance
 arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
@@
 \emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
-how we constructed the training set, before we outline the model that we adopt.
+how we constructed the training set, before we outline the models that we adopt.
 
@@
 surveys.  The fourth internal data release of the \ges\ survey includes 
-giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
-\apogee\ data release \citep{sloan_dr13} reports labels for stars on the
+giant and main-sequence stars, but only 142 overlap with \rave.  The thirteenth data 
+release from \sdss\ \citep{sloan_dr13} includes labels for \apogee\ stars on the
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
@@
 stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
-If we consider lower-resolution studies, there are 2,369 stars that overlap with 
-\lamost\ --- of which 2,213 have positive S/N ratios in the $g$-band 
-(\texttt{snrg}).  However, the labels are expectedly less precise given the lower 
-resolution, no detailed abundance information is available for main-sequence stars
-\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
+If we consider lower-resolution studies as potential training sets, there are 2,369
+stars that overlap with \lamost\ --- of which 2,213 have positive S/N ratios in the 
+$g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
+lower resolution, no detailed abundance information is available for main-sequence 
+stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
 but that sample contains only giant stars}, and similar systematic trends are apparent
@@
 At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
-by optimizing the likelihood function:
+by optimizing the penalized likelihood function:
 \begin{eqnarray}\label{eq:train}
@@
     \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
-    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2)
+    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
     \right]
-  \quad .
+  \quad ,
 \end{eqnarray}
 
-In practice this is achieved by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a
-convex optimization problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
-We can also adopt strict priors on $\vectheta$ such that labels can only make non-zero
-contributions to the data in certain wavelength ranges.  That is to say, we can specify
-that certain labels (e.g., [Al/H]) can only contribute to the model at wavelength pixels
-where we know an aluminium transition exists.  This method, which we call \emph{censoring},
-is described in Section~\ref{sec:censoring} and \citet{Ness_2016b}.
+\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
+for each model in later sections, and $Q(\vectheta)$ is a regularizing function that 
+encourages $\vectheta$ values to take on zero values without breaking convexity:
+\begin{eqnarray}\label{eq:regularization-function}
+	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}|
+\end{eqnarray}
+
+Note that the $d$ subscript here is zero-indexed; the function $Q(\vectheta)$ does not act
+on the (first) $\theta_0$ coefficient, as this is a `pivot point' or offset that we do not
+expect to diminish with increasing regularization (e.g., see equation 
+\ref{eq:vectorizer-three-label-design-matrix}).  In practice equation \ref{eq:train} is 
+optimized by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a convex optimization 
+problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
 
@@
 pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
-the likelihood function is:
+the objective function is:
 \begin{eqnarray}\label{eq:test}
@@
 
-\subsection{Three-label models ($\teff$, $\logg$, [Fe/H])}
+\subsection{Three-label models ($\teff$, $\logg$, $\feh$)}
 \label{sec:a-simple-model}
 
-% Given previous successes, we will first adopt a three-parameter model for giants + dwarfs.
-% Polynomial thing --> expand out \vectheta
+The model we have introduced is generic in the sense that there is no description for how
+complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
+Given previous successes using a second-order polynomial model 
+\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
+model experiments:
+\begin{eqnarray}\label{eq:vectorizer-three-label}
+\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
+\quad ,
+\end{eqnarray}
+
+\noindent{}such that $\vecv(\ell)$ produces the design matrix:
+\begin{tiny}
+\begin{equation}\label{eq:vectorizer-three-label-design-matrix}
+\vecv(\ell) \rightarrow \begin{bmatrix}
+	1 & T_{{\rm eff},0} & \logg_{0} & [{\rm Fe/H}]_{0} & T_{{\rm eff},0}^2 & \logg_{0}\cdot{}T_{{\rm eff},0} & [{\rm Fe/H}]_{0}\cdot{}T_{{\rm eff},0} & \logg_{0}^2 & [{\rm Fe/H}]_{0}\cdot{}\logg_1 & [{\rm Fe/H}]_1^2 \\
+	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
+	1 & T_{{\rm eff},n} & \logg_{n} & [{\rm Fe/H}]_{n} & T_{{\rm eff},n}^2 & \logg_{n}\cdot{}T_{{\rm eff},n} & [{\rm Fe/H}]_{n}\cdot{}T_{{\rm eff},n} & \logg_{n}^2 & [{\rm Fe/H}]_{n}\cdot{}\logg_n & [{\rm Fe/H}]_n^2 \\
+	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
+	1 & T_{{\rm eff},N-1} & \logg_{N-1} & [{\rm Fe/H}]_{N-1} & T_{{\rm eff},N-1}^2 & \logg_{N-1}\cdot{}T_{{\rm eff},N-1} & [{\rm Fe/H}]_{N-1}\cdot{}T_{{\rm eff},N-1} & \logg_{N-1}^2 & [{\rm Fe/H}]_{N-1}\cdot{}\logg_{N-1} & [{\rm Fe/H}]_{N-1}^2
+\end{bmatrix}\quad .
+\end{equation}
+\end{tiny}
+
+\subsubsection{A three-label model for main-sequence and giant stars}\label{sec:a-simple-joint-model}
+
+\subsubsection{A three-label model for main-sequence stars}\label{sec:a-simple-model-ms-stars}
+
 % worked but didn't show the structure in detail.
@@
 \label{sec:giant-abundance-model}
-% Restricted to 3-label model, but the giant set includes abundance labels.
-% The spectral range covered by \rave\ includes atomic transitions of X, Y, Z elements.
-% The strongest lines are indicated in Figure XXX.
-% Censoring masks
 
-% Train
+The giant stars in our training set have stellar parameters and up to 15 
+individual abundances from the \apogee\ \aspcap.  Only a subset of these
+elements have atomic transitions in the \rave\ wavelength region.
+
+% The giant training set has many abundances.
+% Which elements?
+% Tests involving censoring? --> Just indicates that many of these lines are blended.
+% So we will use regularization instead of censoring
+
+% Set hyper-parameters by LOO-CV
+
+% Trained this and rain it on all RAVE spectra.
+
+\subsection{Joining the models}\label{sec:joining-the-models}
 
@@
 % TODO: Replace ZENODO-URL with a data url
-Source code for this project is available at \texttt{\giturl}.
+\noindent{}Source code for this project is available at \texttt{\giturl}.
 This document was compiled from revision hash \texttt{\githash} in that repository.

### 2016-08-31 article/unrave.tex
Some justification of hyperparameters

@@
 
-\subsection{A 10-label model for detailed abundances of giant stars}
+\subsection{A 9-label model for detailed abundances of giant stars}
 \label{sec:giant-abundance-model}
@@
 individual abundances from the \apogee\ \aspcap.  Only a subset of these
-elements have atomic transitions in the \rave\ wavelength region.
+elements have atomic transitions in the \rave\ wavelength region: 
+\ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, 
+\ion{Fe}{1}, and \ion{Ni}{1}.  We excluded [Ti/H] from our abundance list 
+because of systematics in the training set \citep{Holtzman_2015,Hawkins_2016}.
+Therefore we are left with nine labels in our giant star model: $\teff$, 
+$\logg$, and seven individual abundances.
 
-% The giant training set has many abundances.
-% Which elements?
-% Tests involving censoring? --> Just indicates that many of these lines are blended.
-% So we will use regularization instead of censoring
+% ARC asks: Should we discuss any tests done with censoring? 
+%  			Do we feel like we need to defend our choice to use regularization
+%           over censoring? The key result from the censoring tests that I did
+%			is that there are a lot of atomic transitions here, but they are
+% 			mostly blended. We *need* those blended lines because if you
+%			censor all but the strongest line(s) for individual elements then
+% 			you are left with a poor predictor of abundance labels.
 
-% Set hyper-parameters by LOO-CV
+We used a second-order (i.e., $\teff^2$ is the highest term) polynomial 
+vectorizer for the giant star model.  We set the regularization hyperparameter
+by performing a grid search of 30 $\Lambda$ points logarithmically spaced from 
+$10^{-3}$ to $10^{3}$.  We performed leave-one-out cross-validation on the
+training set for each value of $\Lambda$.  Specifically, we removed one star
+from the training set, then trained the model, and inferred labels for the 
+star that was removed from the training set.  We repeated this procedure for
+every star in the training set (for every model with a different $\Lambda$)
+and compared the inferred labels against the training set values.  For
+comparison purposes, we also performed cross-validation on an unregularized 
+($\Lambda = 0$) model using the same training set.  We calculated the bias
+and root-mean-square (RMS) deviation between the input labels and those inferred
+by cross-validation for every model.  The biases were negligible across all
+models.  In Figure \ref{fig:giant-model-hyperparameter-search} we show the
+RMS difference in label recovery compared to the unregularized model as a 
+percentage improvement.  The enveloped region represents the minimum and 
+maximum changes for any label with respect to the unregularized model.
+Depending on the regularization strength $\Lambda$, the RMS in a given label
+can decrease by $\approx30$\% with respect to the unregularized case.  At the
+point of maximum improvement, near $\Lambda = X.XX$, the RMS in all labels has
+decreased by at least $X$\%.  As expected the regularization produces a 
+sparser matrix of $\vectheta$, with a $\approx20$\% increase in the fraction
+of zero-valued entries, particularly for the cross-terms (Figure 
+\ref{fig:giant-model-sparsity}, third panel).  Based on the improvement in 
+RMS by cross-validation, we adopt $\Lambda = X.XX$ for our giant star model.
 
-% Trained this and rain it on all RAVE spectra.
+Results from cross-validation on our regularized model imply that the precision
+in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and varies between X.XX~dex
+and X.XX~dex depending on the elemental abundance.  We made no prior judgement
+on whether a star was likely a giant or not; we used this model (and all other
+models) to test all XXX,XXX \rave\ spectra.  Figure \ref{fig:test-set-hrd} (left
+panel) shows the inferred $\teff$ and $\logg$ values for all \rave\ spectra with
+S/N $\geqslant 50$.  This panel undoubtedly includes dwarf stars that cannot be
+appropriately classified by our giant model.  In Section \ref{sec:joining-the-models}
+we will heuristically combine these results with the main-sequence star model
+described in \ref{sec:main-sequence-model}.
 
 \subsection{Joining the models}\label{sec:joining-the-models}
+
+% All them hax.
 
@@
 
-\subsection{Internal validation}
-We first conducted leave-one-out cross-validation to validate our results.
-For each training set star we removed that star from the training set and
-then re-trained the model.  Using that re-trained model, we then treated
-the excluded star as a test star and compared our inferred labels to the
-test set labels.  We show the results of this experiment in Figure
-\ref{fig:loocv-comparison}.  The bias and root mean square (RMS) error in 
-$\teff$ is XX~K and XX~K, and in $\logg$ is X.XX~dex and X.XX~dex,
-respectively.  The bias in abundance labels is negligible --- between
-X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).
-% Fill this in tomorrow.
+Here we describe a
+
+\subsection{Multi-epoch observations}\label{sec:repeat-observations}
 

### 2016-08-31 article/unrave.tex
Minors

@@
 At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
-$\ell$, which provides us with the formal errors on $\ell$.  We judge the veracity of these
-formal errors in our validation experiments in Section \ref{sec:validation}.
+$\ell$, which provides us with the formal errors on $\ell$.  The formal errors are expected
+to be underestimated, and in Section \ref{sec:validation} we judge the veracity of these 
+errors through validation experiments.
 
@@
 
-% All them hax.
+% All them hax used to combine all the results.
 
-\section{Validation experiments}
-\label{sec:validation}
+% --> Chi-sq of the sample.
+% --> Convex hull? Distance to plane?
 
-Here we describe a
+\section{Validation experiments}\label{sec:validation}
+
+We have already performed leave-one-out cross-validation on all of our models
+as part of the hyperparameter selection.  Here we describe the results from our
+internal, external and astrophysical validation checks.
 
@@
 between the comparison spectrum and a repeat visit, normalized by their 
-formal errors summed in quadrature (e.g., $\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
+formal errors summed in quadrature (e.g., 
+$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
 If our measurements were unbiased by S/N and the errors were representative, 
@@
 % Identify outliers:
-% --> Chi-sq of the sample.
-% --> Convex hull? Distance to plane?
 % --> Stars with morphological classification
@@
 % --> Gaia-ESO Survey
-% --> APOGEE / EPIC (whichever was not used)
 % --> Literature studies presented in Kordopatis+ 2014

### 2016-09-02 article/unrave.tex
De-scoped some aspects: photometry in pixels; censoring; fit velocity at test time; priors on test labels

@@
 % Remove at submission:
-\slugcomment{{\color{red} \textbf{To appear on arXiv on 9 September 2016 (\difftoday{2016}{9}{7}away)}}}
+\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 September 2016 (\difftoday{2016}{9}{7}away)}}}
 
@@
 \author{and the \rave\ collaboration}
-
-% Stubs for discussion & analysis.
-% TODO: Inserting 2MASS photometry as fake pixels
-% TODO: Censoring masks on abundance labels
-% TODO: Priors on astrophysical parameters from isochrones
-% TODO: Include velocities and rotational velocities at test time?
 
@@
 \begin{eqnarray}\label{eq:regularization-function}
-	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}|
+	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}| \quad .
 \end{eqnarray}
@@
 
-The \rave\ survey performed repeat observations for 23,288 stars  with time 
+The \rave\ survey performed repeat observations for 43,918 stars with time 
 intervals ranging from a few hours to up to four years.  This timing was 
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-comparison.png}
+\includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
 \caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:ges-dr4-comparison}}

### 2016-09-02 article/unrave.tex
Explicitly write out our (highest-level) assumptions /ping @davidwhogg

@@
 
+% For AASTeX v6
 %\AuthorCallLimit=10
+%\fullcollaborationName{The \rave\ Collaboration}
 
@@
 radial velocities, effective temperature $\teff$, surface gravity $\logg$, and 
-chemical abundances of up to seven elements (Mg, Al, Si, Ca, Ti, Fe, Ni) for 
+chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) for 
 \Nstars\ stars.  We validate our results internally from repeat visits, and 
@@
 
-Throughout this work we adopt the assumptions and nomenclature first set out in
-\citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
-here with \citet{Casey_2016b} is that in this work the \emph{labelled} set and the
-\emph{training} set will be the same set.  In \citet{Casey_2016b} two subsets of
-the labelled set were defined: the training set and a distinct validation set.
-This allowing us to train on the \emph{training set} and then test on the 
-\emph{validation set}.  However in this work the labelled set is relatively 
-small, about an order of magnitude smaller than what was available in 
-\citet{Casey_2016b}, and therefore no validation set exists. We will use all 
-labelled set stars in the training step and supplement our validation experiments 
-with literature comparisons.  Thus, throughout this work we will use the terms 
-\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
-how we constructed the training set, before we outline the models that we adopt.
+\noindent{}Our work makes the following assumptions about the \rave\ spectra and \thecannon:
+
+\begin{itemize}
+\item We will assume that any fibre- and time-dependent variations in spectral
+resolution in the \rave\ spectra are negligible.
+\item The \rave\ noise variances are approximately correct, independent between
+pixels, and normally distributed.
+\item We assume that the normalization procedure employed by the \rave\ pipeline
+is invariant with respect to the labels, and invariant with respect to the S/N
+ratio.
+\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
+have similar spectra.
+\item A stellar spectrum is a smooth function of the label values for that star,
+and we assume that the function is smooth enough within a sub-space of the labels
+(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
+with a low-order polynomial in label space.
+\item We have constructed a training set of stars with mean accurate labels.  Note
+that we do not assume that \emph{every} label in the training is accurate -- we can
+afford to have a small fraction of inaccurate labels (e.g., unambiguous misclassifications
+present in the training set).
+\item We assume our training set is representative of the test set (the \rave\ survey).
+\end{itemize}
+
+We adopt the nomenclature first set out in \citet{Ness_2015}, and extended in 
+\citet{Casey_2016b}.  However, one distinction here with \citet{Casey_2016b} is 
+that in this work the \emph{labelled} set and the \emph{training} set will be 
+the same set.  In \citet{Casey_2016b} two subsets of the labelled set were 
+defined: the training set and a distinct validation set.  This allowing us to 
+train on the \emph{training set} and then test on the \emph{validation set}.  
+However in this work the labelled set is relatively small, about an order of 
+magnitude smaller than what was available in \citet{Casey_2016b}, and therefore 
+no validation set exists. We will use all labelled set stars in the training 
+step and supplement our validation experiments with literature comparisons.  
+Thus, throughout this work we will use the terms \emph{labelled set} and 
+\emph{training set} interchangeably.
 

### 2016-09-02 article/unrave.tex
Describe the three-label model training and LOOCV

@@
 \usepackage{bm}
+\usepackage{xcolor}
+
+\definecolor{unoffensive-warning}{HTML}{B4DCED}
 
@@
 % ------------------------------------------
-\usepackage{color}
+%\usepackage{color}
 \usepackage{datenumber}
@@
 \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
-model experiments:
+model experiments,
 \begin{eqnarray}\label{eq:vectorizer-three-label}
@@
 \noindent{}such that $\vecv(\ell)$ produces the design matrix:
-\begin{tiny}
-\begin{equation}\label{eq:vectorizer-three-label-design-matrix}
-\vecv(\ell) \rightarrow \begin{bmatrix}
-	1 & T_{{\rm eff},0} & \logg_{0} & [{\rm Fe/H}]_{0} & T_{{\rm eff},0}^2 & \logg_{0}\cdot{}T_{{\rm eff},0} & [{\rm Fe/H}]_{0}\cdot{}T_{{\rm eff},0} & \logg_{0}^2 & [{\rm Fe/H}]_{0}\cdot{}\logg_1 & [{\rm Fe/H}]_1^2 \\
-	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
-	1 & T_{{\rm eff},n} & \logg_{n} & [{\rm Fe/H}]_{n} & T_{{\rm eff},n}^2 & \logg_{n}\cdot{}T_{{\rm eff},n} & [{\rm Fe/H}]_{n}\cdot{}T_{{\rm eff},n} & \logg_{n}^2 & [{\rm Fe/H}]_{n}\cdot{}\logg_n & [{\rm Fe/H}]_n^2 \\
-	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
-	1 & T_{{\rm eff},N-1} & \logg_{N-1} & [{\rm Fe/H}]_{N-1} & T_{{\rm eff},N-1}^2 & \logg_{N-1}\cdot{}T_{{\rm eff},N-1} & [{\rm Fe/H}]_{N-1}\cdot{}T_{{\rm eff},N-1} & \logg_{N-1}^2 & [{\rm Fe/H}]_{N-1}\cdot{}\logg_{N-1} & [{\rm Fe/H}]_{N-1}^2
-\end{bmatrix}\quad .
-\end{equation}
-\end{tiny}
+\begin{eqnarray}
+	\vecv(\ell) \rightarrow \begin{bmatrix} \vecv(\ell_0) \\ \vdots \\ \vecv(\ell_{N-1}) \end{bmatrix} \quad .
+\end{eqnarray}
 
-\subsubsection{A three-label model for main-sequence and giant stars}\label{sec:a-simple-joint-model}
+\stub{We trained the model initially with no regularization.
+Could reproduce general structure of a main-sequence and HRD, but the intricacies of the structure was gone.
+Model appeared to be too simple.
+However we can use this model to identify dwarfs/giants,etc.
+We may revisit this model later}
 
-\subsubsection{A three-label model for main-sequence stars}\label{sec:a-simple-model-ms-stars}
+\subsubsection{A 3-label model for the main-sequence}
+\label{sec:the-ms-model}
 
-% worked but didn't show the structure in detail.
-% Split into two different models with partial overlap.
-% One for giants and one for dwarfs.
-% The labels for stars in the training sets are shown in Figure X.
+We constructed a three-label quadratic model for the main-sequence and sub-giant
+stars in the \epic\ training set.  In order to set the regularization hyperparameter
+$\Lambda$ for this model, we trained 30 models with different regularization strengths,
+spaced evenly in logarithmic space between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
+We then performed leave-one-out cross-validation for each model.  Specifically, for 
+each star in the training set: we removed the star; trained the model; and then 
+inferred labels from the removed star as if it was a test object. We also performed 
+leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
+will use as the basis for comparison.  For every model we calculated the bias and
+root-mean-square (RMS) deviation between the input labels and those inferred through
+cross-validation.
 
-% We trained all three models.
-% Note the differences in metallicity scales.
-% Look at spectral derivatives.
-% Show the model internals are relatable.
-% Map spectral derivatives from giant model to main-sequence model.
-% Update the model with those new spectral derivatives.
+We show the \emph{percentage difference} in the RMS deviation of the labels with respect
+to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
+and lower envelope represent the boundaries across all labels, showing that the RMS
+decreased in \emph{all} labels with increasing regularization.  We found similar
+improvements in the bias values, however these were already minimal in the unregularized
+case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
+we achieve RMS deviations that are about 30\% better than the unregularized model.
+Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
+At this regularization strength, the bias and RMS values found by leave-one-out
+cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
+for $\logg$, with $X.XX$~dex and $X.XX$~dex for [Fe/H].
 
-% How we combined the results into one.
-% Refer to test set HRD.
+We made no \emph{a priori} judgements on whether a star was likely a giant
+or a dwarf star; we inferred labels for all \rave\ spectra using this 
+main-sequence model.  The resulting labels for the survey sample are shown
+in Figure \ref{fig:hrd-from-individual-models}.  Significant structure is 
+visible.  The increased density of Solar-type stars is consistent with \rave\
+observing stars in the local neighbourhood, and the high number of turnoff
+and main-sequence stars relative to the subgiant branch is expected from 
+the relative lifetimes of these evolutionary phases.  An overdensity
+of stars near the giant branch clump is visible, which is likely due to
+giant stars existing in the test set, but not in the training set.  In
+later sections we will describe how our results from multiple models were
+compiled to produce a single catalog.
+
+\subsubsection{A 3-label model for the main-sequence and giant branch}
+\label{sec:the-joint-model}
+
+% Chose the regularization hyperparameter.
 
@@
 % TODO: Replace ZENODO-URL with a data url
-\noindent{}Source code for this project is available at \texttt{\giturl}.
-This document was compiled from revision hash \texttt{\githash} in that repository.
+\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
+\noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
+and this document was compiled from revision hash \texttt{\githash} in that repository.
 Inferred labels and covariance matrices are available electronically with 
@@
 \rave\ collaboration, past and present.
+}}
 
 \acknowledgements
-A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sergey Koposov (Cambridge),
-and Sven Buder (MPIA).
+A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and
+Sergey Koposov (Cambridge).
 This research made use of: Astropy, a community-developed core Python package for

### 2016-09-02 article/figures/plot_ges_comparison.py
Updating figures and shit

@@ -77,8 +77,11 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     # yerr
     c = data_table["snr"]
 
+    uves = np.array([_.startswith("u") for _ in data_table["FILENAME"]])
 
-    scat = ax.scatter(x, y, c=data_table["snr"], s=75, cmap="plasma") 
+    #scat = ax.scatter(x[uves], y[uves], c=data_table["snr"][uves], s=75, marker="s", vmin=np.nanmin(data_table["snr"]), vmax=np.nanmax(data_table["snr"]), cmap="plasma")
+    #ax.scatter(x[~uves], y[~uves], c=data_table["snr"][~uves], s=75, marker="o", vmin=np.nanmin(data_table["snr"]), vmax=np.nanmax(data_table["snr"]), cmap="plasma") 
+    scat = ax.scatter(x, y, c=data_table["snr"], s=75, cmap="plasma")
 
     limits = label_limits[label_name]
     ax.plot(limits, limits, c="#666666", linestyle=":", zorder=-1)

### 2016-09-02 article/figures/rave_io.py

@@ -2,12 +2,25 @@
 import os
 import numpy as np
 from astropy.table import Table
+from scipy.io import readsav
 
 DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
 
 
 def get_rave_kordopatis_dr4():
     return Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
+
+def get_kordopatis_comparisons():
+    data = readsav(os.path.join(DATA_PATH, "RAVE_DR5_calibration_data.save"))
+    
+    return Table(data={
+        "TEFF": data["calibration_data"]["TEFF"][0],
+        "LOGG": data["calibration_data"]["LOGG"][0],
+        "FEH": data["calibration_data"]["FEH"][0],
+        "REF": data["calibration_data"]["REF"][0],
+        "Name": [each.strip() for each in data["calibration_data"]["DR5_OBS_ID"][0]]
+        })
+
 
 
 
@@ -27,7 +40,7 @@ def get_cannon_dr1(filename=None):
         rave_cannon_dr1["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in rave_cannon_dr1["FILENAME"]]
 
     
-    if filename == "rave-tgas-v37.fits.gz":
+    if filename in ("rave-tgas-v37.fits.gz", "rave-tgas-v31.fits.gz"):
         rave_cannon_dr1["TEFF"] = rave_cannon_dr1["EPIC_TEFF"]
         rave_cannon_dr1["LOGG"] = rave_cannon_dr1["EPIC_LOGG"]
         rave_cannon_dr1["FE_H"] = rave_cannon_dr1["EPIC_FEH"]
@@ -38,6 +51,9 @@ def get_cannon_dr1(filename=None):
 
     if "snr" not in rave_cannon_dr1.dtype.names and "snr_ms" in rave_cannon_dr1.dtype.names:
         rave_cannon_dr1["snr"] = rave_cannon_dr1["snr_ms"]
+
+    if "snr" not in rave_cannon_dr1.dtype.names and "SNRK" in rave_cannon_dr1.dtype.names:
+        rave_cannon_dr1["snr"] = rave_cannon_dr1["SNRK"]
 
[…]

### 2016-09-02 stack_repeat_visit_spectra.py

@@ -44,6 +44,7 @@ for group in dr5.group_by("GroupID").groups:
         if not os.path.exists(spectrum_path):
             print("Could not find {} in group {}".format(spectrum_path, group_id))
             subset[i] = False
+            raise WTFError
             continue
 
         with open(spectrum_path, "rb") as fp:
@@ -65,6 +66,11 @@ for group in dr5.group_by("GroupID").groups:
     stacked_flux = np.sum(flux * ivar, axis=0)/stacked_ivar
 
     assert np.any(np.isfinite(stacked_flux))
+    if not np.all(np.isfinite(stacked_ivar)):
+        print("Warning: {} pixels in {} had non-finite inverse variance".format(
+            np.sum(~np.isfinite(stacked_ivar)), group["RAVEID"][0]))
+        stacked_ivar[~np.isfinite(stacked_ivar)] = 0
+
     assert np.all(np.isfinite(stacked_ivar))
 
     stacked_spectrum_path = os.path.join(

### 2016-09-02 article/unrave.tex

@@
 
-\subsection{Three-label models ($\teff$, $\logg$, $\feh$)}
+\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
 \label{sec:a-simple-model}
@@
 complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
-Given previous successes using a second-order polynomial model 
+Given previous successes using a second-order (quadratic) polynomial model 
 \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
@@
 
-\subsubsection{A 3-label model for the main-sequence}
+\subsection{A 3-label model for unevolved stars}
 \label{sec:the-ms-model}
@@
 main-sequence model.  The resulting labels for the survey sample are shown
-in Figure \ref{fig:hrd-from-individual-models}.  Significant structure is 
+in Figure \ref{fig:hrd-from-individual-models}, where significant structure is 
 visible.  The increased density of Solar-type stars is consistent with \rave\
@@
 the relative lifetimes of these evolutionary phases.  An overdensity
-of stars near the giant branch clump is visible, which is likely due to
-giant stars existing in the test set, but not in the training set.  In
-later sections we will describe how our results from multiple models were
-compiled to produce a single catalog.
-
-\subsubsection{A 3-label model for the main-sequence and giant branch}
-\label{sec:the-joint-model}
-
-% Chose the regularization hyperparameter.
+of stars near the giant branch clump is also present.  This artefact is due 
+to having giant stars in the test set, but not in the training set, and the
+model is (poorly) extrapolating outside the convex hull of the training set.
+In Section \ref{sec:joining-the-models} we describe how results from multiple
+models are combined to produce a single catalog.
 
@@
 
-The giant stars in our training set have stellar parameters and up to 15 
-individual abundances from the \apogee\ \aspcap.  Only a subset of these
-elements have atomic transitions in the \rave\ wavelength region: 
-\ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, 
-\ion{Fe}{1}, and \ion{Ni}{1}.  We excluded [Ti/H] from our abundance list 
-because of systematics in the training set \citep{Holtzman_2015,Hawkins_2016}.
-Therefore we are left with nine labels in our giant star model: $\teff$, 
-$\logg$, and seven individual abundances.
+We only have $\teff$, $\logg$, and [Fe/H] for unevolved stars in our training
+set from the \epic\ catalog \citep{Huber_2016}.  However, the giant stars 
+have stellar parameters ($\teff$, $\logg$) and up to 15 individual abundances
+from the \aspcap\ \citep{aspcap_paper}.  A subset of these elements have atomic
+transitions in the \rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, 
+\ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  
+We excluded [Ti/H] from our abundance list because of systematics in the training
+set \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left with nine labels 
+in our giant star model: $\teff$, $\logg$, and seven individual abundances.
 
-% ARC asks: Should we discuss any tests done with censoring? 
-%  			Do we feel like we need to defend our choice to use regularization
-%           over censoring? The key result from the censoring tests that I did
-%			is that there are a lot of atomic transitions here, but they are
-% 			mostly blended. We *need* those blended lines because if you
-%			censor all but the strongest line(s) for individual elements then
-% 			you are left with a poor predictor of abundance labels.
-
-We used a second-order (i.e., $\teff^2$ is the highest term) polynomial 
-vectorizer for the giant star model.  We set the regularization hyperparameter
-by performing a grid search of 30 $\Lambda$ points logarithmically spaced from 
-$10^{-3}$ to $10^{3}$.  We performed leave-one-out cross-validation on the
-training set for each value of $\Lambda$.  Specifically, we removed one star
-from the training set, then trained the model, and inferred labels for the 
-star that was removed from the training set.  We repeated this procedure for
-every star in the training set (for every model with a different $\Lambda$)
-and compared the inferred labels against the training set values.  For
-comparison purposes, we also performed cross-validation on an unregularized 
-($\Lambda = 0$) model using the same training set.  We calculated the bias
-and root-mean-square (RMS) deviation between the input labels and those inferred
-by cross-validation for every model.  The biases were negligible across all
-models.  In Figure \ref{fig:giant-model-hyperparameter-search} we show the
-RMS difference in label recovery compared to the unregularized model as a 
-percentage improvement.  The enveloped region represents the minimum and 
-maximum changes for any label with respect to the unregularized model.
-Depending on the regularization strength $\Lambda$, the RMS in a given label
-can decrease by $\approx30$\% with respect to the unregularized case.  At the
-point of maximum improvement, near $\Lambda = X.XX$, the RMS in all labels has
-decreased by at least $X$\%.  As expected the regularization produces a 
-sparser matrix of $\vectheta$, with a $\approx20$\% increase in the fraction
-of zero-valued entries, particularly for the cross-terms (Figure 
-\ref{fig:giant-model-sparsity}, third panel).  Based on the improvement in 
-RMS by cross-validation, we adopt $\Lambda = X.XX$ for our giant star model.
+We used a quadratic vectorizer for the giant star model, and set the regularization
+hyperparameter $\Lambda$ in the same way described in Section \ref{sec:ms-model}.
+The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
+again the enveloped region represents the minimum and maximum change for any label
+with respect to the unregularized case.  At the point of maximum improvement 
+near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
+and up to $X$\%.  Here the regularization strength also produces a sparser matrix
+of $\vectheta$, with a $\approx20$\% increase in the fraction of zero-valued 
+entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
+third panel).  Based on decreasing RMS deviation in the labels, we adopt 
+$\Lambda = X.XX$ for our giant star model.
 
 Results from cross-validation on our regularized model imply that the precision
-in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and varies between X.XX~dex
-and X.XX~dex depending on the elemental abundance.  We made no prior judgement
-on whether a star was likely a giant or not; we used this model (and all other
-models) to test all XXX,XXX \rave\ spectra.  Figure \ref{fig:test-set-hrd} (left
-panel) shows the inferred $\teff$ and $\logg$ values for all \rave\ spectra with
-S/N $\geqslant 50$.  This panel undoubtedly includes dwarf stars that cannot be
-appropriately classified by our giant model.  In Section \ref{sec:joining-the-models}
-we will heuristically combine these results with the main-sequence star model
-described in \ref{sec:main-sequence-model}.
+in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and X.XX~dex (EL) and 
+X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
+XXX,XXX \rave\ spectra using this model, and these results are summarized in
+Figure \ref{fig:test-set-hrd} (middle panel).  The red clump is clearly visible
+and in the expected location, without requiring any post-analysis calibration.
+Artefacts due to dwarf stars in the test set are also present.
 
-\subsection{Joining the models}\label{sec:joining-the-models}
+\subsection{Joining the models}
+\label{sec:joining-the-models}
 

### 2016-09-02 article/figures/plot_v34_search.py
Add plot of RMS improvement at increasing regularization strength

@@ -1,0 +1,162 @@
+import os
+
+import numpy as np
+import cPickle as pickle
+import matplotlib.pyplot as plt
+
+from astropy.table import Table
+from glob import glob
+from matplotlib.ticker import MaxNLocator
+
+scale_factors = [0.5, 1, 2, 5, 10]
+Lambdas = [0, 0.001, 0.01, 0.1, 1.0, 10.0, 25.0, 50.0, 100.0]
+labels = ("TEFF", "LOGG", "FE_H", "MG_H", "AL_H", "O_H", "NI_H", "SI_H", "CA_H")
+
+
+
+# y axis should be some metric(s)
+# x axis should be Lambda
+
+
+
+
+#labelled_set = Table.read("{}-labelled-set.fits".format(model_name))
+labelled_set = Table.read("/data/gaia-eso/arc/rave/preprocessing/giant_trainingsets_APOGEERAVE_V2.fits")
+
+ok = np.ones(len(labelled_set), dtype=bool)
+for label_name in labels:
+    ok *= np.isfinite(labelled_set[label_name])
+    ok *= (labelled_set[label_name] > -5)
+
+labelled_set = labelled_set[ok]
+labelled_set = np.array([labelled_set[label_name] for label_name in labels]).T
+
+
+
+def metric(labelled_set, test_set):
+    return np.nanstd(labelled_set - test_set, axis=0)
+
+
+color_choices = ["r", "b", "g", "m", "y"]
+color_choices = ["#3498DB"]
+
+files = glob("/data/gaia-eso/arc/rave/rave-tgas-v34-test-labelled-set-?.????????e*-1.00.pkl")
+
+metric_index = 0
+
+offset = 5
+x = []
+colors = []
+y = []
+
+for filename in files:
+
+    with open(filename, "rb") as fp:
+        test_set = pickle.load(fp)
+
+    basename = os.path.basename(filename)
+    Lambda = basename[32:41 + offset]
+    scale_factor = basename[42 + offset:46 + offset]
+
+    x.append(float(Lambda))
+    colors.append(float(scale_factor))
+
+    y.append(metric(labelled_set, test_set))
+
+x = np.array(x)
+y = np.array(y)
+colors = np.array(colors)
+
+fig, ax = plt.subplots()
+N_metrics = len(y[0])
+
[…]

### 2016-09-02 article/unrave.tex

@@
 
-\newcommand{\stub}[1]{\textbf{#1}}
+\newcommand{\stub}[1]{{\color{blue} \textbf{#1}}}
 
@@
 
-\newcommand{\Nstars}{483,330}
+\newcommand{\Nspectra}{520,782}
+\newcommand{\Nstars}{457,589}
 
@@
 for the fainter targets in the first \gaia\ data release. \stub{A one-line 
-science result based on RAVE/Hipparcos overlap}
+science result based on RAVE/Hipparcos overlap.}
 \end{abstract}
@@
 Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
-2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
-made for very efficient observing.  Spectra for up to 150 targets were simultaneously
-acquired, and when observations concluded in April 2013, a total of 574,630 spectra
-had been collected of 483,330 unique objects.
+2003--2013.  A large 5.7~degree field-of-view and robotic fibre positioner 
+made for very efficient observing;  spectra for up to 150 targets could be simultaneously
+acquired.  When observations concluded in April 2013, a total of \Nspectra\ useful spectra
+had been collected of \Nstars\ unique objects.
 
@@
 \begin{eqnarray}\label{eq:vectorizer-three-label}
-\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
-\quad ,
+\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
 \end{eqnarray}
@@
 However we can use this model to identify dwarfs/giants,etc.
-We may revisit this model later}
+Rather than go to a very complex model, we opted to use two models: one for the main-sequence and one for the giant branch.}
 
@@
 $\Lambda$ for this model, we trained 30 models with different regularization strengths,
-spaced evenly in logarithmic space between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
+spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
 We then performed leave-one-out cross-validation for each model.  Specifically, for 
@@
 
-We used a quadratic vectorizer for the giant star model, and set the regularization
-hyperparameter $\Lambda$ in the same way described in Section \ref{sec:ms-model}.
+We used a quadratic vectorizer for the giant star model, where the quadratic and
+cross-terms are expanded in the same way as \ref{eq:vectorizer-three-label}, only
+with nine labels.  The regularization hyperparameter $\Lambda$ was set in the same
+way described in Section \ref{sec:ms-model}, using the same 30 trials of $\Lambda$.
 The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
-again the enveloped region represents the minimum and maximum change for any label
-with respect to the unregularized case.  At the point of maximum improvement 
+again the enveloped region represents the minimum and maximum change in RMS label
+deviation with respect to the unregularized case.  At the point of maximum improvement 
 near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
@@
 entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
-third panel).  Based on decreasing RMS deviation in the labels, we adopt 
-$\Lambda = X.XX$ for our giant star model.
+third panel).  Based on the increased model sparsity and decreasing RMS deviation
+in the labels, we adopt $\Lambda = X.XX$ for our giant star model.
 
@@
 
-\section{Validation experiments}\label{sec:validation}
+\section{Validation experiments}
+\label{sec:validation}
 
-We have already performed leave-one-out cross-validation on all of our models
-as part of the hyperparameter selection.  Here we describe the results from our
-internal, external and astrophysical validation checks.
+In addition to the cross-validation tests that we have previously described, 
+we conducted a number of internal and external validation experiments to 
+test the validity of our results.
 
-\subsection{Multi-epoch observations}\label{sec:repeat-observations}
+\subsection{Multi-epoch observations}
+\label{sec:repeat-observations}
 
@@
 with multiple visits we constructed a high S/N stacked spectrum for that
-star, weighted by the inverse variances in the individual visit spectra,
-which then served as the basis of comparison to all individual visits.  
-Figure \ref{fig:formal-errors-comparison} shows the difference in labels 
+star, weighted by the inverse variances in the individual visit spectra.
+
+We treated the stacked spectra as normal survey stars.  We inferred labels
+using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
+of comparison for all individual visits to that object.  Figure 
+\ref{fig:formal-errors-comparison} shows the difference in labels 
 between the comparison spectrum and a repeat visit, normalized by their 
@@
 $\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
-If our measurements were unbiased by S/N and the errors were representative, 
-these values should be normally distributed with a zero mean and unit variance.
-\stub{Are they? Should we inflate our formal errors?}
+If our measurements were unbiased by S/N and the formal errors were 
+representative, these values should be normally distributed with a zero 
+mean and unit variance.
+\stub{Are they? (Probably). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
+we show these inflated errors only, as we gauge them to be more representative.}
 
-% Internal validation:
-% --> bootstrap resampling of the training set?
+\stub{Covariances between these parameters are accessible through the matrix of correlation coefficients in the associated results file.}
 
-% Identify outliers:
-% --> Stars with morphological classification
-% --> Some rogue gallery of a few high S/N cases that have clearly gone haywire?
+\subsection{Jackknifing the training set}
+\label{sec:jackknife}
+
+\stub{Should we?}
+
+\subsection{Quality control}
+\stub{What should we adopt as recommended QC? Show rogue gallery of a few "high s/n cases" where things went haywire?}
 
@@
 
-\stub{What abundances are available?}
+\stub{Discuss comparisons with: GES; calibration sources used in \citet{Kordopatis_2013}; and `gold-standard' studies: Reddy, Bensby, Valenti \& Fischer}
 
-% This is more relevant for the discussion, not description of the training set.
-We note that the while Ti can be measured, its values in both 
-APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
-
-\stub{Any differences in RAVE abundances?}
-
-% External validation: 
-% --> Gaia-ESO Survey
-% --> Literature studies presented in Kordopatis+ 2014
-% --> Gold standards (incl. exoplanet host star studies): Reddy (2003, 2006), Bensby (2014), Valenti & Fischer (2005)
+\stub{Any differences in RAVE abundances? Corrections or caution required?}
 
 \subsection{Astrophysical validation}
+\label{sec:astrophysical-validation}
 
@@
 globular cluster NGC~3201, and 10 members of NGC~362.  
-\stub{Were any of these stars unintentionally in the training set? \\ KH: Not using my training sets.}
 All of these systems are well-studied and exhibit abundance anti-correlations 
@@
 (e.g., C--N), but those relationships are less relevant for this work because 
-we do not measure C and N abundances here.
+we do not measure C and N abundances here.  We note that \emph{none} of these 
+cluster stars are present in our training set.
 
@@
 
+\stub{Chemistry + Kinematics = Stockholm}
+
+\stub{What do the two provide us with?}
+
+\stub{What is one interesting science result should we show based on these abundances and Hipparcos results, which was not known previously?}
+
 \section{Conclusion}
@@
 and this document was compiled from revision hash \texttt{\githash} in that repository.
-Inferred labels and covariance matrices are available electronically with 
-associated metadata at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  
-Please note that it is a condition of using these results that the fifth \rave\ 
-data release by \citet{Kunder_2016} must (also) be cited, as the work presented 
-here would not have been possible without the tireless efforts of the entire 
-\rave\ collaboration, past and present.
+Inferred labels, covariance matrices, and relevant metadata are available electronically
+at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  Please note that it is a condition
+of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must 
+also be cited, as the work presented here would not have been possible without the 
+tireless efforts of the entire \rave\ collaboration, past and present.
 }}
@@
 \begin{figure}[p]
+\includegraphics[width=\textwidth]{figures/hrd-train-set.pdf}
 \caption{A H-R diagram showing the training set labels.\label{fig:training-set-hrd}}
@@
 \includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
-\caption{A `spectroscopic' Hertzsprung-Russell diagram showing the effective temperature $\teff$ and surface gravity $\logg$ for stars in the test set.
-Only stars with $\chi_r < 3$ are shown, and each panel shows stars that meet a threshold in $S/N$ ratio.\label{fig:test-set-hrd}}
+\caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
+Only stars with $v\sin{i} < 1$~km~s$^{-1}$ and $\chi_{red} < 3$ are shown.
+The top three panels show logarithmic density, and bins in the lower three panels are colored by the median metallicity in each bin.\label{fig:test-set-hrd}}
 \end{figure}
@@
 \begin{figure*}[p]
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fifth data release \citep{Kunder_2016}.\label{fig:rave-dr5-comparison}}
+\includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
+\caption{Stellar label comparison between the \project{unRAVE} catalog and the \ges\ fourth internal data release.\label{fig:ges-stellar-parameters}}
 \end{figure*}
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{Gaia-ESO Survey} fourth internal data release.\label{fig:ges-dr4-comparison}}
+\includegraphics[width=\textwidth]{figures/ges-abundances.pdf}
+\caption{Individual chemical abundances in \project{unRAVE} compared to \ges.\label{fig:ges-abundances}}
 \end{figure*}
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/literature-comparison.png}
-\caption{Stellar label comparisons between the \project{UNRAVE} catalog and the samples discussed in \citep{Kordopatis_2013}}
+\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
+\caption{Stellar parameter comparisons for stars in common between this work and studies using high-resolution, high S/N spectra and \hipparcos\ parallaxes.\label{fig:gold-standard-comparison}}
 \end{figure*}
 
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
+\caption{HRD of gold standard comparisons.\label{fig:gold-standard-hrd}}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
+\caption{Comparison with the calibration (literature) samples used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.\label{fig:kordopatis-calibration}}
+\end{figure*}
+
+\begin{figure}[p]
+\includegraphics[width=0.5\textwidth]{figures/gce.pdf}
+\caption{Giant star abundances ([X/Fe]) with respect to [Fe/H], showing the Galactic chemical evolution trend for individual elements.\label{fig:gce}}
+\end{figure}
+
 \end{document}

### 2016-09-02 article/unrave.tex
Include figure showing RMS improvement wrt regularization

@@
 \begin{figure}[p]
+\includegraphics[width=\textwidth]{figures/set-giant-regularization.pdf}
+\caption{The fractional improvement in RMS deviation between inferred and training labels for the 9-label giant model at different regularization strengths.
+The RMS values were calculated by leave-one-out cross-validation.
+The points and solid line indicate the mean improvement across all labels.
+The filled area represents the minimum and maximum improvements across all labels.
+The maximum improvement (decrease) in RMS label deviations occurs at $\Lambda \approx 0.3$, and this improvement persists until $\Lambda \approx 0.6$.}
+\end{figure}
+
+\begin{figure}[p]
 \includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}

### 2016-09-03 article/unrave.tex
unbalanced
Some text edits and incorporate Kordopatis comments sent by email

@@
 release, constituting the largest overlap of any stellar spectroscopic sample.  
-Here we perform an independent analysis of \rave\ spectra using an implementation
-of \thecannon\ that uses prior probabilities on atomic line formation and stellar
-astrophysical parameters.  Our approach yields precise stellar labels at lower 
-S/N ratios than most \emph{ab initio} approaches, allowing us to deliver improved
-radial velocities, effective temperature $\teff$, surface gravity $\logg$, and 
-chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) for 
-\Nstars\ stars.  We validate our results internally from repeat visits, and 
-externally against existing high-resolution spectroscopic surveys.  The typical
-precision in chemical abundances is X.XX~dex.  We provide spectrophotometric 
-distances based on our labels, improving upon the parallax distances available
-for the fainter targets in the first \gaia\ data release. \stub{A one-line 
-science result based on RAVE/Hipparcos overlap.}
+We have performed an independent analysis of \rave\ spectra using \thecannon,
+and here we report effective temperature $\teff$, surface gravity $\logg$, 
+as well as chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, 
+Ni) for \Nstars\ stars.  We validate our results internally from repeat visits, 
+and externally against existing high-resolution spectroscopic surveys.  The 
+typical precision is XXX~K in $\teff$, X.XX~dex in $\logg$, and X.XX~dex for
+detailed abundances.  We provide spectrophotometric distances based on our
+labels, improving upon the parallax-based distances available for distant 
+giants in the first \gaia\ data release.
+\stub{A one-line science result based on RAVE/Hipparcos overlap.}
+
 \end{abstract}
 
-\keywords{}
+\keywords{\vspace{5em}}
 
@@
 
-The Milky Way is considered to be the best laboratory for understanding galaxy
+The Milky Way is considered to be our best laboratory for understanding galaxy
 formation and evolution.  This premise hinges on the ability to precisely measure 
@@
 \citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
-Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Bovy_2016}.  However, these 
-quantities are not known for even 1\% of stars in the Milky Way.  Stellar 
-distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,
-Madler_2016}, proper motions can be plagued by unquantified systematics from 
-the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
+Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Kordopatis_2015,Bovy_2016}.  
+However, these quantities are not known for even 1\% of stars in the Milky Way.  Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
+Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics 
+from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
 stellar spectroscopists frequently report significantly different chemical 
-abundance patterns for the same star \citep{Smiljanic_2014}.  The impact 
+abundance patterns from the same spectrum \citep{Smiljanic_2014}.  The impact 
 these issues have on scientific inferences cannot be understated.  Imperfect 
@@
 These formed just some of the scientific drivers for the \gaia\ space telescope.
-\gaia\ is primarily an astrometric mission, providing precise positions,
+\gaia\ is primarily an astrometric mission, and will provide precise positions,
 parallaxes and proper motions for more than $10^9$ stars in its final data
@@
 available in the first few data releases. Until those abundances are available,
-astronomers seeking to use chemistry and kinematic measurements are reliant on 
-ground-based spectroscopic surveys in order to complement the available astrometry.
+astronomers seeking to simultaneously use chemical and dynamical information are
+reliant on ground-based spectroscopic surveys in order to complement the 
+available \gaia\ astrometry.
 
-The first \project{Gaia} data release will include positions, proper motions, and 
+The first \gaia\ data release will include positions, proper motions, and 
 parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
@@
 the first \gaia\ data release, and 47\% of stars are likely to have better 
-proper motions (137,211 stars).  This motivated us to examine what chemical 
-abundances were available for \rave, and to evaluate whether we could enable new 
-chemodynamic studies by contributing to this existing set of chemical abundances.
+proper motions (137,211 stars).  Although the \gaia\ universe model assumes 
+end-of-mission uncertainties --- and does not account for systematics in the first
+data release --- this calculation still provides intuition for the relative 
+improvement the first \gaia\ data release can make to ground-based surveys.  The 
+expected improvements for \rave\ motivated us to examine what chemical abundances
+were available from those data, and to evaluate whether we could enable new 
+chemodynamic studies by contributing to the existing set of chemical abundances.
 
@@
 comparisons with literature samples, and investigations to ensure our results
-are consistent with expectations from astrophysics.  We discuss these results in
-Section~\ref{sec:discussion}, and in Section~\ref{sec:conclusion} we provide 
-concluding remarks as well as a description of how to access our results electronically.
+are consistent with expectations from astrophysics.  We discuss the implications
+of these comparisons and our results in Section~\ref{sec:discussion}.  We provide
+concluding remarks in Section~\ref{sec:conclusion}, as well as a description for
+how to access our results electronically.
 
@@
 Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
-2003--2013.  A large 5.7~degree field-of-view and robotic fibre positioner 
-made for very efficient observing;  spectra for up to 150 targets could be simultaneously
-acquired.  When observations concluded in April 2013, a total of \Nspectra\ useful spectra
-had been collected of \Nstars\ unique objects.
+2003--2013.  A large 5.7~degree field-of-view and robotic fibre positioner made for 
+very efficient observing:  spectra for up to 150 targets could be simultaneously
+acquired.  When observations concluded in April 2013, a total of \Nspectra\ useful 
+spectra had been collected of \Nstars\ unique objects.
 
@@
 $9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
-The $I$ band was selected because it overlaps with the wavelength range that \rave\ 
-operates in:  8410--8795~\AA.  This spectral region is very well-studied.  It is one 
-of the key setups used for the ground-based high-resolution \ges\ survey
-\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
-coverage is comparable to the Radial Velocity Spectrometer on board the \gaia\
-space telescope \citep{Recio-Blanco_2016}.  The region includes the \ion{Ca}{2} 
-near infrared triplet lines, strong transitions that are dominated by pressure 
-broadening, which are visible even in very low S/N data or metal-poor stars.  
-Atomic transitions of light-, $\alpha$-, and Fe-peak elements are also present, 
-allowing for detailed chemical abundance studies.
+The $I$ band was used for the target selection because it has a good overlap with the
+wavelength range that \rave\ operates in:  8410--8795~\AA.  This spectral region 
+is very well-studied.  It is one of the key setups used for the ground-based 
+high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}, and the 
+\rave\ resolution and wavelength coverage is comparable to the Radial Velocity 
+Spectrometer on board the \gaia\ space telescope \citep{Kordopatis_2011,Recio-Blanco_2016}.  
+The region includes the \ion{Ca}{2} near infrared triplet lines --- strong 
+transitions that are dominated by pressure broadening --- which are visible even 
+in very low S/N data or metal-poor stars.  Atomic transitions of light-, $\alpha$-, 
+and Fe-peak elements are also present, allowing for detailed chemical abundance studies.
 
@@
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
-that the main-sequence labels suffered from significant systematic effects.  A flat, 
+that the \apogee\ main-sequence labels suffered from significant systematic effects.  A flat, 
 then `up-turning' main-sequence is present, and the metallicity gradient trends in 
 the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
-stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
+stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram).
 If we consider lower-resolution studies as potential training sets, there are 2,369
@@
 Therefore, requiring low uncertainties will decrease the total sample size, but 
-will remove most misclassifications.  The situation is far more favourable for 
+in practice it removes most misclassifications.  The situation is far more favourable for 
 dwarfs and giants.  Only $1-4$\% of giant stars are misclassified as dwarfs, and 
@@
 effect on the detailed abundance or spectrophotometric distance determinations
-between these studies.
+between these studies \citep{Binney_2014}.
 
@@
 
-\stub{Any differences in RAVE abundances? Corrections or caution required?}
+\stub{Any differences in RAVE abundances? Corrections required due to having an unbalanced training set? Any caution required?}
 
@@
 
+\bibitem[Binney et al.(2014)]{Binney_2014} Binney, J., Burnett, B., Kordopatis, G., et al.\ 2014, \mnras, 437, 351
+
 \bibitem[Bensby et al.(2014)]{Bensby_2014} Bensby, T., Feltzing, S., \& Oey, M.~S.\ 2014, \aap, 562, A71
@@
 
+\bibitem[Kordopatis et al.(2011)]{Kordopatis_2011} Kordopatis, G., Recio-Blanco, A., de Laverny, P., et al.\ 2011, \aap, 535, A106
+
 \bibitem[Kordopatis et al.(2013)]{Kordopatis_2013} Kordopatis, G., Gilmore, G., Steinmetz, M., et al.\ 2013, \aj, 146, 134
+
+\bibitem[Kordopatis et al.(2015)]{Kordopatis_2015} Kordopatis, G., Binney, J., Gilmore, G., et al.\ 2015, \mnras, 447, 3526
 

### 2016-09-03 article/unrave.tex
Edited text and rearranged up to sect 3.3

@@
 transitions that are dominated by pressure broadening --- which are visible even 
-in very low S/N data or metal-poor stars.  Atomic transitions of light-, $\alpha$-, 
-and Fe-peak elements are also present, allowing for detailed chemical abundance studies.
+in very low signal-to-noise (S/N) data or metal-poor stars.  Atomic transitions of 
+light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
+chemical abundance studies.
 
@@
 primary motivation.  The \rave\ spectra have an effective resolution 
-$\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
-$\approx$50~per~pixel$^{-1}$.  For comparison, the \galah\ survey 
+$\mathcal{R} \approx 7{,}500$ and the distribution of S/N ratios peaks at 
+$\approx$50~pixel$^{-1}$.  For comparison, the \galah\ survey 
 \citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
 abundance analyses --- includes a wavelength range about 2.5 times larger at 
-resolution $\mathcal{R} \approx 28,000$, and yet the \galah\ project still 
+resolution $\mathcal{R} \approx 28{,}000$, and yet the \galah\ project still 
 targets for S/N $\gtrsim100$.
@@
 as well as revised determinations of stellar radial velocities and morphological
-classifications. At the end of this processing for each survey observation we are 
-provided with: rest-frame wavelengths, continuum-normalized fluxes, fractional 
-uncertainties in continuum-normalized flux values, as well as relevant metadata.  
-We refer the reader to the official fifth data release paper of the \rave\ survey, 
-as presented by \citet{Kunder_2016}, for more details of this re-processing.
+classifications. At the end of this processing for each survey observation we were 
+provided with: rest-frame wavelengths (without resampling), continuum-normalized 
+fluxes, $1\sigma$ uncertainties in the continuum-normalized flux values, as well 
+as relevant metadata for each observation.  We refer the reader to the official 
+fifth data release paper of the \rave\ survey, as presented by \citet{Kunder_2016}, 
+for more details of this re-processing.
 
@@
 there is a limited number of pre-processing steps that we conducted before our
-analysis could begin.  First, we calculated inverse variance arrays from the
-fractional uncertainties provided, and then re-sampled the flux and inverse variance
+analysis began.  First, we calculated inverse variance arrays from the $1\sigma$ 
+uncertainties provided, and then we re-sampled the flux and inverse variance
 arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
@@
 
-We chose to adopt a data-driven model --- rather than a physics-driven model ---
-for this analysis.  Specifically we will use an implementation of \thecannon\
+We chose to adopt a data-driven model --- rather than a physics-based model ---
+for this analysis.  Specifically, we will use an implementation of \thecannon\
 \citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
-of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven model 
+of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven approach 
 makes use of all available information in the spectrum and lowers the S/N ratio 
-at which systematic effects begin to dominate.  In other words, a well-constructed
-data-driven model will yield more precise labels (e.g., stellar parameters and
-chemical abundances) at low S/N ratios than most physics-driven models\footnote{
-However, see \citet{Casey_2016a}}.  This is particularly relevant for the 
-low-resolution \rave\ data analysed here, because about half of the spectra have
-S/N $\lesssim 50$.
+at which systematic effects begin to dominate.  In other words, in the low S/N
+regime, a well-constructed data-driven model will yield more precise labels 
+(e.g., stellar parameters and chemical abundances) than most physics-driven 
+models\footnote{However, see \citet{Casey_2016a}}.  This is particularly relevant 
+for the low-resolution \rave\ data analysed here, because about half of the 
+spectra have S/N $\lesssim 50$.
 
-\noindent{}Our work makes the following assumptions about the \rave\ spectra and \thecannon:
+There are two main analysis steps when using \thecannon: the \emph{training} 
+step and the \emph{test} step.  These stages are briefly described in the
+following section, and a more thorough introduction can be found in \citet{Ness_2015}.
+We make the following explicit assumptions about the \rave\ spectra and \thecannon:
 
 \begin{itemize}
-\item We will assume that any fibre- and time-dependent variations in spectral
+\item We assume that any fibre- and time-dependent variations in spectral
 resolution in the \rave\ spectra are negligible.
@@
 \item We assume that the normalization procedure employed by the \rave\ pipeline
-is invariant with respect to the labels, and invariant with respect to the S/N
-ratio.
+is invariant with respect to the labels we seek to measure, and invariant with 
+respect to the S/N ratio.
 \item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
@@
 that we do not assume that \emph{every} label in the training is accurate -- we can
-afford to have a small fraction of inaccurate labels (e.g., unambiguous misclassifications
-present in the training set).
+afford to have a small fraction of inaccurate labels (e.g., a few obvious 
+misclassifications in the training set are affordable).
 \item We assume our training set is representative of the test set (the \rave\ survey).
@@
 
-We adopt the nomenclature first set out in \citet{Ness_2015}, and extended in 
-\citet{Casey_2016b}.  However, one distinction here with \citet{Casey_2016b} is 
-that in this work the \emph{labelled} set and the \emph{training} set will be 
-the same set.  In \citet{Casey_2016b} two subsets of the labelled set were 
-defined: the training set and a distinct validation set.  This allowing us to 
-train on the \emph{training set} and then test on the \emph{validation set}.  
-However in this work the labelled set is relatively small, about an order of 
-magnitude smaller than what was available in \citet{Casey_2016b}, and therefore 
-no validation set exists. We will use all labelled set stars in the training 
-step and supplement our validation experiments with literature comparisons.  
-Thus, throughout this work we will use the terms \emph{labelled set} and 
-\emph{training set} interchangeably.
+%We adopt the nomenclature first set out in \citet{Ness_2015}, and extended in 
+%\citet{Casey_2016b}.  However, one distinction here with \citet{Casey_2016b} is 
+%that in this work the \emph{labelled} set and the \emph{training} set will be 
+%the same set.  In \citet{Casey_2016b} two subsets of the labelled set were 
+%defined: the training set and a distinct validation set.  This allowing us to 
+%train on the \emph{training set} and then test on the \emph{validation set}.  
+%However in this work the labelled set is relatively small, about an order of 
+%magnitude smaller than what was available in \citet{Casey_2016b}, and therefore 
+%no validation set exists. We will use all labelled set stars in the training 
+%step and supplement our validation experiments with literature comparisons.  
+%Thus, throughout this work we will use the terms \emph{labelled set} and 
+%\emph{training set} interchangeably.
+
+\subsection{The model}
+\label{sec:the-model}
+
+\noindent{}Given our assumptions, the model we adopt is
+\begin{eqnarray}\label{eq:model}
+y_{jn} & = & \vecv(\ell_n)\cdot\vectheta_j + e_{jn}\quad ,
+\end{eqnarray}
+
+\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
+$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
+$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
+$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
+wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
+vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Section 
+\ref{sec:a-simple-model}).  The noise values $e_{jn}$ can be considered to be drawn 
+from a Gaussian distribution with zero mean and variance $\sigma_{jn}^2 + s_j^2$, where
+$\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and $s_j^2$ describes the excess 
+variance at the $j$-th wavelength pixel.
+
+At the \emph{training} step we fix the $K$-lists of labels for all the training set stars $n$.
+At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
+by optimizing the penalized likelihood function:
+\begin{eqnarray}\label{eq:train}
+\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
+    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
+    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
+    \right]
+  \quad ,
+\end{eqnarray}
+
+\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
+for each model in later sections, and $Q(\vectheta)$ is a regularizing function that 
+encourages $\vectheta$ values to take on zero values without breaking convexity:
+\begin{eqnarray}\label{eq:regularization-function}
+	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}| \quad .
+\end{eqnarray}
+
+Note that the $d$ subscript here is zero-indexed; the function $Q(\vectheta)$ does not act
+on the (first) $\theta_0$ coefficient, as this is a `pivot point' or offset that we do not
+expect to diminish with increasing regularization (e.g., see equation 
+\ref{eq:vectorizer-three-label}).  In practice equation \ref{eq:train} is 
+optimized by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a convex optimization 
+problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
+
+The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
+pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
+the objective function is:
+\begin{eqnarray}\label{eq:test}
+  \ell_m &\leftarrow& \argmin{\ell}\left[
+    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
+    \right]
+  \quad .
+\end{eqnarray}
+
+At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
+$\ell$, which provides us with the formal errors on $\ell$.  The formal errors are expected
+to be underestimated, and in Section \ref{sec:validation} we judge the veracity of these 
+errors through validation experiments.
 
@@
 
-We sought to construct a training set consisting of stars across the main-sequence,
-sub-giant, and the red giant branch.  We required stars with precisely measured
+We sought to construct a training set of stars across the main-sequence, the
+sub-giant branch, and the red giant branch.  We demanded stars with precisely measured
 effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
-of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
+of O, Mg, Si, Ca, Al, Fe, and Ni.  This proved to be difficult because the magnitude
 range of \rave\ does not overlap substantially with high-resolution spectroscopic
 surveys.  The fourth internal data release of the \ges\ survey includes 
-giant and main-sequence stars, but only 142 overlap with \rave.  The thirteenth data 
-release from \sdss\ \citep{sloan_dr13} includes labels for \apogee\ stars on the
+giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
+be a useful training set for our purposes.  The thirteenth data release from the Sloan
+Digital Sky Survey \citep{sloan_dr13} includes labels for \apogee\ stars on the
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
-that the \apogee\ main-sequence labels suffered from significant systematic effects.  A flat, 
-then `up-turning' main-sequence is present, and the metallicity gradient trends in 
+that the \apogee\ main-sequence labels suffered from significant systematic effects.  
+A flat, then `up-turning' main-sequence is present, and the metallicity gradient trends in 
 the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
-stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram).
+stars incorrectly sit above an isochrone in a classical Hertzsprung-Russell diagram).
 If we consider lower-resolution studies as potential training sets, there are 2,369
@@
 stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
-but that sample contains only giant stars}, and similar systematic trends are apparent
-in the lower main-sequence.  The \galah\ survey has a good overlap with \rave, but 
-results from that sample are not yet precise enough for our purposes.
+but that sample contains only giant stars}, and the \lamost\ lower main-sequence suffers
+from the same systematic effects seen in \apogee\ data.  The \galah\ survey has a good 
+overlap with \rave, but results from that sample are not yet precise enough for our 
+purposes.
 
@@
 giants according to \apogee.  From this sample we selected stars to have: metallicity
-determinations ([Fe/H] $> -5$); high S/N ratios in both the \apogee\ and \rave\ spectra
-($>200$ in \apogee; $>50$ in \rave); and we further required that the \aspcap\ pipeline 
-did not report any peculiar flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left 
-us with 357 stars along the giant branch, with metallicities ranging from 
-$[{\rm Fe/H}] = -1.62$ to 0.32.  Intermediate tests with globular cluster members showed
-that the metallicity range of the training set needed to extend below 
-$[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be practically useful.
-Therefore, as we will describe in more detail below, we supplemented this sample with 
-giant stars from the \epic\ catalog \citep{Huber_2016}.
+determinations ([Fe/H] $> -5$); S/N ratios of $>20)$ in \apogee\ $>25$ in \rave; and 
+we further required that the \aspcap\ pipeline did not report any peculiar flags 
+(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the giant 
+branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Intermediate 
+tests with globular cluster members showed that the metallicity range of the training 
+set needed to extend below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be 
+practically useful.  Therefore, as we will describe in more detail below, we 
+supplemented this sample with giant stars from the \epic\ catalog \citep{Huber_2016}.
 
-Assembling a training set for the main-sequence and sub-giant branch was more complex.
-There are no spectroscopic studies that have a large enough sample size that overlap 
-with \rave, and which also cover our required range in stellar parameters 
-(e.g., FGKM-type stars).  Moreover, most of the spectroscopic studies we considered 
-exhibit a flat lower main-sequence, a systematic consequence of the analysis method 
-adopted \citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose
-to make use of the \epic\ catalog \citep{Huber_2016} for the training set labels on
-the main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
+Assembling a suitable training set for the main-sequence and sub-giant branch was less
+trivial.  There are no spectroscopic studies that have a large enough sample size that
+overlap with \rave\ and expand our required range in stellar parameters (e.g., FGKM-type 
+stars).  Moreover, most of the spectroscopic studies we considered also showed a flat 
+lower main-sequence, a systematic consequence of the analysis method adopted 
+\citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make 
+use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
+main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
 \project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
@@
 due to limited information available, it is possible that a star has been 
-mis-classified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
+misclassified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
 note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
-probability of misclassification is usually quantified in the uncertainties of 
-each star; most sub-giants that could be dwarfs have large confidence intervals.  
-Therefore, requiring low uncertainties will decrease the total sample size, but 
-in practice it removes most misclassifications.  The situation is far more favourable for 
-dwarfs and giants.  Only $1-4$\% of giant stars are misclassified as dwarfs, and 
-about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
-with narrow confidence intervals are usually of high fidelity, and any spurious 
-misclassifications would be clearly distinguishable if spectra were available.
+probability of misclassification is usually quantified in the uncertainties given
+for each star; most dwarfs that have a higher likelihood of being sub-giants have
+large confidence intervals.  Therefore, requiring low uncertainties will decrease 
+the total sample size, but in practice it removes most misclassifications.  The 
+situation is far more favourable for dwarfs and giants.  Only $1-4$\% of giant 
+stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
+giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
+usually of high fidelity, and any spurious misclassifications could be identified
+if spectra were available.
 
@@
 must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
-30~per~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
+30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
 removed most metal-poor stars, which we later found to cause the test labels to have
 under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
-(ignored) those quality constraints for stars in the \epic\ catalog with 
-$[{\rm Fe/H}] < -1$.  Of the stars we forcibly included with $[{\rm Fe/H}] < -1$,
-none were classified as being sub-giants in \epic, therefore they have a low (7\%)
-likelihood of misclassification.  After training a model based on main-sequence and
-giant stars (Section \ref{sec:the-model}), we found we could further identify 
-misclassifications by leave-one-out cross-validation (iteratively, even).  However,
-we chose not to do this because the number of likely misclassifications in the training
+(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  After training
+a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
+we could identify misclassifications by leave-one-out cross-validation.  However, we 
+chose not to do this because the number of likely misclassifications in the training
 set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
@@
 for all giant stars in our sample, but only imprecise metallicities are available from
-\epic\ for stars on the main-sequence and the sub-giant branch.  To our knowledge this 
-problem has not been addressed with data-driven models in the astrophysics literature.
-In the following section we will construct a simplified model that does not include detailed
-abundances, before demonstrating how we (1) improve upon the imprecise metallicities for 
-main-sequence/sub-giant stars, and (2) translate abundances from the giant branch stars 
-to the main-sequence/sub-giant branch sample.
-
-\subsection{The model}
-\label{sec:the-model}
-
-\noindent{}The model we adopt is
-\begin{eqnarray}\label{eq:model}
-y_{jn} & = & \vecv(\ell_n)\cdot\vectheta_j + e_{jn}\quad ,
-\end{eqnarray}
-
-\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
-$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
-$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
-$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
-wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
-vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Section 
-\ref{sec:a-simple-model}).  The noise values $e_{jn}$ can be considered to be drawn 
-from a Gaussian distribution with zero mean and variance $\sigma_{jn}^2 + s_j^2$, where
-$\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and $s_j^2$ describes the excess 
-variance at the $j$-th wavelength pixel.  We direct the reader to \citet{Ness_2015} for
-a more detailed introduction and justification of \thecannon.
-
-At the training step we fix the $K$-lists of labels for all the training set stars $n$.
-At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
-by optimizing the penalized likelihood function:
-\begin{eqnarray}\label{eq:train}
-\vectheta_j,s^2_j &\leftarrow& \argmin{\vectheta,s}\left[
-    \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
-    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
-    \right]
-  \quad ,
-\end{eqnarray}
-
-\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
-for each model in later sections, and $Q(\vectheta)$ is a regularizing function that 
-encourages $\vectheta$ values to take on zero values without breaking convexity:
-\begin{eqnarray}\label{eq:regularization-function}
-	Q(\vectheta) = \sum_{d=1}^{D-1} |{\theta_d}| \quad .
-\end{eqnarray}
-
-Note that the $d$ subscript here is zero-indexed; the function $Q(\vectheta)$ does not act
-on the (first) $\theta_0$ coefficient, as this is a `pivot point' or offset that we do not
-expect to diminish with increasing regularization (e.g., see equation 
-\ref{eq:vectorizer-three-label-design-matrix}).  In practice equation \ref{eq:train} is 
-optimized by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a convex optimization 
-problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
-
-The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
-pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
-the objective function is:
-\begin{eqnarray}\label{eq:test}
-  \ell_m &\leftarrow& \argmin{\ell}\left[
-    \sum_{j=0}^{J-1} \frac{[y_{jm}-\vecv(\ell)\cdot\vectheta_j]^2}{\sigma_{jm}^2 + s_j^2}
-    \right]
-  \quad .
-\end{eqnarray}
-
-At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
-$\ell$, which provides us with the formal errors on $\ell$.  The formal errors are expected
-to be underestimated, and in Section \ref{sec:validation} we judge the veracity of these 
-errors through validation experiments.
+\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section 
+we will construct separate models for the giant branch and main-sequence.  We will
+demonstrate the abundance labels are on the same scale, before combining the results from
+multiple models to produce a single catalog.
 
@@
 
-The model we have introduced is generic in the sense that there is no description for how
-complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
-Given previous successes using a second-order (quadratic) polynomial model 
-\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
-model experiments,
+The model we introduced in Section~\ref{sec:the-model} is generic in the sense that there 
+is no description for how complex the vectorizing function $\vecv(\ell)$ should be, or how
+many $K$ labels are used.  Given previous successes using a second-order (quadratic) 
+polynomial model \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a 
+three-label ($\teff$, $\logg$, [Fe/H]) model for all (main-sequence and giant) stars. Thus, 
+for our first set of model experiments,
 \begin{eqnarray}\label{eq:vectorizer-three-label}

### 2016-09-03 article/unrave.tex
Text updates

@@
 three-label ($\teff$, $\logg$, [Fe/H]) model for all (main-sequence and giant) stars. Thus, 
-for our first set of model experiments,
+for our first set of model experiments the vectorizer $\vecv(\ell_n)$ expands as,
 \begin{eqnarray}\label{eq:vectorizer-three-label}
@@
 
-\stub{We trained the model initially with no regularization.
-Could reproduce general structure of a main-sequence and HRD, but the intricacies of the structure was gone.
-Model appeared to be too simple.
-However we can use this model to identify dwarfs/giants,etc.
-Rather than go to a very complex model, we opted to use two models: one for the main-sequence and one for the giant branch.}
+We used no regularization ($\Lambda = 0$) for this model.  After training the model we
+treated all \Nspectra\ spectra as test set objects.  In Figure \ref{fig:3label-test-set}
+we show the two-dimensional distribution of effective temperatures $\teff$ and surface 
+gravities $\logg$ for all spectra with $\chi_{r}^2 < 3$ and $S/N > 10$.  The main-sequence
+and red giant branch are clearly visible.  However, the details of stellar evolution are
+no longer present: the sub-giant branch is not discernible, and there are a number of
+systematic artefacts (over-densities) in the label space of the test set.  These artefacts
+disappear when we require additional quality constraints (e.g., no peculiar morphological
+classifications), but the details of the Hertzsprung-Russell diagram are still not
+present.  Thus, we concluded that while this model is useful for discerning stellar
+classifications (e.g., F2-type giant), the labels are too imprecise.
+
+We chose to adopt a separate model for the main-sequence and the giant branch rather
+than switch to a single, higher order model.  This choice allows us to infer
+detailed abundance labels for giant stars, and to infer metallicity labels for stars
+on the main-sequence and sub-giant branch.  However, adopting two separate models
+introduces the challenge of how to combine the results from two models, or how to
+assign one star as being a main-sequence/giant star.  In Section~\ref{sec:joining-the-models}
+we describe how we will use the 3-label model of the main-sequence and giant branch
+(introduced in this section) to discriminate between results from the 3-label 
+main-sequence model in Section \ref{sec:unevolved-star-model} and the 9-label giant 
+star model in Section \ref{sec:evolved-star-model}.
 
 \subsection{A 3-label model for unevolved stars}
-\label{sec:the-ms-model}
+\label{sec:unevolved-star-model}
 
-We constructed a three-label quadratic model for the main-sequence and sub-giant
+We constructed a three-label quadratic model using only main-sequence and sub-giant
 stars in the \epic\ training set.  In order to set the regularization hyperparameter
@@
 will use as the basis for comparison.  For every model we calculated the bias and
-root-mean-square (RMS) deviation between the input labels and those inferred through
-cross-validation.
+root-mean-square (RMS) deviation between the training set labels and those inferred
+through cross-validation.
 
@@
 to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
-and lower envelope represent the boundaries across all labels, showing that the RMS
-decreased in \emph{all} labels with increasing regularization.  We found similar
-improvements in the bias values, however these were already minimal in the unregularized
+and lower envelope represent the boundaries across all labels, showing that with increasing
+regularization, the RMS decreased in \emph{all} labels.  We found similar
+improvements in the biases, however these were already minimal in the unregularized
 case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
-we achieve RMS deviations that are about 30\% better than the unregularized model.
+we achieve RMS deviations that are about 30\% better than the unregularized case.
 Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
@@
 
-We made no \emph{a priori} judgements on whether a star was likely a giant
-or a dwarf star; we inferred labels for all \rave\ spectra using this 
-main-sequence model.  The resulting labels for the survey sample are shown
-in Figure \ref{fig:hrd-from-individual-models}, where significant structure is 
-visible.  The increased density of Solar-type stars is consistent with \rave\
-observing stars in the local neighbourhood, and the high number of turnoff
-and main-sequence stars relative to the subgiant branch is expected from 
-the relative lifetimes of these evolutionary phases.  An overdensity
+We inferred labels for all \rave\ spectra using this main-sequence model.  
+The resulting labels for the survey sample are shown in Figure \ref{fig:hrd-from-individual-models}.  
+The increased density of Solar-type stars is consistent with \rave\
+observing stars in the local neighbourhood, and the high number of turn-off
+and main-sequence stars relative to the sub-giant branch is expected from 
+the relative lifetimes of these evolutionary phases.  An over-density
 of stars near the giant branch clump is also present.  This artefact is due 
@@
 model is (poorly) extrapolating outside the convex hull of the training set.
-In Section \ref{sec:joining-the-models} we describe how results from multiple
-models are combined to produce a single catalog.
 
 \subsection{A 9-label model for detailed abundances of giant stars}
-\label{sec:giant-abundance-model}
+\label{sec:evolved-star-model}
 
@@
 with nine labels.  The regularization hyperparameter $\Lambda$ was set in the same
-way described in Section \ref{sec:ms-model}, using the same 30 trials of $\Lambda$.
+way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
 The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
@@
 
-% All them hax used to combine all the results.
-
-% --> Chi-sq of the sample.
-% --> Convex hull? Distance to plane?
+% The single model.
 
@@
 \stub{Covariances between these parameters are accessible through the matrix of correlation coefficients in the associated results file.}
-
-\subsection{Jackknifing the training set}
-\label{sec:jackknife}
-
-\stub{Should we?}
 
@@
 
-% We have introduced a non-parametric version of The Cannon that uses strict priors on line formation (censoring masks), and prior probability distributions on astrophysical parameters, using isochrones.
-% We ran it on the entire RAVE spectra.
-% We deliver effective temperature, surface gravity, and abundances of up to 8 elements for \Nstars.
-Our internal and external validation tests suggest that we achieve a typical precision of X~K in $\teff$, 0.0X~dex in $\logg$, and $\approx{}X.XX$ in individual chemical abundances.
-This catalog constitutes the largest collection of stellar abundances for stars in the first \project{Gaia} data release.
-When combined with positions and 3D velocities from \project{Gaia}, the \project{UNRAVE} catalog will likely be crucial for understanding our local place in the Milky Way.
+We performed an independent analysis of \rave\ spectra using a data-driven approach.
+We report effective temperature $\teff$, surface gravity $\logg$, and metallicity [Fe/H]
+for X stars on the main-sequence and sub-giant branch.  For another X stars on the red
+giant branch, we deliver $\teff$, $\logg$, and a total of X.XX~million elemental abundances:
+O, Mg, Al, Ca, Si, Fe, and Ni for X stars.
+
+\stub{Typical precision}
+
+\stub{A key science result that we see already}
+
+\stub{Something about exoplanet host stars, TESS, etc}
+
+This catalog constitutes the largest collection of stellar abundances for stars in the
+first \gaia\ data release.  When combined with positions and 3D velocities from \gaia,
+we hope the \project{unRAVE} catalog will help advance understanding of the Milky Way's
+formation and evolution.
 
@@
 \acknowledgements
-A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and
-Sergey Koposov (Cambridge).
+We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and Sergey Koposov (Cambridge).
 This research made use of: Astropy, a community-developed core Python package for

### 2016-09-04 article/figures/plot_joint_model_metrics.py
Add figures relating to joining the models

@@ -1,0 +1,168 @@
+
+import os
+import matplotlib.pyplot as plt
+from astropy.table import Table, vstack
+from matplotlib.colors import LogNorm
+from matplotlib.ticker import MaxNLocator
+from scipy.spatial import Delaunay
+
+
+import AnniesLasso as tc
+
+output_filename, overwrite = \
+    ("/data/gaia-eso/arc/rave-data-files/unrave-v0.8-36-37-39.fits.gz", True)
+
+RESULTS_PATH = "/data/gaia-eso/arc/rave/results/"
+RESULTS_PATH = ""
+
+ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v37.fits.gz"))
+giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v36.fits.gz"))
+joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v41.fits.gz"))
+
+for t in (ms_results, giant_results, joint_results):
+    if "Name" not in t.dtype.names:
+        t["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-result")[0].replace("_result.pkl", "") for each in t["FILENAME"]]
+
+    t.sort("Name")
+
+assert np.all(ms_results["Name"] == joint_results["Name"])
+assert np.all(giant_results["Name"] == joint_results["Name"])
+
+ms_model = tc.load_model("rave-tgas-v37.model")
+
+# HACK MAGIC BEGINS #
+ms_model._labelled_set = Table.read("rave-tgas-v16b-labelled-set-cut.fits")
+ms_model._labelled_set["TEFF"] = ms_model._labelled_set["EPIC_TEFF"]
+ms_model._labelled_set["LOGG"] = ms_model._labelled_set["EPIC_LOGG"]
+ms_model._labelled_set["FE_H"] = ms_model._labelled_set["EPIC_FEH"]
+
+# HACK MAGIC ENDS
+
+
+ms_results["TEFF"] = ms_results["EPIC_TEFF"]
+ms_results["LOGG"] = ms_results["EPIC_LOGG"]
+ms_results["FE_H"] = ms_results["EPIC_FEH"]
+
+
+
+
[…]

### 2016-09-04 article/unrave.tex
Incorporate HWR comments on abstract

@@
 \newcommand{\sdss}{\project{\acronym{SDSS}}}
+\newcommand{\tgas}{\project{\acronym{TGAS}}}
 
@@
 \newcommand{\Nstars}{457,589}
+\newcommand{\Nstarsqc}{434,470}
 
@@
 \begin{abstract}
-The Milky Way may be our greatest laboratory for understanding galaxy formation.
-This premise relies on the ability to precisely determine the astrometry and
-chemistry for many individual stars, and to use those data to make inferences
-about the Galaxy's formation and evolution. On 14 September 2016 the \gaia\ 
-mission will release precise astrometry for $\approx2\times10^6$ stars.  
-However, a comprehensive catalog of detailed chemical abundances to complement
-those data is currently lacking.  The RAdial Velocity Experiment (\rave) survey 
-has acquired spectra for 292,036 stars in common with the first \gaia\ data 
-release, constituting the largest overlap of any stellar spectroscopic sample.  
-We have performed an independent analysis of \rave\ spectra using \thecannon,
-and here we report effective temperature $\teff$, surface gravity $\logg$, 
-as well as chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, 
-Ni) for \Nstars\ stars.  We validate our results internally from repeat visits, 
-and externally against existing high-resolution spectroscopic surveys.  The 
-typical precision is XXX~K in $\teff$, X.XX~dex in $\logg$, and X.XX~dex for
-detailed abundances.  We provide spectrophotometric distances based on our
-labels, improving upon the parallax-based distances available for distant 
-giants in the first \gaia\ data release.
-\stub{A one-line science result based on RAVE/Hipparcos overlap.}
-
+The Milky Way is a powerful laboratory for understanding galaxy formation,
+as it can provide orbits, ages, physical parameters, as well as chemical 
+abundances for vast sets of individual stars.  These inferences require 
+both astrometric and spectroscopic data.  Indeed, the Tycho-Gaia 
+Astrometric Solution (\tgas) sample in the first \gaia\ data release 
+requires a spectroscopic complement that includes radial velocities, 
+stellar parameters, and elemental abundances.  Among existing spectroscopic 
+samples, the RAdial Velocity Experiment (\rave) survey has the largest 
+overlap with \tgas: 292,036 stars.  Here we present a data-driven 
+re-analysis of \rave\ spectra, using an implementation of \thecannon, that
+yields more precise and accurate stellar parameters and abundances.  We 
+derive, and validate, improved effective temperature $\teff$, surface 
+gravity $\logg$, and chemical abundances of up to seven elements (O, Mg, 
+Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in chemical 
+abundances is X.XX~dex for stars on the giant branch.  We provide 
+spectrophotometric distances based on our stellar parameters, which should 
+be superior to the parallax-based distances for faint or distant objects in
+\tgas.  The unRAVE catalog presented here, in conjunction with \tgas, will 
+be the most powerful data set for chemo-orbital analysis at the dawn of 
+broadly available \gaia\ data.
 \end{abstract}
@@
 Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Kordopatis_2015,Bovy_2016}.  
-However, these quantities are not known for even 1\% of stars in the Milky Way.  Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
+However, these quantities are not known for even 1\% of stars in the Milky Way.  
+Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
 Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics 
@@
 with a low-order polynomial in label space.
-\item We have constructed a training set of stars with mean accurate labels.  Note
+\item The training set has mean accurate labels.  Note
 that we do not assume that \emph{every} label in the training is accurate -- we can
@@
 misclassifications in the training set are affordable).
-\item We assume our training set is representative of the test set (the \rave\ survey).
+\item We assume the training set is representative of the test set (the \rave\ survey).
 \end{itemize}
@@
 wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
-vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Section 
-\ref{sec:a-simple-model}).  The noise values $e_{jn}$ can be considered to be drawn 
-from a Gaussian distribution with zero mean and variance $\sigma_{jn}^2 + s_j^2$, where
-$\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and $s_j^2$ describes the excess 
-variance at the $j$-th wavelength pixel.
+vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Sections 
+\ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
+be considered to be drawn from a Gaussian distribution with zero mean and variance 
+$\sigma_{jn}^2 + s_j^2$, where $\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and 
+$s_j^2$ describes the excess variance at the $j$-th wavelength pixel.
 
-At the \emph{training} step we fix the $K$-lists of labels for all the training set stars $n$.
-At each wavelength pixel $j$, we then calculate the parameters $\vectheta_j$ and $s_j^2$
+At the \emph{training} step we fix the $K$-lists of labels for the $n$ training set.
+At each wavelength pixel $j$, we then find the parameters $\vectheta_j$ and $s_j^2$
 by optimizing the penalized likelihood function:
@@
 \noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
-for each model in later sections, and $Q(\vectheta)$ is a regularizing function that 
-encourages $\vectheta$ values to take on zero values without breaking convexity:
+in later sections, and $Q(\vectheta)$ is a regularizing function that encourages 
+$\vectheta$ values to take on zero values without breaking convexity:
 \begin{eqnarray}\label{eq:regularization-function}
@@
 Note that the $d$ subscript here is zero-indexed; the function $Q(\vectheta)$ does not act
-on the (first) $\theta_0$ coefficient, as this is a `pivot point' or offset that we do not
-expect to diminish with increasing regularization (e.g., see equation 
+on the (first) $\theta_0$ coefficient, as this is a `pivot point' (mean flux value) that 
+we do not expect to diminish with increasing regularization (e.g., see equation 
 \ref{eq:vectorizer-three-label}).  In practice equation \ref{eq:train} is 
@@
 
-At the test step optimization we store the covariance matrix $\bm{\Sigma}$ for the labels
-$\ell$, which provides us with the formal errors on $\ell$.  The formal errors are expected
-to be underestimated, and in Section \ref{sec:validation} we judge the veracity of these 
-errors through validation experiments.
+After optimizing equation \ref{eq:test} for the $m$-th star we store the covariance matrix 
+$\bm{\Sigma}_m$ for the labels $\ell_m$, which provides us with the formal errors on $\ell_m$. 
+The formal errors are expected to be underestimated, and in Section~\ref{sec:validation} 
+we judge the veracity of these errors through validation experiments.
 
@@
 giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
-be a useful training set for our purposes.  The thirteenth data release from the Sloan
-Digital Sky Survey \citep{sloan_dr13} includes labels for \apogee\ stars on the
+be a useful training set for our purposes.  The thirteenth data release from the 
+\project{Sloan Digital Sky Survey} \citep{sloan_dr13} includes labels for \apogee\ stars on the
 giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
@@
 giants according to \apogee.  From this sample we selected stars to have: metallicity
-determinations ([Fe/H] $> -5$); S/N ratios of $>20)$ in \apogee\ $>25$ in \rave; and 
-we further required that the \aspcap\ pipeline did not report any peculiar flags 
+determinations ([Fe/H] $> -5$); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
+we further required that \aspcap\ did not report any peculiar flags 
 (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the giant 
@@
 Assembling a suitable training set for the main-sequence and sub-giant branch was less
-trivial.  There are no spectroscopic studies that have a large enough sample size that
-overlap with \rave\ and expand our required range in stellar parameters (e.g., FGKM-type 
-stars).  Moreover, most of the spectroscopic studies we considered also showed a flat 
-lower main-sequence, a systematic consequence of the analysis method adopted 
+trivial.  There are no spectroscopic studies that both extend the range of stellar types
+we are interested in (e.g., FGKM-type stars), and have a large enough sample size that
+overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
+showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
 \citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make 
@@
 sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
-however there is a sparse sampling of stars reaching to $\logg \approx 4$.  For
-the \epic\ main-sequence/sub-giant star training set we required $\logg > 3.5$,
+however there is a sparse sampling of stars reaching to $\logg \approx 4$.  We
+required $\logg > 3.5$ for the \epic\ main-sequence/sub-giant star training set,
 allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
@@
 
+test
+
+\begin{figure*}[t!]
+\includegraphics[width=\textwidth]{figures/test-set-density.pdf}
+\caption{\label{fig:test-set-density}}
+\end{figure*}
+
+\begin{figure}[h!]
+\includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
+\caption{\label{fig:joint-model-differences}}
+\end{figure}
+
 \section{Validation experiments}
@@
 We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and Sergey Koposov (Cambridge).
-This research made use of: Astropy, a community-developed core Python package for
-Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
-and \project{TOPCAT} \citep{Taylor_2005}.
+This research made use of: 
+  Astropy, a community-developed core Python package for Astronomy \citep{astropy};
+  NASA's Astrophysics Data System Bibliographic Services;
+% numpy, scipy, matplotlib.
+and 
+  \project{TOPCAT} \citep{Taylor_2005}.
 This work was partly supported by the European Union FP7 programme through ERC 
-grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
+grant number 320360. Funding for K.~H.~ has been provided through the Simons 
+Foundation Society of Fellows and the Marshall Scholarship.
 Funding for RAVE has been provided by: the Australian Astronomical Observatory; 

### 2016-09-04 article/unrave.tex
some edits

@@
 \newcommand{\tgas}{\project{\acronym{TGAS}}}
+\newcommand{\unrave}{\project{unRAVE}}
 
@@
 % Remove at submission:
-\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 September 2016 (\difftoday{2016}{9}{7}away)}}}
+\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 September 2016 (\difftoday{2016}{9}{9}away)}}}
 
@@
 be superior to the parallax-based distances for faint or distant objects in
-\tgas.  The unRAVE catalog presented here, in conjunction with \tgas, will 
+\tgas.  The \unrave\ catalog presented here, in conjunction with \tgas, will 
 be the most powerful data set for chemo-orbital analysis at the dawn of 
@@
 parallaxes and proper motions for more than $10^9$ stars in its final data
-release in 2022.  While this is a sample size about three orders of magnitude 
+release in 2022.  While this is a sample size about four orders of magnitude 
 larger than its predecessor \hipparcos, both astrometry and chemistry are 
@@
 astronomers seeking to simultaneously use chemical and dynamical information are
-reliant on ground-based spectroscopic surveys in order to complement the 
-available \gaia\ astrometry.
+reliant on ground-based spectroscopic surveys to complement the available 
+\gaia\ astrometry.
 
-The first \gaia\ data release will include positions, proper motions, and 
-parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
-catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
-stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
-catalog against the \apogee\ \citep{Zasowski_2013}, \ges\
-\citep{Gilmore_2012,Randich_2013}, \galah\ \citep{DeSilva_2015},
-\lamost\ \citep{Cui_2012}, and \rave\ \citep{Steinmetz_2006} 
-surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
-have the largest overlap with the first \gaia\ data release: 292,036 
-stars.  We then used the \gaia\ universe model snapshot \citep{Robin_2012} to 
-estimate the precision in parallax and proper motions that will be available in 
-the first \gaia\ data release (DR1) for stars in those overlap samples.  Comparing
-the expected precision to what is currently available, we further found that the
-\rave\ survey will benefit most from \gaia\ DR1.  The distances of 63\% of the 
-\rave--\gaia\ DR1 overlap sample (182,862 stars) are expected to improve with 
-the first \gaia\ data release, and 47\% of stars are likely to have better 
-proper motions (137,211 stars).  Although the \gaia\ universe model assumes 
-end-of-mission uncertainties --- and does not account for systematics in the first
-data release --- this calculation still provides intuition for the relative 
-improvement the first \gaia\ data release can make to ground-based surveys.  The 
-expected improvements for \rave\ motivated us to examine what chemical abundances
+The first \gaia\ data release will include the Tycho-Gaia Astrometric Solution
+\citep[hereafter \tgas;][]{Michalik_2015a,Michalik_2015b}: positions, proper 
+motions, and parallaxes for approximately two million stars in the Tycho-2 
+\citep{Hog_2000} catalog.  After cross-matching all major stellar spectroscopic 
+surveys\footnote{Specifically we cross-matched the Tycho-2 catalog against the 
+\apogee\ \citep{Zasowski_2013}, \ges\ \citep{Gilmore_2012,Randich_2013}, 
+\galah\ \citep{DeSilva_2015}, \lamost\ \citep{Cui_2012}, and \rave\ 
+\citep{Steinmetz_2006} surveys}, we found the RAdial Velocity Experiment 
+(\rave) survey to have the largest overlap with the first \gaia\ data release: 
+292,036 stars.  We then used the \gaia\ universe model snapshot 
+\citep{Robin_2012} to estimate the precision in parallax and proper motions that
+will be available in the first \gaia\ data release (DR1) for stars in those 
+overlap samples.  Comparing the expected precision to what is currently available, 
+we further found that the \rave\ survey will benefit most from \gaia\ DR1.  The 
+distances of 63\% of the \rave--\gaia\ DR1 overlap sample (182,862 stars) are 
+expected to improve with the first \gaia\ data release, and 47\% of stars are 
+likely to have better proper motions (137,211 stars).  Although the \gaia\ universe
+model assumes end-of-mission uncertainties --- and does not account for systematics
+in the first data release --- this calculation still provides intuition for the 
+relative improvement the first \gaia\ data release can make to ground-based surveys.  
+The expected improvements for \rave\ motivated us to examine what chemical abundances
 were available from those data, and to evaluate whether we could enable new 
-chemodynamic studies by contributing to the existing set of chemical abundances.
+chemo-dynamic studies by contributing to the existing set of chemical abundances.
 
@@
 of these comparisons and our results in Section~\ref{sec:discussion}.  We provide
-concluding remarks in Section~\ref{sec:conclusion}, as well as a description for
-how to access our results electronically.
+concluding remarks in Section~\ref{sec:conclusion}, and describe how to access our
+results electronically.
 
@@
 outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
-there is a limited number of pre-processing steps that we conducted before our
-analysis began.  First, we calculated inverse variance arrays from the $1\sigma$ 
+there is a limited number of pre-processing steps that we performed before starting
+our analysis.  First, we calculated inverse variance arrays from the $1\sigma$ 
 uncertainties provided, and then we re-sampled the flux and inverse variance
@@
 There are two main analysis steps when using \thecannon: the \emph{training} 
-step and the \emph{test} step.  These stages are briefly described in the
-following section, and a more thorough introduction can be found in \citet{Ness_2015}.
-We make the following explicit assumptions about the \rave\ spectra and \thecannon:
+step and the \emph{test} step.  We describe these stages in the context of our
+model in the following section, and a more thorough introduction can be found
+in \citet{Ness_2015}.  We make the following explicit assumptions about the 
+\rave\ spectra and \thecannon:
 
@@
 \item We assume that the normalization procedure employed by the \rave\ pipeline
-is invariant with respect to the labels we seek to measure, and invariant with 
-respect to the S/N ratio.
+is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
+or [Fe/H]), and invariant with respect to the S/N ratio.
 \item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
@@
 with a low-order polynomial in label space.
-\item The training set has mean accurate labels.  Note
-that we do not assume that \emph{every} label in the training is accurate -- we can
-afford to have a small fraction of inaccurate labels (e.g., a few obvious 
+\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
+Note that we do not assume that \emph{every} label in the training is accurate -- we
+can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
 misclassifications in the training set are affordable).
@@
 \end{itemize}
-
-%We adopt the nomenclature first set out in \citet{Ness_2015}, and extended in 
-%\citet{Casey_2016b}.  However, one distinction here with \citet{Casey_2016b} is 
-%that in this work the \emph{labelled} set and the \emph{training} set will be 
-%the same set.  In \citet{Casey_2016b} two subsets of the labelled set were 
-%defined: the training set and a distinct validation set.  This allowing us to 
-%train on the \emph{training set} and then test on the \emph{validation set}.  
-%However in this work the labelled set is relatively small, about an order of 
-%magnitude smaller than what was available in \citet{Casey_2016b}, and therefore 
-%no validation set exists. We will use all labelled set stars in the training 
-%step and supplement our validation experiments with literature comparisons.  
-%Thus, throughout this work we will use the terms \emph{labelled set} and 
-%\emph{training set} interchangeably.
 
@@
 At each wavelength pixel $j$, we then find the parameters $\vectheta_j$ and $s_j^2$
-by optimizing the penalized likelihood function:
+by optimizing the penalized likelihood function
 \begin{eqnarray}\label{eq:train}
@@
 we do not expect to diminish with increasing regularization (e.g., see equation 
-\ref{eq:vectorizer-three-label}).  In practice equation \ref{eq:train} is 
-optimized by fixing $s_j^2 = 0$ to make equation~\ref{eq:train} a convex optimization 
-problem, then we optimize for $\vectheta_j$, before solving for $s_j^2$.
+\ref{eq:vectorizer-three-label}).  In practice we first fix $s_j^2 = 0$ to make equation
+\ref{eq:train} a convex optimization problem, then we optimize for $\vectheta_j$, before 
+solving for $s_j^2$.
 
@@
 $g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
-lower resolution, no detailed abundance information is available for main-sequence 
+lower resolution, there are no elemental abundances available for the main-sequence 
 stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
@@
 we further required that \aspcap\ did not report any peculiar flags 
-(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the giant 
-branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Intermediate 
+(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
+branch, with metallicities ranging from $[{\rm Fe/H}] = -1.79$ to 0.26.  Intermediate 
 tests with globular cluster members showed that the metallicity range of the training 
 set needed to extend below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be 
-practically useful.  Therefore, as we will describe in more detail below, we 
-supplemented this sample with giant stars from the \epic\ catalog \citep{Huber_2016}.
+practically useful.  \stub{Supplemented this sample with known metal-poor stars in \rave}.
 
 Assembling a suitable training set for the main-sequence and sub-giant branch was less
-trivial.  There are no spectroscopic studies that both extend the range of stellar types
-we are interested in (e.g., FGKM-type stars), and have a large enough sample size that
+trivial.  There are no spectroscopic studies that extend the range of stellar types we 
+are interested in (e.g., FGKM-type stars), and also have a large enough sample size that
 overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
 showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
-\citep[e.g., see][for a discussion]{Bensby_2014}.  For these reasons we chose to make 
+\citep[see][for discussion on this issue]{Bensby_2014}.  For these reasons we chose to make 
 use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
@@
 giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
-usually of high fidelity, and any spurious misclassifications could be identified
-if spectra were available.
+usually of high fidelity, and given that we have spectra, we can identify any
+spurious misclassifications.
 
@@
 under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
-(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  After training
+(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  
+\stub{Include known metal-poor stars in \rave?}. After training
 a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
@@
 
-We previously alluded to including giant stars from \epic\ in our training set based 
-on \apogee\ labels.  This was principally to ensure an adequate balance of metal-poor
-giant stars in the training set.  Without this inclusion, we again found that the
-abundances of giant stars would be under-predicted at low metallicity, critically 
-affecting any inferences we wish to make with low metallicity stars (e.g., globular
-clusters).  The metallicities in the \epic\ catalog come from a variety of spectroscopic
-sources, or photometrically-selected when no spectroscopic information is available.
-While these heterogeneous metallicities may introduce scaling differences, on balance we
-later found that it was more important to include reasonably accurate (albeit imprecise)
-metallicities of metal-poor stars in the training set rather than excluding metal-poor
-stars entirely.  In Section~\ref{sec:validation} we show that these potential issues
-have a negligible impact on the test set labels, and we further show that our abundances
-are consistent with high-resolution studies in the literature.
+\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
+\label{sec:a-simple-model}
 
@@
 for all giant stars in our sample, but only imprecise metallicities are available from
-\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section 
-we will construct separate models for the giant branch and main-sequence.  We will
-demonstrate the abundance labels are on the same scale, before combining the results from
-multiple models to produce a single catalog.
-
-\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
-\label{sec:a-simple-model}
-
-The model we introduced in Section~\ref{sec:the-model} is generic in the sense that there 
-is no description for how complex the vectorizing function $\vecv(\ell)$ should be, or how
-many $K$ labels are used.  Given previous successes using a second-order (quadratic) 
-polynomial model \citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a 
-three-label ($\teff$, $\logg$, [Fe/H]) model for all (main-sequence and giant) stars. Thus, 
-for our first set of model experiments the vectorizer $\vecv(\ell_n)$ expands as,
+\epic\ for stars on the main-sequence and the sub-giant branch.  Here we will construct 
+a simple model for \emph{all} stars that only makes use of three labels ($\teff$, $\logg$, 
+[Fe/H]), before we outline how we derive abundances for giant branch stars.  The complexity
+for this model will be quadratic ($\teff^2$ is the highest term), where the vectorizer 
+$\vecv(\ell_n)$ expands as,
 \begin{eqnarray}\label{eq:vectorizer-three-label}
@@
 We used no regularization ($\Lambda = 0$) for this model.  After training the model we
-treated all \Nspectra\ spectra as test set objects.  In Figure \ref{fig:3label-test-set}
-we show the two-dimensional distribution of effective temperatures $\teff$ and surface 
-gravities $\logg$ for all spectra with $\chi_{r}^2 < 3$ and $S/N > 10$.  The main-sequence
-and red giant branch are clearly visible.  However, the details of stellar evolution are
-no longer present: the sub-giant branch is not discernible, and there are a number of
-systematic artefacts (over-densities) in the label space of the test set.  These artefacts
-disappear when we require additional quality constraints (e.g., no peculiar morphological
-classifications), but the details of the Hertzsprung-Russell diagram are still not
-present.  Thus, we concluded that while this model is useful for discerning stellar
-classifications (e.g., F2-type giant), the labels are too imprecise.
+treated all \Nspectra\ spectra as test set objects.  In the right-hand panel of Figure 
+\ref{fig:tes-set-density} we show the effective temperature $\teff$ and surface gravity 
+$\logg$ for all spectra.  The main-sequence and red giant branch are clearly visible.  
+However, the details of stellar evolution are no longer present: the sub-giant branch is 
+not discernible, and there are a number of systematic artefacts (over-densities) present
+in label space.  These artefacts disappear when we require additional quality constraints 
+(e.g., no peculiar morphological classifications), but the complexity of the 
+Hertzsprung-Russell diagram is still not present.  Thus, we concluded that while this 
+model could be useful for deriving stellar classifications (e.g., F2-type giant), the 
+labels are too imprecise.
 
-We chose to adopt a separate model for the main-sequence and the giant branch rather
-than switch to a single, higher order model.  This choice allows us to infer
-detailed abundance labels for giant stars, and to infer metallicity labels for stars
-on the main-sequence and sub-giant branch.  However, adopting two separate models
-introduces the challenge of how to combine the results from two models, or how to
-assign one star as being a main-sequence/giant star.  In Section~\ref{sec:joining-the-models}
+We chose to adopt separate models for the main-sequence and the red giant branch rather
+than switch to a single model with more complexity.  This choice allowed us to derive
+stellar parameters for stars on the main-sequence and sub-giant branch, as well as 
+detailed elemental abundances for red giant branch stars.  However, adopting two separate
+models introduces the challenge of how to combine the results from two models, or how to
+assign one star as `belonging' to a single model.  In Section~\ref{sec:joining-the-models}
 we describe how we will use the 3-label model of the main-sequence and giant branch
-(introduced in this section) to discriminate between results from the 3-label 
-main-sequence model in Section \ref{sec:unevolved-star-model} and the 9-label giant 
+(introduced in this section) to discriminate between results from a 3-label 
+main-sequence model in Section \ref{sec:unevolved-star-model} and a 9-label giant 
 star model in Section \ref{sec:evolved-star-model}.
 
-\subsection{A 3-label model for unevolved stars}
+\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for unevolved stars}
 \label{sec:unevolved-star-model}
@@
 We show the \emph{percentage difference} in the RMS deviation of the labels with respect
-to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
+to the unregularized model in Figure \ref{fig:set-hyperparameters}.  The upper 
 and lower envelope represent the boundaries across all labels, showing that with increasing
@@
 case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
+%3.56224789e+01
 we achieve RMS deviations that are about 30\% better than the unregularized case.
-Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
-At this regularization strength, the bias and RMS values found by leave-one-out
+Based on this improvement we set $\Lambda = X.XX$ for this model.  At this regularization 
+strength, the bias and RMS values found by leave-one-out
 cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
@@
 
-We inferred labels for all \rave\ spectra using this main-sequence model.  
-The resulting labels for the survey sample are shown in Figure \ref{fig:hrd-from-individual-models}.  
+We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
+The results for the survey sample are shown in the left panel of Figure \ref{fig:test-set-density}.  
 The increased density of Solar-type stars is consistent with \rave\
@@
 
-We only have $\teff$, $\logg$, and [Fe/H] for unevolved stars in our training
-set from the \epic\ catalog \citep{Huber_2016}.  However, the giant stars 
-have stellar parameters ($\teff$, $\logg$) and up to 15 individual abundances
-from the \aspcap\ \citep{aspcap_paper}.  A subset of these elements have atomic
-transitions in the \rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, 
-\ion{Al}{1}, \ion{Si}{1}, \ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  
-We excluded [Ti/H] from our abundance list because of systematics in the training
+The red giant branch stars in our training set have have stellar parameters 
+($\teff$, $\logg$) and up to 15 individual abundances from the \aspcap\ 
+\citep{aspcap_paper}.  A subset of these elements have atomic transitions in the 
+\rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, 
+\ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  However, we 
+excluded [Ti/H] from our abundance list because of systematics in the training
 set \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left with nine labels 
-in our giant star model: $\teff$, $\logg$, and seven individual abundances.
+in our giant star model: $\teff$, $\logg$, and seven elemental abundances.
 
-We used a quadratic vectorizer for the giant star model, where the quadratic and
-cross-terms are expanded in the same way as \ref{eq:vectorizer-three-label}, only
-with nine labels.  The regularization hyperparameter $\Lambda$ was set in the same
-way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
-The results are shown in Figure \ref{fig:set-giant-model-hyperparameters}, where
+Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
+we used a quadratic vectorizer for the giant star model.  Here the terms are 
+expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
+with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
+in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
+The results are shown in Figure \ref{fig:set-hyperparameters}, where
 again the enveloped region represents the minimum and maximum change in RMS label
@@
 and up to $X$\%.  Here the regularization strength also produces a sparser matrix
-of $\vectheta$, with a $\approx20$\% increase in the fraction of zero-valued 
-entries, particularly for the cross-terms (Figure \ref{fig:giant-model-sparsity},
-third panel).  Based on the increased model sparsity and decreasing RMS deviation
-in the labels, we adopt $\Lambda = X.XX$ for our giant star model.
+of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
+Based on the increased model sparsity and decreasing RMS deviation in the labels, 
+we adopt $\Lambda = X.XX$ for the giant star model.
 
 Results from cross-validation on our regularized model imply that the precision
-in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and X.XX~dex (EL) and 
+in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and between X.XX~dex (EL) and 
 X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
-XXX,XXX \rave\ spectra using this model, and these results are summarized in
-Figure \ref{fig:test-set-hrd} (middle panel).  The red clump is clearly visible
+\Nspectra\ \rave\ spectra using this model, and these results are summarized in
+Figure \ref{fig:test-set-density} (middle panel).  The red clump is clearly visible
 and in the expected location, without requiring any post-analysis calibration.
@@
 
-% The single model.
+% For each spectrum we have results from three different models, described in the previous sections
 
-test
+% the artefacts show that the models poorly extrapolate outside the training set.
+% the joint model shows that a single quadratic model is too simple to apply to the entire rave data
+
+% we cannot just use reduced chi sq to exclude stars, nor can we just exclude things that are
+% outside the range of values in the training set --> stars project back into the label space
+
+% therefore we will use the joint model to inform us as to whether we should adopt
+% results from the giant star model, the dwarf model, or a linear combination of the two.
+
+% Show results compared to the joint model.
+% Central density of points that agree well with the joint model.
+% Sigmas chosen to scale both models so that the central densities are approximately isotropic.
+% The artefacts in the previous figures appear here too, in strange locations.
+
+% Equations for calculating weights
+
+% Summing bsaed on weights
+
+% Show HRD with mean weight fraction
 
@@
 The filled area represents the minimum and maximum improvements across all labels.
-The maximum improvement (decrease) in RMS label deviations occurs at $\Lambda \approx 0.3$, and this improvement persists until $\Lambda \approx 0.6$.}
+The maximum improvement (decrease) in RMS label deviations occurs at $\Lambda \approx 0.3$, and this improvement persists until $\Lambda \approx 0.6$.\label{fig:set-hyperparameters}}
 \end{figure}

### 2016-09-04 article/unrave.tex
Updates

@@
 transitions that are dominated by pressure broadening --- which are visible even 
-in very low signal-to-noise (S/N) data or metal-poor stars.  Atomic transitions of 
+in metal-poor stars or spectra with very low signal-to-noise (S/N) ratios.
+Atomic transitions of 
 light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
@@
 The red giant branch stars in our training set have have stellar parameters 
-($\teff$, $\logg$) and up to 15 individual abundances from the \aspcap\ 
-\citep{aspcap_paper}.  A subset of these elements have atomic transitions in the 
+($\teff$, $\logg$) and up to 15 elemental abundances from the \aspcap\ 
+\citep{Garcia_Perez_2016}.  A subset of these elements have atomic transitions in the 
 \rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, 
 \ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  However, we 
-excluded [Ti/H] from our abundance list because of systematics in the training
-set \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left with nine labels 
-in our giant star model: $\teff$, $\logg$, and seven elemental abundances.
+excluded [Ti/H] from our abundance list because of systematics in the \aspcap\
+[Ti/H] abundances \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left 
+with nine labels in our giant star model: $\teff$, $\logg$, and seven elemental 
+abundances.
 
@@
 
-% For each spectrum we have results from three different models, described in the previous sections
+We have derived labels for all \Nspectra\ \rave\ spectra using the three models
+described in previous sections.  The results from the joint model --- that includes 
+the main-sequence, sub-giant and red giant branch --- shows that a single quadratic
+model is too simple for the \rave\ spectral range.  The other models have problems,
+too: unrealistic over-densities in label space show that the main-sequence model and
+the giant model make very poor extrapolations for stars outside their respective
+training sets.  For these reasons we were forced to exclude or severely penalize 
+incorrect results from both models.
 
-% the artefacts show that the models poorly extrapolate outside the training set.
-% the joint model shows that a single quadratic model is too simple to apply to the entire rave data
+Before attempting to join the results from different models, we excluded results
+in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
+labels that are outside the extent of the training set.  Specifically for the
+results from the giant model we (conservatively) excluded stars with derived 
+$\logg > 3.5$, and for the results from the main-sequence model we excluded 
+sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the 
+two-dimensional ($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  Unfortunately these restrictions did not remove all spurious
+results.  The reason for this can be explained with an example:  consider that our 
+giant star model was trained with only giant stars but tested with both giant 
+stars and dwarf stars.  Some classes of stars (e.g., metal-poor dwarfs) can 
+project into a region of label space that would suggest it is a giant 
+(e.g., a clump star).  These objects could have relatively low $\chi_{r}^2$ values
+(e.g., $\chi_{r}^2 < 3$) and in this example, they would appear as bonafide red clump 
+stars.  These incorrect projections are extrapolation errors in high dimensions that
+project to `normal' parts of the label space in two dimensions.  For these reasons 
+we also made use of the joint model in Section~\ref{sec:a-simple-model}
+to inform whether we should adopt results from: the red giant branch model; the
+main-sequence/sub-giant model; or a linear combination of the two.
 
-% we cannot just use reduced chi sq to exclude stars, nor can we just exclude things that are
-% outside the range of values in the training set --> stars project back into the label space
+In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
+$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
+model; and the differences between the red giant branch model and the joint model. We have scaled
+the differences in $\teff$ and $\logg$ to make the central peak near $(0, 0)$ to 
+% ARC to HOGG: This is a total hack. Should we fit for these and \rho_{\teff,\logg} or are you content with the hack?
+be approximately isotropic by setting
+	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
+	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
+	$\delta_{\log{g}} = 0.15$~dex for both models.
+The stars within the peak at $(0, 0)$ represent objects where
+the joint model and the comparison model both report similar labels.  The artefacts 
+seen in the Hertszprung-Russell diagrams in Figure \ref{fig:test-set-density} are
+also present in Figure \ref{fig:joint-model-differences} as over-densities far away
+from the central peak.  Therefore, we can 
+adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model,
+% ARC to HOGG: Please suggest any nomenclature or definition changes as you see fit!
+\begin{eqnarray}
+	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2   \nonumber \\
+	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2   \quad ,
+\end{eqnarray}
 
-% therefore we will use the joint model to inform us as to whether we should adopt
-% results from the giant star model, the dwarf model, or a linear combination of the two.
+\noindent{}as a metric to derive the weights,
+\begin{eqnarray}
+	w_{ms} = \frac{1}{{d_{ms}}^2} & \text{and} & w_{giant} = \frac{1}{{d_{giant}}^2} \quad ,
+\end{eqnarray}
 
-% Show results compared to the joint model.
-% Central density of points that agree well with the joint model.
-% Sigmas chosen to scale both models so that the central densities are approximately isotropic.
-% The artefacts in the previous figures appear here too, in strange locations.
+\noindent{}and produce the weighted labels $\hat\ell$:
+\begin{eqnarray}
+	\hat\ell = \frac{w_{ms}\ell_{ms} + w_{giant}\ell_{giant}}{w_{ms} + w_{giant}} \quad .
+\end{eqnarray}
 
-% Equations for calculating weights
+We calculate weighted errors of $\hat\ell$ and correlation coefficients in 
+the same manner.  In Figure \ref{fig:model-weights} we show the mean relative
+weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional bin of 
+$\hat\teff$ and $\hat\logg$.  For giant stars the relative weight of 
+the main-sequence model is zero, and vice-versa for main-sequence stars.
+The relative weights smoothly transition from 0 to 1 on the sub-giant branch
+near $\log{g} \approx 3.5$, in the training set overlap region of both models.
+For abundance labels in the giant model that are not in the main-sequence
+model (e.g., [O/H], [Mg/H]), we only report abundances if there was no contribution
+from the main-sequence model: $w_{ms} = 0$.
 
-% Summing bsaed on weights
-
-% Show HRD with mean weight fraction
-
-\begin{figure*}[t!]
-\includegraphics[width=\textwidth]{figures/test-set-density.pdf}
-\caption{\label{fig:test-set-density}}
-\end{figure*}
-
-\begin{figure}[h!]
-\includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
-\caption{\label{fig:joint-model-differences}}
-\end{figure}
+Our catalog now contains entries of $XXX$ stars with effective temperature $\teff$,
+surface gravity $\logg$, iron abundance [Fe/H], and $XXX$ of those stars are giants
+with elemental abundances of [O/H], [Al/H], [Mg/H], [Ca/H], [Si/H], and [Ni/H].
+In total, this leads to a catalog of X.XX million stellar abundances. 
+The weighted $\teff$ and $\logg$ values for stars meeting different
+S/N constraints are shown in Figure \ref{fig:test-set-hrd}, both in logarithmic density
+and mean metallicity.  The artefacts from individual models are no longer 
+apparent, and the rich structure of the Hertzsprung-Russell is visible.
 
@@
 In addition to the cross-validation tests that we have previously described, 
-we conducted a number of internal and external validation experiments to 
-test the validity of our results.
+we have conducted a number of internal and external validation experiments to 
+test the validity of our results.  We will begin by describing internal validation
+tests based on repeat observations, before evaluating our accuracy based on
+high-resolution literature comparisons.
 
@@
 
-\stub{Covariances between these parameters are accessible through the matrix of correlation coefficients in the associated results file.}
-
-\subsection{Quality control}
-\stub{What should we adopt as recommended QC? Show rogue gallery of a few "high s/n cases" where things went haywire?}
-
 \subsection{External validation}
@@
 
+\subsubsection{Comparison with \rave\ DR5}
+
 Here we cross-match our results against literature samples to verify that
-our labels are accurate.  Because this work is an independent analysis, our
-first point of reference is against the official \rave\ data releases.  These
-comparisons are shown in Figures \ref{fig:rave-dr4-comparison} and 
-\ref{fig:rave-dr5-comparison}, where we show our labels with respect to the
-results in the fourth and fifth \rave\ data releases, respectively.  In order
-to provide a fair comparison, we only show stars that meet a number of quality
-flags in both samples. 
+our results are accurate.  Because this work is an independent analysis, our
+first point of reference is against the official \rave\ data release.  This
+comparison is shown in Figure \ref{fig:rave-dr5-comparison}, 
+where we show our results with respect to the fifth \rave\ data release \citep{Kunder_2016}.  In order
+to provide a fair comparison, we only show the $XXX$ stars that meet a number of quality
+flags in both samples: 
 \stub{What selections qualify as a fair comparison in both samples?}
@@
 
+\subsubsection{Comparisons with Reddy, Bensby, and Valenti \& Fischer}
+
+\stub{Caution re MP dwarfs appear as sub-giants in our sample}
+
+\subsubsection{Comparison with the \ges\ survey}
+
 \stub{Discuss comparisons with: GES; calibration sources used in \citet{Kordopatis_2013}; and `gold-standard' studies: Reddy, Bensby, Valenti \& Fischer}
-
-\stub{Any differences in RAVE abundances? Corrections required due to having an unbalanced training set? Any caution required?}
 
@@
 After verifying that our stellar labels are comparable with other
-high-resolution studies, here we verify that the abundance labels we find
+high-resolution studies, here we verify that the abundances we find
 are consistent with expectations from astrophysics.  One example is stars in
 globular clusters, which show anti-correlations in light element abundances 
-\citep[e.g.,][and references therein]{Da_Costa,Carretta_2009}. In the \rave\
+\citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
 survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters.  We have used that sample
-Specifically, this GC sample includes 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
+consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
 globular cluster NGC~3201, and 10 members of NGC~362.  
 All of these systems are well-studied and exhibit abundance anti-correlations 
-in Na--O and Mg--Al \citep{people}.  Other correlations are also present 
+in Na--O and Mg--Al \citep{Marino_2011,Carretta_2009,Carretta_2013,Munoz_2013}.  Other correlations are also present 
 (e.g., C--N), but those relationships are less relevant for this work because 
 we do not measure C and N abundances here.  We note that \emph{none} of these 
-cluster stars are present in our training set.
+globular cluster stars are present in our training set.
 
@@
 \acknowledgements
-We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and Sergey Koposov (Cambridge).
+We thank 
+	Jonathan Bird (Vanderbilt),
+	Sven Buder (MPIA), 
+and 
+	Sergey Koposov (Cambridge).
 This research made use of: 
+  	NASA's Astrophysics Data System Bibliographic Services;
   	Astropy, a community-developed core Python package for Astronomy \citep{astropy};
-  NASA's Astrophysics Data System Bibliographic Services;
-% numpy, scipy, matplotlib.
 and 
@@
 
+\bibitem[Carretta et al.(2009)]{Carretta_2009} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2009, \aap, 505, 117
+
+\bibitem[Carretta et al.(2013)]{Carretta_2013} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2013, \aap, 557, A138
+
 \bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88
@@
 \bibitem[Deason et al.(2011)]{Deason_2011} Deason, A.~J., Belokurov, V., \& Evans, N.~W.\ 2011, \mnras, 416, 2903
+
+\bibitem[Garc{\'{\i}}a P{\'e}rez et al.(2016)]{Garcia_Perez_2016} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., et al.\ 2016, \aj, 151, 144
 
@@
 
+\bibitem[Marino et al.(2011)]{Marino_2011} Marino, A.~F., Milone, A.~P., Piotto, G., et al.\ 2011, \apj, 731, 64
+
 \bibitem[Michalik et al.(2015a)]{Michalik_2015a} Michalik, D., Lindegren, L., \& Hobbs, D.\ 2015, \aap, 574, A115
@@
 \bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
+
+\bibitem[Mu{\~n}oz et al.(2013)]{Munoz_2013} Mu{\~n}oz, C., Geisler, D., \& Villanova, S.\ 2013, \mnras, 433, 2006
 
@@
 
+\bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
+
 \bibitem[Randich et al.(2013)]{Randich_2013} Randich, S., Gilmore, G., \& Gaia-ESO Consortium 2013, The Messenger, 154, 47
@@
 \bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
+
+\bibitem[Reddy et al.(2003)]{Reddy_2003} Reddy, B.~E., Tomkin, J., Lambert, D.~L., \& Allende Prieto, C.\ 2003, \mnras, 340, 304
+
+\bibitem[Reddy et al.(2006)]{Reddy_2006} Reddy, B.~E., Lambert, D.~L., \& Allende Prieto, C.\ 2006, \mnras, 367, 1329
 
@@
 \bibitem[Taylor(2005)]{Taylor_2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+
+\bibitem[Valenti \& Fischer(2005)]{Valenti_Fischer_2005} Valenti, J.~A., \& Fischer, D.~A.\ 2005, \apjs, 159, 141
 
@@
 
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/test-set-density.pdf}
+\caption{\label{fig:test-set-density}}
+\end{figure*}
+
 \begin{figure}[p]
@@
 The maximum improvement (decrease) in RMS label deviations occurs at $\Lambda \approx 0.3$, and this improvement persists until $\Lambda \approx 0.6$.\label{fig:set-hyperparameters}}
+\end{figure}
+
+\begin{figure}[p]
+\center
+\includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
+\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the joint model (top panel), and the giant model and the joint model (bottom panel).
+The density scaling is logarithmic, and the differences in $\teff$ and $\logg$ are scaled to make them approximately isotropic (see text for details).
+The peak at $(0, 0)$ represents good agreement between the joint model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
+\end{figure}
+
+\begin{figure}[p]
+\center
+\includegraphics[width=0.5\textwidth]{figures/model-weights.pdf}
+\caption{The mean relative main-sequence model weight $w_{ms}/(w_{ms} + w_{giant})$ at each hexagonal bin of weighted effective temperature $\teff$ and surface gravity $\logg$.
+The relative weighting illustrates how only results from the main-sequence model are adopted for unevolved stars, and there is a gradual transition to using results from the giant model, before only results from the giant model are used for evolved stars.\label{fig:model-weights}}
 \end{figure}

### 2016-09-04 combine_model_results.py

@@ -236,14 +236,42 @@ ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], gridsize=N,
 
 
 
-N = 70
-fig, ax = plt.subplots()
+# Plot log(density) of the three models.
+K = 1
+factor = 3.5
+lbdim = 0.2 * factor
+tdim = 0.1 * factor
+rdim = 0.25 * factor
+whspace = 0.05
+xspace = factor
+yspace = factor * K + factor * (K - 1) * whspace + lbdim * (K - 1)
+xdim = lbdim + xspace + rdim
+ydim = lbdim + yspace + tdim
+
+fig, ax = plt.subplots(K, 1, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+N = 35
 
 ok = combined_table["snr"] > 10
-ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], gridsize=N,
-    extent=(3000, 7500, 0, 5.5), C=weights2[0][ok], edgecolor="#ffffff", linewidths=0.0)
+_ = ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], gridsize=N,
+    extent=(3000, 8000, 0, 5.5), C=weights2[0][ok], edgecolor="#ffffff", linewidths=0.1,
+    cmap="Spectral")
 
+ax.set_xlim(ax.get_xlim()[::-1])
+ax.set_ylim(ax.get_ylim()[::-1])
+ax.set_xlabel(r"$T_{\rm eff}$ $[{\rm K}]$")
+ax.set_ylabel(r"$\log{g}$")
 
+cbar = plt.colorbar(_, ticks=[0, 0.25, 0.5, 0.75, 1.0])
+cbar.set_label(r"$w_{ms}/(w_{ms} + w_{giant})$")
+
+fig.tight_layout()
+fig.savefig("article/figures/model-weights.pdf", dpi=300)
+fig.savefig("article/figures/model-weights.png", dpi=300)
 
 
 raise a

### 2016-09-04 article/unrave.tex
filler text

@@
 
+\subsection{Spectrophotometric distances}
+\label{sec:distances}
+
+\stub{These will be derived by Jason Sanders when he returns from his honeymoon.
+The timing means that the distances will not be available when we submit the first
+version to arXiv. So we may need to remove this section and the reference to it
+from the abstract.}
+
 \section{Validation experiments}
@@
 
+\subsubsection{Comparison with the \citet{Kordopatis_2013} calibration sample}
+\label{sec:validation-kordopatis}
+
 \subsubsection{Comparisons with Reddy, Bensby, and Valenti \& Fischer}
+\label{sec:validation-gold-standards}
 
@@
 \subsubsection{Comparison with the \ges\ survey}
-
-\stub{Discuss comparisons with: GES; calibration sources used in \citet{Kordopatis_2013}; and `gold-standard' studies: Reddy, Bensby, Valenti \& Fischer}
+\label{sec:validation-ges}
+
+There are 142 stars that overlap between \rave\ and the fourth internal
+data release of the \ges\ survey. These are a mix of main-sequence, 
+sub-giant and red giant branch stars.  Despite most of these stars having
+relatively low S/N ratios in \rave\ ($\approx 25$), there is good 
+agreement in stellar parameters (Figure~\ref{fig:ges-comparison}.  
+The RMS in effective temperature, surface gravity and metallicity is
+236~K, 0.39~dex, and 0.18~dex, respectively. 
+% TODO: Check if those numbers above are still valid when final catalog is used.
+
+\stub{A paragraph about the abundance agreement.}
 

### 2016-09-04 article/figures/plot_dr4_comparison.py
Updated figures except for training set hrd and abundance-related ones

@@ -13,30 +13,29 @@ try:
     combined_table
 
 except NameError:
-    from rave_io import (rave_kordopatis_dr4, rave_cannon_dr1)
+    from rave_io import (get_rave_kordopatis_dr4, get_cannon_dr1)
 
     from astropy.table import join
 
-    combined_table = join(rave_cannon_dr1, rave_kordopatis_dr4, keys=("Name", ))
+    combined_table = join(get_cannon_dr1(), get_rave_kordopatis_dr4(), keys=("Name", ))
 
 else:
     print("Warning: Using pre-loaded data!")
 
-QC = combined_table["OK"] * (combined_table["QK_1"] == 0)
-combined_table = combined_table[QC]
+QC = (combined_table["snr"] > 10) * (combined_table["r_chi_sq"] < 3) * (combined_table["QK"] == 0)
 
 N_bins = 50
 
 all_columns = [
     # RAVE label, DR4 label
-    ("TEFF", "TeffK_1"),
-    ("LOGG", "loggK_1"),
-    ("FE_H", "__M_H_K_1"),
+    ("TEFF", "TeffK"),
+    ("LOGG", "loggK"),
+    ("FE_H", "c_M_H_K"),
 ]
 label_limits = {
     "TEFF": (3000, 8000),
     "LOGG": (0, 5),
-    "FE_H": (-2.5, 0.75)
+    "FE_H": (-2.5, 0.5)
 }
 latex_labels = (r"$T_{\rm eff}$", r"$\log{g}$", r"$[{\rm Fe/H}]$")
 
@@ -59,26 +58,24 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
 
     cannon_label, dr4_label = columns
 
-    x = combined_table[cannon_label]._data
-    y = combined_table[dr4_label]._data
+    x = combined_table[cannon_label]._data[QC]
+    y = combined_table[dr4_label]._data[QC]
     
     finite = np.isfinite(x*y)
     limits = label_limits[cannon_label]
     bins = np.linspace(limits[0], limits[1], N_bins + 1)
-    ax.set_axis_bgcolor("#CCCCCC")
-    ax.hist2d(x[finite], y[finite], bins=(bins, bins), norm=LogNorm(),
-        cmap="viridis")
+    cmap = matplotlib.cm.Blues
+    ax.set_axis_bgcolor(cmap(0))
[…]

### 2016-09-04 article/figures/plot_ges_comparison.py

@@ -108,8 +108,6 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm unRAVE})$"]))
     ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm GES})$"]))
 
-    [_.set_rotation(30) for _ in ax.get_xticklabels()]
-    [_.set_rotation(30) for _ in ax.get_yticklabels()]
 
     diff = y - x
     print(label_name, np.nanmean(diff), np.nanstd(diff))

### 2016-09-04 article/figures/plot_gold_standard_comparisons.py

@@ -79,7 +79,7 @@ def reddy_2003_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "logg",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
     return get_data(reddy_2003, cannon_label, label_name)
@@ -90,7 +90,7 @@ def reddy_2006_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "logg",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
     return get_data(reddy_2006, cannon_label, label_name)
@@ -101,7 +101,7 @@ def valenti_2005_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "log_g_",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
     return get_data(valenti_2005, cannon_label, label_name)
@@ -177,11 +177,11 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
             diffs.extend(y-x)
 
             if ax == axes[0]:
-                handles.append(ax.scatter([-10000], [-10000], facecolor="k", **kwds))
+                handles.append(ax.scatter([-10000], [-10000], facecolor="#cccccc", **kwds))
 
         else:
             scat = ax.scatter(x, y, c=c0, cmap="plasma", **kwds)
-    print(label_name, np.nanmedian(diffs), np.nanstd(diffs))
+    print(label_name, np.nanmedian(diffs), np.nanstd(diffs), np.isfinite(diffs).sum())
 
     if ax.is_first_row():
         ax.set_xticklabels([])
@@ -232,7 +232,7 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
 
 
 fig.legend(handles, comparison_labels, loc="upper center", 
-    frameon=False, ncol=len(comparison_labels), numpoints=1)
+    frameon=False, ncol=len(comparison_labels), numpoints=1, scatterpoints=1)
 
 #divider = make_axes_locatable(axes[-1])
 #cax = divider.append_axes("right", size="5%", pad=0.05)

### 2016-09-04 article/figures/plot_gold_standard_hrd.py

@@ -10,7 +10,7 @@ from matplotlib.ticker import MaxNLocator
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 try:
-    bensby_2014, reddy_2003, reddy_2006, valenti_2005
+    bensby_2014, reddy_2003, reddy_2006, valenti_2005, rave_cannon_dr1
 
 except NameError: # Do you know who I am? That's Jeff Vader!
 
@@ -40,8 +40,8 @@ except NameError: # Do you know who I am? That's Jeff Vader!
     OK = (rave_cannon_dr1["SNRK"] > 10) * (rave_cannon_dr1["R"] > 25) * (rave_cannon_dr1["R_CHI_SQ"] < 3)
     rave_cannon_dr1 = rave_cannon_dr1[OK].filled()
     """
-
-    ok = (rave_cannon_dr1["snr"] > 50) * (rave_cannon_dr1["r_chi_sq"] < 3)
+    
+    ok = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3)
     rave_cannon_dr1 = rave_cannon_dr1[ok].filled()
 
     bensby_2014 = join(rave_cannon_dr1, get_literature_bensby(), keys=("Name", ))
@@ -79,7 +79,7 @@ def reddy_2003_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "logg",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
     return get_data(reddy_2003, cannon_label, label_name)
@@ -90,7 +90,7 @@ def reddy_2006_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "logg",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
     return get_data(reddy_2006, cannon_label, label_name)
@@ -101,7 +101,7 @@ def valenti_2005_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "log_g_",
-        "FE_H": "__Fe_H__2"
+        "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
[…]

### 2016-09-04 article/figures/plot_hrd_test.py

@@ -49,7 +49,7 @@ for i, snr in enumerate(snrs):
     # On top figure show density.
     # On bottom axes show metallicity.
     
-    ok = (t["SNRK"] > snr) * (t["R_CHI_SQ"] < 3) * (t["VSINI"] < 1)
+    ok = (t["snr"] > snr) * (t["r_chi_sq"] < 3) #* (t["VSINI"] < 1)
     
     ax = axes[0, i]
 

### 2016-09-04 article/figures/plot_joint_model_metrics.py

@@ -9,14 +9,12 @@ from scipy.spatial import Delaunay
 
 import AnniesLasso as tc
 
-output_filename, overwrite = \
-    ("/data/gaia-eso/arc/rave-data-files/unrave-v0.8-36-37-39.fits.gz", True)
 
 RESULTS_PATH = "/data/gaia-eso/arc/rave/results/"
-RESULTS_PATH = ""
+RESULTS_PATH = "../../"
 
 ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v37.fits.gz"))
-giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v36.fits.gz"))
+giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
 joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v41.fits.gz"))
 
 for t in (ms_results, giant_results, joint_results):
@@ -28,10 +26,10 @@ for t in (ms_results, giant_results, joint_results):
 assert np.all(ms_results["Name"] == joint_results["Name"])
 assert np.all(giant_results["Name"] == joint_results["Name"])
 
-ms_model = tc.load_model("rave-tgas-v37.model")
+ms_model = tc.load_model(os.path.join(RESULTS_PATH, "rave-tgas-v37.model"))
 
 # HACK MAGIC BEGINS #
-ms_model._labelled_set = Table.read("rave-tgas-v16b-labelled-set-cut.fits")
+ms_model._labelled_set = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v16b-labelled-set-cut.fits"))
 ms_model._labelled_set["TEFF"] = ms_model._labelled_set["EPIC_TEFF"]
 ms_model._labelled_set["LOGG"] = ms_model._labelled_set["EPIC_LOGG"]
 ms_model._labelled_set["FE_H"] = ms_model._labelled_set["EPIC_FEH"]
@@ -66,15 +64,15 @@ fig.subplots_adjust(
 extent = (3000, 8000, 0.5, 5.5)
 
 titles = (
+    r"${\rm Joint}$ ${\rm model}$",
     r"${\rm Main-sequence}$ ${\rm model}$", 
[…]

### 2016-09-04 article/figures/plot_kordopatis_calibration_sample.py

@@ -105,7 +105,7 @@ fig.tight_layout()
 
 cbar = plt.colorbar(_, 
     cax=fig.add_axes([0.93, fig.subplotpars.bottom, 0.02, fig.subplotpars.top - fig.subplotpars.bottom]))
-cbar.set_label(r"${\rm S/N}$ ${\rm unRAVE}$ $[{\rm pixel}^{-1}]$")
+cbar.set_label(r"${\rm S/N}$ ${\rm RAVE}$ $[{\rm pixel}^{-1}]$")
 
 
 fig.subplots_adjust(right=0.90)

### 2016-09-04 article/figures/rave_io.py

@@ -5,7 +5,7 @@ from astropy.table import Table
 from scipy.io import readsav
 
 DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
-
+DATA_PATH = ""
 
 def get_rave_kordopatis_dr4():
     return Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
@@ -32,7 +32,8 @@ def get_cannon_dr1(filename=None):
     #filename = "rave-tgas-v31.fits.gz"
     #filename = "rave-tgas-v29.fits.gz"
     #filename = "rave-tgas-v37.fits.gz"
-    filename = filename or "unrave-v0.7-37_36.fits.gz"
+    #filename = filename or "unrave-v0.7-37_36.fits.gz"
+    filename = filename or "unrave-v0.9.fits.gz"
 
     rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, filename))
 

### 2016-09-04 combine_model_results.py

@@ -14,13 +14,13 @@ from scipy.spatial import Delaunay
 import AnniesLasso as tc
 
 output_filename, overwrite = \
-    ("/data/gaia-eso/arc/rave-data-files/unrave-v0.8-36-37-39.fits.gz", True)
+    ("/data/gaia-eso/arc/rave-data-files/unrave-v0.9-36-42-41.fits.gz", True)
 
 RESULTS_PATH = "/data/gaia-eso/arc/rave/results/"
 RESULTS_PATH = ""
 
 ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v37.fits.gz"))
-giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v36.fits.gz"))
+giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
 joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v41.fits.gz"))
 
 for t in (ms_results, giant_results, joint_results):
@@ -234,8 +234,7 @@ ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], gridsize=N,
     cmap="Blues", norm=LogNorm(), edgecolor="#ffffff", linewidths=0.0)
 
 
-
-
+"""
 # Plot log(density) of the three models.
 K = 1
 factor = 3.5
@@ -273,6 +272,7 @@ fig.tight_layout()
 fig.savefig("article/figures/model-weights.pdf", dpi=300)
 fig.savefig("article/figures/model-weights.png", dpi=300)
 
+"""
 
 raise a
 

### 2016-09-04 article/figures/plot_joint_model_metrics.py
common color scheme for log density; bug fix in figure ref

@@ -132,7 +132,7 @@ y = ((ms_results["LOGG"] - joint_results["LOGG"]) - y_mu)/y_sigma
 ticks = (-10, -5, 0, 5, 10)
 
 kwds = dict(extent=(-10, 10, -10, 10), gridsize=25, linewidths=0.1, 
-    edgecolor="#000000", cmap="plasma", norm=LogNorm())
+    edgecolor="#000000", cmap="Blues", norm=LogNorm())
 
 axes[0].hexbin(x, y, **kwds)
 #axes[0].axhline(0, c="#FFFFFF", linewidth=0.5, linestyle="-")

### 2016-09-05 article/figures/plot_gce.py
Update abundance figures

@@ -19,21 +19,22 @@ except NameError: # Do you know who I am? That's Jeff Vader!
 
     from rave_io import get_cannon_dr1, get_rave_kordopatis_dr4
 
-    rave_cannon_dr1 = get_cannon_dr1("rave-tgas-v36.fits.gz")
-    rave_cannon_dr1["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in rave_cannon_dr1["FILENAME"]]
+    rave_cannon_dr1 = get_cannon_dr1()
 
-    from astropy.table import join
-    rave_cannon_dr1 = join(rave_cannon_dr1, get_rave_kordopatis_dr4(), keys=("Name", ))
+    #from astropy.table import join
+    #rave_cannon_dr1 = join(rave_cannon_dr1, get_rave_kordopatis_dr4(), keys=("Name", ))
 
 
 
 x = rave_cannon_dr1["FE_H"] 
 
 
-ok  = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3) \
-    * (rave_cannon_dr1["R"] > 10) * (rave_cannon_dr1["loggK"] < 3)
+#ok  = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3) \
+#    * (rave_cannon_dr1["R"] > 10) * (rave_cannon_dr1["loggK"] < 3)
 
-default_kwds = dict(gridsize=50,  cmap="viridis", norm=LogNorm())
+ok = (rave_cannon_dr1["snr"] > 10)
+
+default_kwds = dict(gridsize=50,  cmap="plasma", norm=LogNorm())
 elements_and_kwds = OrderedDict([
     ("O", dict(extent=(-1.5, 0.5, -0.05, 0.4))),
     ("AL", dict(extent=(-1.5, 0.5, -0.3, 0.5))),
@@ -63,12 +64,13 @@ for i, (element, updated_kwds) in enumerate(elements_and_kwds.items()):
     #    np.linspace(extent[2], extent[3], 13)
     #    ],
     #    **kwds)
-    cmap = plt.cm.viridis
+    cmap = plt.cm.Blues
     extent = kwds.pop("extent")
[…]

### 2016-09-05 article/figures/plot_ges_abundances.py

@@ -19,8 +19,9 @@ except NameError: # Do you know who I am? That's Jeff Vader!
 
     from rave_io import get_cannon_dr1, get_ges_idr4, get_rave_kordopatis_dr4
 
-    rave_cannon_dr1 = get_cannon_dr1("rave-tgas-v36.fits.gz")
-    rave_cannon_dr1["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in rave_cannon_dr1["FILENAME"]]
+    rave_cannon_dr1 = get_cannon_dr1()
+    assert "Name" in rave_cannon_dr1.dtype.names
+    #rave_cannon_dr1["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in rave_cannon_dr1["FILENAME"]]
 
     from astropy.table import join
     rave_cannon_dr1 = join(rave_cannon_dr1, get_rave_kordopatis_dr4(), keys=("Name", ))
@@ -28,8 +29,10 @@ except NameError: # Do you know who I am? That's Jeff Vader!
 
 
 # Produce error columns from COV matrix.
-for i, label_name in enumerate(rave_cannon_dr1.dtype.names[1:10]):
-    rave_cannon_dr1["E_{}".format(label_name)] = rave_cannon_dr1["COV"][:, i, i]**0.5
+if "COV" in rave_cannon_dr1.dtype.names:
+
+    for i, label_name in enumerate(rave_cannon_dr1.dtype.names[1:10]):
+        rave_cannon_dr1["E_{}".format(label_name)] = rave_cannon_dr1["COV"][:, i, i]**0.5
 
 
 asplund_2009_solar_abundances = {
@@ -44,9 +47,7 @@ asplund_2009_solar_abundances = {
 
 ges_species = ("O_1", "AL_1", "MG_1", "CA_1", "SI_1", "NI_1")
 
-ok  = (rave_cannon_dr1["snr"] > 0) * (rave_cannon_dr1["r_chi_sq"] < 3) \
-    * (rave_cannon_dr1["R"] > 10) * (rave_cannon_dr1["LOGG_2"] < 3.9)
-
+ok  = (rave_cannon_dr1["snr"] > 10)
 
 N, M = (3, 2)
[…]

### 2016-09-05 combine_model_results.py

@@ -9,7 +9,7 @@ import matplotlib.pyplot as plt
 from astropy.table import Table, vstack
 from matplotlib.colors import LogNorm
 from scipy.spatial import Delaunay
-
+from collections import OrderedDict
 
 import AnniesLasso as tc
 
@@ -161,7 +161,7 @@ axes[0].hexbin(x, y, gridsize=100, extent=(-3, +3, -3, +3), norm=LogNorm(), line
 
 
 # TODO WITHOUT THIS PENALISATION TERM THE GOLD STANDARD COMPARISON IS BAD.
-x_mu, y_mu = (0, 1.0)
+x_mu, y_mu = (0, 0)
 x_sigma, y_sigma = (50, 0.15)
 
 x2 = ((giant_results["TEFF"] - joint_results["TEFF"]) - x_mu)/x_sigma
@@ -177,11 +177,22 @@ giant_distance = np.sqrt(x2**2 + y2**2)
 
 
 # Weight by relative distance, or just take the closest of the two?
-combined_data = {
-    "Name": giant_results["Name"],
-    "snr": joint_results["snr"],
-    "r_chi_sq": joint_results["r_chi_sq"]
-}
+combined_data = OrderedDict([
+    ("Name", giant_results["Name"]),
+    ("TEFF", None),
+    ("LOGG", None),
+    ("FE_H", None),
+    ("O_H",  None),
+    ("MG_H", None),
+    ("AL_H", None),
+    ("SI_H", None),
+    ("CA_H", None),
+    ("NI_H", None),
+    #("COV", None),
+    ("snr", joint_results["snr"]),
+    ("r_chi_sq", joint_results["r_chi_sq"])
+])
+    
 
 #w1 = np.exp(-0.5 * ms_distance**2)
 #w2 = np.exp(-0.5 * (giant_distance**2))
@@ -216,6 +227,17 @@ for label_name in label_names:
     weights2[~np.isfinite(foo)] = 0
 
     combined_data[label_name] = np.nansum(foo, axis=0)/np.sum(weights2, axis=0)
+
+
+
+
+
+# Need abundances for things that are *probably* giants.
+is_giant = (weights2[1] > 0.5)
+
+for label_name in ("MG_H", "AL_H", "O_H", "NI_H", "SI_H", "CA_H"):
+    combined_data[label_name] = np.nan * np.ones(len(joint_results))
+    combined_data[label_name][is_giant] = giant_results[label_name][is_giant]
 
 
 
[…]

### 2016-09-05 article/unrave.tex
Remove color box for Traivs
common color scheme for log density; bug fix in figure ref
Update regularization figure

@@
 \usepackage{bm}
-\usepackage{xcolor}
-
-\definecolor{unoffensive-warning}{HTML}{B4DCED}
+
+%\usepackage{xcolor}
+%\definecolor{unoffensive-warning}{HTML}{B4DCED}
 
@@
 % ------------------------------------------
-%\usepackage{color}
+\usepackage{color}
 \usepackage{datenumber}
@@
 % TODO: Replace ZENODO-URL with a data url
-\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
+%\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
 \noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
@@
 tireless efforts of the entire \rave\ collaboration, past and present.
-}}
+%}}
 
@@
 \begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/set-giant-regularization.pdf}
-\caption{The fractional improvement in RMS deviation between inferred and training labels for the 9-label giant model at different regularization strengths.
+\includegraphics[width=\textwidth]{figures/set-hyperparameters.pdf}
+\caption{The percentage change in RMS deviation between inferred and training labels at different regularization strengths.
 The RMS values were calculated by leave-one-out cross-validation.
 The points and solid line indicate the mean improvement across all labels.
-The filled area represents the minimum and maximum improvements across all labels.
-The maximum improvement (decrease) in RMS label deviations occurs at $\Lambda \approx 0.3$, and this improvement persists until $\Lambda \approx 0.6$.\label{fig:set-hyperparameters}}
+The filled area represents the minimum and maximum improvements across all labels.\label{fig:set-hyperparameters}}
 \end{figure}
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/dr4-comparison.png}
+\includegraphics[width=\textwidth]{figures/dr4-comparison.pdf}
 \caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release \citep{Kordopatis_2013}.\label{fig:rave-dr4-comparison}}

### 2016-09-05 article/figures/plot_v34_search.py

@@ -13,6 +13,7 @@ Lambdas = [0, 0.001, 0.01, 0.1, 1.0, 10.0, 25.0, 50.0, 100.0]
 labels = ("TEFF", "LOGG", "FE_H", "MG_H", "AL_H", "O_H", "NI_H", "SI_H", "CA_H")
 
 
+import AnniesLasso as tc
 
 # y axis should be some metric(s)
 # x axis should be Lambda
@@ -20,27 +21,131 @@ labels = ("TEFF", "LOGG", "FE_H", "MG_H", "AL_H", "O_H", "NI_H", "SI_H", "CA_H")
 
 
 
-#labelled_set = Table.read("{}-labelled-set.fits".format(model_name))
-labelled_set = Table.read("/data/gaia-eso/arc/rave/preprocessing/giant_trainingsets_APOGEERAVE_V2.fits")
-
-ok = np.ones(len(labelled_set), dtype=bool)
-for label_name in labels:
-    ok *= np.isfinite(labelled_set[label_name])
-    ok *= (labelled_set[label_name] > -5)
-
-labelled_set = labelled_set[ok]
-labelled_set = np.array([labelled_set[label_name] for label_name in labels]).T
 
 
+#def metric(labelled_set, test_set):
+#    return np.nanstd(labelled_set - test_set, axis=0)
 
-def metric(labelled_set, test_set):
-    return np.nanstd(labelled_set - test_set, axis=0)
+
+comparisons = [
+    (
+        tc.load_model("rave-tgas-v34.model"), 
+        glob("rave-tgas-v34-test-labelled-set-?.????????e*-1.00.pkl"),
+        "rave-tgas-v34-test-labelled-set-0.00000000e+00-1.00.pkl",
+        "#3498DB",
+        r"${\rm Main{-}sequence}$ ${\rm model}$"
+    ),
+    (
+        tc.load_model("rave-tgas-v16b.model"),
+        glob("rave-tgas-v16b-test-labelled-set-*+vsini.pkl"),
+        'rave-tgas-v16b-test-labelled-set-1.00000000e-03+vsini.pkl',
+        "r",
+        r"${\rm Giant}$ ${\rm model}$"
+    )
+]
 
 
-color_choices = ["r", "b", "g", "m", "y"]
-color_choices = ["#3498DB"]
 
-files = glob("/data/gaia-eso/arc/rave/rave-tgas-v34-test-labelled-set-?.????????e*-1.00.pkl")
+# Plot log(density) of the three models.
+K = 1
+factor = 3.5
[…]

### 2016-09-05 article/unrave.tex
Include hyperaprameters figure
Update gold standard hrd to show links, indicating MP dwarf/subgiant problem (urgh :sad:), and update captions in all figures'

@@
 For abundance labels in the giant model that are not in the main-sequence
-model (e.g., [O/H], [Mg/H]), we only report abundances if there was no contribution
-from the main-sequence model: $w_{ms} = 0$.
+model (e.g., [O/H], [Mg/H]), we only report abundances for objects if
+$w_{ms} < 0.05$.
 
@@
 \includegraphics[width=\textwidth]{figures/hrd-train-set.pdf}
-\caption{A H-R diagram showing the training set labels.\label{fig:training-set-hrd}}
+\caption{Effective temperature $\teff$ and surface gravity $\logg$ for all stars in the training sets.
+Stars are colored by their metallicity [Fe/H], and the three panels show stars in the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/test-set-density.pdf}
-\caption{\label{fig:test-set-density}}
+\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
+These panels demonstrate how a single quadratic model is insufficient for all \rave\ stars (left panel), and illustrate some of the systematic artefacts that can result from testing on stars outside of the training set (center and right panel).
+These panels do not represent our final results, which are shown in Figure~\ref{fig:test-set-hrd}.\label{fig:test-set-density}}
 \end{figure*}
@@
 \begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/set-hyperparameters.pdf}
+\center
+\includegraphics[width=0.45\textwidth]{figures/set-hyperparameters.pdf}
 \caption{The percentage change in RMS deviation between inferred and training labels at different regularization strengths.
-The RMS values were calculated by leave-one-out cross-validation.
+The RMS values were calculated by leave-one-out cross-validation, and are shown with respect to an unregularized model ($\Lambda = 0$).
 The points and solid line indicate the mean improvement across all labels.
-The filled area represents the minimum and maximum improvements across all labels.\label{fig:set-hyperparameters}}
+The filled area represents the minimum and maximum improvements over all labels.
+With increasing regularization strength, there is a minimum in the RMS deviation over all labels, which is where we set $\Lambda$ for each model (see text for details).\label{fig:set-hyperparameters}}
 \end{figure}
@@
 \caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
-Only stars with $v\sin{i} < 1$~km~s$^{-1}$ and $\chi_{red} < 3$ are shown.
-The top three panels show logarithmic density, and bins in the lower three panels are colored by the median metallicity in each bin.\label{fig:test-set-hrd}}
+Only results with $\chi_{r}^2 < 3$ and S/N ratios greater than the quoted limit in each panel are shown.
+The top three panels show logarithmic density, and bins in the bottom three panels are colored by the median metallicity in each bin.\label{fig:test-set-hrd}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/dr4-comparison.pdf}
-\caption{Stellar label comparison between the \project{UNRAVE} catalog and the \project{RAVE} fourth data release \citep{Kordopatis_2013}.\label{fig:rave-dr4-comparison}}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth \rave\ data release \citep{Kordopatis_2013} and this work.
+Only stars meeting all the following quality criteria are shown: $\chi_{r}^2 < 3$ (\project{unRAVE} constraint); S/N $> 10$ (\project{unRAVE} constraint); and \texttt{QK = 0} (\rave\ DR4 constraint).
+\label{fig:rave-dr4-comparison}}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
-\caption{Stellar label comparison between the \project{unRAVE} catalog and the \ges\ fourth internal data release.\label{fig:ges-stellar-parameters}}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth internal data release from the \ges\ survey, and this work.
+The number of stars in each panel are shown, as well as the bias and RMS deviation in each label.
+Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+Most of the \ges-\rave\ overlap stars have relatively low S/N ratios in \rave, near $\approx 30$.\label{fig:ges-stellar-parameters}}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{figures/ges-abundances.pdf}
-\caption{Individual chemical abundances in \project{unRAVE} compared to \ges.\label{fig:ges-abundances}}
+\caption{Detailed chemical abundances in the fourth internal data release from the \ges\ survey compared to this work.
+The number of stars shown in each panel is indicated, and the bias and RMS deviations are shown.
+Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.\label{fig:ges-abundances}}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
-\caption{Stellar parameter comparisons for stars in common between this work and studies using high-resolution, high S/N spectra and \hipparcos\ parallaxes.\label{fig:gold-standard-comparison}}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
+Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+\label{fig:gold-standard-comparison}}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
-\caption{HRD of gold standard comparisons.\label{fig:gold-standard-hrd}}
+\caption{Hertsprung-Russell diagrams of stars in common between this work and that of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
+Stars are colored by the metallicity of each study.
+Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
+This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
 \end{figure*}
@@
 \includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
-\caption{Comparison with the calibration (literature) samples used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.\label{fig:kordopatis-calibration}}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison with the literature calibration sources used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.
+Stars are colored by the S/N of the \rave\ spectra.
+Note that this comparison is for illustrative purposes only: it is not an indication of independent agreement with the literature because some metal-poor stars in this literature sample were used in the construction of our training set (see text for details).\label{fig:kordopatis-calibration}}
 \end{figure*}
@@
 \begin{figure}[p]
-\includegraphics[width=0.5\textwidth]{figures/gce.pdf}
-\caption{Giant star abundances ([X/Fe]) with respect to [Fe/H], showing the Galactic chemical evolution trend for individual elements.\label{fig:gce}}
+\center
+\includegraphics[width=0.65\textwidth]{figures/gce.pdf}
+\caption{Detailed chemical abundances ([X/Fe]) for giant stars in \unrave\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
+Note that the y-axes limits vary for each panel, however for clarity we show the scaled-Solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
+Only stars with S/N $> 10$ and $\chi_r^2 < 3$ are shown.\label{fig:gce}}
 \end{figure}

### 2016-09-05 article/figures/plot_gold_standard_hrd.py

@@ -158,7 +158,7 @@ comparison_labels = (
     )
 M = len(comparison_labels)
 
-link = False
+link = True
 
 factor = 3.5
 lbdim = 0.2 * factor

### 2016-09-05 article/figures/plot_hrd_train.py
Updated figure #1

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 from matplotlib.colors import LogNorm
 from matplotlib.ticker import MaxNLocator
 
-
+"""
 try:
     ms_training_set, giant_training_set, joint_training_set
 
@@ -17,14 +17,23 @@ except NameError:
 else:
     print("Using pre-loaded data")
 
-
+"""
 names = (
-    r"${\rm Main-sequence}$ ${\rm star}$ ${\rm model}$", 
+    r"${\rm Joint}$ ${\rm star}$ ${\rm model}$",
+    r"${\rm Main{-}sequence}$ ${\rm star}$ ${\rm model}$", 
     r"${\rm Giant}$ ${\rm star}$ ${\rm model}$", 
-    r"${\rm Joint}$ ${\rm star}$ ${\rm model}$"
 )
 
-labelled_sets = (ms_training_set, giant_training_set, joint_training_set)
+import AnniesLasso as tc
+labelled_sets = (
+    tc.load_model("rave-tgas-v41.model").labelled_set, # joint
+    tc.load_model("rave-tgas-v16b.model").labelled_set, # ms result
+    tc.load_model("rave-tgas-v42.model").labelled_set, # giant result
+)
+labelled_sets[1]["TEFF"] = labelled_sets[1]["EPIC_TEFF"]
+labelled_sets[1]["LOGG"] = labelled_sets[1]["EPIC_LOGG"]
+labelled_sets[1]["FE_H"] = labelled_sets[1]["EPIC_FEH"]
+
 M = len(labelled_sets)
 
 factor = 3.5
@@ -43,7 +52,7 @@ fig.subplots_adjust(
     top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
 
 
-xlim = (7000, 3500)
+xlim = (7500, 3500)
 ylim = (5.5, 0)
 
 vmin, vmax = (-3, 0.5)
@@ -53,9 +62,9 @@ vmin, vmax = (-3, 0.5)
 
 for ax, labelled_set, name in zip(axes, labelled_sets, names):
     
-    scat = ax.scatter(labelled_set["TEFF"], labelled_set["LOGG"],
-        c=labelled_set["FE_H"], cmap="magma",
-        vmin=vmin, vmax=vmax, alpha=0.75, s=50)
+    ax.scatter(labelled_set["TEFF"], labelled_set["LOGG"],
[…]

### 2016-09-05 article/figures/plot_hrd_train.py
Updated some figures (minor things)

@@ -1,6 +1,7 @@
 
 
 import numpy as np
+import matplotlib as mpl
 import matplotlib.pyplot as plt
 from matplotlib.colors import LogNorm
 from matplotlib.ticker import MaxNLocator
@@ -19,9 +20,9 @@ else:
 
 """
 names = (
-    r"${\rm Joint}$ ${\rm star}$ ${\rm model}$",
-    r"${\rm Main{-}sequence}$ ${\rm star}$ ${\rm model}$", 
-    r"${\rm Giant}$ ${\rm star}$ ${\rm model}$", 
+    r"${\rm Joint}$ ${\rm model}$",
+    r"${\rm Main{-}sequence}$ ${\rm model}$", 
+    r"${\rm Giant}$ ${\rm model}$", 
 )
 
 import AnniesLasso as tc
@@ -97,7 +98,6 @@ for ax in axes:
     ax.set_ylim(ylim)
 
 
-import matplotlib
 cax, kw = mpl.colorbar.make_axes(list(axes), fraction=0.075, pad=0.025, aspect=10)
 cbar = plt.colorbar(scat, cax=cax, ticks=[-3, -2, -1, 0])
 

### 2016-09-05 article/figures/plot_joint_model_metrics.py

@@ -65,7 +65,7 @@ extent = (3000, 8000, 0.5, 5.5)
 
 titles = (
     r"${\rm Joint}$ ${\rm model}$",
-    r"${\rm Main-sequence}$ ${\rm model}$", 
+    r"${\rm Main{-}sequence}$ ${\rm model}$", 
     r"${\rm Giant}$ ${\rm branch}$ ${\rm model}$",
 )
 all_results = (joint_results, ms_results, giant_results)

### 2016-09-05 article/unrave.tex
Write about gold standard comparisons

@@
 abundances for vast sets of individual stars.  These inferences require 
-both astrometric and spectroscopic data.  Indeed, the Tycho-Gaia 
-Astrometric Solution (\tgas) sample in the first \gaia\ data release 
-requires a spectroscopic complement that includes radial velocities, 
-stellar parameters, and elemental abundances.  Among existing spectroscopic 
-samples, the RAdial Velocity Experiment (\rave) survey has the largest 
-overlap with \tgas: 292,036 stars.  Here we present a data-driven 
-re-analysis of \rave\ spectra, using an implementation of \thecannon, that
-yields more precise and accurate stellar parameters and abundances.  We 
-derive, and validate, improved effective temperature $\teff$, surface 
-gravity $\logg$, and chemical abundances of up to seven elements (O, Mg, 
-Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in chemical 
-abundances is X.XX~dex for stars on the giant branch.  We provide 
-spectrophotometric distances based on our stellar parameters, which should 
-be superior to the parallax-based distances for faint or distant objects in
-\tgas.  The \unrave\ catalog presented here, in conjunction with \tgas, will 
+both astrometric and spectroscopic data.  Indeed, in order to be maximally
+useful for chemo-dynamic studies, the Tycho-Gaia Astrometric Solution 
+(\tgas) sample in the first \gaia\ data release requires a spectroscopic 
+complement that includes radial velocities, stellar parameters, and 
+elemental abundances.  Among existing spectroscopic samples, the RAdial 
+Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
+292,036 stars.  Here we present a data-driven re-analysis of \rave\ 
+spectra, using an implementation of \thecannon, that yields more precise 
+and accurate stellar parameters and abundances.  We derive, and validate,
+improved effective temperature $\teff$, surface gravity $\logg$, and 
+chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
+for \Nstarsqc\ stars.  The typical precision in chemical abundances is 
+X.XX~dex for stars on the giant branch.  
+The \unrave\ catalog presented here, in conjunction with \tgas, will 
 be the most powerful data set for chemo-orbital analysis at the dawn of 
@@
 \end{abstract}
+
+%We *will* provide 
+%spectrophotometric distances based on our stellar parameters, which should 
+%be superior to the parallax-based distances for faint or distant objects in
+%\tgas.
 
@@
 
-\subsubsection{Comparison with \rave\ DR5}
-
-Here we cross-match our results against literature samples to verify that
-our results are accurate.  Because this work is an independent analysis, our
-first point of reference is against the official \rave\ data release.  This
-comparison is shown in Figure \ref{fig:rave-dr5-comparison}, 
-where we show our results with respect to the fifth \rave\ data release \citep{Kunder_2016}.  In order
-to provide a fair comparison, we only show the $XXX$ stars that meet a number of quality
-flags in both samples: 
-\stub{What selections qualify as a fair comparison in both samples?}
-There is very good agreement in $\teff$, with a bias
-and RMS of just XX~K and XXX~K, respectively.  We find our $\logg$ values are
-closest to the `calibrated' \rave\ surface gravities, which make use of 
-asteroseismic information to correct (increase) the $\logg$ values along the
-giant branch.  The calibrated \rave\ surface gravities also correct the 
-location of the red clump (by $\sim0.5$~dex) to be in the same position in our
-results.  However, while our $\logg$ values generally agree with the calibrated
-\rave\ values along the giant branch, our $\logg$ values do differ slightly 
-along the main-sequence.  This is principally because the main-sequence in 
-this work tends to taper down towards higher $\logg$ values at cooler 
-temperatures, whereas the comparison sample tend to have a slightly flatter
-main-sequence.  This difference is not likely to have a very significant 
-effect on the detailed abundance or spectrophotometric distance determinations
-between these studies \citep{Binney_2014}.
-
-\subsubsection{Comparison with the \citet{Kordopatis_2013} calibration sample}
+\subsubsection{Comparison with \rave\ DR4}
 \label{sec:validation-kordopatis}
+
+As an initial point of comparison, we cross-matched our results against the 
+official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
+In order to provide a fair comparison, we only show stars that meet a number
+of quality flags in \emph{both} samples. Specifically we require that the 
+$S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
+$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag reported by
+\citet{Kordopatis_2013} is zero, indicating no problems with quality control.
+There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
+and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
+\citet{Kordopatis_2013} has been noted in other studies, and this has issue
+has been resolved in the fifth \rave\ data release by correcting $\logg$
+values with a small sample of asteroseismic targets.  There is also a
+slight discrepancy in the $\log$ values along the main-sequence, where our
+work tends to taper down towards higher $\logg$ values at cooler temperatures,
+and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
+This difference is not likely to have a very significant effect on the detailed
+abundance or spectrophotometric distance determinations between these studies 
+\citep{Binney_2014}.
 
@@
 
-\stub{Caution re MP dwarfs appear as sub-giants in our sample}
+We searched the literature for studies that overlap with \rave, and which base
+their analysis on high-resolution, high S/N spectra.  We found four notable
+studies with a sufficient level of overlap: the disk studies \citet{Reddy_2003}
+and \citet{Reddy_2006}, the \citet{Valenti_Fischer_2005} work on exoplanet
+host star candidates, and the work of \citet{Bensby_2014}.  These studies 
+perform a careful (manual; expert) analysis using extremely high-resolution
+\stub{how high?}, high S/N spectra, and make use of \project{Hipparcos}
+parallaxes where possible.  Most of the stars in these samples are main-sequence
+or sub-giant stars.  Therefore, these works constitute an excellent comparison
+to evaluate the accuracy of our results on the main-sequence and sub-giant branch.
+
+We compare the stellar parameters ($\teff$, $\logg$, [Fe/H]) of stars that
+overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
+We only include stars with $\chi_r^2 < 3$ and $S/N > 10$, although the latter
+cut removed only few stars because the average S/N in the \rave\ spectra for
+these stars is relatively high ($\gtrsim{}50$). When considering the relative 
+information content available in \rave\ (945~pixels in the near infrared with
+$\mathcal{R} ~ 7{,}500$) compared to these literature studies (\stub{what snr, resolution, wavelength range and hipparcos parallaxes}, we consider the
+ agreement to be satisfactory.
+
+There are some discrepancies between our results and these literature studies,
+and these differences arise from the manner in which we join results from the
+main-sequence and giant model.  Specifically, for metal-poor stars near
+the turn-off in these literature comparisons, in our weighted results we find 
+these stars to have similar metallicities, but to be slightly more evolved: 
+on the sub-giant branch or the base of the giant branch.  This effect is best
+illustrated in Figure~\ref{fig:gold-standard-hrd}, where we show a 
+Hertzsprung-Russell diagram for all stars in common between this work and that
+of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.  While the
+overall shape of the turn-off in our work is consistent with the literature
+studies, the linked data points show that metal-poor stars in the literature
+appear as more evolved stars in our combined results.  Instead, if we only
+take the results from our main-sequence model --- ignoring the influence of the
+giant model --- and compare those values to these literature studies, no such 
+systematic discrepancies are present in the Hertzsprung-Russell diagram: we 
+find those metal-poor turn-off stars to also be turn-off stars, with comparable
+metallicities.  Our investigation indicates that this issue affects dwarf stars
+near the turn-off that have $[{\rm Fe/H}] \lesssim -1$.  While the agreement
+would improve if we adopted the main-sequence results for these specific stars,
+the underlying issue of metal-poor dwarf stars would still be present.  Thus,
+we use this example to illustrate caution for studies making inferences based
+on main-sequence dwarf stars.  The results from both the main-sequence star model, 
+and the giant star model, are available at \stub{should we put them online somewhere
+for open-science, or just make them available upon request so that we don't
+muddy the waters and confuse people as to which catalog to use?}
 

### 2016-09-05 article/unrave.tex
typo in Wojno, lay out groundwork for discussion

@@
 The target selection for \rave\ is based on the $I$-band apparent magnitude,
-$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
+$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wojno_2016}.  
 The $I$ band was used for the target selection because it has a good overlap with the
@@
 overall shape of the turn-off in our work is consistent with the literature
-studies, the linked data points show that metal-poor stars in the literature
+studies (right panel), the linked data points in the first three panels show that metal-poor stars in the literature
 appear as more evolved stars in our combined results.  Instead, if we only
@@
 
+We note that this issue does not appear to affect our derived metallicities for these
+stars.  For this reason, if stars affected by this issue can be confirmed to be
+dwarfs by external information --- rather than being sub-giants as our analysis
+suggests for some --- then the metallicity we report can still be considered
+useful.  For the more metal-rich dwarf stars, the sample we present here
+will undoubtedly be informative for upcoming studies based on exoplanet host
+star properties.  At present, however, there are just $\approx$30 stars in 
+\rave\ that overlap with the compilations of exoplanet host star properties 
+listed at \texttt{exoplanets.org} and \texttt{exoplanets.eu}.
+
 \subsubsection{Comparison with the \ges\ survey}
@@
 
-\stub{Chemistry + Kinematics = Stockholm}
-
-\stub{What do the two provide us with?}
-
-\stub{What is one interesting science result should we show based on these abundances and Hipparcos results, which was not known previously?}
+% What information have we extracted: almost maximal use of giant branch,
+% lesss so for dwarfs.
+
+% Caveats in the results: metal-poor dwarfs/sub-giants
+% HETEROGENEOUS TRAINING SET
+% UNSURE IF TWO ABUNDANCE SCALES ARE ON THE SAME SCALE --!
+% METAL POOR ABUNDANCES
+% recommended SNR and chi-sq etc
+% weird projections in space. --> use morphological classifications to guide
+
+% despite these limitations, this catalog is the largest??? of abundances.
+
+% in the future we will do spectrophotometric distances to improve on this
 
@@
 
-\bibitem[Wonjo et al.(2016)]{Wonjo_2016} Wonjo, J., et al.\ 2016, submitted
+\bibitem[Wojno et al.(2016)]{Wojno_2016} Wojno, J., et al.\ 2016, submitted
 

### 2016-09-05 article/unrave.tex
Include text about GES comparison

@@
 relatively low S/N ratios in \rave\ ($\approx 25$), there is good 
-agreement in stellar parameters (Figure~\ref{fig:ges-comparison}.  
+agreement in stellar parameters (Figure~\ref{fig:ges-stellar-parameters}.  
 The RMS in effective temperature, surface gravity and metallicity is
-236~K, 0.39~dex, and 0.18~dex, respectively. 
-% TODO: Check if those numbers above are still valid when final catalog is used.
-
-\stub{A paragraph about the abundance agreement.}
+233~K, 0.37~dex, and 0.17~dex, respectively.
+
+Based on this comparison, we find no evidence for a systematic offset 
+between stars on the main-sequence and those on the giant branch. Specifically,
+we do not find any dwarf stars that are misclassified as sub-giants, as we
+found for some metal-poor stars in Section~\ref{sec:validation-gold-standards}.
+In addition we find no evidence for a systematic offset in metallicity (or
+individual abundances) between giant stars and main-sequence stars.  This is
+a crucial inference, as the metallicities for stars in our main-sequence
+training set have a principally different source to those on the giant branch.
+We cannot make these same inferences based on other surveys, like \apogee,
+because (1) \apogee\ stars formed part of the training set, and (2) they do
+not include main-sequence stars.  Even if we found good agreement between
+\epic\ and \apogee\ metallicities, this would not be informative, because
+\apogee\ is the source of metallicity for many stars on the giant branch in
+the \epic\ sample. Therefore, although this is a qualitative comparison only,
+it is reassuring that there is no obvious systematic difference between the
+metallicities of main-sequence and giant branch stars.
+
+The metallicity agreement between this work and \ges\ extends down to low
+metallicity, near $[{\rm Fe/H}] \approx -1.5$.  The scatter increases for
+the four stars in the overlap sample with [Fe/H] $< -1$, however these
+spectra have low S/N ratios, which is reflected in the larger uncertainties
+reported for these metallicities.
+% TODO: Check if that is still the case after we examine errors from repeat
+% 		obs.
+
+The fourth internal data release of the \ges\ includes detailed chemical
+abundances of up to 45~species ($\approx32$ elements at different ionization
+stages).  This provides us with an independent validation for our detailed
+abundances on the giant branch.  These comparisons are shown for six elemental
+abundances in Figure~\ref{fig:ges-abundances}, where markers are colored by the
+S/N of the \rave\ spectrum.  The number of stars available in each abundance
+comparison varies due to what is available in the \ges\ data release, which
+is itself a function of the instrument used, the spectral type, and other
+factors.  The absolute bias for individual elements varies from as low as
+0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
+[Si/H] abundances relative to the \ges\ survey. The RMS deviation in each
+label is small for stars with [X/H] $\sim -0.5$, before increasing at lower
+metallicities.  If we consider all stars, the smallest abundance RMS we see
+with respect to \ges\ is 0.18~dex for [Ca/H].  The increasing RMS at low 
+metallicity is likely a consequence of multiple factors, namely: inaccurate
+abundance labels for metal-poor stars (Section~\ref{sec:the-training-set});
+only weak, blended lines being available in \rave, which cease to be visible
+in hot and/or metal-poor stars; and to a lesser extent, low S/N ratios for 
+those particular stars being compared.  Unfortunately not all of these
+factors are represented by the quoted errors in each label.  For these reasons, 
+although it affects only a small number of stars, we recommend caution when 
+using individual abundances for very metal-poor giant stars in our sample.
 
@@
 	Sven Buder (MPIA), 
+	Sergey Koposov (Cambridge),
 and 
-	Sergey Koposov (Cambridge).
+	Sofia Randich (INAF).
 This research made use of: 

### 2016-09-05 article/unrave.tex
note which column we used in dr4 comparison caption
Include stub for dibs and ocs/gcs

@@
 using individual abundances for very metal-poor giant stars in our sample.
+% TODO: Revise this last sentence after we see the globular cluster comparisons,
+%		because we are saying 'hey be careful' then perhaps saying 'look how good
+%		those same abundances are in GCs`
 
@@
 \label{sec:astrophysical-validation}
+
+\subsubsection{Globular clusters}
+\label{sec:globular-cluster-validation}
 
@@
 we do not measure C and N abundances here.  We note that \emph{none} of these 
-globular cluster stars are present in our training set.
+globular cluster stars were present in our training set.
 
@@
 
-% Astrophysical validation:
-% --> Residuals as a function of galactic position? (DIBS)
+\subsubsection{Open clusters}
+\label{sec:open-cluster-validation}
+
+\subsubsection{Diffuse interstellar bands}
+\label{sec:dibs-validation}
+
+\stub{Look at residuals between model and data as a function of galactic position. This might be out of scope unless someone volunteers to do it}
 
@@
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth \rave\ data release \citep{Kordopatis_2013} and this work.
+Here we show the `calibrated' metallicity (column \texttt{c\_M\_H\_K}) from the \rave\ survey.
 Only stars meeting all the following quality criteria are shown: $\chi_{r}^2 < 3$ (\project{unRAVE} constraint); S/N $> 10$ (\project{unRAVE} constraint); and \texttt{QK = 0} (\rave\ DR4 constraint).

### 2016-09-05 article/unrave.tex
Minor edits before attacking discussion

@@
 at which systematic effects begin to dominate.  In other words, in the low S/N
-regime, a well-constructed data-driven model will yield more precise labels 
+regime, a well-constructed data-driven model will yield more precise \emph{labels}
 (e.g., stellar parameters and chemical abundances) than most physics-driven 
@@
 for the low-resolution \rave\ data analysed here, because about half of the 
-spectra have S/N $\lesssim 50$.
+spectra have S/N $\lesssim 50$~pixel$^{-1}$.
 
@@
 is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
-or [Fe/H]), and invariant with respect to the S/N ratio.
+or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
+that the normalization procedure does not produce different results for high S/N
+spectra compared to low S/N spectra, nor does the normalization procedure vary 
+non-linearly with respect to stellar parameters (e.g., [Fe/H]).
 \item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
@@
 wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
-vectorizing functions with polynomial expansions (e.g., $\teff^2$, see Sections 
+vectorizing functions with second-order polynomial expansions (e.g., $\teff^2$, see Sections 
 \ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
@@
 
-At the \emph{training} step we fix the $K$-lists of labels for the $n$ training set.
+At the \emph{training} step we fix the $K$-lists of labels for the $n$ training set stars.
 At each wavelength pixel $j$, we then find the parameters $\vectheta_j$ and $s_j^2$
@@
 \noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
-in later sections, and $Q(\vectheta)$ is a regularizing function that encourages 
+in later sections, and $Q(\vectheta)$ is a L1 regularizing function that encourages 
 $\vectheta$ values to take on zero values without breaking convexity:
@@
 The \emph{test step} is where we fix the parameters $\vectheta_j,s_j^2$ at all wavelength
-pixels $j$, and optimize the $K$-list of labels $\ell_m$ for each test set star $m$.  Here
+pixels $j$, and optimize the $K$-list of labels $\ell_m$ for the $m$-th test set star.  Here
 the objective function is:
@@
 We sought to construct a training set of stars across the main-sequence, the
-sub-giant branch, and the red giant branch.  We demanded stars with precisely measured
+sub-giant branch, and the red giant branch.  We required stars with precisely measured
 effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
@@
 overlap sample for giant star labels in the training set.  Of these, about 900 are 
-giants according to \apogee.  From this sample we selected stars to have: metallicity
-determinations ([Fe/H] $> -5$); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
+giants according to \apogee.  From this sample we selected stars to have: 
+determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, and Ni); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
 we further required that \aspcap\ did not report any peculiar flags 
@@
 trivial.  There are no spectroscopic studies that extend the range of stellar types we 
-are interested in (e.g., FGKM-type stars), and also have a large enough sample size that
-overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
+are interested in (e.g., FGKM-type stars), and which have a large enough sample size 
+overlapping with \rave.  Moreover, most of the spectroscopic studies we considered also 
 showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
@@
 probability of misclassification is usually quantified in the uncertainties given
-for each star; most dwarfs that have a higher likelihood of being sub-giants have
+for each star; most dwarfs that have a higher possibility of being sub-giants have
 large confidence intervals.  Therefore, requiring low uncertainties will decrease 
@@
 was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
-stars (of 4,611).
+stars (of 4,611).  The full training set for each model (see next sections) is shown
+in Figure~\ref{fig:training-set-hrd}.
 
@@
 treated all \Nspectra\ spectra as test set objects.  In the right-hand panel of Figure 
-\ref{fig:tes-set-density} we show the effective temperature $\teff$ and surface gravity 
+\ref{fig:test-set-density} we show the effective temperature $\teff$ and surface gravity 
 $\logg$ for all spectra.  The main-sequence and red giant branch are clearly visible.  
@@
 We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
-The results for the survey sample are shown in the left panel of Figure \ref{fig:test-set-density}.  
+The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
 The increased density of Solar-type stars is consistent with \rave\
@@
 the relative lifetimes of these evolutionary phases.  An over-density
-of stars near the giant branch clump is also present.  This artefact is due 
+of stars near the base of the giant branch is also present.  This artefact is due 
 to having giant stars in the test set, but not in the training set, and the
@@
 deviation with respect to the unregularized case.  At the point of maximum improvement 
-near $\Lambda = X.XX$, the RMS in all nine labels has decreased by at least $X$\%, 
-and up to $X$\%.  Here the regularization strength also produces a sparser matrix
+near $\Lambda = X.XX$, the RMS in all nine labels has decreased by up to $X$\%.  Here the regularization strength also produces a sparser matrix
 of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
@@
 \Nspectra\ \rave\ spectra using this model, and these results are summarized in
-Figure \ref{fig:test-set-density} (middle panel).  The red clump is clearly visible
+the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible
 and in the expected location, without requiring any post-analysis calibration.
-Artefacts due to dwarf stars in the test set are also present.
+However, artefacts due to dwarf stars in the test set are also present.
 
@@
 described in previous sections.  The results from the joint model --- that includes 
-the main-sequence, sub-giant and red giant branch --- shows that a single quadratic
+the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
 model is too simple for the \rave\ spectral range.  The other models have problems,
@@
 the joint model and the comparison model both report similar labels.  The artefacts 
-seen in the Hertszprung-Russell diagrams in Figure \ref{fig:test-set-density} are
+seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
 also present in Figure \ref{fig:joint-model-differences} as over-densities far away
@@
 mean and unit variance.
-\stub{Are they? (Probably). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
+\stub{Are they? (Probably not!). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
 we show these inflated errors only, as we gauge them to be more representative.}
@@
 $S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
-$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag reported by
-\citet{Kordopatis_2013} is zero, indicating no problems with quality control.
+$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag from
+\citet{Kordopatis_2013} is zero, indicating that no problems were reported by the pipeline.
 There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
@@
 overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
-We only include stars with $\chi_r^2 < 3$ and $S/N > 10$, although the latter
+We only include stars with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter
 cut removed only few stars because the average S/N in the \rave\ spectra for
-these stars is relatively high ($\gtrsim{}50$). When considering the relative 
+these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$). When considering the relative 
 information content available in \rave\ (945~pixels in the near infrared with
-$\mathcal{R} ~ 7{,}500$) compared to these literature studies (\stub{what snr, resolution, wavelength range and hipparcos parallaxes}, we consider the
+$\mathcal{R} \approx 7{,}500$) compared to these literature studies (\stub{what snr, resolution, wavelength range and hipparcos parallaxes}, we consider the
  agreement to be satisfactory.
@@
 the underlying issue of metal-poor dwarf stars would still be present.  Thus,
-we use this example to illustrate caution for studies making inferences based
-on main-sequence dwarf stars.  The results from both the main-sequence star model, 
+we use this example to illustrate caution for studies making inferences from our
+work based on main-sequence dwarf stars.  The results from both the main-sequence star model, 
 and the giant star model, are available at \stub{should we put them online somewhere
@@
 sub-giant and red giant branch stars.  Despite most of these stars having
-relatively low S/N ratios in \rave\ ($\approx 25$), there is good 
-agreement in stellar parameters (Figure~\ref{fig:ges-stellar-parameters}.  
+relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
+agreement in stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
 The RMS in effective temperature, surface gravity and metallicity is
@@
 individual abundances) between giant stars and main-sequence stars.  This is
-a crucial inference, as the metallicities for stars in our main-sequence
+a crucial observation, as the metallicities for stars in our main-sequence
 training set have a principally different source to those on the giant branch.
@@
 
-% What information have we extracted: almost maximal use of giant branch,
-% lesss so for dwarfs.
-
-% Caveats in the results: metal-poor dwarfs/sub-giants
+% What we have presented, an independent analysis that makes maximal use of giant branch
+% less-so for dwarfs
+% the intent is for usefulness with chemo-dynamic studies.
+
+% Caveats in the results:
+% MULTIPLE MODELS
 % HETEROGENEOUS TRAINING SET
+% METAL POOR ABUNDANCES
+
 % UNSURE IF TWO ABUNDANCE SCALES ARE ON THE SAME SCALE --!
-% METAL POOR ABUNDANCES
+% even if they are on the same scale...
+% metal-poor dwarfs/sub-giants
+
+% quality constraints recmomended
 % recommended SNR and chi-sq etc
@@
 % despite these limitations, this catalog is the largest??? of abundances.
-
 % in the future we will do spectrophotometric distances to improve on this
@@
 
+% Cross-validation, and
+% Validate through literature comparisons
+
 \stub{Typical precision}
-
-\stub{A key science result that we see already}
-
 \stub{Something about exoplanet host stars, TESS, etc}
@@
 This work was partly supported by the European Union FP7 programme through ERC 
-grant number 320360. Funding for K.~H.~ has been provided through the Simons 
+grant number 320360. A.~R.~C. thanks King's College of the University of Cambridge
+for their support.  Funding for K.~H.~ has been provided through the Simons 
 Foundation Society of Fellows and the Marshall Scholarship.
@@
 Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
-Most of the \ges-\rave\ overlap stars have relatively low S/N ratios in \rave, near $\approx 30$.\label{fig:ges-stellar-parameters}}
+Most of the \ges/\rave\ overlap stars have relatively low S/N ratios in \rave, near $\approx 30$~pixel$^{-1}$.\label{fig:ges-stellar-parameters}}
 \end{figure*}

### 2016-09-05 article/unrave.tex
some discussion added about how bad we are

@@
 
-% What we have presented, an independent analysis that makes maximal use of giant branch
-% less-so for dwarfs
-% the intent is for usefulness with chemo-dynamic studies.
+We have presented an independent re-analysis of \rave\ spectra that makes near-maximal
+use of the information content for red giant branch star, and provides significant a
+improvement in stellar parameter ($\teff$, $\logg$, [Fe/H]) precision over previous
+\rave\ data releases.  When combined with \gaia/\tgas, these results provide a powerful
+compendium for chemo-dynamic studies of the Milky Way.  However, our results have caveats.
+Any inferences based upon these results should be made with the recognition of those caveats,
+and acknowledge that they are subject to our explicit assumptions.
+
+For practical purposes we adopted separate models: one for the giant branch and one
+for the main-sequence.  A third model was used to derive relative weights for which
+results to use.  The relative weighting we have used does not have any formal
+interpretation as a likelihood or belief (in any sense): it was introduced for
+practical reasons to identify systematic errors and combine results for multiple
+models.  Because the relative weights have no formal interpretation, it is reasonable
+to consider this method is as \emph{ad hoc} as any other approach.  Therefore, the
+relative weighting has no warranty to be (formally) correct, and may introduce 
+inconsistencies or systematic errors rather than
+minimizing them.  Evidence of this is already present in our comparisons to
+`gold standard' works \citep{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005},
+where metal-poor dwarf stars appear as metal-poor sub-giants in our combined results,
+but have reasonable labels (by any available statistic) as dwarfs in the main-sequence
+model.  Clearly, the relative distance from the joint model to the giant model is 
+slightly favoured (over the main-sequence model) for metal-poor dwarf stars. In this
+sense, our approach towards combining labels from multiple models is sub-optimal.
+
+If we only consider the results from individual models, there are a number of cautionary
+remarks that stem from the construction of the training set.  The labels for red giant 
+branch stars primarily come from \apogee, where previous successes with \thecannon\
+have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity.
+However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
+a tapering-off in the test set, where \emph{no} stars had reliably reported metallicities
+below that of the training set, which forced us to construct a heterogeneous training
+set.  The metal-poor stars included in this sample are from high-resolution studies
+\citep{Fulbright,Ruchti,elsewhere}, but it is unknown if the stellar parameters are
+of high fidelity (e.g., we have limited quality metrics available), and there is no
+guarantee that the stellar parameters \emph{or} abundances are on the same scale as
+\apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
 

### 2016-09-05 article/unrave.tex
minor edits to abstract/intro

--- 2016-09-05 article/unrave.tex
@@
 spectra, using an implementation of \thecannon, that yields more precise 
-and accurate stellar parameters and abundances.  We derive, and validate,
+and accurate stellar parameters and abundances.  We derive, and validate
 improved effective temperature $\teff$, surface gravity $\logg$, and 
@@
 X.XX~dex for stars on the giant branch.  
-The \unrave\ catalog presented here, in conjunction with \tgas, will 
-be the most powerful data set for chemo-orbital analysis at the dawn of 
+The \unrave\ catalog presented here, in conjunction with \tgas, represents the most powerful data set for chemo-orbital analysis at the dawn of 
 broadly available \gaia\ data.
@@
 
-These formed just some of the scientific drivers for the \gaia\ space telescope.
+The \gaia\ mission was borne out of the rich scientific potential of the stars of the Milky Way. 
 \gaia\ is primarily an astrometric mission, and will provide precise positions,
 parallaxes and proper motions for more than $10^9$ stars in its final data
-release in 2022.  While this is a sample size about four orders of magnitude 
+release, in 2022.  While this is a sample size about four orders of magnitude 
 larger than its predecessor \hipparcos, both astrometry and chemistry are 
@@
 (\rave) survey to have the largest overlap with the first \gaia\ data release: 
-292,036 stars.  We then used the \gaia\ universe model snapshot 
+292,036 stars.  We used the \gaia\ universe model snapshot 
 \citep{Robin_2012} to estimate the precision in parallax and proper motions that
@@
 we further found that the \rave\ survey will benefit most from \gaia\ DR1.  The 
-distances of 63\% of the \rave--\gaia\ DR1 overlap sample (182,862 stars) are 
+distances of a large fraction, of 63\% of the \rave--\gaia\ DR1 overlap sample, (182,862 stars) are 
 expected to improve with the first \gaia\ data release, and 47\% of stars are 

### 2016-09-05 article/unrave.tex
Some stubs in discussion, some banter, some ranting

--- 2016-09-05 article/unrave.tex
@@
 We have presented an independent re-analysis of \rave\ spectra that makes near-maximal
-use of the information content for red giant branch star, and provides significant a
-improvement in stellar parameter ($\teff$, $\logg$, [Fe/H]) precision over previous
-\rave\ data releases.  When combined with \gaia/\tgas, these results provide a powerful
-compendium for chemo-dynamic studies of the Milky Way.  However, our results have caveats.
+use of the information content for red giant branch stars, and provides a significant
+improvement in stellar parameters ($\teff$, $\logg$, [Fe/H]) over previous \rave\ data
+releases.  When combined with the \tgas\ sample, these results may amount to a powerful
+compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
 Any inferences based upon these results should be made with the recognition of those caveats,
-and acknowledge that they are subject to our explicit assumptions.
+and acknowledge that they are subject to our explicit assumptions, some of which are 
+provably incorrect.
+
+\stub{assumptions that are provably wrong: accurate labels in the training set. lol}
 
@@
 
-% Caveats in the results:
-% MULTIPLE MODELS
-% HETEROGENEOUS TRAINING SET
-% METAL POOR ABUNDANCES
-
-% UNSURE IF TWO ABUNDANCE SCALES ARE ON THE SAME SCALE --!
-% even if they are on the same scale...
-% metal-poor dwarfs/sub-giants
-
-% quality constraints recmomended
-% recommended SNR and chi-sq etc
-% weird projections in space. --> use morphological classifications to guide
-
-% despite these limitations, this catalog is the largest??? of abundances.
-% in the future we will do spectrophotometric distances to improve on this
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-We performed an independent analysis of \rave\ spectra using a data-driven approach.
-We report effective temperature $\teff$, surface gravity $\logg$, and metallicity [Fe/H]
-for X stars on the main-sequence and sub-giant branch.  For another X stars on the red
-giant branch, we deliver $\teff$, $\logg$, and a total of X.XX~million elemental abundances:
-O, Mg, Al, Ca, Si, Fe, and Ni for X stars.
-
-% Cross-validation, and
-% Validate through literature comparisons
-
-\stub{Typical precision}
-\stub{Something about exoplanet host stars, TESS, etc}
+\stub{Metallicities on the main-sequence and giant branch -- we cannot be sure if they
+are correct. Are there any indications that we can get which suggest whether they are
+good or bad?}
+
+For some of the most metal-poor giant stars in \rave\, we \emph{know} the abundances 
+are not on the same scale as \apogee, because we were forced to adopt abundances for 
+specific elements when they were unavailable.  Although we sought to adopt the mean
+level of Galactic chemical enrichment at the overall metallicity, this is not a
+representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
+metallicity, there is no requirement for zero abundance spread.
+More fundamentally, we are incorrectly assuming that the element \emph{must} be 
+detectable in the photosphere of the star.  There may be no transition that is 
+detectable in that star, even with zero noise, because it is too weak to have any 
+effect on the spectrum.  In the most optimistic case, this can be considered to be
+forcing the model to make use of correlated information between abundances.  In a
+more representative (pessimistic) case, we are simply invoking what all abundances
+should be at low metallicity.
+
+This choice is reflected in the abundances of the test set.  While we do recover
+trustworthy metallicities for ultra metal-poor ($[{\rm Fe/H}] \lesssim -4$) stars like 
+CD-38~245, the individual abundances for all extremely metal-poor stars aggregate
+(in [X/Fe] space; Figure~\ref{fig:gce}) at the assumed abundances for the metal-poor
+stars in our sample.  Thus, while the overall metallicities appear reliable, the 
+individual abundances for extremely metal-poor stars in the test set cannot be
+considered trustworthy in any sense.  For this reason we have updated the electronic
+catalog to discard these results, and included a quality control flag to indicate
+the result was discarded (see catalog \texttt{README} file for specifics related to
+the data structure and content).
+
+\stub{assumptions that are provably wrong: RAVE spectral resolution}
+
+\stub{assumptions that are provably wrong: correlated noise in the RAVE spectra -- this is minor and unlikely to cause a real issue}
+
+\stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}
+
+\stub{weird projections in the data space --> use morphological classifications to guide}
+
+\stub{quality constraints, etc}
+
+\stub{despite limitations, this catalog may be useful for science. N of abundances}
+
+\stub{in the future we will do spectrophotometric distances to improve on this}
 
@@
 first \gaia\ data release.  When combined with positions and 3D velocities from \gaia,
-we hope the \project{unRAVE} catalog will help advance understanding of the Milky Way's
-formation and evolution.
+we are optimistic that the \project{unRAVE} catalog will help advance understanding 
+of the Milky Way's formation and evolution.
 

### 2016-09-05 article/unrave.tex
add to discussion

@@
 \end{abstract}
-
-%We *will* provide 
-%spectrophotometric distances based on our stellar parameters, which should 
-%be superior to the parallax-based distances for faint or distant objects in
-%\tgas.
 
@@
 apparent, and the rich structure of the Hertzsprung-Russell is visible.
-
-\subsection{Spectrophotometric distances}
-\label{sec:distances}
-
-\stub{These will be derived by Jason Sanders when he returns from his honeymoon.
-The timing means that the distances will not be available when we submit the first
-version to arXiv. So we may need to remove this section and the reference to it
-from the abstract.}
 
@@
 compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
-Any inferences based upon these results should be made with the recognition of those caveats,
-and acknowledge that they are subject to our explicit assumptions, some of which are 
-provably incorrect.
-
-\stub{assumptions that are provably wrong: accurate labels in the training set. lol}
+Inferences based on these results should recognize those caveats, and acknowledge that 
+these results are subject to our explicit assumptions, some of which are provably
+incorrect.
 
@@
 branch stars primarily come from \apogee, where previous successes with \thecannon\
-have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity.
+have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
+\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016}.
 However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
-a tapering-off in the test set, where \emph{no} stars had reliably reported metallicities
-below that of the training set, which forced us to construct a heterogeneous training
+a tapering-off in the test set --- where \emph{no} stars had reliably reported metallicities
+below that of the training set --- which forced us to construct a heterogeneous training
 set.  The metal-poor stars included in this sample are from high-resolution studies
-\citep{Fulbright,Ruchti,elsewhere}, but it is unknown if the stellar parameters are
+\citep{Fulbright_2010,Ruchti_2011}, but it is unknown if the stellar parameters are
 of high fidelity (e.g., we have limited quality metrics available), and there is no
@@
 
-\stub{Metallicities on the main-sequence and giant branch -- we cannot be sure if they
-are correct. Are there any indications that we can get which suggest whether they are
-good or bad?}
+If the metallicities of metal-poor giant stars are on the same scale as the \apogee\
+abundances, there is a larger issue in verifying that the main-sequence metallicities
+and giant branch metallicities are on the same scale.  The training set for the
+main-sequence stars includes metallicities from a variety of sources --- including
+\lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
+no straightforward manner to ensure that these two models produce metallicities on
+the same scale.  We see no systematic offset in metallicities of dwarf and giant
+stars that overlap between \rave\ and the \ges\ survey, and our examination of
+globular and open clusters suggests that \stub{if there is a systematic offset, it must
+be small.} Nevertheless, these are only verification checks based on $<1$\% of the
+data, and there is currently insufficient data for us to prove both models are on
+the same abundance scale.
 
@@
 are not on the same scale as \apogee, because we were forced to adopt abundances for 
-specific elements when they were unavailable.  Although we sought to adopt the mean
+specific elements when they were unavailable.  In that sense, our assumption that the
+training set contains accurate labels is clearly untrue (for \emph{at least} some stars).
+Although we sought to adopt the mean
 level of Galactic chemical enrichment at the overall metallicity, this is not a
@@
 
-\stub{assumptions that are provably wrong: RAVE spectral resolution}
-
-\stub{assumptions that are provably wrong: correlated noise in the RAVE spectra -- this is minor and unlikely to cause a real issue}
+In Section~\ref{sec:method} we assumed that any fibre- or time-dependent variations
+in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
+\citet{Kordopatis_2013} note that the effective resolution of \rave\
+spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
+effective resolution is a function of time and position on the CCD (e.g., fibre number).  For this reason we ought to expect our derived stellar
+parameters or abundances to be correlated with the fibre number, or over
+time.  If significant, the trend could produce systematically offset
+stellar abundances solely due to the fibre number.  We have not seen
+evidence of these correlations in our results, however, we have only
+performed cursory (non-exhaustive) experiments to investigate this issue.
 
@@
 
+\bibitem[Fulbright et al.(2010)]{Fulbright_2010} Fulbright, J.~P., Wyse, R.~F.~G., Ruchti, G.~R., et al.\ 2010, \apjl, 724, L104
+
 \bibitem[Garc{\'{\i}}a P{\'e}rez et al.(2016)]{Garcia_Perez_2016} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., et al.\ 2016, \aj, 151, 144
@@
 \bibitem[Robin et al.(2012)]{Robin_2012} Robin, A.~C., Luri, X., Reyl{\'e}, C., et al.\ 2012, \aap, 543, A100
+
+\bibitem[Ruchti et al.(2011)]{Ruchti_2011} Ruchti, G.~R., Fulbright, J.~P., Wyse, R.~F.~G., et al.\ 2011, \apj, 743, 107
 

### 2016-09-05 article/unrave.tex
Explain MP dwarfs at low teff #7

@@
 time.  If significant, the trend could produce systematically offset
-stellar abundances solely due to the fibre number.  We have not seen
-evidence of these correlations in our results, however, we have only
+stellar abundances solely due to the fibre number.  \citet{Kordopatis_2013}
+conclude that resolution-based effects on the \rave\ stellar parameters
+should be a second-order effect. We have not seen evidence of these 
+resolution-based correlations in our results, however, we have only
 performed cursory (non-exhaustive) experiments to investigate this issue.
@@
 \stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}
+
+The fibre fringing is another example of how test spectra that are
+unusual from the training set can be projected as peculiar artefacts in the
+label space.  In other words, unusual spectra can appear as `clumps' in
+regions of parameter space that we could consider as being normal 
+(e.g., Solar-type stars).  We addressed this issue for the main-sequence
+and giant models by using a joint model to calculate relative weights.
+However, spectra that are unusual from the joint model training set can
+still project as systematic artefacts.  An interrogation of our results
+has identified one such group of metal-poor ($[{\rm Fe/H}] \approx -0.4$)
+stars near $\teff \approx 3250$~K, which peculiarly sit above the isochrone. 
+The \rave\ morphological classifications \citep{Matijevic_2012} indicate 
+these stars have chromospheric emission, a feature not present in our 
+training set.  Based on this (and other previously discussed) examples,
+we recommend either a thorough investigation of unexpected structure
+\footnote{Before submitting to \emph{Nature}} 
+using all the available 
+information from \rave, or judicious quality controls (e.g., requiring the 
+morphological classification to be `normal') to remove these spurious measurements.
 
@@
 
+\bibitem[Matijevi{\v c} et al.(2012)]{Matijevic_2012} Matijevi{\v c}, G., Zwitter, T., Bienaym{\'e}, O., et al.\ 2012, \apjs, 200, 14
+
 \bibitem[Marino et al.(2011)]{Marino_2011} Marino, A.~F., Milone, A.~P., Piotto, G., et al.\ 2011, \apj, 731, 64

### 2016-09-05 article/unrave.tex
filling out the discussion

@@
 training set.  Based on this (and other previously discussed) examples,
-we recommend either a thorough investigation of unexpected structure
-\footnote{Before submitting to \emph{Nature}} 
+we recommend either a thorough investigation of unexpected structure\footnote{
+Before submitting to \emph{Nature}} 
 using all the available 
 information from \rave, or judicious quality controls (e.g., requiring the 
-morphological classification to be `normal') to remove these spurious measurements.
-
-\stub{weird projections in the data space --> use morphological classifications to guide}
-
-\stub{quality constraints, etc}
-
-\stub{despite limitations, this catalog may be useful for science. N of abundances}
-
-\stub{in the future we will do spectrophotometric distances to improve on this}
-
-This catalog constitutes the largest collection of stellar abundances for stars in the
-first \gaia\ data release.  When combined with positions and 3D velocities from \gaia,
+morphological classification to be `normal') to remove these spurious projections.
+
+\stub{based on our interrogation of the data, we supply a 'QC' column that uses very strict requirements to eliminate weird things: morphological classifications, SNR, R (pipeline), $\chi_r^2$, etc....}
+
+We have presented the largest collection of stellar abundances for stars in the 
+first \gaia\ data release.  In total more than X.XX million abundances are
+available, with effective temperature $\teff$, surface gravity $\logg$ and 
+metallicity available for XXXXXX stars.  Despite the caveats and limitations 
+discussed here, our validation experiments and comparisons with high-resolution
+spectroscopic studies suggests that our results have sufficient accuracy and
+precision to be useful for exoplanet host stars studies 
+\citep[c.f.,][]{Valenti_Fischer_2005}, investigations of the Milky Way disk 
+\citep[c.f.,][]{Reddy_2003,Reddy_2006,Bensby_2014}, and a multitude of 
+chemo-dynamic studies.  When combined with positions and 3D velocities from \gaia,
 we are optimistic that the \project{unRAVE} catalog will help advance understanding 

### 2016-09-05 article/unrave.tex
Use \, as multiply symbol. Closes #20
Clarify 63% text. Closes #18

@@
 overlap samples.  Comparing the expected precision to what is currently available, 
-we further found that the \rave\ survey will benefit most from \gaia\ DR1.  The 
-distances of a large fraction, of 63\% of the \rave--\gaia\ DR1 overlap sample, (182,862 stars) are 
-expected to improve with the first \gaia\ data release, and 47\% of stars are 
-likely to have better proper motions (137,211 stars).  Although the \gaia\ universe
+we further found that the \rave\ survey will benefit most from \gaia\ DR1: 63\%
+of stars in the \rave--\gaia\ DR1 overlap sample (182,862 stars) are 
+expected to improve with the first \gaia\ data release, and 47\% of stars (137,211 stars) are 
+likely to have better proper motions.  Although the \gaia\ universe
 model assumes end-of-mission uncertainties --- and does not account for systematics
@@
     \sum_{n=0}^{N-1} \frac{[y_{jn}-\vecv(\ell_n)\cdot\vectheta]^2}{\sigma^2_{jn}+s^2}
-    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}Q(\vectheta)
+    + \sum_{n=0}^{N-1} \ln(\sigma^2_{jn}+s^2) + \Lambda{}\,Q(\vectheta)
     \right]
@@
 \begin{eqnarray}\label{eq:vectorizer-three-label}
-\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
+\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\,T_{{\rm eff},n}, \feh_n\,T_{{\rm eff},n}, \logg_n^2, \feh_n\,\logg_n, \feh_n^2\right]
 \end{eqnarray}
@@
 \begin{eqnarray}
-	\hat\ell = \frac{w_{ms}\ell_{ms} + w_{giant}\ell_{giant}}{w_{ms} + w_{giant}} \quad .
+	\hat\ell = \frac{w_{ms}\,\ell_{ms} + w_{giant}\,\ell_{giant}}{w_{ms} + w_{giant}} \quad .
 \end{eqnarray}

### 2016-09-05 article/unrave.tex
Clarify representative using mostly Hogg's words - closes #19

@@
 misclassifications in the training set are affordable).
-\item We assume the training set is representative of the test set (the \rave\ survey).
+\item We assume that the training data are similar (in spectra) to the test data 
+where they overlap in label space, and we assume that the training data spans enough
+of the label space to capture the variation in the test-set spectra.
 \end{itemize}

### 2016-09-05 article/unrave.tex
Include training set source in abstract; claim improvement over DR4. Closes #16 #17

@@
 spectra, using an implementation of \thecannon, that yields more precise 
-and accurate stellar parameters and abundances.  We derive, and validate
-improved effective temperature $\teff$, surface gravity $\logg$, and 
-chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
-for \Nstarsqc\ stars.  The typical precision in chemical abundances is 
-X.XX~dex for stars on the giant branch.  
-The \unrave\ catalog presented here, in conjunction with \tgas, represents the most powerful data set for chemo-orbital analysis at the dawn of 
-broadly available \gaia\ data.
+and accurate stellar parameters and abundances than previous \rave\
+data releases.  We constructed our model using high-fidelity \apogee\ 
+stellar parameters and abundances for red giant branch stars that overlap
+with \rave, and stellar parameters from the \project{K2}/\project{Ecliptic 
+Pole Input Catalog} for main-sequence and sub-giant branch stars.
+We derive, and validate, improved effective temperature $\teff$, surface 
+gravity $\logg$, and chemical abundances of up to seven elements (O, Mg, 
+Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in 
+chemical abundances is X.XX~dex for red giant branch stars.  The \unrave\ 
+catalog presented here, in conjunction with \tgas, represents the 
+most powerful data set for chemo-orbital analysis at the dawn of broadly 
+available \gaia\ data.
 \end{abstract}
@@
 chemo-dynamic studies.  When combined with positions and 3D velocities from \gaia,
-we are optimistic that the \project{unRAVE} catalog will help advance understanding 
+we are optimistic that the \unrave\ catalog will help advance understanding 
 of the Milky Way's formation and evolution.

### 2016-09-05 article/unrave.tex
Additions to tex, plot_clusters.py, and padova isochrones, added cluster HRDs

@@
 survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
-globular cluster NGC~3201, and 10 members of NGC~362.  
-All of these systems are well-studied and exhibit abundance anti-correlations 
+consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 members of NGC~362.
+In addition to these systems, NGC~6752 and 47~Tuc were observed \citep{Kunder_2016}.
+In the top panel of Figure~\ref{fig:globular-cluster-HRD} we plot the spectroscopic H-R diagram for the members of these globular clusters.
+For comparison, in the bottom panel of Figure~\ref{fig:globular-cluster-HRD} we show the same as the top panel but for the RAVE DR4.
+For a detailed description of the cluster membership for NGC~5139, NGC~3201, and NGC~362 we refer the reader to \cite{Anguiano_2015}.
+The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
+\stub{check if this is the case in Kunder et al 2016.}
+
+All of these systems (\stub{check if this is the case for the remaining two new GCs added.}) in \cite{Anguiano_2015} are well-studied and exhibit abundance anti-correlations 
 in Na--O and Mg--Al \citep{Marino_2011,Carretta_2009,Carretta_2013,Munoz_2013}.  Other correlations are also present 
@@
 % Some literature notes:
+%check if NGC 6571 has anticorrelations present
 % See http://iopscience.iop.org/article/10.1088/0004-637X/731/1/64/pdf
@@
 \label{sec:open-cluster-validation}
+The RAVE survey has observed several open clusters which can serve as a qualitative sanity check that the stellar parameters match our expectations.
+These include ....
+\stub{add more here about each open cluster plus its members}.
+The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
+\stub{check if this is the case in Kunder et al 2016.}
+
+ In Figure~\ref{fig:globular-cluster-HRD}, we plot the cannon-derived H-R diagram (top panel) and compare it to the RAVE DR4 H-R diagram (bottom panel) for the stars which are probably members of the open clusters.
 
@@
 \begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/HRD_GCs.pdf}
+\caption{Top Panel: Cannon-derived $\logg$ as a function of $\teff$ for globular cluster members.
+The cluster name is indicated in the top left corner.
+The points are color-coded by the metallicity of the member star.
+Bottom Panel: RAVE DR4 $\logg$ as a function of $\teff$ for the same clusters in the top panel.
+The points are color-coded by the metallicity of the member star.
+In both panels, the black line represent the theoretical expectation of a single stellar population from the Padova isochrone set.
+\label{fig:globular-cluster-HRD}}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/HRD_OCs.pdf}
+\caption{Same as Figure~\ref{fig:globular-cluster-HRD} for the open cluster members in the RAVE dataset.  \label{fig:open-cluster-HRD}}
+\end{figure*}
+
+\begin{figure*}[p]
 \includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}

### 2016-09-05 article/unrave.tex
no more conclusion

--- 2016-09-05 article/unrave.tex
@@
 are consistent with expectations from astrophysics.  We discuss the implications
-of these comparisons and our results in Section~\ref{sec:discussion}.  We provide
-concluding remarks in Section~\ref{sec:conclusion}, and describe how to access our
-results electronically.
+of these comparisons in Section~\ref{sec:discussion}, and conclude with instructions
+for how to access our results electronically.
 

### 2016-09-06 article/unrave.tex
Additions to tex file on GC and OCs

@@
 % There is some order already in mind, but final order will be determined by  contribution.
-\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W. Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
+\author{Some combination of: Harry Enke, Kenneth C.
+Freeman, Gerry Gilmore, Keith Hawkins, David W.
+Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
 
@@
 \author{Luca Casagrande, Cristina Chiappini, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
+
+% RAVE builders
+\author{Brad K. Gibson}
+% E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
 
@@
 \citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
-survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
+survey, \cite{Anguiano_2015} found 80 stars with positions and radial velocities that are
 consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 members of NGC~362.
-In addition to these systems, NGC~6752 and 47~Tuc were observed \citep{Kunder_2016}.
+In addition to these systems, NGC~6752 and \stub{47~Tuc} were observed \citep{Kunder_2016}.
+For a detailed description of the cluster membership for NGC~5139, NGC~3201, and NGC~362 we refer the reader to \cite{Anguiano_2015}.
+The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
+\stub{check if this is the case in Kunder et al 2016.} We note here that we have not re-evaluated cluster membership as that is beyond the scope of the paper, however with the results presented here it is possible to study these systems in further detail.
+
 In the top panel of Figure~\ref{fig:globular-cluster-HRD} we plot the spectroscopic H-R diagram for the members of these globular clusters.
 For comparison, in the bottom panel of Figure~\ref{fig:globular-cluster-HRD} we show the same as the top panel but for the RAVE DR4.
-For a detailed description of the cluster membership for NGC~5139, NGC~3201, and NGC~362 we refer the reader to \cite{Anguiano_2015}.
-The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
-\stub{check if this is the case in Kunder et al 2016.}
+In both panels isochrones, we have also plotted the PARSEC isochrones \citep{
+Bressan_2012} for comparison.
+The basic properties of the globular cluster used to generate the isochrones were sourced from Table 2 of \cite{Kunder_2016} and from the 2010 version of the Harris Globular Cluster Catalogue \citep{Harris_1996}.This figure indicates that for some stars in specific clusters clusters, namely NGC3201 and NGC362, are slightly offset from the isochrones most likely because the star's metallicity is significantly higher than the mean cluster metallicity.
+This could be a result of a combination of the inclusion of stars that are not cluster members but appeared to be consistent with the RAVE DR4 stellar parameters.
+In the remaining globular clusters the stars are consistent with the isochrones.
+In the case of NGC5139 ($\omega$ cen), it may be possible to pick out two of the several components that exist within that cluster.
 
@@
 
+%Mean metallicity of NGC3201 cluster = -0.75 +/- 0.20
+%Mean metallicity of NGC362 cluster = -0.92 +/- 0.48
+%Mean metallicity of NGC5139 cluster = -0.85 +/- 0.55
+%Mean metallicity of NGC6752 cluster = -1.17 +/- 0.30
+
 \subsubsection{Open clusters}
@@
 The RAVE survey has observed several open clusters which can serve as a qualitative sanity check that the stellar parameters match our expectations.
-These include ....
-\stub{add more here about each open cluster plus its members}.
-The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
-\stub{check if this is the case in Kunder et al 2016.}
-
- In Figure~\ref{fig:globular-cluster-HRD}, we plot the cannon-derived H-R diagram (top panel) and compare it to the RAVE DR4 H-R diagram (bottom panel) for the stars which are probably members of the open clusters.
+These include $\sim$160 stars within the Pleiades (78 stars), the Hyades (26 stars), M67 (30 stars) and IC4651 (13 stars) open clusters.
+Details on the open cluster membership and selection can be found in \cite{Kunder_2016}.
+The basic properties of the open cluster were sourced from Table 2 of \cite{Kunder_2016} and from \cite{Kharchenko_2013}.
+
+In Figure~\ref{fig:open-cluster-HRD}, we plot the cannon-derived H-R diagram (top panel) and compare it to the RAVE DR4 H-R diagram (bottom panel) for the stars which are likely members of the open clusters.
+In all cases, except the Pleiades, the stellar parameters derived from the Cannon are consistent with the expectations from the isochrones.
+We note that theoretical isochrones struggle to match the observed color-magnitude diagrams of young cluster, such as the Pleiades, and thus there has been an effort to use empirical isochrones for such systems \citep[e.g.][]{Bouy_2015}.
+Additionally, the improvement in the precision of the stellar parameters in the study compared to RAVE DR4 is illustrated in the top and bottom panels of Figure~\ref{fig:open-cluster-HRD} by comparing the tightness of the main-sequence in and sub-giant branch in these open cluster.
+
+We illustrate the precision in the chemical abundance by quantifying the star-to-star dispersion of  the detailed abundances for these open cluster members in
+Figure \ref{fig:open-cluster-abundances}.
+The internal cluster abundance dispersion can be used as a gauge of precision because open clusters are thought to be chemically homogenous \citep[e.g.][]{Bovy_2016}.
+\stub{What do we find?}
+
+%Mean metallicity of Hyades cluster = 0.07 +/- 0.09 (26 stars)
+%Mean metallicity of IC4651 cluster = 0.15 +/- 0.12 (13 stars)
+%Mean metallicity of M67 cluster = -0.02 +/- 0.17 (30 stars)
+%Mean metallicity of Pleiades cluster = -0.02 +/- 0.12 (78 stars)
 
@@
 
+\bibitem[Bressan et al.(2012)]{Bressan_2012} Bressan, A., Marigo, P., Girardi, L., et al.\ 2012, \mnras, 427, 127
+
 \bibitem[Bovy et al.(2016)]{Bovy_2016} Bovy, J., Rix, H.-W., Schlafly, E.~F., et al.\ 2016, \apj, 823, 30
@@
 
+\bibitem[Bouy et al.(2015)]{Bouy_2015} Bouy, H., Bertin, E., Sarro, L.~M., et al.\ 2015, \aap, 577, A148
+
 \bibitem[Carretta et al.(2009)]{Carretta_2009} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2009, \aap, 505, 117
@@
 \bibitem[Gilmore et al.(2012)]{Gilmore_2012} Gilmore, G., Randich, S., Asplund, M., et al.\ 2012, The Messenger, 147, 25
+
+\bibitem[Harris(1996)]{Harris_1996} Harris, W.~E.\ 1996, \aj, 112, 1487
 

### 2016-09-06 combine_model_results.py
Update script to combine spectra with correct DR5 table

@@ -19,7 +19,7 @@ output_filename, overwrite = \
 RESULTS_PATH = "/data/gaia-eso/arc/rave/results/"
 RESULTS_PATH = ""
 
-ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v37.fits.gz"))
+ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v43.fits.gz"))
 giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
 joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v41.fits.gz"))
 
@@ -155,7 +155,7 @@ x_mu, y_mu = (0, 0)
 x_sigma, y_sigma = (90, 0.15)
 x = ((ms_results["TEFF"] - joint_results["TEFF"]) - x_mu)/x_sigma
 y = ((ms_results["LOGG"] - joint_results["LOGG"]) - y_mu)/y_sigma
-
+z = ((ms_results["FE_H"] - joint_results["FE_H"]) - 0)/0.15
 
 axes[0].hexbin(x, y, gridsize=100, extent=(-3, +3, -3, +3), norm=LogNorm(), linewidths=0.1)
 
@@ -166,19 +166,20 @@ x_sigma, y_sigma = (50, 0.15)
 
 x2 = ((giant_results["TEFF"] - joint_results["TEFF"]) - x_mu)/x_sigma
 y2 = ((giant_results["LOGG"] - joint_results["LOGG"]) - y_mu)/y_sigma
-
+z2 = ((giant_results["FE_H"] - joint_results["FE_H"]) - 0)/0.15
 
 axes[1].hexbin(x2, y2, gridsize=100, extent=(-3, +3, -3, +3), norm=LogNorm(), linewidths=0.1)
 
 
 
-ms_distance = np.sqrt(x**2 + y**2)
-giant_distance = np.sqrt(x2**2 + y2**2)
+ms_distance = np.sqrt(x**2 + y**2 + z**2)
+giant_distance = np.sqrt(x2**2 + y2**2 + z2**2)
 
 
 # Weight by relative distance, or just take the closest of the two?
 combined_data = OrderedDict([
-    ("Name", giant_results["Name"]),
+    #("Name", giant_results["Name"]),
+    ("RAVE_OBS_ID", giant_results["Name"]),
     ("TEFF", None),
     ("LOGG", None),
[…]

### 2016-09-06 article/figures/plot_v34_search.py
Fix hyperparameter figure

@@ -33,14 +33,15 @@ comparisons = [
         glob("rave-tgas-v34-test-labelled-set-?.????????e*-1.00.pkl"),
         "rave-tgas-v34-test-labelled-set-0.00000000e+00-1.00.pkl",
         "#3498DB",
-        r"${\rm Main{-}sequence}$ ${\rm model}$"
+        r"${\rm Giant}$ ${\rm branch}$ ${\rm model}$"
+
     ),
     (
         tc.load_model("rave-tgas-v16b.model"),
         glob("rave-tgas-v16b-test-labelled-set-*+vsini.pkl"),
         'rave-tgas-v16b-test-labelled-set-1.00000000e-03+vsini.pkl',
         "r",
-        r"${\rm Giant}$ ${\rm model}$"
+        r"${\rm Main{-}sequence}$ ${\rm model}$"
     )
 ]
 

### 2016-09-06 article/unrave.tex
Add JBH
Caution at the start and end of the joining-the-models section that what we have done is totally unjustified in any probabilistic sense, and it is merely a heuristic

@@
 % RAVE builders
-\author{Brad K. Gibson}
+\author{Brad K. Gibson, Joss Bland-Hawthorn}
 % E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
@@
 training sets.  For these reasons we were forced to exclude or severely penalize 
-incorrect results from both models.
+incorrect results from both models.  We emphasize that the choices here are entirely
+heuristic, and depart from interpreting \thaecannon\ output labels as the maxima
+of individual likelihood functions.  Each model produces estimates of the labels
+for a given star, and we use those estimates to produce a unified estimate, but 
+this is done by disregarding any probabilistic justification.
 
@@
 and mean metallicity.  The artefacts from individual models are no longer 
-apparent, and the rich structure of the Hertzsprung-Russell is visible.
+apparent, and the rich structure of the Hertzsprung-Russell is visible.  Despite
+this, there are a number of caveats introduced by the choices we have made to
+combine estimates from multiple models. We discuss these issues in detail in
+Section \ref{sec:discussion}.
 

### 2016-09-06 article/unrave.tex
Typo

@@
 incorrect results from both models.  We emphasize that the choices here are entirely
-heuristic, and depart from interpreting \thaecannon\ output labels as the maxima
+heuristic, and depart from interpreting \thecannon\ output labels as the maxima
 of individual likelihood functions.  Each model produces estimates of the labels

### 2016-09-06 article/unrave.tex
Address Ness concerns #22 and add Siebert as co-author

@@
 % RAVE builders
-\author{Brad K. Gibson, Joss Bland-Hawthorn}
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert}
+%Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
 % E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
-
+%1Observatoire Astronomique, Universite de Strasbourg, CNRS, 11 rue de l?universit ? e, 67000 Strasbourg, France
 \author{and the \rave\ collaboration}
@@
 
-The \gaia\ mission was borne out of the rich scientific potential of the stars of the Milky Way. 
+The \gaia\ mission represents a critical step forward in understanding the Galaxy.
 \gaia\ is primarily an astrometric mission, and will provide precise positions,
@@
 in later sections, and $Q(\vectheta)$ is a L1 regularizing function that encourages 
-$\vectheta$ values to take on zero values without breaking convexity:
+$\vectheta$ values to take on zero values without breaking convexity \citep{Casey_2016b}:
 \begin{eqnarray}\label{eq:regularization-function}

### 2016-09-06 article/unrave.tex
Add Navarro
Included comments on Intro/abstract #23

@@
 % RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert}
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro}
 %Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
-% E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
-%1Observatoire Astronomique, Universite de Strasbourg, CNRS, 11 rue de l?universit ? e, 67000 Strasbourg, France
+%E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
+%Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
+%Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
+%Senior CIfAR Fellow; Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2
 \author{and the \rave\ collaboration}
@@
 The Milky Way is a powerful laboratory for understanding galaxy formation,
-as it can provide orbits, ages, physical parameters, as well as chemical 
+as it can provide orbits, ages, atmospheric parameters, as well as chemical 
 abundances for vast sets of individual stars.  These inferences require 
@@
 \citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
-Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Kordopatis_2015,Bovy_2016}.  
+Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Boeche_2013,Kordopatis_2015,Bovy_2016}.  
 However, these quantities are not known for even 1\% of stars in the Milky Way.  
@@
 \citep{Steinmetz_2006} surveys}, we found the RAdial Velocity Experiment 
-(\rave) survey to have the largest overlap with the first \gaia\ data release: 
-292,036 stars.  We used the \gaia\ universe model snapshot 
+(\rave) survey is expected to have the largest overlap with the first \gaia\ 
+data release: $\approx$292,036 stars.  We used the \gaia\ universe model snapshot 
 \citep{Robin_2012} to estimate the precision in parallax and proper motions that
-will be available in the first \gaia\ data release (DR1) for stars in those 
+could be available in the first \gaia\ data release (DR1) for stars in those 
 overlap samples.  Comparing the expected precision to what is currently available, 
 we further found that the \rave\ survey will benefit most from \gaia\ DR1: 63\%
-of stars in the \rave--\gaia\ DR1 overlap sample (182,862 stars) are 
-expected to improve with the first \gaia\ data release, and 47\% of stars (137,211 stars) are 
+of stars in the \rave--\gaia\ DR1 overlap sample ($\approx$182,862 stars) are 
+expected to improve with the first \gaia\ data release, and 47\% of stars ($\approx$137,211 stars) are 
 likely to have better proper motions.  Although the \gaia\ universe
@@
 
+\bibitem[Bensby et al.(2014)]{Bensby_2014} Bensby, T., Feltzing, S., \& Oey, M.~S.\ 2014, \aap, 562, A71
+
 \bibitem[Binney et al.(2014)]{Binney_2014} Binney, J., Burnett, B., Kordopatis, G., et al.\ 2014, \mnras, 437, 351
 
-\bibitem[Bensby et al.(2014)]{Bensby_2014} Bensby, T., Feltzing, S., \& Oey, M.~S.\ 2014, \aap, 562, A71
+\bibitem[Boeche et al.(2013)]{Boeche_2013} Boeche, C., Chiappini, C., Minchev, I., et al.\ 2013, \aap, 553, A19
 

### 2016-09-06 article/unrave.tex
Include additional comments from Steinmetz #23

@@
 very efficient observing:  spectra for up to 150 targets could be simultaneously
-acquired.  When observations concluded in April 2013, a total of \Nspectra\ useful 
-spectra had been collected of \Nstars\ unique objects.
+acquired.  When observations concluded in April 2013, at least \Nspectra\ useful 
+spectra had been collected of more than \Nstars\ unique objects.
 
@@
 The $I$ band was used for the target selection because it has a good overlap with the
-wavelength range that \rave\ operates in:  8410--8795~\AA.  This spectral region 
-is very well-studied.  It is one of the key setups used for the ground-based 
-high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}, and the 
-\rave\ resolution and wavelength coverage is comparable to the Radial Velocity 
-Spectrometer on board the \gaia\ space telescope \citep{Kordopatis_2011,Recio-Blanco_2016}.  
-The region includes the \ion{Ca}{2} near infrared triplet lines --- strong 
-transitions that are dominated by pressure broadening --- which are visible even 
-in metal-poor stars or spectra with very low signal-to-noise (S/N) ratios.
-Atomic transitions of 
+wavelength range that \rave\ operates in:  8410--8795~\AA.  The resolution and 
+wavelength coverage is comparable to the Radial Velocity Spectrometer on board 
+the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011,Recio-Blanco_2016}, 
+and the wavelength range overlaps with one of the key setups used for the ground-based 
+high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}.  The region 
+includes the \ion{Ca}{2} near infrared triplet lines --- strong transitions that 
+are dominated by pressure broadening --- which are visible even in metal-poor stars
+or spectra with very low signal-to-noise (S/N) ratios.  Atomic transitions of 
 light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
@@
 Despite the relatively low resolution and S/N of the spectra compared to other
-surveys, the \rave\ data releases have demonstrated the team's enormous success 
-in deriving radial velocities, stellar astrophysical parameters ($\teff$, $\logg$),
-as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
-Siebert_2011,Kordopatis_2013,Kunder_2016}.  In this work we make use of spectra
+surveys, the \rave\ data releases have provided excellent radial velocities, 
+stellar astrophysical parameters ($\teff$, $\logg$), as well as individual 
+chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,Boeche_2011,
+Kordopatis_2013,Kunder_2016}.  In this work we make use of spectra
 that has been reprocessed for the fifth \rave\ data release.  These re-processing
@@
 
-We chose to adopt a data-driven model --- rather than a physics-based model ---
-for this analysis.  Specifically, we will use an implementation of \thecannon\
+We chose to adopt a data-driven model for this analysis, rather than the
+physics-based models used in \rave\ data releases 2--5.
+ Specifically, we will use an implementation of \thecannon\
 \citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
@@
 but that sample contains only giant stars}, and the \lamost\ lower main-sequence suffers
-from the same systematic effects seen in \apogee\ data.  The \galah\ survey has a good 
-overlap with \rave, but results from that sample are not yet precise enough for our 
-purposes.
+from the same systematic effects seen in \apogee\ data.
 
@@
 and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
-\citet{Kordopatis_2013} has been noted in other studies, and this has issue
+\citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this has issue
 has been resolved in the fifth \rave\ data release by correcting $\logg$
-values with a small sample of asteroseismic targets.  There is also a
+values with a calibration sample consisting of asteroseismic targets and the
+\gaia\ benchmark stars.  There is also a
 slight discrepancy in the $\log$ values along the main-sequence, where our
@@
 work based on main-sequence dwarf stars.  The results from both the main-sequence star model, 
-and the giant star model, are available at \stub{should we put them online somewhere
+and the giant star model, are available on request. \stub{should we put them online somewhere
 for open-science, or just make them available upon request so that we don't
@@
 relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
-agreement in stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
+agreement in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
 The RMS in effective temperature, surface gravity and metallicity is
@@
 spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
-effective resolution is a function of time and position on the CCD (e.g., fibre number).  For this reason we ought to expect our derived stellar
+effective resolution is a function of temperature variations, fibre-to-fibre
+variations, and thus position on the CCD \citep{Steinmetz_2006}.
+ For this reason we ought to expect our derived stellar
 parameters or abundances to be correlated with the fibre number, or over
@@
 and this document was compiled from revision hash \texttt{\githash} in that repository.
-Inferred labels, covariance matrices, and relevant metadata are available electronically
-at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  Please note that it is a condition
+Inferred labels, associated errors, and relevant metadata are available electronically
+at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  These results are also available
+through the \rave\ DR5 database \citep{Kunder_2016}. Please note that it is a condition
 of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must 
@@
 
+\bibitem[Boeche et al.(2011)]{Boeche_2011} Boeche, C., Siebert, A., Williams, M., et al.\ 2011, \aj, 142, 193
+
 \bibitem[Boeche et al.(2013)]{Boeche_2013} Boeche, C., Chiappini, C., Minchev, I., et al.\ 2013, \aap, 553, A19
@@
 \bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
+
+\bibitem[Munari et al.(2005)]{Munari_2005} Munari, U., Sordo, R., Castelli, F., \& Zwitter, T.\ 2005, \aap, 442, 1127
 

### 2016-09-06 article/unrave.tex
Edits up to Sec 3.2

@@
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-292,036 stars.  Here we present a data-driven re-analysis of \rave\ 
+up to 292,036 stars.  Here we present a data-driven re-analysis of \rave\ 
 spectra, using an implementation of \thecannon, that yields more precise 
@@
 Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in 
+% TODO
 chemical abundances is X.XX~dex for red giant branch stars.  The \unrave\ 
@@
 
-\keywords{\vspace{5em}}
+\keywords{}
 
@@
 parallaxes and proper motions for more than $10^9$ stars in its final data
-release, in 2022.  While this is a sample size about four orders of magnitude 
+release in 2022.  While this is a sample size about four orders of magnitude 
 larger than its predecessor \hipparcos, both astrometry and chemistry are 
@@
 \galah\ \citep{DeSilva_2015}, \lamost\ \citep{Cui_2012}, and \rave\ 
-\citep{Steinmetz_2006} surveys}, we found the RAdial Velocity Experiment 
+\citep{Steinmetz_2006} surveys}, we found that the RAdial Velocity Experiment 
 (\rave) survey is expected to have the largest overlap with the first \gaia\ 
-data release: $\approx$292,036 stars.  We used the \gaia\ universe model snapshot 
+data release: up to 292,036 stars.  We used the \gaia\ universe model snapshot 
 \citep{Robin_2012} to estimate the precision in parallax and proper motions that
@@
 We chose to adopt a data-driven model for this analysis, rather than the
-physics-based models used in \rave\ data releases 2--5.
- Specifically, we will use an implementation of \thecannon\
+physics-based models used in the \rave\ data releases.  Specifically, we will
+use an implementation of \thecannon\
 \citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
@@
 We used no regularization ($\Lambda = 0$) for this model.  After training the model we
-treated all \Nspectra\ spectra as test set objects.  In the right-hand panel of Figure 
+treated all \Nspectra\ spectra as test set objects.  In the left-hand panel of Figure 
 \ref{fig:test-set-density} we show the effective temperature $\teff$ and surface gravity 
@@
 The stars within the peak at $(0, 0)$ represent objects where
-the joint model and the comparison model both report similar labels.  The artefacts 
+the joint model and the comparison model both report similar labels.  
+\stub{State these are empirically chosen normalization factors, and how}
+The artefacts 
 seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
@@
 
+\stub{Include text about fast rotators; blue stragglers}
+
 \stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}
@@
 morphological classification to be `normal') to remove these spurious projections.
+
+\stub{Sub-giant artefact; comment as projection}
 
@@
 \begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
+Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+\label{fig:gold-standard-comparison}}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
+\caption{Hertsprung-Russell diagrams of stars in common between this work and that of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
+Stars are colored by the metallicity of each study.
+Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
+This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
+\end{figure*}
+
+\begin{figure*}[p]
 \includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
-\label{fig:gold-standard-comparison}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
-\caption{Hertsprung-Russell diagrams of stars in common between this work and that of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the metallicity of each study.
-Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
-This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
-\end{figure*}
-
-\begin{figure*}[p]
 \includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}

### 2016-09-06 article/unrave.tex
Additions to GC plot including NGC1851 and to tex file

@@
 \citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
-survey, \cite{Anguiano_2015} found 80 stars with positions and radial velocities that are
+survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
 consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 members of NGC~362.
-In addition to these systems, NGC~6752 and \stub{47~Tuc} were observed \citep{Kunder_2016}.
+In addition to these systems, 12 stars thought to belong to NGC~1851 were found in the RAVE data in \cite{Kunder_2014} and 10 stars in NGC~6752 and XX stars in \stub{47~Tuc} were discussed in \citep{Kunder_2016}.
 For a detailed description of the cluster membership for NGC~5139, NGC~3201, and NGC~362 we refer the reader to \cite{Anguiano_2015}.
-The membership selection of the remaining clusters can be found in \cite{Kunder_2016}.
-\stub{check if this is the case in Kunder et al 2016.} We note here that we have not re-evaluated cluster membership as that is beyond the scope of the paper, however with the results presented here it is possible to study these systems in further detail.
+The membership selection for NGC~1851 can be found in \citep{Kunder_2014} and for NGC~6752 and 47~Tuc can be found in \cite{Kunder_2016}.
+We note here that we have not re-evaluated cluster membership as that is beyond the scope of the paper, however with the results presented here it is possible to study these systems in further detail.
 
@@
 Bressan_2012} for comparison.
-The basic properties of the globular cluster used to generate the isochrones were sourced from Table 2 of \cite{Kunder_2016} and from the 2010 version of the Harris Globular Cluster Catalogue \citep{Harris_1996}.This figure indicates that for some stars in specific clusters clusters, namely NGC3201 and NGC362, are slightly offset from the isochrones most likely because the star's metallicity is significantly higher than the mean cluster metallicity.
+The basic properties of the globular cluster used to generate the isochrones were sourced from Table 2 of \cite{Kunder_2016}, \cite{Marin-Franch_2009}, and the 2010 version of the Harris Globular Cluster Catalogue \citep{Harris_1996}.This figure indicates that for some stars in specific clusters clusters, namely NGC3201 and NGC362, are slightly offset from the isochrones most likely because the star's metallicity is significantly higher than the mean cluster metallicity.
 This could be a result of a combination of the inclusion of stars that are not cluster members but appeared to be consistent with the RAVE DR4 stellar parameters.
@@
 
+\bibitem[Kunder et al.(2014)]{Kunder_2014} Kunder, A., Bono, G., Piffl, T., et al.\ 2014, \aap, 572, A30
+
 \bibitem[Kuzma et al.(2016)]{Kuzma_2016} Kuzma, P.~B., Da Costa, G.~S., Mackey, A.~D., \& Roderick, T.~A.\ 2016, \mnras, 461, 3639
@@
 \bibitem[Marino et al.(2011)]{Marino_2011} Marino, A.~F., Milone, A.~P., Piotto, G., et al.\ 2011, \apj, 731, 64
+
+\bibitem[Mar{\'{\i}}n-Franch et al.(2009)]{Marin-Franch_2009} Mar{\'{\i}}n-Franch, A., Aparicio, A., Piotto, G., et al.\ 2009, \apj, 694, 1498
 

### 2016-09-07 article/unrave.tex
De-stubbing

--- 2016-09-06 article/unrave.tex
@@
 tests with globular cluster members showed that the metallicity range of the training 
-set needed to extend below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog to be 
-practically useful.  \stub{Supplemented this sample with known metal-poor stars in \rave}.
+set needed to extend at least below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog 
+to be practically useful.  Without additional metal-poor stars, the lowest metallicity
+labels reported by our model would be around $[{\rm Fe/H}] \approx -2$, even for well
+studied stars with $[{\rm Fe/H}] \sim -4$ (e.g., CD~38-245).  For this reason we
+supplemented our sample of \apogee\ giant stars with 176 known metal-poor giant stars 
+observed by \rave.  The effective temperature $\teff$, surface gravity $\logg$ and
+iron abundance [Fe/H] labels were adopted from \citet{Fulbright_2010, Ruchti_2011}.
+The elemental abundances for O, Mg, Al, Si, Ca, and Ni were assumed to follow typical
+trends of Galactic chemical evolution, such that we assumed $[{\rm Mg/Fe}] = +0.4$,
+$[{\rm O/Fe}] = +0.4$, $[{\rm Al/Fe}] = -0.5$, $[{\rm Ca/Fe}] = +0.4$, 
+$[{\rm Si/Fe}] = +0.4$, and $[{\rm Ni/Fe}] = -0.25$.  This decision is made solely 
+to ensure that our overall metallicity scale extends that of the \rave\ survey, down 
+to $[{\rm Fe/H}] \sim -4$.  In other words, we do not recommend the use of individual 
+abundance labels at $[{\rm Fe/H}] \sim -4$; in practice the abundances of most of these
+elements are largely unrecoverable in the \rave\ wavelength region at $[{\rm Fe/H}] \sim -4$.
+We discuss this issue in more detail in Section~\ref{sec:discussion}.
 
@@
 under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
-(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$.  
-\stub{Include known metal-poor stars in \rave?}. After training
+(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$, and included
+an additional 12 turn-off stars with $-1.6 \gtrsim [{\rm Fe/H}] \gtrsim -2.1$ from 
+\citet{Ruchti_2011}.  After training
 a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
@@
 set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
-was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
-stars (of 4,611).  The full training set for each model (see next sections) is shown
+was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 595 
+stars (583 of 4,611 from \epic).  The full training set for each model (see next sections) is shown
 in Figure~\ref{fig:training-set-hrd}.

### 2016-09-07 article/unrave.tex
fill in them numbers

@@
 We chose to adopt separate models for the main-sequence and the red giant branch rather
-than switch to a single model with more complexity.  This choice allowed us to derive
+than switch to a single model with higher complexity.  This choice allowed us to derive
 stellar parameters for stars on the main-sequence and sub-giant branch, as well as 
@@
 We constructed a three-label quadratic model using only main-sequence and sub-giant
-stars in the \epic\ training set.  In order to set the regularization hyperparameter
+stars. In order to set the regularization hyperparameter
 $\Lambda$ for this model, we trained 30 models with different regularization strengths,
@@
 leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
-will use as the basis for comparison.  For every model we calculated the bias and
-root-mean-square (RMS) deviation between the training set labels and those inferred
-through cross-validation.
+will use as the basis for comparison.  For the unregularized case, we calculated the 
+bias and root-mean-square (RMS) deviation between: the training set labels, and the 
+labels we derive by cross-validation, where one star is removed at a time and the
+model is re-trained.  We repeated this calculation of bias and RMS deviation for all
+30 models with different regularization strengths $\Lambda$.
 
@@
 improvements in the biases, however these were already minimal in the unregularized
-case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
-%3.56224789e+01
-we achieve RMS deviations that are about 30\% better than the unregularized case.
-Based on this improvement we set $\Lambda = X.XX$ for this model.  At this regularization 
+case.  The improvement in RMS reaches a minimum value near 
+$\Lambda = 35.6$ ($\approx10^{1.5}$), where
+we achieve RMS deviations that are about 10\% better than the unregularized case.
+Based on this improvement we set $\Lambda = 35.6$ for this model.  At this regularization 
 strength, the bias and RMS values found by leave-one-out
-cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
-for $\logg$, with $X.XX$~dex and $X.XX$~dex for [Fe/H].
+cross validation are, respectively: $38$~K and $256$~K for $\teff$, $0.05$~dex and 
+$0.29$~dex for $\logg$, with $0.03$~dex and $0.17$~dex for [Fe/H].
 
@@
 deviation with respect to the unregularized case.  At the point of maximum improvement 
-near $\Lambda = X.XX$, the RMS in all nine labels has decreased by up to $X$\%.  Here the regularization strength also produces a sparser matrix
+near $\Lambda = 0.13$ $(\approx10^{-0.9})$, the RMS in all nine labels has decreased by up to $30$\%,
+with all labels showing an improvement $>5$\%, and the mean improvement over all labels is about 10\%.
+Near $\Lambda \approx 10^{-0.9}$ to $10^{-0.3}$, the regularization also produces a sparser matrix
 of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
 Based on the increased model sparsity and decreasing RMS deviation in the labels, 
-we adopt $\Lambda = X.XX$ for the giant star model.
-
-Results from cross-validation on our regularized model imply that the precision
-in labels is XXX~K in $\teff$, X.XX~dex in $\logg$, and between X.XX~dex (EL) and 
-X.XX~dex (EL) depending on the elemental abundance.  We inferred labels for all
-\Nspectra\ \rave\ spectra using this model, and these results are summarized in
+we adopt $\Lambda = 0.57$ ($10^{-0.25}$) for the giant star model.  The bias in 
+labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
+$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elements.  The RMS
+at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
+between $0.07-0.09$~dex depending on the elemental abundance.  We inferred labels 
+for all \Nspectra\ \rave\ spectra using this model, and these results are summarized in
 the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible

### 2016-09-07 article/unrave.tex
Remove Chiappini; Add Munari

@@
 % RAVE DR5 core team:
-\author{Luca Casagrande, Cristina Chiappini, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
+\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
 
 % RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro}
-%Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
-%E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
-%Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
-%Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
-%Senior CIfAR Fellow; Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari}
+% Bland-Hawthorn: Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
+% Gibson: E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
+% Siebert: Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
+% Bienayme: Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
+% Navarro: Senior CIfAR Fellow; Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2
+% Munari: INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy
+
 \author{and the \rave\ collaboration}
@@
 for a given star, and we use those estimates to produce a unified estimate, but 
-this is done by disregarding any probabilistic justification.
+this joint estimate is calculated by disregarding the probabilistic attributes
+of individual estimates.
 
@@
 $\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
-model; and the differences between the red giant branch model and the joint model. We have scaled
-the differences in $\teff$ and $\logg$ to make the central peak near $(0, 0)$ to 
-% ARC to HOGG: This is a total hack. Should we fit for these and \rho_{\teff,\logg} or are you content with the hack?
-be approximately isotropic by setting
+model; and the differences between the red giant branch model and the joint model. 
+We have scaled the differences in $\teff$ and $\logg$ to make the central peak 
+near $(0, 0)$ to be approximately isotropic by setting
 	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
@@
 	$\delta_{\log{g}} = 0.15$~dex for both models.
-The stars within the peak at $(0, 0)$ represent objects where
+It is important to note that these $\delta$ values do not represent any kind of intrinsic
+uncertainty or precision: they are merely normalization factors.  Adopting
+substantially different scaling factors would produce in clear inconsistencies in our
+results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
+factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
+approximately isotropic, and to some extent, comparable.  In Figure 
+\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent objects where
 the joint model and the comparison model both report similar labels.  
-\stub{State these are empirically chosen normalization factors, and how}
 The artefacts 
@@
 from the central peak.  Therefore, we can 
-adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model,
-% ARC to HOGG: Please suggest any nomenclature or definition changes as you see fit!
+adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model to the 
+main-sequence model $d_{ms}$,
 \begin{eqnarray}
-	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2   \nonumber \\
-	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2   \quad ,
+	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
 \end{eqnarray}
 
-\noindent{}as a metric to derive the weights,
+However when adopting a similar distance metric for the giant model, we found that the
+giant model and the joint model produced similar label estimates (with low $\chi_r^2$
+values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
+in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
+could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
+as a term that would benefit the main-sequence model.  As a result, our joint estimate 
+for metal-poor dwarf stars that overlap with high-resolution studies
+\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
+would appear as metal-poor giant stars, because the giant star model would have a much
+smaller distance metric than the main-sequence model.  It is reasonable to assert that
+this issue may be complicated by the increased complexity, or flexibility, of the giant
+star model relative to the main-sequence model.  In any case, we found that by empirically
+penalizing the giant model distance metric by absolute metallicity,
+\begin{eqnarray}
+	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
+\end{eqnarray}
+
+\noindent{}the weighting of dwarfs was no longer a function of joint metallicity, and this
+issue was resolved without introducing new artefacts. Using these distance metrics we 
+then derived the weights,
 \begin{eqnarray}
@@
 
-We calculate weighted errors of $\hat\ell$ and correlation coefficients in 
-the same manner.  In Figure \ref{fig:model-weights} we show the mean relative
+We calculate weighted errors of $\hat\ell$ in the same manner.  
+In Figure \ref{fig:model-weights} we show the mean relative
 weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional bin of 
@@
 
-Our catalog now contains entries of $XXX$ stars with effective temperature $\teff$,
-surface gravity $\logg$, iron abundance [Fe/H], and $XXX$ of those stars are giants
-with elemental abundances of [O/H], [Al/H], [Mg/H], [Ca/H], [Si/H], and [Ni/H].
-In total, this leads to a catalog of X.XX million stellar abundances. 
 The weighted $\teff$ and $\logg$ values for stars meeting different
@@
 
+\stub{In Figure \ref{fig:test-set-repeats} we look at the difference as a function of SNR}
+
 \subsection{External validation}
@@
 In order to provide a fair comparison, we only show stars that meet a number
-of quality flags in \emph{both} samples. Specifically we require that the 
-$S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
-$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag from
-\citet{Kordopatis_2013} is zero, indicating that no problems were reported by the pipeline.
-There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
-and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
+of quality flags in \emph{both} samples.  Our constraints require that the
+S/N ratio exceeds 10~pixel$^{-1}$, and $\chi_r^2 < 3$.  For this comparison
+we further required that:
+the \texttt{QK} flag from \citet{Kordopatis_2013} is zero, indicating no
+problems were reported by the pipeline; $T_{{\rm eff},DR4} > 4000$~K;
+the error in radial velocity \texttt{e\_HRV} is $<$8~km~s$^{-1}$; and the three
+principal morphological flags \texttt{c1}, \texttt{c2}, \texttt{c3}, from 
+\citet{Matijevic_2012} all indicate `n' for a normal FGK-type star.
+There is good agreement in $\teff$, with a bias and RMS of just 4~K and 240~K,
+respectively. The offset in $\logg$ on the giant branch between this study and 
 \citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this has issue
@@
 their analysis on high-resolution, high S/N spectra.  We found four notable
-studies with a sufficient level of overlap: the disk studies \citet{Reddy_2003}
-and \citet{Reddy_2006}, the \citet{Valenti_Fischer_2005} work on exoplanet
-host star candidates, and the work of \citet{Bensby_2014}.  These studies 
+studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,Reddy_2006}
+and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} work on exoplanet
+host star candidates.  These studies 
 perform a careful (manual; expert) analysis using extremely high-resolution
-\stub{how high?}, high S/N spectra, and make use of \project{Hipparcos}
+
+, high S/N spectra, and make use of \project{Hipparcos}
 parallaxes where possible.  Most of the stars in these samples are main-sequence
@@
 \begin{figure*}[p]
+\caption{Distribution of differences in label estimates from multiple visits, divided by their formal errors summed in quadrature.
+If our measurements were unbiased and the formal errors were representative, these distribution should be normally distributed with zero mean and unit variance.\label{fig:formal-errors-comparison}}
+\end{figure*}
+
+\begin{figure*}[p]
 \caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:test-set-repeats}}
@@
 Here we show the `calibrated' metallicity (column \texttt{c\_M\_H\_K}) from the \rave\ survey.
-Only stars meeting all the following quality criteria are shown: $\chi_{r}^2 < 3$ (\project{unRAVE} constraint); S/N $> 10$ (\project{unRAVE} constraint); and \texttt{QK = 0} (\rave\ DR4 constraint).
-\label{fig:rave-dr4-comparison}}
+Only stars meeting quality constraints in \emph{both} studies are shown (see text for details).\label{fig:rave-dr4-comparison}}
 \end{figure*}

### 2016-09-07 article/unrave.tex
Updated text in gold standard comparison, since MP dwarf/subgiant issue now fixed

@@
 host star candidates.  These studies 
-perform a careful (manual; expert) analysis using extremely high-resolution
-
-, high S/N spectra, and make use of \project{Hipparcos}
+perform a careful (manual; expert) analysis using extremely high-resolution, high S/N spectra, and make use of \project{Hipparcos}
 parallaxes where possible.  Most of the stars in these samples are main-sequence
@@
 information content available in \rave\ (945~pixels in the near infrared with
-$\mathcal{R} \approx 7{,}500$) compared to these literature studies (\stub{what snr, resolution, wavelength range and hipparcos parallaxes}, we consider the
- agreement to be satisfactory.
-
-There are some discrepancies between our results and these literature studies,
-and these differences arise from the manner in which we join results from the
-main-sequence and giant model.  Specifically, for metal-poor stars near
-the turn-off in these literature comparisons, in our weighted results we find 
-these stars to have similar metallicities, but to be slightly more evolved: 
-on the sub-giant branch or the base of the giant branch.  This effect is best
-illustrated in Figure~\ref{fig:gold-standard-hrd}, where we show a 
-Hertzsprung-Russell diagram for all stars in common between this work and that
-of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.  While the
-overall shape of the turn-off in our work is consistent with the literature
-studies (right panel), the linked data points in the first three panels show that metal-poor stars in the literature
-appear as more evolved stars in our combined results.  Instead, if we only
-take the results from our main-sequence model --- ignoring the influence of the
-giant model --- and compare those values to these literature studies, no such 
-systematic discrepancies are present in the Hertzsprung-Russell diagram: we 
-find those metal-poor turn-off stars to also be turn-off stars, with comparable
-metallicities.  Our investigation indicates that this issue affects dwarf stars
-near the turn-off that have $[{\rm Fe/H}] \lesssim -1$.  While the agreement
-would improve if we adopted the main-sequence results for these specific stars,
-the underlying issue of metal-poor dwarf stars would still be present.  Thus,
-we use this example to illustrate caution for studies making inferences from our
-work based on main-sequence dwarf stars.  The results from both the main-sequence star model, 
-and the giant star model, are available on request. \stub{should we put them online somewhere
-for open-science, or just make them available upon request so that we don't
-muddy the waters and confuse people as to which catalog to use?}
-
-We note that this issue does not appear to affect our derived metallicities for these
-stars.  For this reason, if stars affected by this issue can be confirmed to be
-dwarfs by external information --- rather than being sub-giants as our analysis
-suggests for some --- then the metallicity we report can still be considered
-useful.  For the more metal-rich dwarf stars, the sample we present here
-will undoubtedly be informative for upcoming studies based on exoplanet host
-star properties.  At present, however, there are just $\approx$30 stars in 
-\rave\ that overlap with the compilations of exoplanet host star properties 
-listed at \texttt{exoplanets.org} and \texttt{exoplanets.eu}.
+$\mathcal{R} \approx 7{,}500$) compared to these literature studies that use
+\project{Hipparcos} parallaxes where possible, and base their inferences on
+spectra with resolving power $\mathcal{R}$ between 40,000 to 110,000, and
+S/N ratios exceeding 150, we consider the agreement to be very satisfactory.
+Indeed, given the metallicity precision available in the \unrave\ catalog, 
+these results will likely be useful for future studies based on exoplanet
+host star properties (e.g., \project{TESS}). At present, however, there are 
+just $\approx$30 stars in \rave\ that overlap with the compilations of 
+exoplanet host star properties listed at \texttt{exoplanets.org} and 
+\texttt{exoplanets.eu}.
 

### 2016-09-07 article/unrave.tex
Clean up all of the cluster text

@@
 
-After verifying that our stellar labels are comparable with other
+After verifying that our atmospheric parameters and abundances are comparable with other
 high-resolution studies, here we verify that the abundances we find
-are consistent with expectations from astrophysics.  One example is stars in
-globular clusters, which show anti-correlations in light element abundances 
+are consistent with expectations from astrophysics.  One example is globular clusters, 
+which show anti-correlations in light element abundances 
 \citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
-survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters: 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 members of NGC~362.
-In addition to these systems, 12 stars thought to belong to NGC~1851 were found in the RAVE data in \cite{Kunder_2014} and 10 stars in NGC~6752 and XX stars in \stub{47~Tuc} were discussed in \citep{Kunder_2016}.
-For a detailed description of the cluster membership for NGC~5139, NGC~3201, and NGC~362 we refer the reader to \cite{Anguiano_2015}.
-The membership selection for NGC~1851 can be found in \citep{Kunder_2014} and for NGC~6752 and 47~Tuc can be found in \cite{Kunder_2016}.
-We note here that we have not re-evaluated cluster membership as that is beyond the scope of the paper, however with the results presented here it is possible to study these systems in further detail.
-
-In the top panel of Figure~\ref{fig:globular-cluster-HRD} we plot the spectroscopic H-R diagram for the members of these globular clusters.
-For comparison, in the bottom panel of Figure~\ref{fig:globular-cluster-HRD} we show the same as the top panel but for the RAVE DR4.
-In both panels isochrones, we have also plotted the PARSEC isochrones \citep{
-Bressan_2012} for comparison.
-The basic properties of the globular cluster used to generate the isochrones were sourced from Table 2 of \cite{Kunder_2016}, \cite{Marin-Franch_2009}, and the 2010 version of the Harris Globular Cluster Catalogue \citep{Harris_1996}.This figure indicates that for some stars in specific clusters clusters, namely NGC3201 and NGC362, are slightly offset from the isochrones most likely because the star's metallicity is significantly higher than the mean cluster metallicity.
-This could be a result of a combination of the inclusion of stars that are not cluster members but appeared to be consistent with the RAVE DR4 stellar parameters.
-In the remaining globular clusters the stars are consistent with the isochrones.
-In the case of NGC5139 ($\omega$ cen), it may be possible to pick out two of the several components that exist within that cluster.
-
-All of these systems (\stub{check if this is the case for the remaining two new GCs added.}) in \cite{Anguiano_2015} are well-studied and exhibit abundance anti-correlations 
-in Na--O and Mg--Al \citep{Marino_2011,Carretta_2009,Carretta_2013,Munoz_2013}.  Other correlations are also present 
-(e.g., C--N), but those relationships are less relevant for this work because 
-we do not measure C and N abundances here.  We note that \emph{none} of these 
-globular cluster stars were present in our training set.
-
-We show the detailed abundances for these globular cluster members in
-Figure \ref{fig:globular-cluster-abundances}.  For comparison purposes we
-have included abundances from the high-resolution studies of \citep{people}.
-\stub{What do we find?}
-
-% Assuming we recover these relationships...
+survey, \cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
+consistent with being members of globular clusters: 49 stars belonging to NGC~5139
+($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 
+members of NGC~362. In addition to these systems, 12 stars thought to belong to NGC~1851
+were found in the \rave\ data by \cite{Kunder_2014}, and a further 10 stars in NGC~6752.  
+We refer the reader to those studies for details regarding the membership selection.
+
+In Figure~\ref{fig:globular-cluster-HRD} we show the effective temperature
+$\teff$ and surface gravity $\logg$ for pre-defined members of globular clusters in \rave.
+The top panels indicate measurements made in this work, and for comparison purposes we have included
+the results from the fourth \rave\ data release.  We show representative \project{PARSEC} isochrones 
+\citep{Bressan_2012} in all panels, where the isochrone ages and metallicities are adopted from
+\citet{Kunder_2016,Marin-Franch_2009}, and the \citet[][accessed 6 September 2016]{Harris_1996} 
+online catalog of globular cluster properties.  The globular cluster with the most number of
+\rave\ members is NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
+consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}.  We
+find the mean metallicity of prescribed $\omega$-Centauri members to be $[{\rm Fe/H}] = -0.85$.
+However, it is clear that the membership criteria could be improved with our
+revised metallicities and detailed chemical abundances. 
 Indeed, this demonstrates that our abundance labels can be used to identify
@@
 
-% Some literature notes:
-%check if NGC 6571 has anticorrelations present
-% See http://iopscience.iop.org/article/10.1088/0004-637X/731/1/64/pdf
-% NGC 3201 has Na-O anti-correlation: Carretta et al. (2009)
-% NGC 362 has "two discrete groups along the Na-O anti-correlation": https://arxiv.org/abs/1307.4085
-% NGC 362 has Mg-Al relationship (low Mg scatter, large Al scatter) --> Carretta (2013)
-% NGC 362 has a large metallicity spread in RAVE (Anguiano_2015) but not the case in high-res studies (e.g.
-https://arxiv.org/abs/1307.4085 has a metallicity spread of 0.05/0.08 dex) may indicate better cluster selection needed and can be done with cannon results.
-% NGC 3201 has Mg-Al relationship (http://arxiv.org/pdf/1305.3645.pdf and references therein)
-
-%Mean metallicity of NGC3201 cluster = -0.75 +/- 0.20
-%Mean metallicity of NGC362 cluster = -0.92 +/- 0.48
-%Mean metallicity of NGC5139 cluster = -0.85 +/- 0.55
-%Mean metallicity of NGC6752 cluster = -1.17 +/- 0.30
-
 \subsubsection{Open clusters}
 \label{sec:open-cluster-validation}
-The RAVE survey has observed several open clusters which can serve as a qualitative sanity check that the stellar parameters match our expectations.
-These include $\sim$160 stars within the Pleiades (78 stars), the Hyades (26 stars), M67 (30 stars) and IC4651 (13 stars) open clusters.
-Details on the open cluster membership and selection can be found in \cite{Kunder_2016}.
-The basic properties of the open cluster were sourced from Table 2 of \cite{Kunder_2016} and from \cite{Kharchenko_2013}.
-
-In Figure~\ref{fig:open-cluster-HRD}, we plot the cannon-derived H-R diagram (top panel) and compare it to the RAVE DR4 H-R diagram (bottom panel) for the stars which are likely members of the open clusters.
-In all cases, except the Pleiades, the stellar parameters derived from the Cannon are consistent with the expectations from the isochrones.
-We note that theoretical isochrones struggle to match the observed color-magnitude diagrams of young cluster, such as the Pleiades, and thus there has been an effort to use empirical isochrones for such systems \citep[e.g.][]{Bouy_2015}.
-Additionally, the improvement in the precision of the stellar parameters in the study compared to RAVE DR4 is illustrated in the top and bottom panels of Figure~\ref{fig:open-cluster-HRD} by comparing the tightness of the main-sequence in and sub-giant branch in these open cluster.
-
-We illustrate the precision in the chemical abundance by quantifying the star-to-star dispersion of  the detailed abundances for these open cluster members in
-Figure \ref{fig:open-cluster-abundances}.
-The internal cluster abundance dispersion can be used as a gauge of precision because open clusters are thought to be chemically homogenous \citep[e.g.][]{Bovy_2016}.
-\stub{What do we find?}
-
-%Mean metallicity of Hyades cluster = 0.07 +/- 0.09 (26 stars)
-%Mean metallicity of IC4651 cluster = 0.15 +/- 0.12 (13 stars)
-%Mean metallicity of M67 cluster = -0.02 +/- 0.17 (30 stars)
-%Mean metallicity of Pleiades cluster = -0.02 +/- 0.12 (78 stars)
-
-\subsubsection{Diffuse interstellar bands}
-\label{sec:dibs-validation}
-
-\stub{Look at residuals between model and data as a function of galactic position. This might be out of scope unless someone volunteers to do it}
+
+The \rave\ survey observed $\sim$160 highly probable members of several open clusters:
+78 stars in the Pleiades, 26 in Hyades, 30 stars in the Solar-metallicity cluster
+M67 and 13 stars in IC4651. Details on the open cluster membership and selection can 
+be found in \cite{Kunder_2016}.  We show the effective temperature $\teff$ and surface
+gravity $\logg$ for these stars in Figure \ref{fig:open-cluster-HRD}.  The isochrones
+are sourced from \citet{Bressan_2012}, with cluster properties adopted from 
+\cite{Kunder_2016} and from \cite{Kharchenko_2013}.  We find good agreement between
+our atmospheric parameters (top panels) and the isochrones shown.  The position
+of the red clump in IC4651 and M67 are perfectly matched to the isochrone, and the
+Hyades main-sequence is in good agreement down to $T_{\rm eff} \approx 4000$~K.
+
+The most pronounced discrepancy is seen in the main-sequence of the Pleiades.
+Like \citet{Kordopatis_2013}, we find lower $\logg$ values than what the isochrone
+shown predicts.  There are two potential reasons for this discrepancy.  Firstly,
+we note that theoretical isochrones often struggle to match the observed
+color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.
+Secondly, the lower $\logg$ values we find near the turn-off for the Pleiades could
+be explained if those stars are fast rotators. As seen in Figure \ref{fig:test-set-hrd},
+for stars with S/N ratios $>10$~pixel$^{-1}$ or $>25$~pixel$^{-1}$, we find many
+stars to sit slightly above the turn-off point.  The \rave\ pipeline identifies
+these stars to be fast rotators, consistent with the idea that they are blue
+stragglers.  Nevertheless, they are slightly higher from the turnoff point, which
+may indicate a tendency for fast-rotating turn-off stars to have systematically
+lower $\logg$ values, by $\approx$0.2~dex.
 
@@
 
-\bibitem[Mu{\~n}oz et al.(2013)]{Munoz_2013} Mu{\~n}oz, C., Geisler, D., \& Villanova, S.\ 2013, \mnras, 433, 2006
-
 \bibitem[Navin et al.(2016)]{Navin_2016} Navin, C.~A., Martell, S.~L., \& Zucker, D.~B.\ 2016, arXiv:1606.06430
@@
 \includegraphics[width=\textwidth]{figures/HRD_GCs.pdf}
-\caption{Top Panel: Cannon-derived $\logg$ as a function of $\teff$ for globular cluster members.
-The cluster name is indicated in the top left corner.
-The points are color-coded by the metallicity of the member star.
-Bottom Panel: RAVE DR4 $\logg$ as a function of $\teff$ for the same clusters in the top panel.
-The points are color-coded by the metallicity of the member star.
-In both panels, the black line represent the theoretical expectation of a single stellar population from the Padova isochrone set.
+\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of globular clusters from \citet{Anguino_2014,Kunder_2014}.
+Bottom Panels: The same stars are shown as per the top panels, where $\teff$ and $\logg$ have been sourced from the fourth \rave\ data release.
+A representative Padova \citep{Bressan_2012} isochrone is shown for each cluster in each panel.
 \label{fig:globular-cluster-HRD}}
@@
 \includegraphics[width=\textwidth]{figures/HRD_OCs.pdf}
-\caption{Same as Figure~\ref{fig:globular-cluster-HRD} for the open cluster members in the RAVE dataset.  \label{fig:open-cluster-HRD}}
+\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of open clusters from \citep{Kunder_2016}.
+The same stars are shown for each open cluster in the bottom panels, where the $\teff$ and $\logg$ entries are sourced from the fourth \rave\ data release.\label{fig:open-cluster-HRD}}
 \end{figure*}

### 2016-09-07 article/unrave.tex
Incorporate Casagrande comments. Closes #25

@@
 % RAVE DR5 core team:
-\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Marica Valentini, Jennifer Wojno, Toma{\^z} Zwitter}
+\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Jennifer Wojno, Toma{\^z} Zwitter}
 
@@
 to infer the structure, kinematics, and chemical enrichment of the Galaxy 
-\citep[e.g.,][]{Schlaufman_2009,Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,
+\citep[e.g.,][]{Nordstrom_2004,Schlaufman_2009,Deason_2011,Casagrande_2011,Ness_2012,Ness_2013a,Ness_2013b,
 Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Boeche_2013,Kordopatis_2015,Bovy_2016}.  
@@
 astrophysics, including the properties of exoplanet host stars, the formation 
-(and destruction) of stars and clusters, as well as the structure of the 
-Galactic disk, to name a few.
+(and destruction) of stars and clusters, as well as the study of stellar
+populations and Galactic structure, to name a few.
 
@@
 are sourced from \citet{Bressan_2012}, with cluster properties adopted from 
-\cite{Kunder_2016} and from \cite{Kharchenko_2013}.  We find good agreement between
-our atmospheric parameters (top panels) and the isochrones shown.  The position
-of the red clump in IC4651 and M67 are perfectly matched to the isochrone, and the
-Hyades main-sequence is in good agreement down to $T_{\rm eff} \approx 4000$~K.
+\cite{Kunder_2016} and \cite{Kharchenko_2013}.
+
+We find good agreement between our atmospheric parameters (top panels) and the 
+isochrones shown.  The position of the red clump in IC4651 and M67 are perfectly matched
+to the isochrone, and the Hyades main-sequence is in good agreement down to 
+$T_{\rm eff} \approx 4000$~K.  Similarly, we find consistency with the literature and
+our metallicity scale.  We find the mean metallicity of M67 stars to be 
+$[{\rm Fe/H}] = -0.02 \pm 0.17$~dex (30 stars), in excellent agreement with the
+expected $[{\rm Fe/H}] = 0.00$ value (not accounting for atomic diffusion).
+We further find the Hyades mean metallicity to be $[{\rm Fe/H}] = 0.07 \pm 0.09$~dex,
+consistent with \citet{Paulson_2003}: $[{\rm Fe/H}] \approx 0.13$.
+For IC4651 from 13 stars we find a mean $[{\rm Fe/H}] = 0.15 \pm 0.12$~dex, matching the
+high-resolution, high S/N study of \citet{Pasquini_2004}, where they find 
+$[{\rm Fe/H}] = 0.10 \pm 0.03$~dex. Finally, from 78 stars we find the mean
+metallicity of the Pleiades to be $[{\rm Fe/H}] = -0.02 \pm 0.12$~dex, in agreement
+with the $[{\rm Fe/H}] = -0.034 \pm 0.024$~dex measurement reported by 
+\citet{Friel_Boesgaard_1990}.
 
@@
 performed cursory (non-exhaustive) experiments to investigate this issue.
-
-\stub{Include text about fast rotators; blue stragglers}
 
@@
 
+\bibitem[Casagrande et al.(2011)]{Casagrande_2011} Casagrande, L., Sch{\"o}nrich, R., Asplund, M., et al.\ 2011, \aap, 530, A138
+
 \bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88
@@
 
+\bibitem[Friel \& Boesgaard(1990)]{Friel_Boesgaard_1990} Friel, E.~D., \& Boesgaard, A.~M.\ 1990, \apj, 351, 480
+
 \bibitem[Fulbright et al.(2010)]{Fulbright_2010} Fulbright, J.~P., Wyse, R.~F.~G., Ruchti, G.~R., et al.\ 2010, \apjl, 724, L104
@@
 
+\bibitem[Kharchenko et al.(2013)]{Kharchenko_2013} Kharchenko, N.~V., Piskunov, A.~E., Schilbach, E., R{\"o}ser, S., \& Scholz, R.-D.\ 2013, \aap, 558, A53
+
 \bibitem[Kordopatis et al.(2011)]{Kordopatis_2011} Kordopatis, G., Recio-Blanco, A., de Laverny, P., et al.\ 2011, \aap, 535, A106
@@
 
+\bibitem[Nordstr{\"o}m et al.(2004)]{Nordstrom_2004} Nordstr{\"o}m, B., Mayor, M., Andersen, J., et al.\ 2004, \aap, 418, 989
+
 \bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
+
+\bibitem[Pasquini et al.(2004)]{Pasquini_2004} Pasquini, L., Randich, S., Zoccali, M., et al.\ 2004, \aap, 424, 951
+
+\bibitem[Paulson et al.(2003)]{Paulson_2003} Paulson, D.~B., Sneden, C., \& Cochran, W.~D.\ 2003, \aj, 125, 3185
 

### 2016-09-07 article/unrave.tex
To satisfy @dfm
de-stub

@@
   \addtocounter{datetwo}{-\thedateone}%
-  \the\numexpr(\thedatetwo)\relax\space days
-}
+  \the\numexpr(\thedatetwo)\relax\space days}
 % ------------------------------------------
@@
 % Remove at submission:
-\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 September 2016 (\difftoday{2016}{9}{9}away)}}}
+\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 Sep 2016 (submission in \difftoday{2016}{9}{9})}}}
 
@@
 the same scale.  We see no systematic offset in metallicities of dwarf and giant
-stars that overlap between \rave\ and the \ges\ survey, and our examination of
-globular and open clusters suggests that \stub{if there is a systematic offset, it must
-be small.} Nevertheless, these are only verification checks based on $<1$\% of the
+stars that overlap between \rave\ and the \ges\ survey, suggesting that if there is
+a systematic offset, it must be small. Nevertheless, these are only verification checks based on $<1$\% of the
 data, and there is currently insufficient data for us to prove both models are on

### 2016-09-07 article/unrave.tex
SNR and instrument for GES overlap
Comment on [Si/H] bias in GES

@@
 data release of the \ges\ survey. These are a mix of main-sequence, 
-sub-giant and red giant branch stars.  Despite most of these stars having
+sub-giant and red giant branch stars.  About half (67) of the sample 
+were acquired with the \acronym{UVES} instrument, the other with the \acronym{GIRAFFE} spectrograph, and
+the S/N of the \ges\ spectra peaks at $\approx140$~pixel$^{-1}$.  Despite 
+most of these stars having
 relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
@@
 0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
-[Si/H] abundances relative to the \ges\ survey. The RMS deviation in each
+[Si/H] abundances relative to the \ges\ survey.  The large bias in [Si/H]
+is likely a consequence of an offset between [Si/H] abundances in \ges\ and
+\apogee, the source of our training set for giant star abundances.  
+The RMS deviation in each
 label is small for stars with [X/H] $\sim -0.5$, before increasing at lower
@@
 using individual abundances for very metal-poor giant stars in our sample.
-% TODO: Revise this last sentence after we see the globular cluster comparisons,
-%		because we are saying 'hey be careful' then perhaps saying 'look how good
-%		those same abundances are in GCs`
 

### 2016-09-07 article/unrave.tex
Add Reid; updates to text, etc

@@
 % RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari}
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid}
 % Bland-Hawthorn: Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
@@
 % Munari: INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy
+% Reid (1/2): Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia
+% Reid (1/2): Western Sydney University, Penrith South DC, NSW 1797
 
@@
 Like \citet{Kordopatis_2013}, we find lower $\logg$ values than what the isochrone
-shown predicts.  There are two potential reasons for this discrepancy.  Firstly,
-we note that theoretical isochrones often struggle to match the observed
-color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.
-Secondly, the lower $\logg$ values we find near the turn-off for the Pleiades could
-be explained if those stars are fast rotators. As seen in Figure \ref{fig:test-set-hrd},
-for stars with S/N ratios $>10$~pixel$^{-1}$ or $>25$~pixel$^{-1}$, we find many
-stars to sit slightly above the turn-off point.  The \rave\ pipeline identifies
-these stars to be fast rotators, consistent with the idea that they are blue
-stragglers.  Nevertheless, they are slightly higher from the turnoff point, which
-may indicate a tendency for fast-rotating turn-off stars to have systematically
-lower $\logg$ values, by $\approx$0.2~dex.
+shown predicts.  \stub{Hawkins: use correct isochrone from Bouy to see if this issue is resolved.}
+%  Firstly,
+%we note that theoretical isochrones often struggle to match the observed
+%color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.
 
@@
 
-We have presented an independent re-analysis of \rave\ spectra that makes near-maximal
-use of the information content for red giant branch stars, and provides a significant
-improvement in stellar parameters ($\teff$, $\logg$, [Fe/H]) over previous \rave\ data
-releases.  When combined with the \tgas\ sample, these results may amount to a powerful
+We have derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all \rave\ 
+spectra, as well as detailed abundances for red giant branch stars, based on an
+independent re-analysis of \rave\ spectra.  Our catalog provides an excellent
+complement to the current \rave\ data release \citep{Kunder_2016}.  
+When combined with the \tgas\ sample, these results may amount to a powerful
 compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
@@
 have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
-\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016}.
+\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}.
 However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
@@
 \stub{Sub-giant artefact; comment as projection}
+
+\stub{Hot stars!}
 

### 2016-09-07 article/unrave.tex
Hot stars #24; abundance limits #13; sub-giant artefact #24

@@
 However, spectra that are unusual from the joint model training set can
-still project as systematic artefacts.  An interrogation of our results
-has identified one such group of metal-poor ($[{\rm Fe/H}] \approx -0.4$)
-stars near $\teff \approx 3250$~K, which peculiarly sit above the isochrone. 
-The \rave\ morphological classifications \citep{Matijevic_2012} indicate 
-these stars have chromospheric emission, a feature not present in our 
-training set.  Based on this (and other previously discussed) examples,
-we recommend either a thorough investigation of unexpected structure\footnote{
-Before submitting to \emph{Nature}} 
-using all the available 
-information from \rave, or judicious quality controls (e.g., requiring the 
-morphological classification to be `normal') to remove these spurious projections.
-
-\stub{Sub-giant artefact; comment as projection}
-
-\stub{Hot stars!}
-
-\stub{based on our interrogation of the data, we supply a 'QC' column that uses very strict requirements to eliminate weird things: morphological classifications, SNR, R (pipeline), $\chi_r^2$, etc....}
+still project as systematic artefacts.
+
+Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
+first is a small over-density at the base of the giant branch, which is
+likely a consequence of joining the 9-label and 3-label models.  The
+second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
+present in our training set, yet there are many in the \rave\ survey.
+However, the \rave\ pre-processing pipeline (\texttt{SPARV}) performs
+template matching against grids of cool \emph{and} hot stars, and therefore
+we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
+we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
+where each hexagonal bin is colored by the \emph{maximum} temperature
+reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
+at just above the turn-off point in our results.
+
+\stub{this following paragraph needs rewording}
+Based on a thorough interrogation of our results, we have adopted two boolean
+columns to signify whether our labels meet specific quality constraints:
+one marked \texttt{QC} to indicate reliable stellar parameters ($\teff$, $\logg$, [Fe/H]), and \texttt{QC\_ABUNDANCE} to indicate reliable chemical abundances.
+For stellar atmospheric parameters to meet our quality constraints, we
+require: $\chi_r^2 < 3$; $S/N >10$~pixel$^{-1}$; and \texttt{TeffS} $< 8000$~K
+in order to remove hot stars that are outside the training set.  We
+specified the quality constraint in chemical abundances to further
+require that the iron abundance $[{\rm Fe/H}] > -1.5$, since this approximately
+corresponds to the point where individual abundances will become undetectable
+in \rave\ spectra, and the individual abundance labels will be driven by our
+constraint to place them on the mean galactic chemical evolution trend.
+These quality constraints should not be taken verbatim: judicious use of
+the morphological classifications \citep{Matijevic_2012} or metadata from
+the pre-processing pipelines is encouraged.
 

### 2016-09-07 article/unrave.tex
caption for hot star figure

@@
 
+\begin{figure}[p]
+\center
+%\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
+\caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV}.
+Hot stars are not included in our training set, and appear above the turnoff in our labels.
+For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars with $\teff > 8000$~K (as reported by \texttt{SPARV}), and other results that are known to be erroneous.\label{fig:hot-stars}}
+\end{figure}
+
 \end{document}

### 2016-09-08 article/unrave.tex
Add Helmi; Wise

@@
 % RAVE DR5 core team:
-\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Jennifer Wojno, Toma{\^z} Zwitter}
+\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Jennifer Wojno, Toma\v{z} Zwitter}
+% Zwitter:  University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia
 
 % RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid}
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid, Rosemary Wyse, Amina Helmi}
 % Bland-Hawthorn: Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
@@
 % Reid (1/2): Western Sydney University, Penrith South DC, NSW 1797
-
+% Wyse: Johns Hopkins University, Baltimore, MD, USA
+% Helmi: Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands
 \author{and the \rave\ collaboration}
@@
 resolution $\mathcal{R} \approx 28{,}000$, and yet the \galah\ project still 
-targets for S/N $\gtrsim100$.
+targets for S/N $\gtrsim100$ per resolution element.
 
@@
 \includegraphics[width=\textwidth]{figures/HRD_GCs.pdf}
-\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of globular clusters from \citet{Anguino_2014,Kunder_2014}.
+\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of globular clusters from \citet{Kunder_2014,Anguino_2015}.
 Bottom Panels: The same stars are shown as per the top panels, where $\teff$ and $\logg$ have been sourced from the fourth \rave\ data release.

### 2016-09-08 article/unrave.tex
Add repeat visits figure

@@
 \affil{Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
+
+% Gilmore, Matijevic, McMillan, Siviero, Wojno
 
@@
 
-\stub{In Figure \ref{fig:test-set-repeats} we look at the difference as a function of SNR}
+\stub{In Figure \ref{fig:repeat-visits} we look at the difference as a function of SNR}
 
@@
 \begin{figure*}[p]
-\caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:test-set-repeats}}
+\includegraphics[width=\textwidth]{figures/repeat-visits.pdf}
+\caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:repeat-visits}}
 \end{figure*}

### 2016-09-08 article/figures/plot_individual_visits.py

@@ -5,109 +5,158 @@ Compare the results from stacked spectra to individual visits.
 """
 
 import os
+import cPickle as pickle
 from astropy.table import Table, join
+from matplotlib.ticker import MaxNLocator
 
-DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
+DATA_PATH = "../../"
 
-stacked_results = Table.read(os.path.join(DATA_PATH, "stacked-spectra-v44.fits.gz"))
-individual_results = Table.read(os.path.join(DATA_PATH, "rave-tgas-v44.fits.gz"))
-stacked_results["RAVEID"] \
-    = [os.path.basename(p).split("-result-")[0] for p in stacked_results["FILENAME"]]
-
-individual_results["RAVE_OBS_ID"] \
-    = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in individual_results["FILENAME"]]
+stacked_results = Table.read(os.path.join(DATA_PATH, "stacked-unrave-v0.95.fits.gz"))
+individual_results = Table.read(os.path.join(DATA_PATH, "unrave-v0.95.fits.gz"))
 
 
-# Match the individual results to RAVE DR5.
-dr5_catalog = Table.read(
-    os.path.join(DATA_PATH, "RAVEDR5_PublicCut_20160905.csv.gz"), format="csv")
-individual_results = join(individual_results, dr5_catalog, keys=("RAVE_OBS_ID", ))
+pickled_filename = "individual-visits.pkl"
+label_names = ("TEFF", "LOGG", "FE_H", "O_H", "MG_H", "AL_H", "SI_H", "CA_H", "NI_H")
+
+if not os.path.exists(pickled_filename):
 
 
+    # Match the individual results to RAVE DR5.
+    dr5_catalog = Table.read(
+        os.path.join(DATA_PATH, "RAVEDR5_PublicCut_20160905.csv.gz"), format="csv")
+    individual_results = join(individual_results, dr5_catalog, keys=("RAVE_OBS_ID", ))
 
-label_names = ("TEFF", "LOGG", "FE_H")
+    individual_results = individual_results[individual_results["QC"]]
 
 
-# Get the list of results for the stacked spectra.
-snrs = []
-differences = {}
-for label_name in label_names:
[…]

### 2016-09-08 combine_model_results.py
Add hot stars figure

@@ -317,7 +317,7 @@ for label_name in ["TEFF", "LOGG", "FE_H"] + abundance_label_names:
 combined_table = Table(data=combined_data)
 
 
-dr5_catalog_with_correct_raveids = Table.read("../rave/RAVEDR5_PublicCut_20160905.csv.gz", format="csv")
+dr5_catalog_with_correct_raveids = Table.read("RAVEDR5_PublicCut_20160905.csv.gz", format="csv")
 keep_columns = (
     # Positional information
     "RAVE_OBS_ID", "RAdeg", "DEdeg",
@@ -336,8 +336,6 @@ keep_columns = (
     "Vrot_SPARV",
     "ZeroPointFLAG",
 
-    "logg_k",
-
     # Morphological information.
     "c1", "c2", "c3", "c4", "c5", "c6", "c7", "c8", "c9", "c10", "c11", "c12", 
     "c13", "c14", "c15", "c16", "c17", "c18", "c19", "c20")
@@ -346,7 +344,7 @@ keep_columns = (
 
 # Fix string columns.
 from astropy.table import Column
-for column in ("StdDev_HRV", "MAD_HRV", "Vrot_SPARV", "Teff_SPARV", "logg_k"):
+for column in ("StdDev_HRV", "MAD_HRV", "Vrot_SPARV", "Teff_SPARV"):
     data = np.array([[e, np.nan][e == "NULL"] for e in dr5_catalog_with_correct_raveids[column].copy()], dtype=float)
 
     index = list(dr5_catalog_with_correct_raveids.dtype.names).index(column)
@@ -395,6 +393,28 @@ for label_name in abundance_label_names:
     combined_table["E_{}".format(label_name)][metal_poor] = np.nan
     
 
+# Make a figure
+ok = (combined_table["SNR"] >= 10) * (combined_table["R_CHI_SQ"] < 3)
+fig, ax = plt.subplots()
+hexbin = ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], 
+    C=combined_table["Teff_SPARV"][ok], reduce_C_function=np.nanmax,
+    gridsize=50, vmin=4000, vmax=10000, extent=(3500, 7500, 0, 5.5),
+    cmap="plasma",)
+
+ax.set_xlim(ax.get_xlim()[::-1])
+ax.set_ylim(ax.get_ylim()[::-1])
+
[…]

### 2016-09-08 article/figures/rave_io.py
Update test set to use v0.95 w/ QC

@@ -26,55 +26,14 @@ def get_kordopatis_comparisons():
 
 def get_cannon_dr1(filename=None):
 
-    #filename = "unrave-v0.2-16b_23giant.fits.gz"
-    #filename = "unrave-v0.3-27_23giant.fits.gz"
-    #filename = "unrave-v0.4-29_23giant.fits.gz"
-    #filename = "rave-tgas-v31.fits.gz"
-    #filename = "rave-tgas-v29.fits.gz"
-    #filename = "rave-tgas-v37.fits.gz"
-    #filename = filename or "unrave-v0.7-37_36.fits.gz"
-    filename = filename or "unrave-v0.9.fits.gz"
+    filename = filename or os.path.join(DATA_PATH, "unrave-v0.95.fits.gz")
 
     rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, filename))
 
-    if "Name" not in rave_cannon_dr1.dtype.names:
-        rave_cannon_dr1["Name"] = [each.split("/")[-2] + "_" + each.split("/")[-1].split(".rvsun.")[0] + "_" + each.split(".rvsun.")[1].split("-")[0] for each in rave_cannon_dr1["FILENAME"]]
-
-    
-    if filename in ("rave-tgas-v37.fits.gz", "rave-tgas-v31.fits.gz"):
-        rave_cannon_dr1["TEFF"] = rave_cannon_dr1["EPIC_TEFF"]
-        rave_cannon_dr1["LOGG"] = rave_cannon_dr1["EPIC_LOGG"]
-        rave_cannon_dr1["FE_H"] = rave_cannon_dr1["EPIC_FEH"]
-
-        from astropy.table import join
-        rave_cannon_dr1 = join(rave_cannon_dr1, get_rave_kordopatis_dr4(), keys=("Name", ))
-
-
-    if "snr" not in rave_cannon_dr1.dtype.names and "snr_ms" in rave_cannon_dr1.dtype.names:
-        rave_cannon_dr1["snr"] = rave_cannon_dr1["snr_ms"]
-
-    if "snr" not in rave_cannon_dr1.dtype.names and "SNRK" in rave_cannon_dr1.dtype.names:
-        rave_cannon_dr1["snr"] = rave_cannon_dr1["SNRK"]
-
[…]

### 2016-09-08 article/figures/plot_dr4_comparison.py
Updating figures to v0.95 with QC

@@ -22,7 +22,9 @@ except NameError:
 else:
     print("Warning: Using pre-loaded data!")
 
-QC = (combined_table["snr"] > 10) * (combined_table["r_chi_sq"] < 3) * (combined_table["QK"] == 0)
+QC =  (combined_table["QC"]) \
+    * (combined_table["TeffK"] > 4000) * (combined_table["c1_1"] == "n") * (combined_table["c2_1"] == "n") \
+    * (combined_table["c3_1"] == "n") * (combined_table["e_HRV"] < 8)
 
 N_bins = 50
 
@@ -33,8 +35,8 @@ all_columns = [
     ("FE_H", "c_M_H_K"),
 ]
 label_limits = {
-    "TEFF": (3000, 8000),
-    "LOGG": (0, 5),
+    "TEFF": (3500, 7500),
+    "LOGG": (0, 5.5),
     "FE_H": (-2.5, 0.5)
 }
 latex_labels = (r"$T_{\rm eff}$", r"$\log{g}$", r"$[{\rm Fe/H}]$")
@@ -69,6 +71,7 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
     ax.hist2d(y[finite], x[finite], bins=(bins, bins), norm=LogNorm(),
         cmap=cmap)
 
+    ax.plot(limits, limits, c="#CCCCCC", linestyle=":")
     # Common limits, labels.
     ax.set_xlim(limits)
     ax.set_ylim(limits)
@@ -80,6 +83,9 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
     diff = y - x
     print(latex_labels, np.nanmean(diff), np.nanstd(diff))
 
+axes[0].set_xticks([4000, 5000, 6000, 7000])
+axes[0].set_yticks([4000, 5000, 6000, 7000])
+axes[-1].set_xlabel(" ".join([r"$[{\rm M/H}]$", r"$({\rm RAVE}$ ${\rm DR4})$"]))
 fig.tight_layout()
 
 fig.savefig("dr4-comparison.pdf", dpi=300)

### 2016-09-08 article/figures/plot_gold_standard_comparisons.py

@@ -41,7 +41,7 @@ except NameError: # Do you know who I am? That's Jeff Vader!
     rave_cannon_dr1 = rave_cannon_dr1[OK].filled()
     """
 
-    ok = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3)
+    ok = rave_cannon_dr1["QC"]
     rave_cannon_dr1 = rave_cannon_dr1[ok].filled()
 
     bensby_2014 = join(rave_cannon_dr1, get_literature_bensby(), keys=("Name", ))
@@ -120,7 +120,7 @@ comparison_labels = (
     r"${\rm Bensby}$ ${\rm et}$ ${\rm al.}$ $(2014)$",
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2003)$", 
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2006)$", 
-    r"${\rm Valenti}$ ${\rm \&}$ ${\rm Fischer}$ $(2005)$") 
+    r"${\rm Valenti}$ ${\rm +}$ ${\rm Fischer}$ $(2005)$") 
 
 label_names = ("TEFF", "LOGG", "FE_H")
 N = len(label_names)
@@ -158,7 +158,7 @@ ylims = {
     "FE_H": 1
 }
 limits = {
-    "TEFF": (4000, 7000),
+    "TEFF": (3500, 7500),
     "LOGG": (2.5, 5.5),
     "FE_H": (-2.5, 0.5)
 }
@@ -227,8 +227,8 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
         ax.set_ylabel(latexify(label_name, "y"))
         
 
-        ax.xaxis.set_major_locator(MaxNLocator(5))
-        ax.yaxis.set_major_locator(MaxNLocator(5))
+        ax.xaxis.set_major_locator(MaxNLocator(6))
+        ax.yaxis.set_major_locator(MaxNLocator(6))
 
 
 fig.legend(handles, comparison_labels, loc="upper center", 

### 2016-09-08 article/figures/plot_gold_standard_hrd.py

@@ -41,7 +41,8 @@ except NameError: # Do you know who I am? That's Jeff Vader!
     rave_cannon_dr1 = rave_cannon_dr1[OK].filled()
     """
     
-    ok = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3)
+    #ok = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3)
+    ok = rave_cannon_dr1["QC"]
     rave_cannon_dr1 = rave_cannon_dr1[ok].filled()
 
     bensby_2014 = join(rave_cannon_dr1, get_literature_bensby(), keys=("Name", ))
@@ -122,7 +123,7 @@ comparison_labels = (
     r"${\rm Bensby}$ ${\rm et}$ ${\rm al.}$ $(2014)$",
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2003)$", 
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2006)$", 
-    r"${\rm Valenti}$ ${\rm \&}$ ${\rm Fischer}$ $(2005)$") 
+    r"${\rm Valenti}$ ${\rm +}$ ${\rm Fischer}$ $(2005)$") 
 
 label_names = ("TEFF", "LOGG", "FE_H")
 N = len(label_names)
@@ -152,7 +153,7 @@ def latexify(label_name, axis):
 comparison_labels = (
     r"${\rm Bensby}$ ${\rm et}$ ${\rm al.}$ $(2014)$",
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2003,2006)$", 
-    r"${\rm Valenti}$ ${\rm \&}$ ${\rm Fischer}$ $(2005)$",
+    r"${\rm Valenti}$ ${\rm +}$ ${\rm Fischer}$ $(2005)$",
     #r"${\rm Kordopatis}$ ${\rm et}$ ${\rm al.}$ $(2013{\rm ;}$ ${\rm RAVE}$ ${\rm DR4})$",
     r"${\rm unRAVE}$"
     )
@@ -252,7 +253,7 @@ for ax, label in zip(axes, comparison_labels):
     #ax.set_xlim(xlim.max(), xlim.min())
     #ax.set_ylim(ylim.max(), ylim.min())
     
-    ax.set_xlim(7000, 4250)
[…]

### 2016-09-08 article/figures/plot_ges_comparison.py
Update GES comparison v0.95 w/ QC and errs

@@ -47,8 +47,7 @@ latex_labels = {
 
 # Exclude ones we think are bad.
 #ok = data_table["OK"] 
-ok = (data_table["r_chi_sq"] < 3) * (data_table["snr"] > 15)
-
+ok = data_table["QC"]
 #ok *= np.isfinite(data_table["TEFF_1"] * data_table["TEFF_2"] * data_table["LOGG_1"] * data_table["LOGG_2"] * data_table["FE_H"] * data_table["FEH"])
 
 data_table = data_table[ok]
@@ -72,8 +71,11 @@ fig.subplots_adjust(
 for i, (ax, label_name) in enumerate(zip(axes, label_names)):
 
     y = data_table[label_name]
-    #xerr
+    xerr = data_table["E_{}".format(label_name)]
+    xerr[xerr == 0] = np.nan
+
     x = data_table[ges_label_names[label_name]]
+    yerr = data_table["E_{}".format(ges_label_names[label_name])]
     # yerr
     c = data_table["snr"]
 
@@ -81,6 +83,7 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
 
     #scat = ax.scatter(x[uves], y[uves], c=data_table["snr"][uves], s=75, marker="s", vmin=np.nanmin(data_table["snr"]), vmax=np.nanmax(data_table["snr"]), cmap="plasma")
     #ax.scatter(x[~uves], y[~uves], c=data_table["snr"][~uves], s=75, marker="o", vmin=np.nanmin(data_table["snr"]), vmax=np.nanmax(data_table["snr"]), cmap="plasma") 
+    ax.errorbar(x, y, xerr=xerr, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
     scat = ax.scatter(x, y, c=data_table["snr"], s=75, cmap="plasma")
 
     limits = label_limits[label_name]

### 2016-09-08 article/figures/plot_ges_abundances.py
Update figures

@@ -28,12 +28,6 @@ except NameError: # Do you know who I am? That's Jeff Vader!
     rave_cannon_dr1 = join(rave_cannon_dr1, get_ges_idr4(), keys=("Name", ))
 
 
-# Produce error columns from COV matrix.
-if "COV" in rave_cannon_dr1.dtype.names:
-
-    for i, label_name in enumerate(rave_cannon_dr1.dtype.names[1:10]):
-        rave_cannon_dr1["E_{}".format(label_name)] = rave_cannon_dr1["COV"][:, i, i]**0.5
-
 
 asplund_2009_solar_abundances = {
     "O": 8.69, 
@@ -47,7 +41,8 @@ asplund_2009_solar_abundances = {
 
 ges_species = ("O_1", "AL_1", "MG_1", "CA_1", "SI_1", "NI_1")
 
-ok  = (rave_cannon_dr1["snr"] > 10)
+ok  = rave_cannon_dr1["QC"]
+
 
 N, M = (3, 2)
 assert N * M == len(ges_species)
@@ -78,9 +73,9 @@ for i, (ax, species) in enumerate(zip(axes.flatten(), ges_species)):
     y = rave_cannon_dr1["{}_H".format(element)]
 
     xerr = rave_cannon_dr1["E_{}".format(species.replace("_", ""))]
-    #yerr = rave_cannon_dr1["E_{}_H".format(element)]
-    #ax.errorbar(x[ok], y[ok], xerr[ok], yerr[ok], 
-    #    fmt=None, ecolor="#666666", zorder=-1)
+    yerr = rave_cannon_dr1["E_{}_H".format(element)]
+    ax.errorbar(x[ok], y[ok], xerr[ok], yerr[ok], 
+        fmt=None, ecolor="#666666", zorder=-1)
     scat = ax.scatter(x[ok], y[ok], c=rave_cannon_dr1["snr"][ok], s=50,
         vmin=np.nanmin(rave_cannon_dr1["snr"]),
         vmax=np.nanmax(rave_cannon_dr1["snr"]),

### 2016-09-08 article/figures/plot_gold_standard_comparisons.py

@@ -68,7 +68,10 @@ def bensby_2014_data(cannon_label):
     label_name = {
         "TEFF": "TeffH",
         "LOGG": "loggH",
-        "FE_H": "Fe_H"
+        "FE_H": "Fe_H",
+        "E_TEFF": "e_Teff",
+        "E_LOGG": "e_logg",
+        "E_FE_H": "e_Fe_H"
     }.get(cannon_label, cannon_label)
 
     return get_data(bensby_2014, cannon_label, label_name)
@@ -79,10 +82,14 @@ def reddy_2003_data(cannon_label):
     label_name = {
         "TEFF": "Teff",
         "LOGG": "logg",
-        "FE_H": "__Fe_H_"
+        "FE_H": "__Fe_H_",
     }.get(cannon_label, cannon_label)
+    # No errors
+    a, b =  get_data(reddy_2003, cannon_label, label_name)
 
-    return get_data(reddy_2003, cannon_label, label_name)
+    if cannon_label.startswith("E_"):
+        a = np.nan * np.ones(len(a))
+    return (a, b)
 
 
 def reddy_2006_data(cannon_label):
@@ -93,7 +100,11 @@ def reddy_2006_data(cannon_label):
         "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
 
-    return get_data(reddy_2006, cannon_label, label_name)
+    # No errors..
+    a, b = get_data(reddy_2006, cannon_label, label_name)
+    if cannon_label.startswith("E_"):
+        a = np.nan * np.ones(len(a))
+    return a, b
 
 
 def valenti_2005_data(cannon_label):
@@ -103,8 +114,12 @@ def valenti_2005_data(cannon_label):
         "LOGG": "log_g_",
         "FE_H": "__Fe_H_"
     }.get(cannon_label, cannon_label)
+    # No errors...
 
-    return get_data(valenti_2005, cannon_label, label_name)
+    a, b =  get_data(valenti_2005, cannon_label, label_name)
+    if cannon_label.startswith("E_"):
+        a = np.nan * np.ones(len(a))
+    return a, b
 
 
 
@@ -169,10 +184,12 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
     diffs = []
     for comparison, kwds in comparisons.items():
         x, y = comparison(label_name) 
+        xerr, yerr = comparison("E_{}".format(label_name))
         c0, c1 = comparison("snr")
[…]

### 2016-09-08 article/figures/plot_gold_standard_hrd.py

@@ -244,6 +244,10 @@ for comp in (bensby_2014_data, reddy_2003_data, reddy_2006_data, valenti_2005_da
     _, y = comp("LOGG")
     _, c = comp("FE_H")
 
+    _, xerr = comp("E_TEFF")
+    _, yerr = comp("E_LOGG")
+    print(comp, np.nanmax(yerr))
+    axes[-1].errorbar(x, y, xerr=xerr, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
     axes[-1].scatter(x, y, c=c, **kwds)
 
 

### 2016-09-08 article/figures/plot_hrd_test.py

@@ -49,8 +49,8 @@ for i, snr in enumerate(snrs):
     # On top figure show density.
     # On bottom axes show metallicity.
     
-    ok = (t["snr"] > snr) * (t["r_chi_sq"] < 3) #* (t["VSINI"] < 1)
-    
+    ok = cannon_dr1["QC"]
+
     ax = axes[0, i]
 
     top_hexbin = ax.hexbin(t["TEFF"][ok], t["LOGG"][ok], 

### 2016-09-08 article/unrave.tex
Add hot stars figure
Update globular cluster figure #29

@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/HRD_GCs.pdf}
-\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of globular clusters from \citet{Kunder_2014,Anguino_2015}.
-Bottom Panels: The same stars are shown as per the top panels, where $\teff$ and $\logg$ have been sourced from the fourth \rave\ data release.
-A representative Padova \citep{Bressan_2012} isochrone is shown for each cluster in each panel.
-\label{fig:globular-cluster-HRD}}
+\center
+\includegraphics[height=\textheight]{figures/globular-clusters.pdf}
+\caption{Effective temperature $\teff$ and surface gravity $\logg$ for globular cluster members identified by \citet{Kunder_2014,Anguiano_2015}.
+Left-hand panels indicate results from the fourth \rave\ data release \citep{Kordopatis_2013}, and the right-hand panels show results from this work.
+A representative isochrone is shown for each cluster \citep{Bressan_2012}.\label{fig:globular-cluster-HRD}}
 \end{figure*}
@@
 \center
-%\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
+\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
 \caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV}.

### 2016-09-08 article/figures/plot_clusters_properly.py

@@ -1,0 +1,160 @@
+
+
+"""
+Plot clusters (properly).
+"""
+
+import os
+from astropy.table import Table, join
+import matplotlib.pyplot as plt
+
+try:
+    unrave, dr4
+
+except NameError:
+    from rave_io import get_cannon_dr1, get_rave_kordopatis_dr4
+
+    clusters = Table.read("RAVEDR4_GC_v3.fits")
+
+    unrave = join(get_cannon_dr1(), clusters, keys=("Name",))
+    dr4 = join(get_rave_kordopatis_dr4(), clusters, keys=("Name",))
+
+else:
+    print("Warning: Using pre-loaded data")
+
+
+
+# QC and labels for DR4
+dr4_ok = np.ones(len(dr4), dtype=bool)
+dr4_teff, dr4_logg, dr4_feh, dr4_e_teff, dr4_e_logg, dr4_e_feh \
+    = ("TeffK_1", "loggK_1", "c_M_H_K_1", "e_TeffK_1", "e_loggK_1", "e__M_H_K_1")
+
+
+unrave_ok = unrave["QC"]
+unrave_teff, unrave_logg, unrave_feh, unrave_e_teff, unrave_e_logg, unrave_e_feh \
+    = ("TEFF", "LOGG", "FE_H", "E_TEFF", "E_LOGG", "E_FE_H")
+
+iso_teff, iso_logg = ("Teff", "logG")
+
+
+
+
+
+
+kwds = dict(cmap="plasma", vmin=-2.5, vmax=0.5, s=80)
+
+cluster_names = sorted(set(clusters["Cluster"]))
+K = len(cluster_names)
+
+
+factor = 3.5
+lbdim = 0.2 * factor
+rdim = 0.1 * factor
+tdim = -0.20 * factor
+whspace = 0.05
+xspace = factor * 2 + whspace
+yspace = factor * K + factor * (K - 1) * whspace + lbdim * (K - 1)
+xdim = lbdim + xspace + rdim
+ydim = lbdim + yspace + tdim
+
+fig, axes = plt.subplots(K, 2, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
+
[…]

### 2016-09-08 article/unrave.tex
Updated open cluster figures

@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/HRD_OCs.pdf}
-\caption{Top panels: Effective temperature $\teff$ and surface gravity $\logg$ from this work for prescribed members of open clusters from \citep{Kunder_2016}.
-The same stars are shown for each open cluster in the bottom panels, where the $\teff$ and $\logg$ entries are sourced from the fourth \rave\ data release.\label{fig:open-cluster-HRD}}
+\center
+\includegraphics[height=\textheight]{figures/open-clusters.pdf}
+\caption{Effective temperature $\teff$ and surface gravity $\logg$ for open cluster members identified by \citep{Kunder_2016}.
+Left-hand panels indicate results from the fourth \rave\ data release \citep{Kordopatis_2013}, and the right-hand panels show results from this work.
+A representative isochrone is shown for each cluster \citep{Bressan_2012}.\label{fig:open-cluster-HRD}}
 \end{figure*}

### 2016-09-08 article/figures/plot_globular_clusters.py

@@ -75,11 +75,11 @@ for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
 
     # Show LHS RAVE DR4
     ax = row[0]
-    ax.text(0.05, 0.92, r"${{\rm {0}}}$".format(cluster_name.strip().replace("NGC", "NGC\,")),
-        horizontalalignment="left", transform=ax.transAxes)
+    ax.text(0.15, 0.82, r"${{\rm {0}}}$".format(cluster_name.strip().replace("NGC", "NGC\,")),
+        horizontalalignment="left", transform=ax.transAxes, fontsize=16)
     Nstars = np.isfinite(dr4[dr4_teff][match] * dr4[dr4_logg][match]).sum()
-    ax.text(0.05, 0.84, r"${0}$ ${{\rm stars}}$".format(Nstars),
-        horizontalalignment="left", transform=ax.transAxes)
+    ax.text(0.15, 0.74, r"${0}$ ${{\rm stars}}$".format(Nstars),
+        horizontalalignment="left", transform=ax.transAxes, fontsize=16)
     
     ax.errorbar(dr4[dr4_teff][match], dr4[dr4_logg][match],
         xerr=dr4[dr4_e_teff][match], yerr=dr4[dr4_e_logg][match], 
@@ -96,11 +96,11 @@ for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
 
     # Show RHS unRAVE
     ax = row[1]
-    ax.text(0.05, 0.92, r"${{\rm {0}}}$".format(cluster_name.strip().replace("NGC", "NGC\,")),
-        horizontalalignment="left", transform=ax.transAxes)
+    ax.text(0.15, 0.82, r"${{\rm {0}}}$".format(cluster_name.strip().replace("NGC", "NGC\,")),
+        horizontalalignment="left", transform=ax.transAxes, fontsize=16)
     Nstars = np.isfinite(unrave[unrave_teff][match] * unrave[unrave_logg][match]).sum()
-    ax.text(0.05, 0.84, r"${0}$ ${{\rm stars}}$".format(Nstars),
-        horizontalalignment="left", transform=ax.transAxes)
+    ax.text(0.15, 0.74, r"${0}$ ${{\rm stars}}$".format(Nstars),
+        horizontalalignment="left", transform=ax.transAxes, fontsize=16)
 
[…]

### 2016-09-08 article/figures/plot_open_clusters.py

@@ -1,0 +1,160 @@
+
+
+"""
+Plot clusters (properly).
+"""
+
+import os
+from astropy.table import Table, join
+import matplotlib.pyplot as plt
+
+try:
+    unrave, dr4
+
+except NameError:
+    from rave_io import get_cannon_dr1, get_rave_kordopatis_dr4
+
+    clusters = Table.read("RAVEDR4_OC.fits")
+
+    unrave = join(get_cannon_dr1(), clusters, keys=("Name",))
+    dr4 = join(get_rave_kordopatis_dr4(), clusters, keys=("Name",))
+
+else:
+    print("Warning: Using pre-loaded data")
+
+
+
+# QC and labels for DR4
+dr4_ok = np.ones(len(dr4), dtype=bool)
+dr4_teff, dr4_logg, dr4_feh, dr4_e_teff, dr4_e_logg, dr4_e_feh \
+    = ("TeffK_1", "loggK_1", "c_M_H_K_1", "e_TeffK_1", "e_loggK_1", "e__M_H_K_1")
+
+
+unrave_ok = unrave["QC"]
+unrave_teff, unrave_logg, unrave_feh, unrave_e_teff, unrave_e_logg, unrave_e_feh \
+    = ("TEFF", "LOGG", "FE_H", "E_TEFF", "E_LOGG", "E_FE_H")
+
+iso_teff, iso_logg = ("Teff", "logG")
+
+
+
+
+
+
+kwds = dict(cmap="plasma", vmin=-0.5, vmax=0.5, s=80)
+
+cluster_names = sorted(set(clusters["Cluster"]))
+K = len(cluster_names)
+
+
+factor = 3.5
+lbdim = 0.2 * factor
+rdim = 0.1 * factor
+tdim = -0.20 * factor
+whspace = 0.05
+xspace = factor * 2 + whspace
+yspace = factor * K + factor * (K - 1) * whspace + lbdim * (K - 1)
+xdim = lbdim + xspace + rdim
+ydim = lbdim + yspace + tdim
+
+fig, axes = plt.subplots(K, 2, figsize=(xdim, ydim))
+fig.subplots_adjust(
+    left=lbdim/xdim, bottom=lbdim/ydim, right=(xspace + lbdim)/xdim,
+    top=(yspace + lbdim)/ydim, wspace=whspace, hspace=whspace)
+
+
+for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
+
[…]

### 2016-09-08 article/figures/plot_kordopatis_calibration_sample.py
Update kordopatis calibration sample figure

@@ -30,7 +30,7 @@ else:
     print("Warning: Using pre-loaded data.")
 
 
-ok = (data_table["r_chi_sq"] < 3)# * (data_table["R"] > 10)
+ok = data_table["QC"]# * (data_table["R"] > 10)
 
 latex_labels = {
     "TEFF_2": r"$T_{\rm eff}$ $[{\rm K}]$ $({\rm Literature})$",
@@ -46,9 +46,9 @@ cannon_labels = ("TEFF_1", "LOGG_1", "FE_H")
 literature_labels = ("TEFF_2", "LOGG_2", "FEH")
 
 limits = {
-    "TEFF_1": [3500, 7000],
-    "LOGG_1": [0, 5],
-    "FE_H": [-4.5, 0.75]
+    "TEFF_1": [3500, 7500],
+    "LOGG_1": [0, 5.5],
+    "FE_H": [-3.5, 0.75]
 }
 
 kwds = dict(cmap="plasma", vmin=np.nanmin(data_table["snr"]), vmax=np.nanmax(data_table["snr"]))
@@ -78,7 +78,12 @@ in enumerate(zip(axes, cannon_labels, literature_labels)):
     y = data_table[cannon_label]
     c = data_table["snr"]
 
-    scat = ax.scatter(x[ok], y[ok], c=c[ok], s=50, alpha=0.5, **kwds)
+    #xerr = data_table["e_{}".format(literature_label)]
+    yerr = data_table["E_{}".format(cannon_label).strip("_1")]
+
+    ax.errorbar(x[ok], y[ok], yerr=yerr[ok], fmt=None, ecolor="#666666",
+        zorder=-1)
+    scat = ax.scatter(x[ok], y[ok], c=c[ok], s=50, **kwds)
 
 
 _ = ax.scatter([-999], [-9999], c=[0], **kwds)
@@ -99,6 +104,12 @@ for ax, cannon_label, literature_label in zip(axes, cannon_labels, literature_la
     ax.set_xlabel(latex_labels[literature_label])
     ax.set_ylabel(latex_labels[cannon_label])
 
+
+axes[0].set_xticks([4000, 5000, 6000, 7000])
+axes[0].set_yticks([4000, 5000, 6000, 7000])
+
+axes[-1].set_xticks([-3.5, -2.5, -1.5, -0.5, 0.5])
+axes[-1].set_yticks([-3.5, -2.5, -1.5, -0.5, 0.5])
 
 fig.tight_layout()
 

### 2016-09-08 article/figures/plot_gce.py
Update GCE figure with v0.95 and QC

@@ -32,16 +32,17 @@ x = rave_cannon_dr1["FE_H"]
 #ok  = (rave_cannon_dr1["snr"] > 10) * (rave_cannon_dr1["r_chi_sq"] < 3) \
 #    * (rave_cannon_dr1["R"] > 10) * (rave_cannon_dr1["loggK"] < 3)
 
-ok = (rave_cannon_dr1["snr"] > 10)
+ok = rave_cannon_dr1["QC"]
 
-default_kwds = dict(gridsize=50,  cmap="plasma", norm=LogNorm())
+default_kwds = dict(gridsize=(31, 9), cmap="Blues", norm=LogNorm(),
+    linewidths=0.1, edgecolor="#ffffff", rasterized=True)
 elements_and_kwds = OrderedDict([
-    ("O", dict(extent=(-1.5, 0.5, -0.05, 0.4))),
-    ("AL", dict(extent=(-1.5, 0.5, -0.3, 0.5))),
-    ("MG", dict(extent=(-1.5, 0.5, -0.3, 0.4))),
-    ("CA", dict(extent=(-1.5, 0.5, -0.3, 0.4))),
-    ("SI", dict(extent=(-1.5, 0.5, -0.1, 0.6))),
-    ("NI", dict(extent=(-1.5, 0.5, -0.3, 0.2)))
+    ("O", dict(extent=(-1.2, 0.5, -0.05, 0.4))),
+    ("AL", dict(extent=(-1.2, 0.5, -0.3, 0.5))),
+    ("MG", dict(extent=(-1.2, 0.5, -0.1, 0.4))),
+    ("CA", dict(extent=(-1.2, 0.5, -0.3, 0.4))),
+    ("SI", dict(extent=(-1.2, 0.5, -0.1, 0.6))),
+    ("NI", dict(extent=(-1.2, 0.5, -0.3, 0.2)))
 ])
 
 
@@ -65,15 +66,15 @@ for i, (element, updated_kwds) in enumerate(elements_and_kwds.items()):
     #    ],
     #    **kwds)
     cmap = plt.cm.Blues
-    extent = kwds.pop("extent")
+    extent = kwds.get("extent")
     xbins = np.linspace(extent[0], extent[1], 50)#np.ptp(extent[:2])/0.04)
     ybins = np.linspace(extent[2], extent[3], 20)#np.ptp(extent[2:])/0.04)
 
     show = np.isfinite(x * y)
[…]

### 2016-09-08 article/unrave.tex
Give up on hexbin for GCE; add SPARV citations

@@
 present in our training set, yet there are many in the \rave\ survey.
-However, the \rave\ pre-processing pipeline (\texttt{SPARV}) performs
+However, the \rave\ pre-processing pipeline \citep[\texttt{SPARV};][]{Steinmetz_2006,Zwitter_2008} performs
 template matching against grids of cool \emph{and} hot stars, and therefore
@@
 \includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
-\caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV}.
+\caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV} \citep{Steinmetz_2006,Zwitter_2008}.
 Hot stars are not included in our training set, and appear above the turnoff in our labels.

### 2016-09-08 article/figures/plot_gce.py

@@ -34,8 +34,7 @@ x = rave_cannon_dr1["FE_H"]
 
 ok = rave_cannon_dr1["QC"]
 
-default_kwds = dict(gridsize=(31, 9), cmap="Blues", norm=LogNorm(),
-    linewidths=0.1, edgecolor="#ffffff", rasterized=True)
+default_kwds = dict(gridsize=(31, 9), cmap="Blues", norm=LogNorm())
 elements_and_kwds = OrderedDict([
     ("O", dict(extent=(-1.2, 0.5, -0.05, 0.4))),
     ("AL", dict(extent=(-1.2, 0.5, -0.3, 0.5))),
@@ -71,8 +70,8 @@ for i, (element, updated_kwds) in enumerate(elements_and_kwds.items()):
     ybins = np.linspace(extent[2], extent[3], 20)#np.ptp(extent[2:])/0.04)
 
     show = np.isfinite(x * y)
-    #ax.hist2d(x[ok * show], y[ok * show], bins=(xbins, ybins), norm=LogNorm(), cmap=cmap)
-    ax.hexbin(x[ok], y[ok], **kwds)
+    ax.hist2d(x[ok * show], y[ok * show], bins=(xbins, ybins), norm=LogNorm(), cmap=cmap)
+    #ax.hexbin(x[ok], y[ok], **kwds)
     ax.axhline(0, c="k", linewidth=2, linestyle="--")
     ax.axvline(0, c="k", linewidth=2, linestyle="--")
 

### 2016-09-08 article/unrave.tex
Update numbers with v0.95

@@
 \newcommand{\Nstars}{457,589}
-\newcommand{\Nstarsqc}{434,470}
+\newcommand{\Nstarsqc}{447,207} % v0.95 finite values
+\newcommand{\Nabundances}{1,692,825} %v0.95 finite values
 
@@
 We have presented the largest collection of stellar abundances for stars in the 
-first \gaia\ data release.  In total more than X.XX million abundances are
+first \gaia\ data release.  In total more than 1.69 million abundances are
 available, with effective temperature $\teff$, surface gravity $\logg$ and 
-metallicity available for XXXXXX stars.  Despite the caveats and limitations 
+metallicity available for \Nstarsqc\ stars.  Despite the caveats and limitations 
 discussed here, our validation experiments and comparisons with high-resolution

### 2016-09-08 article/unrave.tex
Corrected typos

--- 2016-09-08 article/unrave.tex
@@
 \galah\ \citep{DeSilva_2015}, \lamost\ \citep{Cui_2012}, and \rave\ 
-\citep{Steinmetz_2006} surveys}, we found that the RAdial Velocity Experiment 
+\citep{Steinmetz_2006} surveys.}, we found that the RAdial Velocity Experiment 
 (\rave) survey is expected to have the largest overlap with the first \gaia\ 
@@
 Observations were conducted on the 1.2~m UK Schmidt telescope at the Australian 
-Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
+Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory.} from 
 2003--2013.  A large 5.7~degree field-of-view and robotic fibre positioner made for 
@@
 (e.g., stellar parameters and chemical abundances) than most physics-driven 
-models\footnote{However, see \citet{Casey_2016a}}.  This is particularly relevant 
+models\footnote{However, see \citet{Casey_2016a}.}.  This is particularly relevant 
 for the low-resolution \rave\ data analysed here, because about half of the 
@@
 stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
-but that sample contains only giant stars}, and the \lamost\ lower main-sequence suffers
+but that sample contains only giant stars.}, and the \lamost\ lower main-sequence suffers
 from the same systematic effects seen in \apogee\ data.
@@
 The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
-The increased density of Solar-type stars is consistent with \rave\
+The increased density of solar-type stars is consistent with \rave\
 observing stars in the local neighbourhood, and the high number of turn-off
@@
 
-The red giant branch stars in our training set have have stellar parameters 
+The red giant branch stars in our training set have stellar parameters 
 ($\teff$, $\logg$) and up to 15 elemental abundances from the \aspcap\ 
@@
 respectively. The offset in $\logg$ on the giant branch between this study and 
-\citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this has issue
+\citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this issue
 has been resolved in the fifth \rave\ data release by correcting $\logg$
@@
 \gaia\ benchmark stars.  There is also a
-slight discrepancy in the $\log$ values along the main-sequence, where our
+slight discrepancy in the $\logg$ values along the main-sequence, where our
 work tends to taper down towards higher $\logg$ values at cooler temperatures,
@@
 We only include stars with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter
-cut removed only few stars because the average S/N in the \rave\ spectra for
+cut removed only a few stars because the average S/N in the \rave\ spectra for
 these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$). When considering the relative 
@@
 The RMS deviation in each
-label is small for stars with [X/H] $\sim -0.5$, before increasing at lower
+label is small for stars with [X/H] $> -0.5$, before increasing at lower
 metallicities.  If we consider all stars, the smallest abundance RMS we see
@@
 The \rave\ survey observed $\sim$160 highly probable members of several open clusters:
-78 stars in the Pleiades, 26 in Hyades, 30 stars in the Solar-metallicity cluster
+78 stars in the Pleiades, 26 in Hyades, 30 stars in the solar-metallicity cluster
 M67 and 13 stars in IC4651. Details on the open cluster membership and selection can 
@@
 regions of parameter space that we could consider as being normal 
-(e.g., Solar-type stars).  We addressed this issue for the main-sequence
+(e.g., solar-type stars).  We addressed this issue for the main-sequence
 and giant models by using a joint model to calculate relative weights.
@@
 \caption{Detailed chemical abundances ([X/Fe]) for giant stars in \unrave\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
-Note that the y-axes limits vary for each panel, however for clarity we show the scaled-Solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
+Note that the y-axis limits vary for each panel, however for clarity we show the scaled-solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
 Only stars with S/N $> 10$ and $\chi_r^2 < 3$ are shown.\label{fig:gce}}

### 2016-09-08 article/unrave.tex
Implement MKN suggestion. Closes #27

@@
 with a low-order polynomial in label space.
-\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
-Note that we do not assume that \emph{every} label in the training is accurate -- we
+\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels
+for most, but not all stars. That is to say that we do not assume that \emph{every} 
+label in the training is accurate -- we
 can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
@@
 
-We constructed a three-label quadratic model using only main-sequence and sub-giant
-stars. In order to set the regularization hyperparameter
-$\Lambda$ for this model, we trained 30 models with different regularization strengths,
-spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
-We then performed leave-one-out cross-validation for each model.  Specifically, for 
-each star in the training set: we removed the star; trained the model; and then 
-inferred labels from the removed star as if it was a test object. We also performed 
-leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
-will use as the basis for comparison.  For the unregularized case, we calculated the 
-bias and root-mean-square (RMS) deviation between: the training set labels, and the 
-labels we derive by cross-validation, where one star is removed at a time and the
-model is re-trained.  We repeated this calculation of bias and RMS deviation for all
-30 models with different regularization strengths $\Lambda$.
-
-We show the \emph{percentage difference} in the RMS deviation of the labels with respect
-to the unregularized model in Figure \ref{fig:set-hyperparameters}.  The upper 
-and lower envelope represent the boundaries across all labels, showing that with increasing
-regularization, the RMS decreased in \emph{all} labels.  We found similar
-improvements in the biases, however these were already minimal in the unregularized
-case.  The improvement in RMS reaches a minimum value near 
-$\Lambda = 35.6$ ($\approx10^{1.5}$), where
-we achieve RMS deviations that are about 10\% better than the unregularized case.
-Based on this improvement we set $\Lambda = 35.6$ for this model.  At this regularization 
-strength, the bias and RMS values found by leave-one-out
-cross validation are, respectively: $38$~K and $256$~K for $\teff$, $0.05$~dex and 
-$0.29$~dex for $\logg$, with $0.03$~dex and $0.17$~dex for [Fe/H].
-
-We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
-The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
-The increased density of solar-type stars is consistent with \rave\
-observing stars in the local neighbourhood, and the high number of turn-off
-and main-sequence stars relative to the sub-giant branch is expected from 
-the relative lifetimes of these evolutionary phases.  An over-density
-of stars near the base of the giant branch is also present.  This artefact is due 
-to having giant stars in the test set, but not in the training set, and the
-model is (poorly) extrapolating outside the convex hull of the training set.
-
-\subsection{A 9-label model for detailed abundances of giant stars}
-\label{sec:evolved-star-model}
-
-The red giant branch stars in our training set have stellar parameters 
-($\teff$, $\logg$) and up to 15 elemental abundances from the \aspcap\ 
-\citep{Garcia_Perez_2016}.  A subset of these elements have atomic transitions in the 
-\rave\ wavelength region: \ion{O}{1}, \ion{Mg}{1}, \ion{Al}{1}, \ion{Si}{1}, 
-\ion{Ca}{2}, \ion{Ti}{1}, \ion{Fe}{1}, and \ion{Ni}{1}.  However, we 
-excluded [Ti/H] from our abundance list because of systematics in the \aspcap\
-[Ti/H] abundances \citep{Holtzman_2015,Hawkins_2016}.  Therefore we are left 
-with nine labels in our giant star model: $\teff$, $\logg$, and seven elemental 
-abundances.
-
-Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
-we used a quadratic vectorizer for the giant star model.  Here the terms are 
-expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
-with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
-in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
-The results are shown in Figure \ref{fig:set-hyperparameters}, where
-again the enveloped region represents the minimum and maximum change in RMS label
-deviation with respect to the unregularized case.  At the point of maximum improvement 
-near $\Lambda = 0.13$ $(\approx10^{-0.9})$, the RMS in all nine labels has decreased by up to $30$\%,
-with all labels showing an improvement $>5$\%, and the mean improvement over all labels is about 10\%.
-Near $\Lambda \approx 10^{-0.9}$ to $10^{-0.3}$, the regularization also produces a sparser matrix
-of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
-Based on the increased model sparsity and decreasing RMS deviation in the labels, 
-we adopt $\Lambda = 0.57$ ($10^{-0.25}$) for the giant star model.  The bias in 
-labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
-$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elements.  The RMS
-at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
-between $0.07-0.09$~dex depending on the elemental abundance.  We inferred labels 
-for all \Nspectra\ \rave\ spectra using this model, and these results are summarized in
-the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible
-and in the expected location, without requiring any post-analysis calibration.
-However, artefacts due to dwarf stars in the test set are also present.
-
-\subsection{Joining the models}
-\label{sec:joining-the-models}
-
-We have derived labels for all \Nspectra\ \rave\ spectra using the three models
-described in previous sections.  The results from the joint model --- that includes 
-the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
-model is too simple for the \rave\ spectral range.  The other models have problems,
-too: unrealistic over-densities in label space show that the main-sequence model and
-the giant model make very poor extrapolations for stars outside their respective
-training sets.  For these reasons we were forced to exclude or severely penalize 
-incorrect results from both models.  We emphasize that the choices here are entirely
-heuristic, and depart from interpreting \thecannon\ output labels as the maxima
-of individual likelihood functions.  Each model produces estimates of the labels
-for a given star, and we use those estimates to produce a unified estimate, but 
-this joint estimate is calculated by disregarding the probabilistic attributes
-of individual estimates.
-
-Before attempting to join the results from different models, we excluded results
-in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
-labels that are outside the extent of the training set.  Specifically for the
-results from the giant model we (conservatively) excluded stars with derived 
-$\logg > 3.5$, and for the results from the main-sequence model we excluded 
-sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the 
-two-dimensional ($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  Unfortunately these restrictions did not remove all spurious
-results.  The reason for this can be explained with an example:  consider that our 
-giant star model was trained with only giant stars but tested with both giant 
-stars and dwarf stars.  Some classes of stars (e.g., metal-poor dwarfs) can 
-project into a region of label space that would suggest it is a giant 
-(e.g., a clump star).  These objects could have relatively low $\chi_{r}^2$ values
-(e.g., $\chi_{r}^2 < 3$) and in this example, they would appear as bonafide red clump 
-stars.  These incorrect projections are extrapolation errors in high dimensions that
-project to `normal' parts of the label space in two dimensions.  For these reasons 
-we also made use of the joint model in Section~\ref{sec:a-simple-model}
-to inform whether we should adopt results from: the red giant branch model; the
-main-sequence/sub-giant model; or a linear combination of the two.
-
-In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
-$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
-model; and the differences between the red giant branch model and the joint model. 
-We have scaled the differences in $\teff$ and $\logg$ to make the central peak 
-near $(0, 0)$ to be approximately isotropic by setting
-	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
-	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
-	$\delta_{\log{g}} = 0.15$~dex for both models.
-It is important to note that these $\delta$ values do not represent any kind of intrinsic
-uncertainty or precision: they are merely normalization factors.  Adopting
-substantially different scaling factors would produce in clear inconsistencies in our
-results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
-factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
-approximately isotropic, and to some extent, comparable.  In Figure 
-\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent objects where
-the joint model and the comparison model both report similar labels.  
-The artefacts 
-seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
-also present in Figure \ref{fig:joint-model-differences} as over-densities far away
-from the central peak.  Therefore, we can 
-adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model to the 
-main-sequence model $d_{ms}$,
-\begin{eqnarray}
-	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
-\end{eqnarray}
-
-However when adopting a similar distance metric for the giant model, we found that the
-giant model and the joint model produced similar label estimates (with low $\chi_r^2$
-values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
-in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
-could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
-as a term that would benefit the main-sequence model.  As a result, our joint estimate 
-for metal-poor dwarf stars that overlap with high-resolution studies
-\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
-would appear as metal-poor giant stars, because the giant star model would have a much
-smaller distance metric than the main-sequence model.  It is reasonable to assert that
-this issue may be complicated by the increased complexity, or flexibility, of the giant
-star model relative to the main-sequence model.  In any case, we found that by empirically
-penalizing the giant model distance metric by absolute metallicity,
-\begin{eqnarray}
-	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
-\end{eqnarray}
-
-\noindent{}the weighting of dwarfs was no longer a function of joint metallicity, and this
-issue was resolved without introducing new artefacts. Using these distance metrics we 
-then derived the weights,
-\begin{eqnarray}
-	w_{ms} = \frac{1}{{d_{ms}}^2} & \text{and} & w_{giant} = \frac{1}{{d_{giant}}^2} \quad ,
-\end{eqnarray}
-
-\noindent{}and produce the weighted labels $\hat\ell$:
-\begin{eqnarray}
-	\hat\ell = \frac{w_{ms}\,\ell_{ms} + w_{giant}\,\ell_{giant}}{w_{ms} + w_{giant}} \quad .
-\end{eqnarray}
-
-We calculate weighted errors of $\hat\ell$ in the same manner.  
-In Figure \ref{fig:model-weights} we show the mean relative
-weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional bin of 
-$\hat\teff$ and $\hat\logg$.  For giant stars the relative weight of 
-the main-sequence model is zero, and vice-versa for main-sequence stars.
-The relative weights smoothly transition from 0 to 1 on the sub-giant branch
-near $\log{g} \approx 3.5$, in the training set overlap region of both models.
-For abundance labels in the giant model that are not in the main-sequence
-model (e.g., [O/H], [Mg/H]), we only report abundances for objects if
-$w_{ms} < 0.05$.
-
-The weighted $\teff$ and $\logg$ values for stars meeting different
-S/N constraints are shown in Figure \ref{fig:test-set-hrd}, both in logarithmic density
-and mean metallicity.  The artefacts from individual models are no longer 
-apparent, and the rich structure of the Hertzsprung-Russell is visible.  Despite
-this, there are a number of caveats introduced by the choices we have made to
-combine estimates from multiple models. We discuss these issues in detail in
-Section \ref{sec:discussion}.
-
-\section{Validation experiments}
-\label{sec:validation}
-
-In addition to the cross-validation tests that we have previously described, 
-we have conducted a number of internal and external validation experiments to 
-test the validity of our results.  We will begin by describing internal validation
-tests based on repeat observations, before evaluating our accuracy based on
-high-resolution literature comparisons.
-
-\subsection{Multi-epoch observations}
-\label{sec:repeat-observations}
-
-The \rave\ survey performed repeat observations for 43,918 stars with time 
-intervals ranging from a few hours to up to four years.  This timing was 
-constructed to be quasi-logarithmic such that spectroscopic binaries could
-be optimally identified. Most of the stars that were observed multiple times
-were only observed twice, with thirteen visits being the maximum number 
-of observations for any target.  These repeat observations allow us to 
-quantify the level of (in)correctness in our formal errors.  For every star
-with multiple visits we constructed a high S/N stacked spectrum for that
-star, weighted by the inverse variances in the individual visit spectra.
-
-We treated the stacked spectra as normal survey stars.  We inferred labels
-using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
-of comparison for all individual visits to that object.  Figure 
-\ref{fig:formal-errors-comparison} shows the difference in labels 
-between the comparison spectrum and a repeat visit, normalized by their 
-formal errors summed in quadrature (e.g., 
-$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
-If our measurements were unbiased by S/N and the formal errors were 
-representative, these values should be normally distributed with a zero 
-mean and unit variance.
-\stub{Are they? (Probably not!). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
-we show these inflated errors only, as we gauge them to be more representative.}
-
-\stub{In Figure \ref{fig:repeat-visits} we look at the difference as a function of SNR}
-
-\subsection{External validation}
-\label{sec:external-validation}
-
-\subsubsection{Comparison with \rave\ DR4}
-\label{sec:validation-kordopatis}
-
-As an initial point of comparison, we cross-matched our results against the 
-official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
-In order to provide a fair comparison, we only show stars that meet a number
-of quality flags in \emph{both} samples.  Our constraints require that the
-S/N ratio exceeds 10~pixel$^{-1}$, and $\chi_r^2 < 3$.  For this comparison
-we further required that:
-the \texttt{QK} flag from \citet{Kordopatis_2013} is zero, indicating no
-problems were reported by the pipeline; $T_{{\rm eff},DR4} > 4000$~K;
-the error in radial velocity \texttt{e\_HRV} is $<$8~km~s$^{-1}$; and the three
-principal morphological flags \texttt{c1}, \texttt{c2}, \texttt{c3}, from 
-\citet{Matijevic_2012} all indicate `n' for a normal FGK-type star.
-There is good agreement in $\teff$, with a bias and RMS of just 4~K and 240~K,
-respectively. The offset in $\logg$ on the giant branch between this study and 
-\citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this issue
-has been resolved in the fifth \rave\ data release by correcting $\logg$
-values with a calibration sample consisting of asteroseismic targets and the
-\gaia\ benchmark stars.  There is also a
-slight discrepancy in the $\logg$ values along the main-sequence, where our
-work tends to taper down towards higher $\logg$ values at cooler temperatures,
-and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
-This difference is not likely to have a very significant effect on the detailed
-abundance or spectrophotometric distance determinations between these studies 
-\citep{Binney_2014}.
-
-\subsubsection{Comparisons with Reddy, Bensby, and Valenti \& Fischer}
-\label{sec:validation-gold-standards}
-
-We searched the literature for studies that overlap with \rave, and which base
-their analysis on high-resolution, high S/N spectra.  We found four notable
-studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,Reddy_2006}
-and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} work on exoplanet
-host star candidates.  These studies 
-perform a careful (manual; expert) analysis using extremely high-resolution, high S/N spectra, and make use of \project{Hipparcos}
-parallaxes where possible.  Most of the stars in these samples are main-sequence
-or sub-giant stars.  Therefore, these works constitute an excellent comparison
-to evaluate the accuracy of our results on the main-sequence and sub-giant branch.
-
-We compare the stellar parameters ($\teff$, $\logg$, [Fe/H]) of stars that
-overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
-We only include stars with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter
-cut removed only a few stars because the average S/N in the \rave\ spectra for
-these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$). When considering the relative 
-information content available in \rave\ (945~pixels in the near infrared with
-$\mathcal{R} \approx 7{,}500$) compared to these literature studies that use
-\project{Hipparcos} parallaxes where possible, and base their inferences on
-spectra with resolving power $\mathcal{R}$ between 40,000 to 110,000, and
-S/N ratios exceeding 150, we consider the agreement to be very satisfactory.
-Indeed, given the metallicity precision available in the \unrave\ catalog, 
-these results will likely be useful for future studies based on exoplanet
-host star properties (e.g., \project{TESS}). At present, however, there are 
-just $\approx$30 stars in \rave\ that overlap with the compilations of 
-exoplanet host star properties listed at \texttt{exoplanets.org} and 
-\texttt{exoplanets.eu}.
-
-\subsubsection{Comparison with the \ges\ survey}
-\label{sec:validation-ges}
-
-There are 142 stars that overlap between \rave\ and the fourth internal
-data release of the \ges\ survey. These are a mix of main-sequence, 
-sub-giant and red giant branch stars.  About half (67) of the sample 
-were acquired with the \acronym{UVES} instrument, the other with the \acronym{GIRAFFE} spectrograph, and
-the S/N of the \ges\ spectra peaks at $\approx140$~pixel$^{-1}$.  Despite 
-most of these stars having
-relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
-agreement in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
-The RMS in effective temperature, surface gravity and metallicity is
-233~K, 0.37~dex, and 0.17~dex, respectively.
-
-Based on this comparison, we find no evidence for a systematic offset 
-between stars on the main-sequence and those on the giant branch. Specifically,
-we do not find any dwarf stars that are misclassified as sub-giants, as we
-found for some metal-poor stars in Section~\ref{sec:validation-gold-standards}.
-In addition we find no evidence for a systematic offset in metallicity (or
-individual abundances) between giant stars and main-sequence stars.  This is
-a crucial observation, as the metallicities for stars in our main-sequence
-training set have a principally different source to those on the giant branch.
-We cannot make these same inferences based on other surveys, like \apogee,
-because (1) \apogee\ stars formed part of the training set, and (2) they do
-not include main-sequence stars.  Even if we found good agreement between
-\epic\ and \apogee\ metallicities, this would not be informative, because
-\apogee\ is the source of metallicity for many stars on the giant branch in
-the \epic\ sample. Therefore, although this is a qualitative comparison only,
-it is reassuring that there is no obvious systematic difference between the
-metallicities of main-sequence and giant branch stars.
-
-The metallicity agreement between this work and \ges\ extends down to low
-metallicity, near $[{\rm Fe/H}] \approx -1.5$.  The scatter increases for
-the four stars in the overlap sample with [Fe/H] $< -1$, however these
-spectra have low S/N ratios, which is reflected in the larger uncertainties
-reported for these metallicities. 
-% TODO: Check if that is still the case after we examine errors from repeat
-% 		obs.
-
-The fourth internal data release of the \ges\ includes detailed chemical
-abundances of up to 45~species ($\approx32$ elements at different ionization
-stages).  This provides us with an independent validation for our detailed
-abundances on the giant branch.  These comparisons are shown for six elemental
-abundances in Figure~\ref{fig:ges-abundances}, where markers are colored by the
-S/N of the \rave\ spectrum.  The number of stars available in each abundance
-comparison varies due to what is available in the \ges\ data release, which
-is itself a function of the instrument used, the spectral type, and other
-factors.  The absolute bias for individual elements varies from as low as
-0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
-[Si/H] abundances relative to the \ges\ survey.  The large bias in [Si/H]
-is likely a consequence of an offset between [Si/H] abundances in \ges\ and
-\apogee, the source of our training set for giant star abundances.  
-The RMS deviation in each
-label is small for stars with [X/H] $> -0.5$, before increasing at lower
-metallicities.  If we consider all stars, the smallest abundance RMS we see
-with respect to \ges\ is 0.18~dex for [Ca/H].  The increasing RMS at low 
-metallicity is likely a consequence of multiple factors, namely: inaccurate
-abundance labels for metal-poor stars (Section~\ref{sec:the-training-set});
-only weak, blended lines being available in \rave, which cease to be visible
-in hot and/or metal-poor stars; and to a lesser extent, low S/N ratios for 
-those particular stars being compared.  Unfortunately not all of these
-factors are represented by the quoted errors in each label.  For these reasons, 
-although it affects only a small number of stars, we recommend caution when 
-using individual abundances for very metal-poor giant stars in our sample.
-
-\subsection{Astrophysical validation}
-\label{sec:astrophysical-validation}
-
-\subsubsection{Globular clusters}
-\label{sec:globular-cluster-validation}
-
-After verifying that our atmospheric parameters and abundances are comparable with other
-high-resolution studies, here we verify that the abundances we find
-are consistent with expectations from astrophysics.  One example is globular clusters, 
-which show anti-correlations in light element abundances 
-\citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
-survey, \cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters: 49 stars belonging to NGC~5139
-($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 
-members of NGC~362. In addition to these systems, 12 stars thought to belong to NGC~1851
-were found in the \rave\ data by \cite{Kunder_2014}, and a further 10 stars in NGC~6752.  
-We refer the reader to those studies for details regarding the membership selection.
-
-In Figure~\ref{fig:globular-cluster-HRD} we show the effective temperature
-$\teff$ and surface gravity $\logg$ for pre-defined members of globular clusters in \rave.
-The top panels indicate measurements made in this work, and for comparison purposes we have included
-the results from the fourth \rave\ data release.  We show representative \project{PARSEC} isochrones 
-\citep{Bressan_2012} in all panels, where the isochrone ages and metallicities are adopted from
-\citet{Kunder_2016,Marin-Franch_2009}, and the \citet[][accessed 6 September 2016]{Harris_1996} 
-online catalog of globular cluster properties.  The globular cluster with the most number of
-\rave\ members is NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
-consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}.  We
-find the mean metallicity of prescribed $\omega$-Centauri members to be $[{\rm Fe/H}] = -0.85$.
-However, it is clear that the membership criteria could be improved with our
-revised metallicities and detailed chemical abundances. 
-Indeed, this demonstrates that our abundance labels can be used to identify
-globular cluster members that are now tidally disrupted \citep{Anguiano_2016,Kuzma_2016,Navin_2016},
-even for stars with low S/N ratios.
-
-\subsubsection{Open clusters}
-\label{sec:open-cluster-validation}
-
-The \rave\ survey observed $\sim$160 highly probable members of several open clusters:
-78 stars in the Pleiades, 26 in Hyades, 30 stars in the solar-metallicity cluster
-M67 and 13 stars in IC4651. Details on the open cluster membership and selection can 
-be found in \cite{Kunder_2016}.  We show the effective temperature $\teff$ and surface
-gravity $\logg$ for these stars in Figure \ref{fig:open-cluster-HRD}.  The isochrones
-are sourced from \citet{Bressan_2012}, with cluster properties adopted from 
-\cite{Kunder_2016} and \cite{Kharchenko_2013}.
-
-We find good agreement between our atmospheric parameters (top panels) and the 
-isochrones shown.  The position of the red clump in IC4651 and M67 are perfectly matched
-to the isochrone, and the Hyades main-sequence is in good agreement down to 
-$T_{\rm eff} \approx 4000$~K.  Similarly, we find consistency with the literature and
-our metallicity scale.  We find the mean metallicity of M67 stars to be 
-$[{\rm Fe/H}] = -0.02 \pm 0.17$~dex (30 stars), in excellent agreement with the
-expected $[{\rm Fe/H}] = 0.00$ value (not accounting for atomic diffusion).
-We further find the Hyades mean metallicity to be $[{\rm Fe/H}] = 0.07 \pm 0.09$~dex,
-consistent with \citet{Paulson_2003}: $[{\rm Fe/H}] \approx 0.13$.
-For IC4651 from 13 stars we find a mean $[{\rm Fe/H}] = 0.15 \pm 0.12$~dex, matching the
-high-resolution, high S/N study of \citet{Pasquini_2004}, where they find 
-$[{\rm Fe/H}] = 0.10 \pm 0.03$~dex. Finally, from 78 stars we find the mean
-metallicity of the Pleiades to be $[{\rm Fe/H}] = -0.02 \pm 0.12$~dex, in agreement
-with the $[{\rm Fe/H}] = -0.034 \pm 0.024$~dex measurement reported by 
-\citet{Friel_Boesgaard_1990}.
-
-The most pronounced discrepancy is seen in the main-sequence of the Pleiades.
-Like \citet{Kordopatis_2013}, we find lower $\logg$ values than what the isochrone
-shown predicts.  \stub{Hawkins: use correct isochrone from Bouy to see if this issue is resolved.}
-%  Firstly,
-%we note that theoretical isochrones often struggle to match the observed
-%color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.
-
-\section{Discussion}
-\label{sec:discussion}
-
-We have derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all \rave\ 
-spectra, as well as detailed abundances for red giant branch stars, based on an
-independent re-analysis of \rave\ spectra.  Our catalog provides an excellent
-complement to the current \rave\ data release \citep{Kunder_2016}.  
-When combined with the \tgas\ sample, these results may amount to a powerful
-compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
-Inferences based on these results should recognize those caveats, and acknowledge that 
-these results are subject to our explicit assumptions, some of which are provably
-incorrect.
-
-For practical purposes we adopted separate models: one for the giant branch and one
-for the main-sequence.  A third model was used to derive relative weights for which
-results to use.  The relative weighting we have used does not have any formal
-interpretation as a likelihood or belief (in any sense): it was introduced for
-practical reasons to identify systematic errors and combine results for multiple
-models.  Because the relative weights have no formal interpretation, it is reasonable
-to consider this method is as \emph{ad hoc} as any other approach.  Therefore, the
-relative weighting has no warranty to be (formally) correct, and may introduce 
-inconsistencies or systematic errors rather than
-minimizing them.  Evidence of this is already present in our comparisons to
-`gold standard' works \citep{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005},
-where metal-poor dwarf stars appear as metal-poor sub-giants in our combined results,
-but have reasonable labels (by any available statistic) as dwarfs in the main-sequence
-model.  Clearly, the relative distance from the joint model to the giant model is 
-slightly favoured (over the main-sequence model) for metal-poor dwarf stars. In this
-sense, our approach towards combining labels from multiple models is sub-optimal.
-
-If we only consider the results from individual models, there are a number of cautionary
-remarks that stem from the construction of the training set.  The labels for red giant 
-branch stars primarily come from \apogee, where previous successes with \thecannon\
-have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
-\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}.
-However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
-a tapering-off in the test set --- where \emph{no} stars had reliably reported metallicities
-below that of the training set --- which forced us to construct a heterogeneous training
-set.  The metal-poor stars included in this sample are from high-resolution studies
-\citep{Fulbright_2010,Ruchti_2011}, but it is unknown if the stellar parameters are
-of high fidelity (e.g., we have limited quality metrics available), and there is no
-guarantee that the stellar parameters \emph{or} abundances are on the same scale as
-\apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
-
-If the metallicities of metal-poor giant stars are on the same scale as the \apogee\
-abundances, there is a larger issue in verifying that the main-sequence metallicities
-and giant branch metallicities are on the same scale.  The training set for the
-main-sequence stars includes metallicities from a variety of sources --- including
-\lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
-no straightforward manner to ensure that these two models produce metallicities on
-the same scale.  We see no systematic offset in metallicities of dwarf and giant
-stars that overlap between \rave\ and the \ges\ survey, suggesting that if there is
-a systematic offset, it must be small. Nevertheless, these are only verification checks based on $<1$\% of the
-data, and there is currently insufficient data for us to prove both models are on
-the same abundance scale.
-
-For some of the most metal-poor giant stars in \rave\, we \emph{know} the abundances 
-are not on the same scale as \apogee, because we were forced to adopt abundances for 
-specific elements when they were unavailable.  In that sense, our assumption that the
-training set contains accurate labels is clearly untrue (for \emph{at least} some stars).
-Although we sought to adopt the mean
-level of Galactic chemical enrichment at the overall metallicity, this is not a
-representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
-metallicity, there is no requirement for zero abundance spread.
-More fundamentally, we are incorrectly assuming that the element \emph{must} be 
-detectable in the photosphere of the star.  There may be no transition that is 
-detectable in that star, even with zero noise, because it is too weak to have any 
-effect on the spectrum.  In the most optimistic case, this can be considered to be
-forcing the model to make use of correlated information between abundances.  In a
-more representative (pessimistic) case, we are simply invoking what all abundances
-should be at low metallicity.
-
-This choice is reflected in the abundances of the test set.  While we do recover
-trustworthy metallicities for ultra metal-poor ($[{\rm Fe/H}] \lesssim -4$) stars like 
-CD-38~245, the individual abundances for all extremely metal-poor stars aggregate
-(in [X/Fe] space; Figure~\ref{fig:gce}) at the assumed abundances for the metal-poor
-stars in our sample.  Thus, while the overall metallicities appear reliable, the 
-individual abundances for extremely metal-poor stars in the test set cannot be
-considered trustworthy in any sense.  For this reason we have updated the electronic
-catalog to discard these results, and included a quality control flag to indicate
-the result was discarded (see catalog \texttt{README} file for specifics related to
-the data structure and content).
-
-In Section~\ref{sec:method} we assumed that any fibre- or time-dependent variations
-in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
-\citet{Kordopatis_2013} note that the effective resolution of \rave\
-spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
-effective resolution is a function of temperature variations, fibre-to-fibre
-variations, and thus position on the CCD \citep{Steinmetz_2006}.
- For this reason we ought to expect our derived stellar
-parameters or abundances to be correlated with the fibre number, or over
-time.  If significant, the trend could produce systematically offset
-stellar abundances solely due to the fibre number.  \citet{Kordopatis_2013}
-conclude that resolution-based effects on the \rave\ stellar parameters
-should be a second-order effect. We have not seen evidence of these 
-resolution-based correlations in our results, however, we have only
-performed cursory (non-exhaustive) experiments to investigate this issue.
-
-\stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}
-
-The fibre fringing is another example of how test spectra that are
-unusual from the training set can be projected as peculiar artefacts in the
-label space.  In other words, unusual spectra can appear as `clumps' in
-regions of parameter space that we could consider as being normal 
-(e.g., solar-type stars).  We addressed this issue for the main-sequence
-and giant models by using a joint model to calculate relative weights.
-However, spectra that are unusual from the joint model training set can
-still project as systematic artefacts.
-
-Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
-first is a small over-density at the base of the giant branch, which is
-likely a consequence of joining the 9-label and 3-label models.  The
-second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
-present in our training set, yet there are many in the \rave\ survey.
-However, the \rave\ pre-processing pipeline (\texttt{SPARV}) performs
-template matching against grids of cool \emph{and} hot stars, and therefore
-we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
-we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
-where each hexagonal bin is colored by the \emph{maximum} temperature
-reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
-at just above the turn-off point in our results.
-
-\stub{this following paragraph needs rewording}
-Based on a thorough interrogation of our results, we have adopted two boolean
-columns to signify whether our labels meet specific quality constraints:
-one marked \texttt{QC} to indicate reliable stellar parameters ($\teff$, $\logg$, [Fe/H]), and \texttt{QC\_ABUNDANCE} to indicate reliable chemical abundances.
-For stellar atmospheric parameters to meet our quality constraints, we
-require: $\chi_r^2 < 3$; $S/N >10$~pixel$^{-1}$; and \texttt{TeffS} $< 8000$~K
-in order to remove hot stars that are outside the training set.  We
-specified the quality constraint in chemical abundances to further
-require that the iron abundance $[{\rm Fe/H}] > -1.5$, since this approximately
-corresponds to the point where individual abundances will become undetectable
-in \rave\ spectra, and the individual abundance labels will be driven by our
-constraint to place them on the mean galactic chemical evolution trend.
-These quality constraints should not be taken verbatim: judicious use of
-the morphological classifications \citep{Matijevic_2012} or metadata from
-the pre-processing pipelines is encouraged.
-
-We have presented the largest collection of stellar abundances for stars in the 
-first \gaia\ data release.  In total more than X.XX million abundances are
-available, with effective temperature $\teff$, surface gravity $\logg$ and 
-metallicity available for XXXXXX stars.  Despite the caveats and limitations 
-discussed here, our validation experiments and comparisons with high-resolution
-spectroscopic studies suggests that our results have sufficient accuracy and
-precision to be useful for exoplanet host stars studies 
-\citep[c.f.,][]{Valenti_Fischer_2005}, investigations of the Milky Way disk 
-\citep[c.f.,][]{Reddy_2003,Reddy_2006,Bensby_2014}, and a multitude of 
-chemo-dynamic studies.  When combined with positions and 3D velocities from \gaia,
-we are optimistic that the \unrave\ catalog will help advance understanding 
-of the Milky Way's formation and evolution.
-
-\subsection*{Access the results electronically}
-
-% TODO: Replace ZENODO-URL with a data url
-%\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
-\noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
-and this document was compiled from revision hash \texttt{\githash} in that repository.
-Inferred labels, associated errors, and relevant metadata are available electronically
-at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  These results are also available
-through the \rave\ DR5 database \citep{Kunder_2016}. Please note that it is a condition
-of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must 
-also be cited, as the work presented here would not have been possible without the 
-tireless efforts of the entire \rave\ collaboration, past and present.
-%}}
-
-\acknowledgements
-We thank 
-	Jonathan Bird (Vanderbilt),
-	Sven Buder (MPIA), 
-	Sergey Koposov (Cambridge),
-and 
-	Sofia Randich (INAF).
-This research made use of: 
-  	NASA's Astrophysics Data System Bibliographic Services;
-  	Astropy, a community-developed core Python package for Astronomy \citep{astropy};
-and 
-  	\project{TOPCAT} \citep{Taylor_2005}.
-This work was partly supported by the European Union FP7 programme through ERC 
-grant number 320360. A.~R.~C. thanks King's College of the University of Cambridge
-for their support.  Funding for K.~H.~ has been provided through the Simons 
-Foundation Society of Fellows and the Marshall Scholarship.
-Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
-the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
-University; the Australian Research Council; the French National Research Agency;
-the German Research Foundation (SPP 1177 and SFB 881); the European Research 
-Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
-Padova; The Johns Hopkins University; the National Science Foundation of the USA
-(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
-Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
-Research Council of Canada; the Slovenian Research Agency; the Swiss National 
-Science Foundation; the Science \& Technology Facilities Council of the UK; 
-Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
-and Sydney. The RAVE web site is https://www.rave-survey.org.
-
-\begin{thebibliography}{dummy}
-\bibitem[Anguiano et al.(2015)]{Anguiano_2015} Anguiano, B., Zucker, D.~B., Scholz, R.-D., et al.\ 2015, \mnras, 451, 1229
-
-\bibitem[Anguiano et al.(2016)]{Anguiano_2016} Anguiano, B., De Silva, G.~M., Freeman, K., et al.\ 2016, \mnras, 457, 2078
-
-\bibitem[Astropy Collaboration et 
-al.(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, Astronomy \& Astrophysics, 558, AA33
-
-\bibitem[Bensby et al.(2014)]{Bensby_2014} Bensby, T., Feltzing, S., \& Oey, M.~S.\ 2014, \aap, 562, A71
-
-\bibitem[Binney et al.(2014)]{Binney_2014} Binney, J., Burnett, B., Kordopatis, G., et al.\ 2014, \mnras, 437, 351
-
-\bibitem[Boeche et al.(2011)]{Boeche_2011} Boeche, C., Siebert, A., Williams, M., et al.\ 2011, \aj, 142, 193
-
-\bibitem[Boeche et al.(2013)]{Boeche_2013} Boeche, C., Chiappini, C., Minchev, I., et al.\ 2013, \aap, 553, A19
-
-\bibitem[Bressan et al.(2012)]{Bressan_2012} Bressan, A., Marigo, P., Girardi, L., et al.\ 2012, \mnras, 427, 127
-
-\bibitem[Bovy et al.(2016)]{Bovy_2016} Bovy, J., Rix, H.-W., Schlafly, E.~F., et al.\ 2016, \apj, 823, 30
-
-\bibitem[Brown et al.(2011)]{Brown_2011} Brown, T.~M., Latham, D.~W., Everett, M.~E., \& Esquerdo, G.~A.\ 2011, \aj, 142, 112
-
-\bibitem[Bouy et al.(2015)]{Bouy_2015} Bouy, H., Bertin, E., Sarro, L.~M., et al.\ 2015, \aap, 577, A148
-
-\bibitem[Carretta et al.(2009)]{Carretta_2009} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2009, \aap, 505, 117
-
-\bibitem[Carretta et al.(2013)]{Carretta_2013} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2013, \aap, 557, A138
-
-\bibitem[Casagrande et al.(2011)]{Casagrande_2011} Casagrande, L., Sch{\"o}nrich, R., Asplund, M., et al.\ 2011, \aap, 530, A138
-
-\bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88
-
-\bibitem[Casey et al.(2013)]{Casey_2013} Casey, A.~R., Da Costa, G., Keller, S.~C., \& Maunder, E.\ 2013, \apj, 764, 39
-
-\bibitem[Casey et al.(2014a)]{Casey_2014a} Casey, A.~R., Keller, S.~C., Da Costa, G., Frebel, A., \& Maunder, E.\ 2014, \apj, 784, 19
-
-\bibitem[Casey et al.(2014b)]{Casey_2014b} Casey, A.~R., Keller, S.~C., Alves-Brito, A., et al.\ 2014, \mnras, 443, 828
-
-\bibitem[Casey \& Schlaufman(2015)]{Casey_Schlaufman_2015} Casey, A.~R., \& Schlaufman, K.~C.\ 2015, \apj, 809, 110
-
-\bibitem[Casey(2016)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
-
-\bibitem[Casey et al.(2016)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
-
-\bibitem[Cui et al.(2012)]{Cui_2012} Cui, X.-Q., Zhao, Y.-H., Chu, Y.-Q., et al.\ 2012, Research in Astronomy and Astrophysics, 12, 1197
-
-\bibitem[De Silva et al.(2015)]{DeSilva_2015} De Silva, G.~M., Freeman, K.~C., Bland-Hawthorn, J., et al.\ 2015, \mnras, 449, 2604
-
-\bibitem[Deason et al.(2011)]{Deason_2011} Deason, A.~J., Belokurov, V., \& Evans, N.~W.\ 2011, \mnras, 416, 2903
-
-\bibitem[Friel \& Boesgaard(1990)]{Friel_Boesgaard_1990} Friel, E.~D., \& Boesgaard, A.~M.\ 1990, \apj, 351, 480
-
-\bibitem[Fulbright et al.(2010)]{Fulbright_2010} Fulbright, J.~P., Wyse, R.~F.~G., Ruchti, G.~R., et al.\ 2010, \apjl, 724, L104
-
-\bibitem[Garc{\'{\i}}a P{\'e}rez et al.(2016)]{Garcia_Perez_2016} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., et al.\ 2016, \aj, 151, 144
-
-\bibitem[Gilmore et al.(2012)]{Gilmore_2012} Gilmore, G., Randich, S., Asplund, M., et al.\ 2012, The Messenger, 147, 25
-
-\bibitem[Harris(1996)]{Harris_1996} Harris, W.~E.\ 1996, \aj, 112, 1487
-
-\bibitem[Hawkins et al.(2016)]{Hawkins_2016} Hawkins, K., Masseron, T., Jofre, P., et al.\ 2016, arXiv:1604.08800
-
-\bibitem[Holtzman et al.(2015)]{Holtzman_2015} Holtzman, J.~A., Shetrone, M., Johnson, J.~A., et al.\ 2015, \aj, 150, 148
-
-\bibitem[Ho et al.(2016)]{Ho_2016} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., et al.\ 2016, arXiv:1602.00303
-
-\bibitem[H{\o}g et al.(2000)]{Hog_2000} H{\o}g, E., Fabricius, C., Makarov, V.~V., et al.\ 2000, \aap, 355, L27
-
-\bibitem[Huber et al.(2016)]{Huber_2016} Huber, D., Bryson, S.~T., Haas, M.~R., et al.\ 2016, \apjs, 224, 2
-
-\bibitem[Jofr{\'e} et al.(2015)]{Jofre_2015} Jofr{\'e}, P., M{\"a}dler, T., Gilmore, G., et al.\ 2015, \mnras, 453, 1428
-
-\bibitem[Kharchenko et al.(2013)]{Kharchenko_2013} Kharchenko, N.~V., Piskunov, A.~E., Schilbach, E., R{\"o}ser, S., \& Scholz, R.-D.\ 2013, \aap, 558, A53
-
-\bibitem[Kordopatis et al.(2011)]{Kordopatis_2011} Kordopatis, G., Recio-Blanco, A., de Laverny, P., et al.\ 2011, \aap, 535, A106
-
-\bibitem[Kordopatis et al.(2013)]{Kordopatis_2013} Kordopatis, G., Gilmore, G., Steinmetz, M., et al.\ 2013, \aj, 146, 134
-
-\bibitem[Kordopatis et al.(2015)]{Kordopatis_2015} Kordopatis, G., Binney, J., Gilmore, G., et al.\ 2015, \mnras, 447, 3526
-
-\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, submitted
-
-\bibitem[Kunder et al.(2014)]{Kunder_2014} Kunder, A., Bono, G., Piffl, T., et al.\ 2014, \aap, 572, A30
-
-\bibitem[Kuzma et al.(2016)]{Kuzma_2016} Kuzma, P.~B., Da Costa, G.~S., Mackey, A.~D., \& Roderick, T.~A.\ 2016, \mnras, 461, 3639
-
-\bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, arXiv:1606.03015
-
-\bibitem[Matijevi{\v c} et al.(2012)]{Matijevic_2012} Matijevi{\v c}, G., Zwitter, T., Bienaym{\'e}, O., et al.\ 2012, \apjs, 200, 14
-
-\bibitem[Marino et al.(2011)]{Marino_2011} Marino, A.~F., Milone, A.~P., Piotto, G., et al.\ 2011, \apj, 731, 64
-
-\bibitem[Mar{\'{\i}}n-Franch et al.(2009)]{Marin-Franch_2009} Mar{\'{\i}}n-Franch, A., Aparicio, A., Piotto, G., et al.\ 2009, \apj, 694, 1498
-
-\bibitem[Michalik et al.(2015a)]{Michalik_2015a} Michalik, D., Lindegren, L., \& Hobbs, D.\ 2015, \aap, 574, A115
-
-\bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
-
-\bibitem[Munari et al.(2005)]{Munari_2005} Munari, U., Sordo, R., Castelli, F., \& Zwitter, T.\ 2005, \aap, 442, 1127
-
-\bibitem[Navin et al.(2016)]{Navin_2016} Navin, C.~A., Martell, S.~L., \& Zucker, D.~B.\ 2016, arXiv:1606.06430
-
-\bibitem[Ness et al.(2012)]{Ness_2012} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2012, \apj, 756, 22
-
-\bibitem[Ness et al.(2013a)]{Ness_2013a} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 430, 836
-
-\bibitem[Ness et al.(2013b)]{Ness_2013b} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 432, 2092
-
-\bibitem[Ness et al.(2015)]{Ness_2015} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16
-
-\bibitem[Ness et al.(2016)]{Ness_2016} Ness, M., Hogg, D.~W., Rix, H.-W., et al.\ 2016, \apj, 823, 114
-
-\bibitem[Nordstr{\"o}m et al.(2004)]{Nordstrom_2004} Nordstr{\"o}m, B., Mayor, M., Andersen, J., et al.\ 2004, \aap, 418, 989
-
-\bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
-
-\bibitem[Pasquini et al.(2004)]{Pasquini_2004} Pasquini, L., Randich, S., Zoccali, M., et al.\ 2004, \aap, 424, 951
-
-\bibitem[Paulson et al.(2003)]{Paulson_2003} Paulson, D.~B., Sneden, C., \& Cochran, W.~D.\ 2003, \aj, 125, 3185
-
-\bibitem[Randich et al.(2013)]{Randich_2013} Randich, S., Gilmore, G., \& Gaia-ESO Consortium 2013, The Messenger, 154, 47
-
-\bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
-
-\bibitem[Reddy et al.(2003)]{Reddy_2003} Reddy, B.~E., Tomkin, J., Lambert, D.~L., \& Allende Prieto, C.\ 2003, \mnras, 340, 304
-
-\bibitem[Reddy et al.(2006)]{Reddy_2006} Reddy, B.~E., Lambert, D.~L., \& Allende Prieto, C.\ 2006, \mnras, 367, 1329
-
-\bibitem[Robin et al.(2012)]{Robin_2012} Robin, A.~C., Luri, X., Reyl{\'e}, C., et al.\ 2012, \aap, 543, A100
-
-\bibitem[Ruchti et al.(2011)]{Ruchti_2011} Ruchti, G.~R., Fulbright, J.~P., Wyse, R.~F.~G., et al.\ 2011, \apj, 743, 107
-
-\bibitem[Schlaufman et al.(2009)]{Schlaufman_2009} Schlaufman, K.~C., Rockosi, C.~M., Allende Prieto, C., et al.\ 2009, \apj, 703, 2177
-
-\bibitem[Siebert et al.(2011)]{Siebert_2011} Siebert, A., Williams, M.~E.~K., Siviero, A., et al.\ 2011, \aj, 141, 187
-
-\bibitem[SDSS Collaboration et al.(2016)]{sloan_dr13} SDSS Collaboration, Albareti, F.~D., Allende Prieto, C., et al.\ 2016, arXiv:1608.02013
-
-\bibitem[Smiljanic et al.(2014)]{Smiljanic_2014} Smiljanic, R., Korn, A.~J., Bergemann, M., et al.\ 2014, \aap, 570, A122
-
-\bibitem[Steinmetz et al.(2006)]{Steinmetz_2006} Steinmetz, M., Zwitter, T., Siebert, A., et al.\ 2006, \aj, 132, 1645
-
-\bibitem[Taylor(2005)]{Taylor_2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
-
-\bibitem[Valenti \& Fischer(2005)]{Valenti_Fischer_2005} Valenti, J.~A., \& Fischer, D.~A.\ 2005, \apjs, 159, 141
-
-\bibitem[van Leeuwen(2007)]{van_Leeuwen_2007} van Leeuwen, F.\ 2007, \aap, 474, 653
-
-\bibitem[Wojno et al.(2016)]{Wojno_2016} Wojno, J., et al.\ 2016, submitted
-
-\bibitem[Zasowski et al.(2013)]{Zasowski_2013} Zasowski, G., Johnson, J.~A., Frinchaboy, P.~M., et al.\ 2013, \aj, 146, 81
-
-\bibitem[Zwitter et al.(2008)]{Zwitter_2008} Zwitter, T., Siebert, A., Munari, U., et al.\ 2008, \aj, 136, 421
-
-\end{thebibliography}
-
-\clearpage
-
-\begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/hrd-train-set.pdf}
-\caption{Effective temperature $\teff$ and surface gravity $\logg$ for all stars in the training sets.
-Stars are colored by their metallicity [Fe/H], and the three panels show stars in the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
-\end{figure}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/test-set-density.pdf}
-\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
-These panels demonstrate how a single quadratic model is insufficient for all \rave\ stars (left panel), and illustrate some of the systematic artefacts that can result from testing on stars outside of the training set (center and right panel).
-These panels do not represent our final results, which are shown in Figure~\ref{fig:test-set-hrd}.\label{fig:test-set-density}}
-\end{figure*}
-
-\begin{figure}[p]
-\center
-\includegraphics[width=0.45\textwidth]{figures/set-hyperparameters.pdf}
-\caption{The percentage change in RMS deviation between inferred and training labels at different regularization strengths.
-The RMS values were calculated by leave-one-out cross-validation, and are shown with respect to an unregularized model ($\Lambda = 0$).
-The points and solid line indicate the mean improvement across all labels.
-The filled area represents the minimum and maximum improvements over all labels.
-With increasing regularization strength, there is a minimum in the RMS deviation over all labels, which is where we set $\Lambda$ for each model (see text for details).\label{fig:set-hyperparameters}}
-\end{figure}
-
-\begin{figure}[p]
-\center
-\includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
-\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the joint model (top panel), and the giant model and the joint model (bottom panel).
-The density scaling is logarithmic, and the differences in $\teff$ and $\logg$ are scaled to make them approximately isotropic (see text for details).
-The peak at $(0, 0)$ represents good agreement between the joint model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
-\end{figure}
-
-\begin{figure}[p]
-\center
-\includegraphics[width=0.5\textwidth]{figures/model-weights.pdf}
-\caption{The mean relative main-sequence model weight $w_{ms}/(w_{ms} + w_{giant})$ at each hexagonal bin of weighted effective temperature $\teff$ and surface gravity $\logg$.
-The relative weighting illustrates how only results from the main-sequence model are adopted for unevolved stars, and there is a gradual transition to using results from the giant model, before only results from the giant model are used for evolved stars.\label{fig:model-weights}}
-\end{figure}
-
-\begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
-\caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
-Only results with $\chi_{r}^2 < 3$ and S/N ratios greater than the quoted limit in each panel are shown.
-The top three panels show logarithmic density, and bins in the bottom three panels are colored by the median metallicity in each bin.\label{fig:test-set-hrd}}
-\end{figure}
-
-\begin{figure*}[p]
-\caption{Distribution of differences in label estimates from multiple visits, divided by their formal errors summed in quadrature.
-If our measurements were unbiased and the formal errors were representative, these distribution should be normally distributed with zero mean and unit variance.\label{fig:formal-errors-comparison}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/repeat-visits.pdf}
-\caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:repeat-visits}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/dr4-comparison.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth \rave\ data release \citep{Kordopatis_2013} and this work.
-Here we show the `calibrated' metallicity (column \texttt{c\_M\_H\_K}) from the \rave\ survey.
-Only stars meeting quality constraints in \emph{both} studies are shown (see text for details).\label{fig:rave-dr4-comparison}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
-\label{fig:gold-standard-comparison}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
-\caption{Hertsprung-Russell diagrams of stars in common between this work and that of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the metallicity of each study.
-Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
-This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth internal data release from the \ges\ survey, and this work.
-The number of stars in each panel are shown, as well as the bias and RMS deviation in each label.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
-Most of the \ges/\rave\ overlap stars have relatively low S/N ratios in \rave, near $\approx 30$~pixel$^{-1}$.\label{fig:ges-stellar-parameters}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-abundances.pdf}
-\caption{Detailed chemical abundances in the fourth internal data release from the \ges\ survey compared to this work.
-The number of stars shown in each panel is indicated, and the bias and RMS deviations are shown.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.\label{fig:ges-abundances}}
-\end{figure*}
-
-\begin{figure*}[p]
-\center
-\includegraphics[height=\textheight]{figures/globular-clusters.pdf}
-\caption{Effective temperature $\teff$ and surface gravity $\logg$ for globular cluster members identified by \citet{Kunder_2014,Anguiano_2015}.
-Left-hand panels indicate results from the fourth \rave\ data release \citep{Kordopatis_2013}, and the right-hand panels show results from this work.
-A representative isochrone is shown for each cluster \citep{Bressan_2012}.\label{fig:globular-cluster-HRD}}
-\end{figure*}
-
-\begin{figure*}[p]
-\center
-\includegraphics[height=\textheight]{figures/open-clusters.pdf}
-\caption{Effective temperature $\teff$ and surface gravity $\logg$ for open cluster members identified by \citep{Kunder_2016}.
-Left-hand panels indicate results from the fourth \rave\ data release \citep{Kordopatis_2013}, and the right-hand panels show results from this work.
-A representative isochrone is shown for each cluster \citep{Bressan_2012}.\label{fig:open-cluster-HRD}}
-\end{figure*}
-
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison with the literature calibration sources used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.
-Stars are colored by the S/N of the \rave\ spectra.
-Note that this comparison is for illustrative purposes only: it is not an indication of independent agreement with the literature because some metal-poor stars in this literature sample were used in the construction of our training set (see text for details).\label{fig:kordopatis-calibration}}
-\end{figure*}
-
-\begin{figure}[p]
-\center
-\includegraphics[width=0.65\textwidth]{figures/gce.pdf}
-\caption{Detailed chemical abundances ([X/Fe]) for giant stars in \unrave\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
-Note that the y-axis limits vary for each panel, however for clarity we show the scaled-solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
-Only stars with S/N $> 10$ and $\chi_r^2 < 3$ are shown.\label{fig:gce}}
-\end{figure}
-
-\begin{figure}[p]
-\center
-\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
-\caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV}.
-Hot stars are not included in our training set, and appear above the turnoff in our labels.
-For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars with $\teff > 8000$~K (as reported by \texttt{SPARV}), and other results that are known to be erroneous.\label{fig:hot-stars}}
-\end{figure}
-
-\end{document}
+We

### 2016-09-08 article/unrave.tex
Fix fuckup

@@
 with a low-order polynomial in label space.
-\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels.
-Note that we do not assume that \emph{every} label in the training is accurate -- we
+\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels
+for most, but not all stars. That is to say that we do not assume that \emph{every} 
+label in the training is accurate --- we
 can afford to have a small fraction of inaccurate labels (e.g., a few obvious 

### 2016-09-08 article/unrave.tex
Add BA to ack
Implement Seabroke suggestions #30

@@
 % RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid, Rosemary Wyse, Amina Helmi}
+\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid, Rosemary Wyse, Amina Helmi, George Seabroke}
 % Bland-Hawthorn: Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
@@
 % Helmi: Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands
+% Seabroke: Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK
 \author{and the \rave\ collaboration}
@@
 overlap samples.  Comparing the expected precision to what is currently available, 
-we further found that the \rave\ survey will benefit most from \gaia\ DR1: 63\%
+we further found that the \rave\ survey will benefit most from \gaia\ DR1: the
+parallaxes of 63\%
 of stars in the \rave--\gaia\ DR1 overlap sample ($\approx$182,862 stars) are 
@@
 
+\stub{Reference Figure \ref{fig:gold-standard-hrd}}
+
 \subsubsection{Comparison with the \ges\ survey}
@@
 
+\subsubsection{Comparison with the \rave\ DR4 calibration sample}
+
+\stub{Reference Figure\ref{fig:kordopatis-calibration}}
+
 \subsection{Astrophysical validation}
@@
 We thank 
+	Borja Anguiano (Macquarie),
 	Jonathan Bird (Vanderbilt),
@@
 \begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison with the literature calibration sources used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.
+Stars are colored by the S/N of the \rave\ spectra.
+Note that this comparison is for illustrative purposes only: it is not an indication of independent agreement with the literature because some metal-poor stars in this literature sample were used in the construction of our training set (see text for details).\label{fig:kordopatis-calibration}}
+\end{figure*}
+
+\begin{figure*}[p]
 \center
@@
 
-\begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison with the literature calibration sources used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.
-Stars are colored by the S/N of the \rave\ spectra.
-Note that this comparison is for illustrative purposes only: it is not an indication of independent agreement with the literature because some metal-poor stars in this literature sample were used in the construction of our training set (see text for details).\label{fig:kordopatis-calibration}}
-\end{figure*}
-
 \begin{figure}[p]

### 2016-09-08 article/unrave.tex
Some text edits up to sec 3.2

@@
 Freeman, Gerry Gilmore, Keith Hawkins, David W.
-Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
+Hogg, Sergey Koposov, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
 
@@
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-up to 292,036 stars.  Here we present a data-driven re-analysis of \rave\ 
-spectra, using an implementation of \thecannon, that yields more precise 
-and accurate stellar parameters and abundances than previous \rave\
-data releases.  We constructed our model using high-fidelity \apogee\ 
-stellar parameters and abundances for red giant branch stars that overlap
-with \rave, and stellar parameters from the \project{K2}/\project{Ecliptic 
-Pole Input Catalog} for main-sequence and sub-giant branch stars.
-We derive, and validate, improved effective temperature $\teff$, surface 
-gravity $\logg$, and chemical abundances of up to seven elements (O, Mg, 
-Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in 
-% TODO
-chemical abundances is X.XX~dex for red giant branch stars.  The \unrave\ 
-catalog presented here, in conjunction with \tgas, represents the 
-most powerful data set for chemo-orbital analysis at the dawn of broadly 
-available \gaia\ data.
+up to 292,036 stars.  We present a data-driven re-analysis of \rave\ 
+spectra using \thecannon, and we provide more precise and accurate stellar
+parameters and abundances than available in previous \rave\ data releases. 
+We constructed our model using high-fidelity \apogee\ stellar parameters 
+and abundances for red giant branch stars that overlap with \rave, and 
+stellar parameters from the \project{K2}/\project{Ecliptic Pole Input 
+Catalog} for main-sequence and sub-giant branch stars.  We derive, and 
+validate, improved effective temperature $\teff$, surface gravity $\logg$, 
+and chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
+for \Nstarsqc\ stars.  In total we report \Nabundances\ elemental 
+abundances, with a typical precision of 0.07~dex.  The \unrave\ catalog 
+presented here, in conjunction with \tgas, represents the most powerful 
+data set for chemo-orbital analysis at the dawn of broadly available 
+\gaia\ data.
 \end{abstract}
 
-\keywords{}
+\keywords{stars: fundamental parameters --- stars: abundances}
 
@@
 to infer the structure, kinematics, and chemical enrichment of the Galaxy 
-\citep[e.g.,][]{Nordstrom_2004,Schlaufman_2009,Deason_2011,Casagrande_2011,Ness_2012,Ness_2013a,Ness_2013b,
-Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,Boeche_2013,Kordopatis_2015,Bovy_2016}.  
-However, these quantities are not known for even 1\% of stars in the Milky Way.  
-Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
-Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics 
-from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
-stellar spectroscopists frequently report significantly different chemical 
-abundance patterns from the same spectrum \citep{Smiljanic_2014}.  The impact 
-these issues have on scientific inferences cannot be understated.  Imperfect 
-astrometry or chemistry limits understanding in a number of sub-fields in
-astrophysics, including the properties of exoplanet host stars, the formation 
-(and destruction) of stars and clusters, as well as the study of stellar
-populations and Galactic structure, to name a few.
+\citep[e.g.,][]{Nordstrom_2004,Schlaufman_2009,Deason_2011,Casagrande_2011,
+Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,Casey_2014a,Casey_2014b,
+Boeche_2013,Kordopatis_2015,Bovy_2016}.  However, these quantities are not known
+for even 1\% of stars in the Milky Way.  Stellar distances are famously imprecise 
+\citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,Madler_2016}, proper motions can be
+plagued by unquantified systematics from the first epoch observations 
+\citep[e.g.,][]{Casey_Schlaufman_2015}, and stellar spectroscopists frequently 
+report significantly different chemical abundance patterns from the same spectrum 
+\citep{Smiljanic_2014}.  The impact these issues have on scientific inferences 
+cannot be understated.  Imperfect astrometry or chemistry limits understanding 
+in a number of sub-fields in astrophysics, including the properties of exoplanet 
+host stars, the formation (and destruction) of star clusters, as well as studies
+of stellar populations and Galactic structure, to name a few.
 
@@
 larger than its predecessor \hipparcos, both astrometry and chemistry are 
-required to fully characterize the formation and evolution of our Galaxy. 
+required to fully characterize the formation and evolution of the Milky Way. 
 \gaia\ will also provide radial velocities, stellar parameters and chemical 
@@
 we further found that the \rave\ survey will benefit most from \gaia\ DR1: the
-parallaxes of 63\%
-of stars in the \rave--\gaia\ DR1 overlap sample ($\approx$182,862 stars) are 
-expected to improve with the first \gaia\ data release, and 47\% of stars ($\approx$137,211 stars) are 
-likely to have better proper motions.  Although the \gaia\ universe
-model assumes end-of-mission uncertainties --- and does not account for systematics
-in the first data release --- this calculation still provides intuition for the 
-relative improvement the first \gaia\ data release can make to ground-based surveys.  
-The expected improvements for \rave\ motivated us to examine what chemical abundances
-were available from those data, and to evaluate whether we could enable new 
-chemo-dynamic studies by contributing to the existing set of chemical abundances.
+distances of 63\%of stars in the \rave--\gaia\ DR1 overlap sample 
+($\approx$182,862 stars) are expected to improve with the first \gaia\ data release, 
+and 47\% of stars ($\approx$137,211 stars) are likely to have better proper motions. 
+Although the \gaia\ universe model assumes end-of-mission uncertainties --- and does 
+not account for systematics in the first data release --- this calculation still 
+provides intuition for the relative improvement that the first \gaia\ data release 
+can make to ground-based surveys.  The expected improvements for \rave\ motivated 
+us to examine what chemical abundances were available from those data, and to 
+evaluate whether we could enable new chemo-dynamic studies by contributing to the
+existing set of chemical abundances.
 
@@
 of these comparisons in Section~\ref{sec:discussion}, and conclude with instructions
-for how to access our results electronically.
+on how to access our results electronically.
 
@@
 wavelength range that \rave\ operates in:  8410--8795~\AA.  The resolution and 
-wavelength coverage is comparable to the Radial Velocity Spectrometer on board 
-the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011,Recio-Blanco_2016}, 
+wavelength coverage of \rave\ is comparable to the Radial Velocity Spectrometer on
+board the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011,Recio-Blanco_2016}, 
 and the wavelength range overlaps with one of the key setups used for the ground-based 
-high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}.  The region 
-includes the \ion{Ca}{2} near infrared triplet lines --- strong transitions that 
+high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}.  The spectral region 
+also includes the \ion{Ca}{2} near infrared triplet lines --- strong transitions that 
 are dominated by pressure broadening --- which are visible even in metal-poor stars
 or spectra with very low signal-to-noise (S/N) ratios.  Atomic transitions of 
-light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
+light-, $\alpha$-, and Fe-peak elements are present as well, allowing for detailed 
 chemical abundance studies.
@@
 surveys, the \rave\ data releases have provided excellent radial velocities, 
-stellar astrophysical parameters ($\teff$, $\logg$), as well as individual 
-chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,Boeche_2011,
-Kordopatis_2013,Kunder_2016}.  In this work we make use of spectra
-that has been reprocessed for the fifth \rave\ data release.  These re-processing
-steps include: a detailed re-reduction of all the original data frames, with flux
-variances propagated at every step; an updated continuum-normalization procedure;
-as well as revised determinations of stellar radial velocities and morphological
-classifications. At the end of this processing for each survey observation we were 
-provided with: rest-frame wavelengths (without resampling), continuum-normalized 
-fluxes, $1\sigma$ uncertainties in the continuum-normalized flux values, as well 
-as relevant metadata for each observation.  We refer the reader to the official 
-fifth data release paper of the \rave\ survey, as presented by \citet{Kunder_2016}, 
-for more details of this re-processing.
+stellar atmospheric parameters ($\teff$, $\logg$), and detailed chemical abundances
+\citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,Boeche_2011,Kordopatis_2013,
+Kunder_2016}.  In this work we make use of spectra that has been reprocessed for 
+the fifth \rave\ data release.  These re-processing steps include: a detailed 
+re-reduction of all the original data frames, with flux variances propagated at 
+every step; an updated continuum-normalization procedure; as well as revised 
+determinations of stellar radial velocities and morphological classifications. 
+At the end of this processing for each survey observation we were provided with:
+rest-frame wavelengths (without resampling), continuum-normalized fluxes, $1\sigma$
+uncertainties in the continuum-normalized flux values, as well as relevant metadata
+for each observation.  We refer the reader to the official fifth data release paper
+of the \rave\ survey, as presented by \citet{Kunder_2016}, for more details of this
+re-processing.
 
@@
 outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
-there is a limited number of pre-processing steps that we performed before starting
+there were a limited number of pre-processing steps that we performed before starting
 our analysis.  First, we calculated inverse variance arrays from the $1\sigma$ 
@@
 
-We chose to adopt a data-driven model for this analysis, rather than the
-physics-based models used in the \rave\ data releases.  Specifically, we will
-use an implementation of \thecannon\
-\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
-of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven approach 
-makes use of all available information in the spectrum and lowers the S/N ratio 
-at which systematic effects begin to dominate.  In other words, in the low S/N
-regime, a well-constructed data-driven model will yield more precise \emph{labels}
-(e.g., stellar parameters and chemical abundances) than most physics-driven 
-models\footnote{However, see \citet{Casey_2016a}.}.  This is particularly relevant 
-for the low-resolution \rave\ data analysed here, because about half of the 
-spectra have S/N $\lesssim 50$~pixel$^{-1}$.
+We chose to adopt a data-driven model for this analysis, in contrast to the
+physics-based models used in \rave\ data releases to date.  Specifically, we will
+use an implementation of \thecannon\ \citep{Ness_2015,Ness_2016}.  Although this 
+choice complicated the construction of our model (e.g., see Section 
+\ref{sec:the-training-set}), a data-driven approach makes use of all available 
+information in the spectrum and lowers the S/N ratio at which systematic effects 
+begin to dominate.  In other words, in the low S/N regime, a well-constructed 
+data-driven model will yield more precise \emph{labels} (e.g., stellar parameters 
+and chemical abundances) than most physics-driven models\footnote{However, see 
+\citet{Casey_2016a}.}.  This is particularly relevant for the low-resolution 
+\rave\ data analysed here, because about half of the spectra have S/N 
+$\lesssim 50$~pixel$^{-1}$.
 
@@
 for most, but not all stars. That is to say that we do not assume that \emph{every} 
-label in the training is accurate --- we
-can afford to have a small fraction of inaccurate labels (e.g., a few obvious 
-misclassifications in the training set are affordable).
+label in the training is accurate. We can afford to have a small fraction of 
+inaccurate labels; few obvious misclassifications in the training set are affordable.
 \item We assume that the training data are similar (in spectra) to the test data 
@@
 giants according to \apogee.  From this sample we selected stars to have: 
-determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, and Ni); S/N ratios of $>$200 in \apogee\ and $>$25 in \rave; and 
-we further required that \aspcap\ did not report any peculiar flags 
+determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, 
+and Ni); S/N ratios of $>$200~pixel$^{-1}$ in \apogee\ and $>$25~pixel$^{-1}$ in \rave; 
+and we further required that \aspcap\ did not report any peculiar flags 
 (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
@@
 iron abundance [Fe/H] labels were adopted from \citet{Fulbright_2010, Ruchti_2011}.
-The elemental abundances for O, Mg, Al, Si, Ca, and Ni were assumed to follow typical
-trends of Galactic chemical evolution, such that we assumed $[{\rm Mg/Fe}] = +0.4$,
-$[{\rm O/Fe}] = +0.4$, $[{\rm Al/Fe}] = -0.5$, $[{\rm Ca/Fe}] = +0.4$, 
-$[{\rm Si/Fe}] = +0.4$, and $[{\rm Ni/Fe}] = -0.25$.  This decision is made solely 
-to ensure that our overall metallicity scale extends that of the \rave\ survey, down 
-to $[{\rm Fe/H}] \sim -4$.  In other words, we do not recommend the use of individual 
-abundance labels at $[{\rm Fe/H}] \sim -4$; in practice the abundances of most of these
-elements are largely unrecoverable in the \rave\ wavelength region at $[{\rm Fe/H}] \sim -4$.
-We discuss this issue in more detail in Section~\ref{sec:discussion}.
+For this sample of metal-poor stars, we assumed that the elemental abundances of O, 
+Mg, Al, Si, Ca, and Ni followed typical trends of Galactic chemical evolution: we
+asserted $[{\rm Mg/Fe}] = +0.4$, $[{\rm O/Fe}] = +0.4$, $[{\rm Al/Fe}] = -0.5$, 
+$[{\rm Ca/Fe}] = +0.4$, $[{\rm Si/Fe}] = +0.4$, and $[{\rm Ni/Fe}] = -0.25$.  We stress
+that this decision is made solely to ensure that our overall metallicity scale reflects
+that of the \rave\ survey, down to $[{\rm Fe/H}] \sim -4$.  Indeed, it is likely that 
+for most of these elements, these abundances cannot be measured from \rave\ spectra for 
+ultra metal-poor stars: the atomic transitions in the \rave\ spectral region are simply
+too weak.  For this reason, our adopted abundances for these very metal-poor stars
+represent an `anchor point' in order to ensure our overall metallicity scale is
+correct.  We do not recommend the use of our individual abundance labels at 
+$[{\rm Fe/H}] \sim -4$.  We discuss this issue in more detail in Section~\ref{sec:discussion}.
 
@@
 are interested in (e.g., FGKM-type stars), and which have a large enough sample size 
-overlapping with \rave.  Moreover, most of the spectroscopic studies we considered also 
+that overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
 showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
@@
 \epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
-benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
-classification.  Despite this limitation the labels in the \epic\ catalog have 
+benefit from having narrow-band $DDO_{51}$ photometry to aid dwarf/giant 
+classification.  Despite this limitation, the labels in the \epic\ catalog have 
 already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
@@
 allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
-employed the following quality constraints: the upper and lower confidence intervals 
-in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
-must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
-30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
-removed most metal-poor stars, which we later found to cause the test labels to have
-under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
-(ignored) those quality constraints for stars with $[{\rm Fe/H}] < -1$, and included
-an additional 12 turn-off stars with $-1.6 \gtrsim [{\rm Fe/H}] \gtrsim -2.1$ from 
-\citet{Ruchti_2011}.  After training
+employed the following quality constraints on the \epic\ catalog: the upper and lower 
+confidence intervals in $\teff$ must be below 150~K; the upper and lower confidence 
+intervals in $\logg$ must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must
+exceed 30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict 
+constraints removed most metal-poor stars, which we later found to cause the test 
+labels to have under-predicted abundances for dwarfs of low metallicity.  For this 
+reason we relaxed (ignored) those quality constraints for stars with 
+$[{\rm Fe/H}] < -1$, and included an additional 12 turn-off stars with 
+$-1.6 \gtrsim [{\rm Fe/H}] \gtrsim -2.1$ from \citet{Ruchti_2011}.  After training
 a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
@@
 was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 595 
-stars (583 of 4,611 from \epic).  The full training set for each model (see next sections) is shown
-in Figure~\ref{fig:training-set-hrd}.
+stars (583 of 4,611 from \epic).  The full training set for each model (see next 
+sections) is shown in Figure~\ref{fig:training-set-hrd}.
 
@@
 we describe how we will use the 3-label model of the main-sequence and giant branch
-(introduced in this section) to discriminate between results from a 3-label 
-main-sequence model in Section \ref{sec:unevolved-star-model} and a 9-label giant 
-star model in Section \ref{sec:evolved-star-model}.
+(introduced in this section) to discriminate between results from a 3-label main-sequence 
+model in Section \ref{sec:unevolved-star-model} and a 9-label giant star model in Section 
+\ref{sec:evolved-star-model}.
 
@@
 We constructed a three-label quadratic model using only main-sequence and sub-giant
-stars. In order to set the regularization hyperparameter
-$\Lambda$ for this model, we trained 30 models with different regularization strengths,
-spaced evenly in logarithmic steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
-We then performed leave-one-out cross-validation for each model.  Specifically, for 
-each star in the training set: we removed the star; trained the model; and then 
-inferred labels from the removed star as if it was a test object. We also performed 
-leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
-will use as the basis for comparison.  For the unregularized case, we calculated the 
-bias and root-mean-square (RMS) deviation between: the training set labels, and the 
-labels we derive by cross-validation, where one star is removed at a time and the
-model is re-trained.  We repeated this calculation of bias and RMS deviation for all
-30 models with different regularization strengths $\Lambda$.
+stars. In order to set the regularization hyperparameter $\Lambda$ for this model, we 
+trained 30 models with different regularization strengths, spaced evenly in logarithmic
+steps between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.  We then performed leave-one-out 
+cross-validation for each model.  Specifically, for each star in the training set: we 
+removed the star; trained the model; and then inferred labels from the removed star as 
+if it was a test object.  We also performed leave-one-out cross-validation on an 
+unregularized ($\Lambda = 0$) model, which we will use as the basis for comparison.  
+For the unregularized case, we calculated the bias and root-mean-square (RMS) deviation 
+between: the training set labels, and the labels we derived by cross-validation, where 
+one star was removed at a time and the model was re-trained.  We repeated this calculation
+of bias and RMS deviation for all 30 models with different regularization strengths $\Lambda$.
 
@@
 
-We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  
+We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  \stub{from Zwitter: ensure it is clear that we say we are running this model on all spectra}
 The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  

### 2016-09-08 article/unrave.tex
Finalzie Zwitter comments. Closes #28

@@
 
-We inferred labels for all \rave\ spectra using this main-sequence/sub-giant star model.  \stub{from Zwitter: ensure it is clear that we say we are running this model on all spectra}
-The results for the survey sample are shown in the centre panel of Figure \ref{fig:test-set-density}.  
-The increased density of solar-type stars is consistent with \rave\
-observing stars in the local neighbourhood, and the high number of turn-off
-and main-sequence stars relative to the sub-giant branch is expected from 
-the relative lifetimes of these evolutionary phases.  An over-density
+We inferred labels for all \Nspectra\ \rave\ spectra using this main-sequence/sub-giant
+star model; we made no \emph{a priori} decisions as to whether a star was likely a
+main-sequence/sub-giant star or not.  The results for the entire survey sample are shown
+in the centre panel of Figure \ref{fig:test-set-density}.  The increased density of 
+solar-type stars is consistent with \rave\ observing stars in the local neighbourhood, 
+and the high number of turn-off and main-sequence stars relative to the sub-giant branch 
+is expected from the relative lifetimes of these evolutionary phases.  An over-density
 of stars near the base of the giant branch is also present.  This artefact is due 
-to having giant stars in the test set, but not in the training set, and the
-model is (poorly) extrapolating outside the convex hull of the training set.
+to having giant stars in the test set, but not in the training set, and the model is 
+(poorly) extrapolating outside the convex hull of the training set.
 
@@
 at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
-between $0.07-0.09$~dex depending on the elemental abundance.  We inferred labels 
-for all \Nspectra\ \rave\ spectra using this model, and these results are summarized in
-the right panel of Figure \ref{fig:test-set-density}.  The red clump is clearly visible
-and in the expected location, without requiring any post-analysis calibration.
-However, artefacts due to dwarf stars in the test set are also present.
+between $0.07-0.09$~dex depending on the elemental abundance.
+
+We inferred labels for all \Nspectra\ \rave\ spectra using this model, again without
+regard for whether a star was likely a giant or not.  The results for all survey stars
+are summarized in the right panel of Figure \ref{fig:test-set-density}.  The red clump
+is clearly visible and in the expected location, without requiring any post-analysis 
+calibration.  However, artefacts due to dwarf stars being present in the test set,
+and not in the training set, are also present.
 

### 2016-09-08 article/unrave.tex
text edits

@@
 
-\subsection{Joining the models}
+\subsection{Deriving joint estimates from all models}
 \label{sec:joining-the-models}
@@
 We have derived labels for all \Nspectra\ \rave\ spectra using the three models
-described in previous sections.  The results from the joint model --- that includes 
-the main-sequence, sub-giant and red giant branch --- shows that a single 3-label quadratic
-model is too simple for the \rave\ spectral range.  The other models have problems,
-too: unrealistic over-densities in label space show that the main-sequence model and
-the giant model make very poor extrapolations for stars outside their respective
-training sets.  For these reasons we were forced to exclude or severely penalize 
-incorrect results from both models.  We emphasize that the choices here are entirely
-heuristic, and depart from interpreting \thecannon\ output labels as the maxima
-of individual likelihood functions.  Each model produces estimates of the labels
-for a given star, and we use those estimates to produce a unified estimate, but 
-this joint estimate is calculated by disregarding the probabilistic attributes
-of individual estimates.
-
-Before attempting to join the results from different models, we excluded results
+described in previous sections.  The results from our first model 
+(Section~\ref{sec:a-simple-model}) --- which includes the main-sequence, sub-giant 
+and red giant branch --- shows that a single 3-label quadratic model is too simple 
+for the \rave\ spectral range.   The other models have problems, too: unrealistic 
+over-densities in label space show that the main-sequence model and the giant model
+make very poor extrapolations for stars outside their respective training sets.  
+For these reasons we were forced to exclude or severely penalize incorrect results
+from both models.  We emphasize that the choices here are entirely heuristic, 
+and depart from interpreting \thecannon\ output labels as the maxima of individual 
+likelihood functions.  Each model produces estimates of the labels for a given star, 
+and we use those estimates to produce a unified estimate, but this joint estimate 
+is calculated by disregarding the probabilistic attributes of individual estimates.
+
+Before attempting to join the results from the models in Sections 
+\ref{sec:unevolved-star-model} and \ref{sec:evolved-star-model}, we excluded results
 in either model that had a reduced $\chi_{r}^2 > 3$.  We further discarded stars with
@@
 $\logg > 3.5$, and for the results from the main-sequence model we excluded 
-sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the 
-two-dimensional ($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  Unfortunately these restrictions did not remove all spurious
-results.  The reason for this can be explained with an example:  consider that our 
-giant star model was trained with only giant stars but tested with both giant 
-stars and dwarf stars.  Some classes of stars (e.g., metal-poor dwarfs) can 
-project into a region of label space that would suggest it is a giant 
-(e.g., a clump star).  These objects could have relatively low $\chi_{r}^2$ values
-(e.g., $\chi_{r}^2 < 3$) and in this example, they would appear as bonafide red clump 
-stars.  These incorrect projections are extrapolation errors in high dimensions that
-project to `normal' parts of the label space in two dimensions.  For these reasons 
-we also made use of the joint model in Section~\ref{sec:a-simple-model}
-to inform whether we should adopt results from: the red giant branch model; the
-main-sequence/sub-giant model; or a linear combination of the two.
-
-In Figure \ref{fig:joint-model-differences} we show the differences in effective temperature
-$\teff$ and surface gravity $\logg$ between: the main-sequence model and the joint
-model; and the differences between the red giant branch model and the joint model. 
-We have scaled the differences in $\teff$ and $\logg$ to make the central peak 
-near $(0, 0)$ to be approximately isotropic by setting
+sub-giant stars ($\logg < 4$ and $\teff < 5000$~K) that were outside the two-dimensional 
+($\teff$, $\logg$) convex hull of the training set used for the main-sequence model.  
+Unfortunately these restrictions did not remove all spurious results.  The reason 
+for this can be explained with an example:  consider that our giant star model was 
+trained with only giant stars but tested with both giant stars and dwarf stars.  
+Some classes of stars (e.g., metal-poor dwarfs) can project into a region of label 
+space that would suggest it is a giant (e.g., a clump star).  These objects could 
+have relatively low $\chi_{r}^2$ values (e.g., $\chi_{r}^2 < 3$) and in this example,
+they would appear as bonafide red clump stars.  These incorrect projections are 
+extrapolation errors in high dimensions that project to `normal' parts of the label 
+space in two dimensions.  For these reasons we also made use of the simple model in 
+Section~\ref{sec:a-simple-model} to inform whether we should adopt results from: 
+the red giant branch model; the main-sequence/sub-giant model; or a linear combination 
+of the two.
+
+In Figure \ref{fig:joint-model-differences} we show the differences in effective 
+temperature $\teff$ and surface gravity $\logg$ between: the main-sequence model 
+(Section~\ref{sec:unevolved-star-model}) and the simplest model (Section~
+\ref{sec:a-simple-model}); and the differences between the red giant branch model 
+(Section~\ref{sec:evolved-star-model}) and the simplest model (Section 
+\ref{sec:a-simple-model}).  We have scaled the differences in $\teff$ and $\logg$ 
+to make the central peak near $(0, 0)$ to be approximately isotropic by setting
 	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
@@
 	$\delta_{\log{g}} = 0.15$~dex for both models.
-It is important to note that these $\delta$ values do not represent any kind of intrinsic
-uncertainty or precision: they are merely normalization factors.  Adopting
+It is important to note that these $\delta$ values do not represent any kind of 
+intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
 substantially different scaling factors would produce in clear inconsistencies in our
@@
 approximately isotropic, and to some extent, comparable.  In Figure 
-\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent objects where
-the joint model and the comparison model both report similar labels.  
-The artefacts 
-seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} are
-also present in Figure \ref{fig:joint-model-differences} as over-densities far away
-from the central peak.  Therefore, we can 
-adopt the scaled distance in labels $\teff$ and $\logg$ from the joint model to the 
-main-sequence model $d_{ms}$,
+\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
+objects where the simple model and the comparison model both report similar labels.  
+The artefacts seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} 
+are also present in Figure \ref{fig:joint-model-differences} as over-densities far away
+from the central peak.  Therefore, we can adopt the scaled distance in labels 
+$\teff$ and $\logg$ from the joint model to the main-sequence model $d_{ms}$,
 \begin{eqnarray}
@@
 
+\stub{Warning: The metallicity correction may disappear -- I am trying to fix that}
 However when adopting a similar distance metric for the giant model, we found that the

### 2016-09-08 article/unrave.tex
Update text and figures to v0.96

@@
 from the central peak.  Therefore, we can adopt the scaled distance in labels 
-$\teff$ and $\logg$ from the joint model to the main-sequence model $d_{ms}$,
+$\teff$ and $\logg$ from the simplest model to the main-sequence model $d_{ms}$,
 \begin{eqnarray}
-	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad .
+	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad ,
 \end{eqnarray}
 
-\stub{Warning: The metallicity correction may disappear -- I am trying to fix that}
-However when adopting a similar distance metric for the giant model, we found that the
-giant model and the joint model produced similar label estimates (with low $\chi_r^2$
-values) for metal-poor ($[{\rm Fe/H]} < -1$) dwarf stars near the turn-off.  The similarity
-in estimated labels occurred in $\teff$, $\logg$, \emph{and} [Fe/H], implying that we
-could not use the difference between $[{\rm Fe/H}]_{giant} - [{\rm Fe/H]}]_{joint}$
-as a term that would benefit the main-sequence model.  As a result, our joint estimate 
-for metal-poor dwarf stars that overlap with high-resolution studies
-\citep{Reddy_2003,Reddy_2006,Bensby_2014,Valenti_Fischer_2005},
-would appear as metal-poor giant stars, because the giant star model would have a much
-smaller distance metric than the main-sequence model.  It is reasonable to assert that
-this issue may be complicated by the increased complexity, or flexibility, of the giant
-star model relative to the main-sequence model.  In any case, we found that by empirically
-penalizing the giant model distance metric by absolute metallicity,
+\noindent{}and the scaled distance from the simplest model to the giant model $d_{giant}$,
 \begin{eqnarray}
-	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  + \lvert[{\rm Fe/H}]_{joint}\rvert  \quad ,
+	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  \quad ,
+\label{eq:d_giant}
 \end{eqnarray}
 
-\noindent{}the weighting of dwarfs was no longer a function of joint metallicity, and this
-issue was resolved without introducing new artefacts. Using these distance metrics we 
-then derived the weights,
+\noindent{}to derive the weights,
 \begin{eqnarray}
@@
 and mean metallicity.  The artefacts from individual models are no longer 
-apparent, and the rich structure of the Hertzsprung-Russell is visible.  Despite
-this, there are a number of caveats introduced by the choices we have made to
+apparent, and stars across the full Hertzsprung-Russell diagram are visible.
+However, there are a number of caveats introduced by the choices we have made to
 combine estimates from multiple models.  We discuss these issues in detail in
@@
 of observations for any target.  These repeat observations allow us to 
-quantify the level of (in)correctness in our formal errors.  For every star
+quantify the level of (in)correctness in our formal errors.
+
+For every star
 with multiple visits we constructed a high S/N stacked spectrum for that

### 2016-09-08 article/figures/rave_io.py

@@ -26,7 +26,7 @@ def get_kordopatis_comparisons():
 
 def get_cannon_dr1(filename=None):
 
-    filename = filename or os.path.join(DATA_PATH, "unrave-v0.95.fits.gz")
+    filename = filename or os.path.join(DATA_PATH, "../../unrave-v0.96.fits.gz")
 
     rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, filename))
 

### 2016-09-08 combine_model_results.py

@@ -19,7 +19,7 @@ RESULTS_PATH = ""
 
 ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v43.fits.gz"))
 giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
-joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v44.fits.gz"))
+joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v45.fits.gz"))
 
 
 for t in (ms_results, giant_results, joint_results):
@@ -206,7 +206,7 @@ axes[1].hexbin(x2, y2, gridsize=100, extent=(-3, +3, -3, +3), norm=LogNorm(), li
 
 
 ms_distance = np.sqrt(x**2 + y**2)
-giant_distance = np.sqrt(x2**2 + y2**2 + z2**2)
+giant_distance = np.sqrt(x2**2 + y2**2)
 
 
 # Weight by relative distance, or just take the closest of the two?
@@ -393,11 +393,13 @@ for label_name in abundance_label_names:
     combined_table["E_{}".format(label_name)][metal_poor] = np.nan
     
 
+
+"""
 # Make a figure
 ok = (combined_table["SNR"] >= 10) * (combined_table["R_CHI_SQ"] < 3)
 fig, ax = plt.subplots()
 hexbin = ax.hexbin(combined_table["TEFF"][ok], combined_table["LOGG"][ok], 
-    C=combined_table["Teff_SPARV"][ok], reduce_C_function=np.nanmax,
+    C=combined_tpwable["Teff_SPARV"][ok], reduce_C_function=np.nanmax,
     gridsize=50, vmin=4000, vmax=10000, extent=(3500, 7500, 0, 5.5),
     cmap="plasma",)
 
@@ -413,7 +415,7 @@ cbar.set_label(r"$T_{{\rm eff},SPARV}$ $[{\rm K}]$")
 fig.tight_layout()
 fig.savefig("article/figures/hot-stars.pdf", dpi=300)
 fig.savefig("article/figures/hot-stars.png")
-
+"""
 
 # Remove the SPARV temperature.
 del combined_table["Teff_SPARV"]
@@ -424,7 +426,7 @@ del combined_table["Teff_SPARV"]
 #    combined_table[column][hot] = np.nan
 #    combined_table["E_{}".format(column)] = np.nan
 
[…]

### 2016-09-08 article/unrave.tex
Add text about errors

@@
 
-The weighted $\teff$ and $\logg$ values for stars meeting different
-S/N constraints are shown in Figure \ref{fig:test-set-hrd}, both in logarithmic density
-and mean metallicity.  The artefacts from individual models are no longer 
-apparent, and stars across the full Hertzsprung-Russell diagram are visible.
-However, there are a number of caveats introduced by the choices we have made to
-combine estimates from multiple models.  We discuss these issues in detail in
-Section \ref{sec:discussion}.
+The weighted $\teff$ and $\logg$ values for stars meeting different S/N 
+constraints are shown in Figure \ref{fig:test-set-hrd}, both in logarithmic 
+density and mean metallicity.  The artefacts from individual models are no 
+longer apparent, and the complete structure of the Hertzsprung-Russell diagram
+is visible.  However, there are a number of caveats introduced by the choices 
+we have made to combine estimates from multiple models.  We discuss these 
+issues in detail in Section \ref{sec:discussion}.
 
@@
 
-\subsection{Multi-epoch observations}
+\subsection{Internal validation}
+\subsubsection{Repeat observations}
 \label{sec:repeat-observations}
@@
 
-For every star
-with multiple visits we constructed a high S/N stacked spectrum for that
-star, weighted by the inverse variances in the individual visit spectra.
-
-We treated the stacked spectra as normal survey stars.  We inferred labels
-using all models and joined the results as per Section \ref{sec:joining-the-models}.  The labels inferred from the stacked spectra served as the basis
-of comparison for all individual visits to that object.  Figure 
-\ref{fig:formal-errors-comparison} shows the difference in labels 
-between the comparison spectrum and a repeat visit, normalized by their 
-formal errors summed in quadrature (e.g., 
-$\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
-If our measurements were unbiased by S/N and the formal errors were 
-representative, these values should be normally distributed with a zero 
-mean and unit variance.
-\stub{Are they? (Probably not!). How much should we inflate our formal errors, and should this be as a function of S/N? Note that in all Figures and tables,
-we show these inflated errors only, as we gauge them to be more representative.}
-
-\stub{In Figure \ref{fig:repeat-visits} we look at the difference as a function of SNR}
+We calculated all possible pair-wise differences between the labels we
+derived from multiple visits.  If \rave\ observed a star $H$ times, there 
+are $H!/2(H-2)!$ possible $a$-$b$ pair-wise combinations where we can calculate 
+the difference between the derived label (for example, $\logg$) over the
+quadrature sum of their formal uncertainties,
+\begin{eqnarray}
+\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{\sigma_{\logg,a}^2 + \sigma_{\logg,b}^2}} \quad .
+\end{eqnarray}
+
+If our derived labels and formal errors were correct, the distribution of
+$\eta$ values would represent a normal distribution with zero mean and
+unity variance.  However, our formal errors are likely to be under-estimated,
+and therefore we introduce a systematic error floor for each label, which
+is added in quadrature to every observation such that for $\logg$,
+\begin{eqnarray}
+	\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{(\sigma_{\logg,a} + \sigma_{\logg,floor})^2 + (\sigma_{\logg,b} + \sigma_{\logg,floor})^2}} \quad .
+\end{eqnarray}
+
+We increased the minimum label error until the distribution of $\eta$ 
+values for a given label approximately reached unity.  The minimum error
+values for each label are shown with the distribution of (new) $\eta$
+values in Figure \ref{fig:pairwise-metrics}.  These minimum values form
+part of our error model, and \emph{have} been added in quadrature with 
+the formal errors; the quoted errors in our labels already incorporate
+these minimum errors.
+
+\subsubsection{Precision as a function of S/N ratio}
+\label{sec:precision-wrt-snr}
+
+We also used the repeat observations to build intuition for the label
+precision that was achievable as a function of S/N ratio.  Specifically
+we stacked all spectra for a given star by summing the fluxes weighted
+by the inverse variances, then treated the stacked spectra as a normal
+survey star.  We inferred labels for the stacked spectra using all models,
+and derived a joint estimate as per Section~\ref{sec:joining-the-models}.
+The labels we inferred from the stacked spectrum serve as a basis of
+comparison for the labels we derived from the individual visit spectra,
+which are of lower S/N ratios.
+
+In Figure \ref{fig:formal-errors-comparison} we show the RMS difference
+in labels between the stacked spectrum and a single visit, binned by the
+S/N ratio of the individual visit spectrum.  Here we only show stars
+where the \emph{stacked} spectrum had S/N $>100$~pixel$^{-1}$ to ensure
+that our baseline comparisons were in a region where we are dominated
+by systematic uncertainties.  The precision in all labels tends to
+flatten out past S/N $> 40$~pixel$^{-1}$, and the precision at high S/N
+ratios is comparable to the minimum error floors we adopted in Section
+\ref{fig:repeat-observations}.
 

### 2016-09-08 article/unrave.tex
Fix caption of Fig 8

@@
 \includegraphics[width=\textwidth]{figures/repeat-visits.pdf}
-\caption{The RMS of the test set labels as a function of S/N ratio for repeated stars in the test set.\label{fig:repeat-visits}}
+\caption{The RMS deviation of labels for repeated observations in the test set.  The RMS deviation is binned as a function of the S/N ratio of the individual visit spectra.  The
+precision flattens out at S/N $\gtrsim 40$~pixel$^{-1}$, where our results become systematics-limited.
+\label{fig:repeat-visits}}
 \end{figure*}

### 2016-09-08 article/unrave.tex
Add words

@@
 values for each label are shown with the distribution of (new) $\eta$
-values in Figure \ref{fig:pairwise-metrics}.  These minimum values form
+values in Figure \ref{fig:pairwise-comparison}.  These minimum values form
 part of our error model, and \emph{have} been added in quadrature with 
@@
 
-In Figure \ref{fig:formal-errors-comparison} we show the RMS difference
+In Figure \ref{fig:repeat-visits} we show the RMS difference
 in labels between the stacked spectrum and a single visit, binned by the
@@
 ratios is comparable to the minimum error floors we adopted in Section
-\ref{fig:repeat-observations}.
+\ref{sec:repeat-observations}.  The median S/N of \rave\ spectra is
+50~pixel$^{-1}$, at which point our abundance precision is $\approx0.1$~dex,
+varying a few tenths of a dex between elements.
 
@@
 \begin{figure*}[p]
-\caption{Distribution of differences in label estimates from multiple visits, divided by their formal errors summed in quadrature.
-If our measurements were unbiased and the formal errors were representative, these distribution should be normally distributed with zero mean and unit variance.\label{fig:formal-errors-comparison}}
+\caption{Distribution of differences in label estimates from multiple visits, divided by the quadrature sum of their formal errors, and a minimum error value for each observation.
+If our measurements were unbiased and our errors were representative, no minimum error floor would be required and this distribution would by normally-distributed with zero mean and unity variance.
+We increased the error floor for a given label until the distribution variance approximately reached unity. The error floors are given for each label.\label{fig:pairwise-comparison}}
 \end{figure*}

### 2016-09-08 article/unrave.tex
Include error floor figures

@@
 derived from multiple visits.  If \rave\ observed a star $H$ times, there 
-are $H!/2(H-2)!$ possible $a$-$b$ pair-wise combinations where we can calculate 
+are $H!/2(H-2)!$ possible $a$-to-$b$ pair-wise combinations where we can calculate 
 the difference between the derived label (for example, $\logg$) over the
-quadrature sum of their formal uncertainties,
-\begin{eqnarray}
-\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{\sigma_{\logg,a}^2 + \sigma_{\logg,b}^2}} \quad .
-\end{eqnarray}
-
-If our derived labels and formal errors were correct, the distribution of
-$\eta$ values would represent a normal distribution with zero mean and
-unity variance.  However, our formal errors are likely to be under-estimated,
+quadrature sum of their formal uncertainties $(\logg_a - \logg_b)/\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}$.  If our derived labels and formal errors were 
+correct, the distribution of these pair-wise comparisons would be
+well-represented by a Gaussian distribution with zero mean and variance
+of unity.  However, our formal errors are likely to be under-estimated,
 and therefore we introduce a systematic error floor for each label, which
-is added in quadrature to every observation such that for $\logg$,
+is added in quadrature to every observation such that (for example, $\logg$),
 \begin{eqnarray}
@@
 
+%\begin{eqnarray}
+%\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}} \quad .
+%\end{eqnarray}
+
 We increased the minimum label error until the distribution of $\eta$ 
 values for a given label approximately reached unity.  The minimum error
-values for each label are shown with the distribution of (new) $\eta$
+values for each label are shown with the distribution of $\eta$
 values in Figure \ref{fig:pairwise-comparison}.  These minimum values form
@@
 
-We also used the repeat observations to build intuition for the label
-precision that was achievable as a function of S/N ratio.  Specifically
-we stacked all spectra for a given star by summing the fluxes weighted
-by the inverse variances, then treated the stacked spectra as a normal
-survey star.  We inferred labels for the stacked spectra using all models,
-and derived a joint estimate as per Section~\ref{sec:joining-the-models}.
-The labels we inferred from the stacked spectrum serve as a basis of
-comparison for the labels we derived from the individual visit spectra,
-which are of lower S/N ratios.
+We further used the repeat observations in \rave\ to build intuition for
+the label precision that was achievable as a function of S/N ratio.  
+Specifically we stacked all spectra for a given star by summing the 
+fluxes weighted by the inverse variances, then treated the stacked spectra
+as a normal survey star.  We inferred labels for the stacked spectra 
+using all three models, and derived a joint estimate as per Section 
+\ref{sec:joining-the-models}.  The labels we inferred from the stacked 
+spectrum then served as a basis of comparison for the labels we derived 
+from the individual visit spectra of the same star, which are of lower 
+S/N ratios.
 
@@
 \begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/pairwise-metrics.pdf}
 \caption{Distribution of differences in label estimates from multiple visits, divided by the quadrature sum of their formal errors, and a minimum error value for each observation.

### 2016-09-08 article/unrave.tex
calibration comparison

@@
 the difference between the derived label (for example, $\logg$) over the
-quadrature sum of their formal uncertainties $(\logg_a - \logg_b)/\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}$.  If our derived labels and formal errors were 
-correct, the distribution of these pair-wise comparisons would be
+quadrature sum of their formal uncertainties $(\logg_a - \logg_b)/\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}$.  If our derived labels were unbiased and our formal
+errors were correct, the distribution of these pair-wise comparisons would be
 well-represented by a Gaussian distribution with zero mean and variance
@@
 \begin{eqnarray}
-	\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{(\sigma_{\logg,a} + \sigma_{\logg,floor})^2 + (\sigma_{\logg,b} + \sigma_{\logg,floor})^2}} \quad .
+	\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2 + 2{\sigma_{\logg,floor}}^2}} \quad .
 \end{eqnarray}
@@
 fluxes weighted by the inverse variances, then treated the stacked spectra
-as a normal survey star.  We inferred labels for the stacked spectra 
+as normal survey stars.  We inferred labels for the stacked spectra 
 using all three models, and derived a joint estimate as per Section 
-\ref{sec:joining-the-models}.  The labels we inferred from the stacked 
+\ref{sec:joining-the-models}.  The labels we inferred from each stacked 
 spectrum then served as a basis of comparison for the labels we derived 
@@
 In Figure \ref{fig:repeat-visits} we show the RMS difference
-in labels between the stacked spectrum and a single visit, binned by the
+in labels between the stacked spectra and single visit, binned by the
 S/N ratio of the individual visit spectrum.  Here we only show stars
@@
 their analysis on high-resolution, high S/N spectra.  We found four notable
-studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,Reddy_2006}
-and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} work on exoplanet
-host star candidates.  These studies 
-perform a careful (manual; expert) analysis using extremely high-resolution, high S/N spectra, and make use of \project{Hipparcos}
-parallaxes where possible.  Most of the stars in these samples are main-sequence
-or sub-giant stars.  Therefore, these works constitute an excellent comparison
-to evaluate the accuracy of our results on the main-sequence and sub-giant branch.
-
-We compare the stellar parameters ($\teff$, $\logg$, [Fe/H]) of stars that
-overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
-We only include stars with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter
-cut removed only a few stars because the average S/N in the \rave\ spectra for
-these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$). When considering the relative 
-information content available in \rave\ (945~pixels in the near infrared with
-$\mathcal{R} \approx 7{,}500$) compared to these literature studies that use
-\project{Hipparcos} parallaxes where possible, and base their inferences on
-spectra with resolving power $\mathcal{R}$ between 40,000 to 110,000, and
-S/N ratios exceeding 150, we consider the agreement to be very satisfactory.
-Indeed, given the metallicity precision available in the \unrave\ catalog, 
-these results will likely be useful for future studies based on exoplanet
-host star properties (e.g., \project{TESS}). At present, however, there are 
-just $\approx$30 stars in \rave\ that overlap with the compilations of 
-exoplanet host star properties listed at \texttt{exoplanets.org} and 
-\texttt{exoplanets.eu}.
-
-\stub{Reference Figure \ref{fig:gold-standard-hrd}}
+studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,
+Reddy_2006} and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} 
+work on exoplanet host star candidates.  These studies perform a careful (manual;
+expert) analysis using extremely high-resolution, high S/N spectra, and make use 
+of \project{Hipparcos} parallaxes where possible.  Most of the stars in these 
+samples are main-sequence or sub-giant stars.  Therefore, these works constitute 
+an excellent comparison to evaluate the accuracy of our results on the 
+main-sequence and sub-giant branch.
+
+In Figure \ref{fig:gold-standard-hrd} we show Hertzsprung-Russell diagrams 
+for the \rave\ stars that overlap with these studies.  We only include stars 
+with $\chi_r^2 < 3$ and $S/N > 10$~pixel$^{-1}$, although the latter cut 
+removed only a few stars because the average S/N in the \rave\ spectra for
+these stars is relatively high ($\gtrsim{}50$~pixel$^{-1}$).  The literature 
+data points in Figure \ref{fig:gold-standard-hrd} are linked to our derived
+labels for the same stars, illustrating good qualitative agreement across 
+the turn-off and sub-giant branch in all studies.  If we treat all three 
+studies as a single point of comparison, the bias between our work and these 
+studies is $-89$~K in $\teff$, just $-0.06$~dex in $\logg$, and $-0.03$~dex 
+in [Fe/H] (see Figure~\ref{fig:gold-standard-comparison}).  The RMS deviation
+in labels is 237~K, 0.30~dex, and 0.15~dex, respectively.  When considering 
+the relative information content available in \rave\ (945~pixels in the near 
+infrared with $\mathcal{R} \approx 7{,}500$) compared to these literature 
+studies that use \project{Hipparcos} parallaxes where possible, and base 
+their inferences on spectra with resolving power $\mathcal{R}$ between 
+40,000 to 110,000, and S/N ratios exceeding 150, we consider the agreement 
+to be very satisfactory.  Indeed, given the metallicity precision available
+in the \unrave\ catalog, these results will likely be useful for future 
+studies based on exoplanet host star properties (e.g., 
+\project{TESS}\footnote{At present, however, there are just $\approx$30 
+stars in \rave\ that overlap with the compilations of exoplanet host star 
+properties listed at \texttt{exoplanets.org} and \texttt{exoplanets.eu}.}).
 
@@
 sub-giant and red giant branch stars.  About half (67) of the sample 
-were acquired with the \acronym{UVES} instrument, the other with the \acronym{GIRAFFE} spectrograph, and
-the S/N of the \ges\ spectra peaks at $\approx140$~pixel$^{-1}$.  Despite 
-most of these stars having
-relatively low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good 
-agreement in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
-The RMS in effective temperature, surface gravity and metallicity is
-233~K, 0.37~dex, and 0.17~dex, respectively.
-
-Based on this comparison, we find no evidence for a systematic offset 
-between stars on the main-sequence and those on the giant branch. Specifically,
-we do not find any dwarf stars that are misclassified as sub-giants, as we
-found for some metal-poor stars in Section~\ref{sec:validation-gold-standards}.
-In addition we find no evidence for a systematic offset in metallicity (or
-individual abundances) between giant stars and main-sequence stars.  This is
-a crucial observation, as the metallicities for stars in our main-sequence
-training set have a principally different source to those on the giant branch.
-We cannot make these same inferences based on other surveys, like \apogee,
-because (1) \apogee\ stars formed part of the training set, and (2) they do
-not include main-sequence stars.  Even if we found good agreement between
-\epic\ and \apogee\ metallicities, this would not be informative, because
-\apogee\ is the source of metallicity for many stars on the giant branch in
-the \epic\ sample. Therefore, although this is a qualitative comparison only,
-it is reassuring that there is no obvious systematic difference between the
-metallicities of main-sequence and giant branch stars.
+were acquired with the \acronym{UVES} instrument --- the other with the 
+\acronym{GIRAFFE} spectrograph --- and the S/N of the \ges\ spectra peaks 
+at $\approx140$~pixel$^{-1}$.  Despite most of these stars having relatively 
+low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good agreement
+in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
+The RMS in effective temperature, surface gravity and metallicity is 233~K, 
+0.37~dex, and 0.17~dex, respectively.
+
+Based on this comparison, we find no evidence for a systematic offset in
+metallicities between stars on the main-sequence and those on the giant branch. 
+This is a crucial observation, as the metallicities for stars in our 
+main-sequence training set have a principally different source to those on 
+the giant branch.  We cannot make these same inferences based on other 
+surveys, like \apogee, because (1) \apogee\ stars formed part of the 
+training set, and (2) they do not include main-sequence stars.  Even if we 
+found good agreement between \epic\ and \apogee\ metallicities, this would 
+not be informative, because \apogee\ is the source of metallicity for many 
+stars on the giant branch in the \epic\ sample. Therefore, although this is
+a qualitative comparison only, it is reassuring that there is no obvious 
+systematic difference between the metallicities of main-sequence and giant
+branch stars.
 
@@
 reported for these metallicities.
-% TODO: Check if that is still the case after we examine errors from repeat
-% 		obs.
 
@@
 stages).  This provides us with an independent validation for our detailed
-abundances on the giant branch.  These comparisons are shown for six elemental
-abundances in Figure~\ref{fig:ges-abundances}, where markers are colored by the
+abundances on the giant branch.  These comparisons are shown in 
+Figure~\ref{fig:ges-abundances}, where markers are colored by the
 S/N of the \rave\ spectrum.  The number of stars available in each abundance
@@
 \subsubsection{Comparison with the \rave\ DR4 calibration sample}
-
-\stub{Reference Figure\ref{fig:kordopatis-calibration}}
+\label{sec:dr4-calibration-sample}
+
+The fourth \rave\ data release made use of a number of high-resolution studies
+to verify the accuracy of the reported stellar atmospheric parameters.  These
+samples include main-sequence stars, giant stars, with a particular focus to
+include metal-poor stars to identify (and correct) any deviations at low
+and high metallicities.  We refer the reader to \citet{Kordopatis_2013} for
+the full compilation of literature sources.  Although the stellar atmospheric 
+parameters in this compilation come from multiple (heterogeneous) sources, we 
+find generally good agreement with these works (Figure 
+\ref{fig:kordopatis-calibration}).  However, some reservation is warranted 
+when evaluating this comparison, as some of the metal-poor stars in this 
+calibration sample formed part of our training set.
 
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
-\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
-\label{fig:gold-standard-comparison}}
-\end{figure*}
-
-\begin{figure*}[p]
 \includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
@@
 This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
+\caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
+Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+\label{fig:gold-standard-comparison}}
 \end{figure*}

### 2016-09-08 article/unrave.tex
some text

@@
 
-After verifying that our atmospheric parameters and abundances are comparable with other
-high-resolution studies, here we verify that the abundances we find
-are consistent with expectations from astrophysics.  One example is globular clusters, 
-which show anti-correlations in light element abundances 
-\citep[e.g.,][and references therein]{Norris_Da_Costa_1995,Carretta_2009}.  In the \rave\
-survey, \cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
+After verifying that our atmospheric parameters and abundances are comparable 
+with other high-resolution studies, here we verify that the abundances we find
+are consistent with expectations from astrophysics.  In the \rave\ survey, 
+\cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
 consistent with being members of globular clusters: 49 stars belonging to NGC~5139
 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 
-members of NGC~362. In addition to these systems, 12 stars thought to belong to NGC~1851
-were found in the \rave\ data by \cite{Kunder_2014}, and a further 10 stars in NGC~6752.  
-We refer the reader to those studies for details regarding the membership selection.
-
-In Figure~\ref{fig:globular-cluster-HRD} we show the effective temperature
-$\teff$ and surface gravity $\logg$ for pre-defined members of globular clusters in \rave.
-The top panels indicate measurements made in this work, and for comparison purposes we have included
-the results from the fourth \rave\ data release.  We show representative \project{PARSEC} isochrones 
-\citep{Bressan_2012} in all panels, where the isochrone ages and metallicities are adopted from
-\citet{Kunder_2016,Marin-Franch_2009}, and the \citet[][accessed 6 September 2016]{Harris_1996} 
-online catalog of globular cluster properties.  The globular cluster with the most number of
-\rave\ members is NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
-consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}.  We
-find the mean metallicity of prescribed $\omega$-Centauri members to be $[{\rm Fe/H}] = -0.85$.
-However, it is clear that the membership criteria could be improved with our
-revised metallicities and detailed chemical abundances. 
-Indeed, this demonstrates that our abundance labels can be used to identify
-globular cluster members that are now tidally disrupted \citep{Anguiano_2016,Kuzma_2016,Navin_2016},
-even for stars with low S/N ratios.
+members of NGC~362. In addition, \citet{Kunder_2014} compiled 12 stars thought to 
+belong to NGC~1851, and a further 10 stars in NGC~6752.  We refer the reader to 
+those studies for details regarding the membership selection.
+
+In Figure~\ref{fig:globular-cluster-HRD} we show our effective temperature
+$\teff$ and surface gravity $\logg$ for these prescribed globular cluster members.
+The right-hand panels indicate measurements made in this work, and for comparison 
+purposes we have included the results from the fourth \rave\ data release in the 
+left-hand panels.  We show representative \project{PARSEC} isochrones 
+\citep{Bressan_2012} in all panels, where the isochrone ages and metallicities 
+are adopted from \citet{Kunder_2016,Marin-Franch_2009}, and the 
+\citet[][accessed 6 September 2016]{Harris_1996} online catalog of globular 
+cluster properties.  The globular cluster with the most number of members is 
+NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
+consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}. 
+Based on the pre-defined membership criteria, we find the mean metallicity
+of $\omega$-Centauri to be $[{\rm Fe/H}] = -0.85$.  However, it is clear that the 
+membership criteria could be improved with our revised metallicities and 
+detailed chemical abundances.  Indeed, our individual abundance labels could 
+be used to identify globular cluster members that are now tidally disrupted 
+\citep{Anguiano_2016,Kuzma_2016,Navin_2016}, even for stars with low S/N ratios.
 
@@
 78 stars in the Pleiades, 26 in Hyades, 30 stars in the solar-metallicity cluster
-M67 and 13 stars in IC4651. Details on the open cluster membership and selection can 
+M67, and 13 stars in IC4651. Details on the open cluster membership and selection can 
 be found in \cite{Kunder_2016}.  We show the effective temperature $\teff$ and surface
@@
 
-We find good agreement between our atmospheric parameters (top panels) and the 
+We find good agreement between our atmospheric parameters (right-hand panels) and the 
 isochrones shown.  The position of the red clump in IC4651 and M67 are perfectly matched
@@
 our metallicity scale.  We find the mean metallicity of M67 stars to be 
-$[{\rm Fe/H}] = -0.02 \pm 0.17$~dex (30 stars), in excellent agreement with the
-expected $[{\rm Fe/H}] = 0.00$ value (not accounting for atomic diffusion).
+$[{\rm Fe/H}] = -0.02 \pm 0.03$~dex, in excellent agreement with the expected 
+$[{\rm Fe/H}] = 0.00$ value (not accounting for atomic diffusion).
 We further find the Hyades mean metallicity to be $[{\rm Fe/H}] = 0.07 \pm 0.09$~dex,
 consistent with \citet{Paulson_2003}: $[{\rm Fe/H}] \approx 0.13$.
-For IC4651 from 13 stars we find a mean $[{\rm Fe/H}] = 0.15 \pm 0.12$~dex, matching the
+For IC4651 from 13 stars we find a mean $[{\rm Fe/H}] = 0.15 \pm 0.03$~dex, matching the
 high-resolution, high S/N study of \citet{Pasquini_2004}, where they find 
 $[{\rm Fe/H}] = 0.10 \pm 0.03$~dex. Finally, from 78 stars we find the mean
-metallicity of the Pleiades to be $[{\rm Fe/H}] = -0.02 \pm 0.12$~dex, in agreement
-with the $[{\rm Fe/H}] = -0.034 \pm 0.024$~dex measurement reported by 
+metallicity of the Pleiades to be $[{\rm Fe/H}] = -0.02 \pm 0.01$~dex, in very 
+good agreement with the $[{\rm Fe/H}] = -0.034 \pm 0.024$~dex measurement reported by 
 \citet{Friel_Boesgaard_1990}.
 
-The most pronounced discrepancy is seen in the main-sequence of the Pleiades.
-Like \citet{Kordopatis_2013}, we find lower $\logg$ values than what the isochrone
-shown predicts.  \stub{Hawkins: use correct isochrone from Bouy to see if this issue is resolved.}
-%  Firstly,
-%we note that theoretical isochrones often struggle to match the observed
-%color-magnitude diagrams for some young clusters \citep[e.g., the Pleiades, see][]{Bouy_2015}.
+Despite the discrepancies between the isochrone and our derived labels for stars
+in the Pleiades, we have made no attempt to refine the membership selection in
+any of the aforementioned clusters.  We note however, that the same discrepancy
+with the isochrone appears present in the fourth \rave\ data release \citet{Kordopatis_2013}.
 
@@
 
-\bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
+%\bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
 

### 2016-09-08 article/unrave.tex
Finish conc

@@
 
-We have derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all \rave\ 
-spectra, as well as detailed abundances for red giant branch stars, based on an
-independent re-analysis of \rave\ spectra.  Our catalog provides an excellent
-complement to the current \rave\ data release \citep{Kunder_2016}.  
-When combined with the \tgas\ sample, these results may amount to a powerful
-compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
-Inferences based on these results should recognize those caveats, and acknowledge that 
-these results are subject to our explicit assumptions, some of which are provably
+We have performed an independent re-analysis of \Nspectra\ \rave\ spectra,
+having derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all stars,
+as well as detailed chemical abundances for red giant branch stars.  When 
+combined with the \tgas\ sample, these results amount to a powerful compendium 
+for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
+Inferences based on these results should recognize those caveats, and acknowledge 
+that these results are subject to our explicit assumptions, some of which are provably
 incorrect.
@@
 models.  Because the relative weights have no formal interpretation, it is reasonable
-to consider this method is as \emph{ad hoc} as any other approach.  Therefore, the
-relative weighting has no warranty to be (formally) correct, and may introduce 
-inconsistencies or systematic errors rather than
-minimizing them.  Evidence of this is already present in our comparisons to
-`gold standard' works \citep{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005},
-where metal-poor dwarf stars appear as metal-poor sub-giants in our combined results,
-but have reasonable labels (by any available statistic) as dwarfs in the main-sequence
-model.  Clearly, the relative distance from the joint model to the giant model is 
-slightly favoured (over the main-sequence model) for metal-poor dwarf stars. In this
-sense, our approach towards combining labels from multiple models is sub-optimal.
+to consider this method is as \emph{ad hoc} as any other approach.  The relative 
+weighting has no warranty to be (formally) correct, and may introduce inconsistencies
+or systematic errors rather than minimizing them.
 
@@
 set.  The metal-poor stars included in this sample are from high-resolution studies
-\citep{Fulbright_2010,Ruchti_2011}, but it is unknown if the stellar parameters are
-of high fidelity (e.g., we have limited quality metrics available), and there is no
-guarantee that the stellar parameters \emph{or} abundances are on the same scale as
-\apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
-
-If the metallicities of metal-poor giant stars are on the same scale as the \apogee\
+\citep{Fulbright_2010,Ruchti_2011}, but it is not known if the stellar parameters are
+of high fidelity because we have limited quality metrics available.  Moreover, there
+is no guarantee that the stellar parameters \emph{or} abundances are on the same scale
+as \apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
+
+If the metallicities of metal-poor giant stars were on the same scale as the \apogee\
 abundances, there is a larger issue in verifying that the main-sequence metallicities
 and giant branch metallicities are on the same scale.  The training set for the
-main-sequence stars includes metallicities from a variety of sources --- including
+main-sequence stars includes metallicities from a variety of sources, including
 \lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
-no straightforward manner to ensure that these two models produce metallicities on
-the same scale.  We see no systematic offset in metallicities of dwarf and giant
-stars that overlap between \rave\ and the \ges\ survey, suggesting that if there is
-a systematic offset, it must be small. Nevertheless, these are only verification checks based on $<1$\% of the
-data, and there is currently insufficient data for us to prove both models are on
-the same abundance scale.
+no straightforward manner to ensure that the main-sequence model and the red giant
+branch model produce metallicities on the same scale.  We see no systematic offset
+in metallicities of dwarf and giant stars that overlap between \rave\ and the \ges\
+survey, suggesting that if there is a systematic offset, it must be small. 
+Nevertheless, these are only verification checks based on $<1$\% of the data, and 
+there is currently insufficient data for us to prove both models are on the same 
+abundance scale.
 
@@
 are not on the same scale as \apogee, because we were forced to adopt abundances for 
-specific elements when they were unavailable.  In that sense, our assumption that the
-training set contains accurate labels is clearly untrue (for \emph{at least} some stars).
-Although we sought to adopt the mean
-level of Galactic chemical enrichment at the overall metallicity, this is not a
+specific elements when they were unavailable.  Although we sought to adopt mean
+level of Galactic chemical enrichment at a given overall metallicity, this is not a
 representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
-metallicity, there is no requirement for zero abundance spread.
-More fundamentally, we are incorrectly assuming that the element \emph{must} be 
-detectable in the photosphere of the star.  There may be no transition that is 
-detectable in that star, even with zero noise, because it is too weak to have any 
-effect on the spectrum.  In the most optimistic case, this can be considered to be
-forcing the model to make use of correlated information between abundances.  In a
-more representative (pessimistic) case, we are simply invoking what all abundances
-should be at low metallicity.
+metallicity, there is no requirement for zero abundance spread.  More fundamentally, 
+we are incorrectly asserting that the element \emph{must} be detectable in the 
+photosphere of the star.  There may be no transition that is detectable in that 
+star, even with zero noise, because it is too weak to have any effect on the 
+spectrum.  In the most optimistic case, this could be considered to be forcing the
+model to make use of correlated information between abundances.  In a more 
+representative (pessimistic) case, we are simply invoking what all abundances should 
+be at low metallicity.
 
@@
 in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
-\citet{Kordopatis_2013} note that the effective resolution of \rave\
-spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
-effective resolution is a function of temperature variations, fibre-to-fibre
-variations, and thus position on the CCD \citep{Steinmetz_2006}.
- For this reason we ought to expect our derived stellar
-parameters or abundances to be correlated with the fibre number, or over
-time.  If significant, the trend could produce systematically offset
-stellar abundances solely due to the fibre number.  \citet{Kordopatis_2013}
-conclude that resolution-based effects on the \rave\ stellar parameters
-should be a second-order effect. We have not seen evidence of these 
-resolution-based correlations in our results, however, we have only
-performed cursory (non-exhaustive) experiments to investigate this issue.
-
-\stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}
-
-The fibre fringing is another example of how test spectra that are
-unusual from the training set can be projected as peculiar artefacts in the
+\citet{Kordopatis_2013} note that the effective resolution of \rave\ spectra
+varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the effective 
+resolution is a function of temperature variations, fibre-to-fibre variations,
+and thus position on the CCD \citep{Steinmetz_2006}. For this reason we ought
+to expect our derived stellar parameters or abundances to be correlated either
+with the fibre number, with the observation date, or both.  If significant, 
+the trend could produce systematically offset stellar abundances solely due to
+the fibre used.  \citet{Kordopatis_2013} conclude that resolution-based effects 
+on the \rave\ stellar parameters should be a second-order effect. We have not 
+seen evidence of these resolution-based correlations in our results, however, 
+we have only performed cursory (non-exhaustive) experiments to investigate 
+this issue.
+
+We have shown some potential outcomes when the test set spectra differ 
+significantly from the spectra in the training set.  Test set spectra that is
+`unusual' from the training set can be projected as peculiar artefacts in 
 label space.  In other words, unusual spectra can appear as `clumps' in
-regions of parameter space that we could consider as being normal 
-(e.g., solar-type stars).  We addressed this issue for the main-sequence
-and giant models by using a joint model to calculate relative weights.
-However, spectra that are unusual from the joint model training set can
-still project as systematic artefacts.
-
-Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
-first is a small over-density at the base of the giant branch, which is
-likely a consequence of joining the 9-label and 3-label models.  The
+regions of parameter space that we could consider as being normal (e.g.,
+an over-density of solar-type stars).  We addressed this issue for the main-sequence
+and giant models by using a third model (Section \ref{sec:a-simple-model}) to
+calculate relative weights.  However, spectra that are unusual from the 
+training set used in the simplest model still project as systematic artefacts.
+
+Indeed, there are two known artefacts in our data that are relevant to this 
+discussion. The first is a small over-density at the base of the giant branch, 
+which is likely a consequence of joining the 9-label and 3-label models.  The
 second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
 present in our training set, yet there are many in the \rave\ survey.
-However, the \rave\ pre-processing pipeline \citep[\texttt{SPARV};][]{Steinmetz_2006,Zwitter_2008} performs
-template matching against grids of cool \emph{and} hot stars, and therefore
-we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
+However, the \rave\ pre-processing pipeline 
+\citep[\texttt{SPARV};][]{Steinmetz_2006,Zwitter_2008} performs template 
+matching against grids of cool \emph{and} hot stars, and therefore we can use 
+that information to identify hot stars.  In Figure \ref{fig:hot-stars} 
 we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
 where each hexagonal bin is colored by the \emph{maximum} temperature
-reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
-at just above the turn-off point in our results.
-
-\stub{this following paragraph needs rewording}
-Based on a thorough interrogation of our results, we have adopted two boolean
-columns to signify whether our labels meet specific quality constraints:
-one marked \texttt{QC} to indicate reliable stellar parameters ($\teff$, $\logg$, [Fe/H]), and \texttt{QC\_ABUNDANCE} to indicate reliable chemical abundances.
-For stellar atmospheric parameters to meet our quality constraints, we
-require: $\chi_r^2 < 3$; $S/N >10$~pixel$^{-1}$; and \texttt{TeffS} $< 8000$~K
-in order to remove hot stars that are outside the training set.  We
-specified the quality constraint in chemical abundances to further
-require that the iron abundance $[{\rm Fe/H}] > -1.5$, since this approximately
-corresponds to the point where individual abundances will become undetectable
-in \rave\ spectra, and the individual abundance labels will be driven by our
-constraint to place them on the mean galactic chemical evolution trend.
-These quality constraints should not be taken verbatim: judicious use of
-the morphological classifications \citep{Matijevic_2012} or metadata from
-the pre-processing pipelines is encouraged.
+reported by \texttt{SPARV} for any star in that bin.  We show the maximum
+temperature reported by \texttt{SPARV} to demonstrate that hot stars project
+into a single clump in our label space --- near the turn-off --- in a region
+where we may otherwise be deceived into thinking the observed over-density
+is consistent with expectations from astrophysics.
+
+We have included a boolean flag column in our catalog for quality control,
+to signify whether a result is thought to be reasonable.  Specifically, 
+stars where \texttt{SPARV} indicates a $\teff > 8000$~K do not meet our quality 
+control flag.  For all other stars, we further require $\chi_r^2 < 3$ and
+$S/N >10$~pixel$^{-1}$.  These quality constraints represent the minimum
+acceptable constraints and should not be taken verbatim: judicious use of
+the morphological classifications \citep{Matijevic_2012} or additional 
+metadata from the \rave\ pre-processing pipelines is strongly encouraged.
 
 We have presented the largest collection of stellar abundances for stars in the 
-first \gaia\ data release.  In total more than 1.69 million abundances are
-available, with effective temperature $\teff$, surface gravity $\logg$ and 
-metallicity available for \Nstarsqc\ stars.  Despite the caveats and limitations 
-discussed here, our validation experiments and comparisons with high-resolution
-spectroscopic studies suggests that our results have sufficient accuracy and
-precision to be useful for exoplanet host stars studies 
-\citep[c.f.,][]{Valenti_Fischer_2005}, investigations of the Milky Way disk 
-\citep[c.f.,][]{Reddy_2003,Reddy_2006,Bensby_2014}, and a multitude of 
-chemo-dynamic studies.  When combined with positions and 3D velocities from \gaia,
-we are optimistic that the \unrave\ catalog will help advance understanding 
-of the Milky Way's formation and evolution.
+first \gaia\ data release.  In total we derive reliable stellar atmospheric 
+parameters for \Nstarsqc, and report more than 1.69 million abundances.  Despite 
+the caveats and limitations discussed here, our validation experiments and 
+comparisons with high-resolution spectroscopic studies suggests that our results 
+have sufficient accuracy and precision to be useful for chemo-dynamic studies 
+that become imminently feasible only in the era of \gaia\ data.  Indeed, we
+are optimistic that the \unrave\ catalog will advance understanding of the
+Milky Way's formation and evolution.
 

### 2016-09-08 article/unrave.tex
fix ack

@@
 Inferred labels, associated errors, and relevant metadata are available electronically
-at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  These results are also available
+at \texttt{\url{ZENODO-URL}} \citep{ZENODO-CITATION}.  These results are also available
 through the \rave\ DR5 database \citep{Kunder_2016}. Please note that it is a condition
@@
 	Sven Buder (MPIA), 
-	Sergey Koposov (Cambridge),
 and 

### 2016-09-08 article/unrave.tex
Update captions

@@
 \caption{Effective temperature $\teff$ and surface gravity $\logg$ for all stars in the training sets.
-Stars are colored by their metallicity [Fe/H], and the three panels show stars in the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
+Stars are colored by their metallicity [Fe/H], and the three panels show stars in the simplest model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/test-set-density.pdf}
-\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the joint model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
+\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the simplest model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
 These panels demonstrate how a single quadratic model is insufficient for all \rave\ stars (left panel), and illustrate some of the systematic artefacts that can result from testing on stars outside of the training set (center and right panel).
@@
 \includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
-\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the joint model (top panel), and the giant model and the joint model (bottom panel).
+\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the simplest model (top panel), and the giant model and the simplest model (bottom panel).
 The density scaling is logarithmic, and the differences in $\teff$ and $\logg$ are scaled to make them approximately isotropic (see text for details).
-The peak at $(0, 0)$ represents good agreement between the joint model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
+The peak at $(0, 0)$ represents good agreement between the simplest model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
 \end{figure}
@@
 \caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
-Only results with $\chi_{r}^2 < 3$ and S/N ratios greater than the quoted limit in each panel are shown.
+Only results meeting our quality constraints are shown (see Section~\ref{sec:discussion}).
 The top three panels show logarithmic density, and bins in the bottom three panels are colored by the median metallicity in each bin.\label{fig:test-set-hrd}}
@@
 Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
-This figure illustrates the good overall agreement in the shape of the turnoff and sub-giant branch, yet also demonstrates how some literature metal-poor ($[{\rm Fe/H}] \lesssim -1$) dwarfs near the turn-off appear as metal-poor sub-giants in \project{unRAVE}.\label{fig:gold-standard-hrd}}
+This figure illustrates the good qualitative agreement in the shape of the turnoff and sub-giant branch.\label{fig:gold-standard-hrd}}
 \end{figure*}
@@
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+Stars are colored by the S/N of the \rave\ spectra.
 \label{fig:gold-standard-comparison}}
@@
 The number of stars in each panel are shown, as well as the bias and RMS deviation in each label.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.
+Stars are colored by the S/N of the \rave\ spectra.
 Most of the \ges/\rave\ overlap stars have relatively low S/N ratios in \rave, near $\approx 30$~pixel$^{-1}$.\label{fig:ges-stellar-parameters}}
@@
 The number of stars shown in each panel is indicated, and the bias and RMS deviations are shown.
-Stars are colored by the S/N of the \rave\ spectra, and only stars with $\chi_r^2 < 3$ and S/N $> 10$ are shown.\label{fig:ges-abundances}}
+Stars are colored by the S/N of the \rave\ spectra.\label{fig:ges-abundances}}
 \end{figure*}
@@
 Note that the y-axis limits vary for each panel, however for clarity we show the scaled-solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
-Only stars with S/N $> 10$ and $\chi_r^2 < 3$ are shown.\label{fig:gce}}
+Only stars meeting our quality constraints are shown (see Section \ref{sec:discussion}).\label{fig:gce}}
 \end{figure}

### 2016-09-09 article/figures/rave_io.py
Figres updated

@@ -26,7 +26,7 @@ def get_kordopatis_comparisons():
 
 def get_cannon_dr1(filename=None):
 
-    filename = filename or os.path.join(DATA_PATH, "../../unrave-v0.96.fits.gz")
+    filename = filename or os.path.join(DATA_PATH, "../../unrave-v0.97.fits.gz")
 
     rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, filename))
 

### 2016-09-09 combine_model_results.py

@@ -19,7 +19,7 @@ RESULTS_PATH = ""
 
 ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v43.fits.gz"))
 giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
-joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v45.fits.gz"))
+joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v46.fits.gz"))
 
 
 for t in (ms_results, giant_results, joint_results):
@@ -426,7 +426,24 @@ del combined_table["Teff_SPARV"]
 #    combined_table[column][hot] = np.nan
 #    combined_table["E_{}".format(column)] = np.nan
 
-combined_table.write("unrave-v0.96.fits.gz", overwrite=True)
+
+
+error_floor = {
+    "TEFF": 70, #100,
+    "LOGG": 0.12, #0.15,
+    "FE_H": 0.06, #0.08,
+    "MG_H": 0.07, #0.08,
+    "O_H": 0.07, #0.08,
+    "AL_H": 0.08, #0.08,
+    "CA_H": 0.07, #0.08,
+    "SI_H": 0.07, #0.08,
+    "NI_H": 0.06, #0.08,
+}
+
+for label_name, floor in error_floor.items():
+    combined_table["E_{}".format(label_name)] = np.sqrt(combined_table["E_{}".format(label_name)]**2 + floor**2)
+
+combined_table.write("unrave-v0.97.fits.gz", overwrite=True)
 
 
 

### 2016-09-09 article/unrave.tex
update author list and affiliations, etc

@@
-%% This file is part of the UNRAVE Project 
+%% This file is part of the unRAVE Project 
 %% Copyright 2016 the authors.  All rights reserved.
 
-%% TODO:
-%% Co-authors: Look for \stub{}s. If you can expand it, do so!
-
-\documentclass[preprint,trackchanges]{aastex}
+\documentclass[preprint]{aastex}
 
@@
 \usepackage{bm}
-
-%\usepackage{xcolor}
-%\definecolor{unoffensive-warning}{HTML}{B4DCED}
-
-% This section can be removed at submission.
-% ------------------------------------------
-\usepackage{color}
-\usepackage{datenumber}
-
-\newcounter{dateone}
-\newcounter{datetwo}
-
-\newcommand{\difftoday}[3]{%
-  \setmydatenumber{dateone}{\the\year}{\the\month}{\the\day}%
-  \setmydatenumber{datetwo}{#1}{#2}{#3}%
-  \addtocounter{datetwo}{-\thedateone}%
-  \the\numexpr(\thedatetwo)\relax\space days}
-% ------------------------------------------
 
@@
 
-\newcommand{\Nspectra}{520,782}
-\newcommand{\Nstars}{457,589}
+\newcommand{\Nspectra}{520,781}
+\newcommand{\Nstars}{457,588}
+
+% TODO: Final checks based on released version.
 \newcommand{\Nstarsqc}{447,207} % v0.95 finite values
@@
 
-% For AASTeX v6
-%\AuthorCallLimit=10
-%\fullcollaborationName{The \rave\ Collaboration}
-
 \begin{document}
-% Remove at submission:
-\slugcomment{{\color{red} \textbf{To appear on arXiv on 13 Sep 2016 (submission in \difftoday{2016}{9}{9})}}}
-
-\title{The \project{unRAVE} catalog}
-
-\author{Andrew R. Casey}
-\affil{Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
-
-% Gilmore, Matijevic, McMillan, Siviero, Wojno
-
-% Specific contributors to this work:
-% There is some order already in mind, but final order will be determined by  contribution.
-\author{Some combination of: Harry Enke, Kenneth C.
-Freeman, Gerry Gilmore, Keith Hawkins, David W.
-Hogg, Sergey Koposov, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix}
-
-% RAVE DR5 core team:
-\author{Luca Casagrande, Andrea Kunder, Paul McMillan, Alessandro Siviero, Jennifer Wojno, Toma\v{z} Zwitter}
-% Zwitter:  University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia
-
-% RAVE builders
-\author{Joss Bland-Hawthorn, Brad K. Gibson, Arnaud Siebert, Olivier Bienayme, Julio F. Navarro, Ulisse Munari, Warren Reid, Rosemary Wyse, Amina Helmi, George Seabroke}
-% Bland-Hawthorn: Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia.
-% Gibson: E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom.
-% Siebert: Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
-% Bienayme: Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France
-% Navarro: Senior CIfAR Fellow; Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2
-% Munari: INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy
-% Reid (1/2): Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia
-% Reid (1/2): Western Sydney University, Penrith South DC, NSW 1797
-% Wyse: Johns Hopkins University, Baltimore, MD, USA
-% Helmi: Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands
-% Seabroke: Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK
-\author{and the \rave\ collaboration}
+\title{The \project{unRAVE} catalog of stellar atmospheric parameters and\\
+chemical abundances for chemo-dynamic studies in the \gaia\ era}
+
+\author{Andrew R.
+Casey\altaffilmark{1}, Keith Hawkins\altaffilmark{1}, David W.
+Hogg\altaffilmark{2,3,4,5}, Melissa Ness\altaffilmark{5}, Hans Walter-Rix\altaffilmark{5}, Georges Kordopatis\altaffilmark{6}, Andrea Kunder\altaffilmark{6}, Matthias Steinmetz\altaffilmark{6}, Harry Enke\altaffilmark{6}, Jason Sanders\altaffilmark{1}, Sergey Koposov\altaffilmark{1}, Gerry Gilmore\altaffilmark{1}, Toma\v{z} Zwitter\altaffilmark{6}, Kenneth C.
+Freeman\altaffilmark{7}, Luca Casagrande\altaffilmark{7},
+Gal Matijevic\altaffilmark{6}, George Seabroke\altaffilmark{8}, Olivier Bienayme\altaffilmark{9}, 
+Joss Bland-Hawthorn\altaffilmark{10}, Brad K.
+Gibson\altaffilmark{11}, Eva~K.~Grebel\altaffilmark{12}, Amina Helmi\altaffilmark{13}, Ulisse Munari\altaffilmark{14}, Julio F.
+Navarro\altaffilmark{15,16}, Warren Reid\altaffilmark{17,18}, Arnaud Siebert\altaffilmark{19}, Rosemary Wyse\altaffilmark{20}\\
+\vspace{2em}
+\emph{(Affiliations can be found at the end of the article)}
+}
 
@@
 (\tgas) sample in the first \gaia\ data release requires a spectroscopic 
-complement that includes radial velocities, stellar parameters, and 
+complement that includes radial velocities, stellar atmospheric parameters, and 
 elemental abundances.  Among existing spectroscopic samples, the RAdial 
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-up to 292,036 stars.  We present a data-driven re-analysis of \rave\ 
+$\approx$250,000 stars.  We present a data-driven re-analysis of \rave\ 
 spectra using \thecannon, and we provide more precise and accurate stellar
@@
 presented here, in conjunction with \tgas, represents the most powerful 
-data set for chemo-orbital analysis at the dawn of broadly available 
+data set for chemo-dynamic analysis at the dawn of broadly available 
 \gaia\ data.
@@
 
-\keywords{stars: fundamental parameters --- stars: abundances}
+\keywords{stars: fundamental parameters --- stars: abundances
+% Clear rest of page
+\clearpage}
 
@@
 surveys\footnote{Specifically we cross-matched the Tycho-2 catalog against the 
-\apogee\ \citep{Zasowski_2013}, \ges\ \citep{Gilmore_2012,Randich_2013}, 
-\galah\ \citep{DeSilva_2015}, \lamost\ \citep{Cui_2012}, and \rave\ 
-\citep{Steinmetz_2006} surveys.}, we found that the RAdial Velocity Experiment 
-(\rave) survey is expected to have the largest overlap with the first \gaia\ 
-data release: up to 292,036 stars.  We used the \gaia\ universe model snapshot 
+\apogee\ DR13 \citep{Zasowski_2013}, \ges\ internal DR4 
+\citep{Gilmore_2012,Randich_2013}, \galah\ internal DR1 \citep{DeSilva_2015}, 
+\lamost\ DR1 \citep{Cui_2012}, and \rave\ DR4 \citep{Steinmetz_2006} catalogs.}, 
+we found that the RAdial Velocity Experiment \citep[\rave;][]{Steinmetz_2006} 
+survey is expected to have the largest overlap with the first \gaia\ 
+% TODO: Sanders: 292,036; Kunder: 264,276
+data release: up to 264,276 stars.  We used the \gaia\ universe model snapshot 
 \citep{Robin_2012} to estimate the precision in parallax and proper motions that
@@
 we further found that the \rave\ survey will benefit most from \gaia\ DR1: the
-distances of 63\%of stars in the \rave--\gaia\ DR1 overlap sample 
-($\approx$182,862 stars) are expected to improve with the first \gaia\ data release, 
-and 47\% of stars ($\approx$137,211 stars) are likely to have better proper motions. 
-Although the \gaia\ universe model assumes end-of-mission uncertainties --- and does 
-not account for systematics in the first data release --- this calculation still 
-provides intuition for the relative improvement that the first \gaia\ data release 
-can make to ground-based surveys.  The expected improvements for \rave\ motivated 
-us to examine what chemical abundances were available from those data, and to 
-evaluate whether we could enable new chemo-dynamic studies by contributing to the
-existing set of chemical abundances.
+distances of 63\% of stars in the \rave--\gaia\ DR1 overlap sample are expected 
+to improve with the first \gaia\ data release, and 47\% of stars are likely to
+have better proper motions.  Although the \gaia\ universe model assumes 
+end-of-mission uncertainties --- and does not account for systematics in the 
+first data release --- this calculation still provides intuition for the relative
+improvement that the first \gaia\ data release can make to ground-based surveys.  
+The expected improvements for \rave\ motivated us to examine what chemical 
+abundances were available from those data, and to evaluate whether we could 
+enable new chemo-dynamic studies by contributing to the existing set of chemical
+abundances.
 
@@
 high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}.  The spectral region 
-also includes the \ion{Ca}{2} near infrared triplet lines --- strong transitions that 
+includes the \ion{Ca}{2} near infrared triplet lines --- strong transitions that 
 are dominated by pressure broadening --- which are visible even in metal-poor stars
 or spectra with very low signal-to-noise (S/N) ratios.  Atomic transitions of 
-light-, $\alpha$-, and Fe-peak elements are present as well, allowing for detailed 
+light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
 chemical abundance studies.
@@
 
-%\begin{eqnarray}
-%\eta_{\logg} & = & \frac{\logg_a - \logg_b}{\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}} \quad .
-%\end{eqnarray}
-
 We increased the minimum label error until the distribution of $\eta$ 
@@
 \citet{Kordopatis_2013} has been noted in other studies (e.g., \apogee), and this issue
-has been resolved in the fifth \rave\ data release by correcting $\logg$
+has been minimized in the fifth \rave\ data release by correcting $\logg$
 values with a calibration sample consisting of asteroseismic targets and the
@@
 
-We have presented the largest collection of stellar abundances for stars in the 
-first \gaia\ data release.  In total we derive reliable stellar atmospheric 
-parameters for \Nstarsqc, and report more than 1.69 million abundances.  Despite 
-the caveats and limitations discussed here, our validation experiments and 
-comparisons with high-resolution spectroscopic studies suggests that our results 
-have sufficient accuracy and precision to be useful for chemo-dynamic studies 
-that become imminently feasible only in the era of \gaia\ data.  Indeed, we
-are optimistic that the \unrave\ catalog will advance understanding of the
-Milky Way's formation and evolution.
+We have presented a comprehensive collection of precise stellar abundances 
+for stars in the first \gaia\ data release.  In total we derive reliable 
+stellar atmospheric parameters for \Nstarsqc, and report more than 1.69 
+million abundances.  Despite the caveats and limitations discussed here, 
+our validation experiments and comparisons with high-resolution 
+spectroscopic studies suggests that our results have sufficient accuracy 
+and precision to be useful for chemo-dynamic studies that become imminently 
+feasible only in the era of \gaia\ data.  Indeed, we are optimistic that 
+the \unrave\ catalog will advance understanding of the Milky Way's 
+formation and evolution.
 
@@
 % TODO: Replace ZENODO-URL with a data url
-%\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
 \noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
@@
 tireless efforts of the entire \rave\ collaboration, past and present.
-%}}
 
@@
 \caption{Detailed chemical abundances ([X/Fe]) for giant stars in \unrave\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
+Bin densities are scaled logarithmically.
 Note that the y-axis limits vary for each panel, however for clarity we show the scaled-solar position by dashed lines, and have common tick mark spacing on the y-axis for all panels.
@@
 Hot stars are not included in our training set, and appear above the turnoff in our labels.
-For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars with $\teff > 8000$~K (as reported by \texttt{SPARV}), and other results that are known to be erroneous.\label{fig:hot-stars}}
+For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars with $\teff > 8000$~K (as reported by \texttt{SPARV}), and other results that are of poor quality.\label{fig:hot-stars}}
 \end{figure}
 
+\clearpage
+
+\altaffiltext{1}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\altaffiltext{3}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
+\altaffiltext{4}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
+\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{6}{University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia}
+\altaffiltext{7}{Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia}
+\altaffiltext{8}{Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK}
+\altaffiltext{9}{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France}
+\altaffiltext{10}{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia}
+\altaffiltext{11}{E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom}
+\altaffiltext{12}{Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany}
+\altaffiltext{13}{Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands}
+\altaffiltext{14}{INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy}
+\altaffiltext{15}{Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2}
+\altaffiltext{16}{Senior CIfAR Fellow}
+\altaffiltext{17}{Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia}
+\altaffiltext{18}{Western Sydney University, Penrith South DC, NSW 1797}
+\altaffiltext{19}{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France}
+\altaffiltext{20}{Johns Hopkins University, Baltimore, MD, USA}
+
 \end{document}

### 2016-09-09 article/unrave.tex
fix order

@@
 
-\author{Andrew R.
-Casey\altaffilmark{1}, Keith Hawkins\altaffilmark{1}, David W.
-Hogg\altaffilmark{2,3,4,5}, Melissa Ness\altaffilmark{5}, Hans Walter-Rix\altaffilmark{5}, Georges Kordopatis\altaffilmark{6}, Andrea Kunder\altaffilmark{6}, Matthias Steinmetz\altaffilmark{6}, Harry Enke\altaffilmark{6}, Jason Sanders\altaffilmark{1}, Sergey Koposov\altaffilmark{1}, Gerry Gilmore\altaffilmark{1}, Toma\v{z} Zwitter\altaffilmark{6}, Kenneth C.
-Freeman\altaffilmark{7}, Luca Casagrande\altaffilmark{7},
-Gal Matijevic\altaffilmark{6}, George Seabroke\altaffilmark{8}, Olivier Bienayme\altaffilmark{9}, 
-Joss Bland-Hawthorn\altaffilmark{10}, Brad K.
-Gibson\altaffilmark{11}, Eva~K.~Grebel\altaffilmark{12}, Amina Helmi\altaffilmark{13}, Ulisse Munari\altaffilmark{14}, Julio F.
-Navarro\altaffilmark{15,16}, Warren Reid\altaffilmark{17,18}, Arnaud Siebert\altaffilmark{19}, Rosemary Wyse\altaffilmark{20}\\
+\author{
+Andrew R. Casey\altaffilmark{1}, 
+Keith Hawkins\altaffilmark{1}, 
+David W. Hogg\altaffilmark{2,3,4,5}, 
+Melissa Ness\altaffilmark{5}, 
+Hans Walter-Rix\altaffilmark{5}, 
+Georges Kordopatis\altaffilmark{6}, 
+Andrea Kunder\altaffilmark{6}, 
+Matthias Steinmetz\altaffilmark{6}, 
+Sergey Koposov\altaffilmark{1}, 
+Harry Enke\altaffilmark{6}, 
+Jason Sanders\altaffilmark{1}, 
+Gerry Gilmore\altaffilmark{1}, 
+Toma\v{z} Zwitter\altaffilmark{6}, 
+Kenneth C. Freeman\altaffilmark{7}, 
+Luca Casagrande\altaffilmark{7}, 
+Gal~Matijevic\altaffilmark{6}, 
+George Seabroke\altaffilmark{8}, 
+Olivier Bienayme\altaffilmark{9}, 
+Joss Bland-Hawthorn\altaffilmark{10}, 
+Brad~K.~Gibson\altaffilmark{11}, 
+Eva~K.~Grebel\altaffilmark{12}, 
+Amina Helmi\altaffilmark{13}, 
+Ulisse Munari\altaffilmark{14}, 
+Julio F. Navarro\altaffilmark{15,16}, 
+Warren Reid\altaffilmark{17,18}, 
+Arnaud Siebert\altaffilmark{19},
+Rosemary Wyse\altaffilmark{20}\\
 \vspace{2em}

### 2016-09-09 article/unrave.tex
Update figures and text to gold standard sect

@@
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-$\approx$250,000 stars.  We present a data-driven re-analysis of \rave\ 
+up to about 264,000 stars.  We present a data-driven re-analysis of \rave\ 
 spectra using \thecannon, and we provide more precise and accurate stellar
@@
 label in the training is accurate. We can afford to have a small fraction of 
-inaccurate labels; few obvious misclassifications in the training set are affordable.
+inaccurate labels; a few obvious misclassifications in the training set are affordable.
 \item We assume that the training data are similar (in spectra) to the test data 
@@
 vectorizing functions with second-order polynomial expansions (e.g., $\teff^2$, see Sections 
-\ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
+\ref{sec:a-simple-model}--\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
 be considered to be drawn from a Gaussian distribution with zero mean and variance 
@@
 but that sample contains only giant stars.}, and the \lamost\ lower main-sequence suffers
-from the same systematic effects seen in \apogee\ data.
+from the same systematic effects seen in the \apogee\ data.
 
@@
 and Ni); S/N ratios of $>$200~pixel$^{-1}$ in \apogee\ and $>$25~pixel$^{-1}$ in \rave; 
-and we further required that \aspcap\ did not report any peculiar flags 
+and we further required that the \aspcap\ did not report any peculiar flags 
 (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
@@
 ultra metal-poor stars: the atomic transitions in the \rave\ spectral region are simply
-too weak.  For this reason, our adopted abundances for these very metal-poor stars
-represent an `anchor point' in order to ensure our overall metallicity scale is
-correct.  We do not recommend the use of our individual abundance labels at 
-$[{\rm Fe/H}] \sim -4$.  We discuss this issue in more detail in Section~\ref{sec:discussion}.
+too weak to influence the spectrum.  For this reason, our adopted abundances for these
+very metal-poor stars represent an `anchor point' in order to ensure our overall 
+metallicity scale is correct.  We do not recommend the use of our individual abundance
+labels at $[{\rm Fe/H}] \sim -4$.  We discuss this issue in more detail in Section~\ref{sec:discussion}.
 
@@
 trivial.  There are no spectroscopic studies that extend the range of stellar types we 
-are interested in (e.g., FGKM-type stars), and which have a large enough sample size 
+are interested in (e.g., FGKM-type stars), and which also have a large enough sample size 
 that overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
@@
 the total sample size, but in practice it removes most misclassifications.  The 
-situation is far more favourable for dwarfs and giants.  Only $1-4$\% of giant 
+situation is far more favourable for dwarfs and giants.  Only 1--4\% of giant 
 stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
@@
 usually of high fidelity, and given that we have spectra, we can identify any
-spurious misclassifications.
+obvious misclassifications.
 
@@
 confidence intervals in $\teff$ must be below 150~K; the upper and lower confidence 
-intervals in $\logg$ must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must
+intervals in $\logg$ must be less than 0.15~dex; the S/N of the \rave\ spectra must
 exceed 30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict 
@@
 chose not to do this because the number of likely misclassifications in the training
-set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
+set was negligible ($\approx$1\%), and the improvement in main-sequence test set labels
 was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 595 
@@
 
-\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
+\subsection{The simple model: a 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
 \label{sec:a-simple-model}
@@
 resulted in a somewhat peculiar situation.  Detailed abundances are available from \apogee\
-for all giant stars in our sample, but only imprecise metallicities are available from
-\epic\ for stars on the main-sequence and the sub-giant branch.  Here we will construct 
-a simple model for \emph{all} stars that only makes use of three labels ($\teff$, $\logg$, 
-[Fe/H]), before we outline how we derive abundances for giant branch stars.  The complexity
-for this model will be quadratic ($\teff^2$ is the highest term), where the vectorizer 
-$\vecv(\ell_n)$ expands as,
+for all giant stars in our sample, however only imprecise (but accurate on expectation)  
+metallicities are available from \epic\ for stars on the main-sequence and the sub-giant 
+branch.  Here we will construct a simple model for \emph{all} stars that only makes use 
+of three labels ($\teff$, $\logg$, [Fe/H]), before we outline how we derive abundances 
+for giant branch stars.  The complexity for this model will be quadratic ($\teff^2$
+is the highest term), where the vectorizer $\vecv(\ell_n)$ expands as,
 \begin{eqnarray}\label{eq:vectorizer-three-label}
@@
 assign one star as `belonging' to a single model.  In Section~\ref{sec:joining-the-models}
-we describe how we will use the 3-label model of the main-sequence and giant branch
-(introduced in this section) to discriminate between results from a 3-label main-sequence 
-model in Section \ref{sec:unevolved-star-model} and a 9-label giant star model in Section 
-\ref{sec:evolved-star-model}.
+we describe how we will use the simple model introduced in this section to discriminate
+between results from a 3-label main-sequence model in Section \ref{sec:unevolved-star-model} 
+and a 9-label giant star model in Section \ref{sec:evolved-star-model}.
 
@@
 labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
-$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elements.  The RMS
-at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
-between $0.07-0.09$~dex depending on the elemental abundance.
+$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elemental
+abundances.  The RMS at this regularization strength is 69~K in $\teff$, 0.18~dex 
+in $\logg$, and varies between $0.07-0.09$~dex depending on the elemental abundance.
 
@@
 
-\subsection{Deriving joint estimates from all models}
+\subsection{Deriving joint estimates from multiple models}
 \label{sec:joining-the-models}
@@
 temperature $\teff$ and surface gravity $\logg$ between: the main-sequence model 
-(Section~\ref{sec:unevolved-star-model}) and the simplest model (Section~
+(Section~\ref{sec:unevolved-star-model}) and the simple model (Section 
 \ref{sec:a-simple-model}); and the differences between the red giant branch model 
-(Section~\ref{sec:evolved-star-model}) and the simplest model (Section 
+(Section~\ref{sec:evolved-star-model}) and the simple model (Section 
 \ref{sec:a-simple-model}).  We have scaled the differences in $\teff$ and $\logg$ 
@@
 from the central peak.  Therefore, we can adopt the scaled distance in labels 
-$\teff$ and $\logg$ from the simplest model to the main-sequence model $d_{ms}$,
+$\teff$ and $\logg$ from the simple model to the main-sequence model $d_{ms}$,
 \begin{eqnarray}
-	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad ,
+	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},simple}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{simple}}{\delta_{\log{g},ms}}\right)^2 \quad ,
 \end{eqnarray}
 
-\noindent{}and the scaled distance from the simplest model to the giant model $d_{giant}$,
+\noindent{}and the scaled distance from the simple model to the giant model $d_{giant}$,
 \begin{eqnarray}
-	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  \quad ,
+	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},simple}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{simple}}{\delta_{\log{g},giant}}\right)^2  \quad ,
 \label{eq:d_giant}
@@
 In Figure \ref{fig:model-weights} we show the mean relative
-weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional bin of 
-$\hat\teff$ and $\hat\logg$.  For giant stars the relative weight of 
+weight $w_{ms}/(w_{ms} + w_{giant})$ within each two-dimensional hexagonal bin of 
+$\hat\teff$ and $\hat\logg$.  Hereafter when we refer to labels (e.g., $\teff$),
+we refer to those from the joint estimate $\hat\ell$, not individual estimates 
+from separate models.  For giant stars the relative weight of 
 the main-sequence model is zero, and vice-versa for main-sequence stars.
@@
 longer apparent, and the complete structure of the Hertzsprung-Russell diagram
-is visible.  However, there are a number of caveats introduced by the choices 
-we have made to combine estimates from multiple models.  We discuss these 
+is visible.  However, there are a number of caveats introduced by the decisions
+we have made on how to combine estimates from multiple models. We discuss these 
 issues in detail in Section \ref{sec:discussion}.
@@
 the difference between the derived label (for example, $\logg$) over the
-quadrature sum of their formal uncertainties $(\logg_a - \logg_b)/\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}$.  If our derived labels were unbiased and our formal
+quadrature sum of their formal errors $(\logg_a - \logg_b)/\sqrt{{\sigma_{\logg,a}}^2 + {\sigma_{\logg,b}}^2}$.  If our derived labels were unbiased and our formal
 errors were correct, the distribution of these pair-wise comparisons would be
@@
 
-We increased the minimum label error until the distribution of $\eta$ 
-values for a given label approximately reached unity.  The minimum error
-values for each label are shown with the distribution of $\eta$
-values in Figure \ref{fig:pairwise-comparison}.  These minimum values form
-part of our error model, and \emph{have} been added in quadrature with 
-the formal errors; the quoted errors in our labels already incorporate
-these minimum errors.
+We increased the minimum label error until the \emph{variance} of the $\eta$ 
+distribution approximately reached unity.  We found the minimum error in
+$\teff$ to be 70~K, 0.12~dex in $\logg$, and varied between 0.06--0.08~dex
+for individual elements.  The minimum errors are given with the 
+distributions of $\eta$ for each label in Figure \ref{fig:pairwise-comparison}.
+These minimum values form part of our error model, such that they \emph{have} 
+been added in quadrature with the formal errors; the quoted label errors 
+in our catalog include these minimum errors.
 
@@
 \ref{sec:repeat-observations}.  The median S/N of \rave\ spectra is
-50~pixel$^{-1}$, at which point our abundance precision is $\approx0.1$~dex,
-varying a few tenths of a dex between elements.
+50~pixel$^{-1}$, at which point our abundance precision is about 
+0.07~dex, varying a few tenths of a dex between different elements.
 
@@
 
-As an initial point of comparison, we cross-matched our results against the 
+As an initial point of external comparison, we cross-matched our results against the 
 official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
@@
 calculate relative weights.  However, spectra that are unusual from the 
-training set used in the simplest model still project as systematic artefacts.
+training set used in the simple model still project as systematic artefacts.
 
@@
 \caption{Effective temperature $\teff$ and surface gravity $\logg$ for all stars in the training sets.
-Stars are colored by their metallicity [Fe/H], and the three panels show stars in the simplest model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
+Stars are colored by their metallicity [Fe/H], and the three panels show stars in the simple model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (middle panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).\label{fig:training-set-hrd}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/test-set-density.pdf}
-\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the simplest model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
+\caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the simple model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
 These panels demonstrate how a single quadratic model is insufficient for all \rave\ stars (left panel), and illustrate some of the systematic artefacts that can result from testing on stars outside of the training set (center and right panel).
@@
 \includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
-\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the simplest model (top panel), and the giant model and the simplest model (bottom panel).
+\caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the simple model (top panel), and the giant model and the simple model (bottom panel).
 The density scaling is logarithmic, and the differences in $\teff$ and $\logg$ are scaled to make them approximately isotropic (see text for details).
-The peak at $(0, 0)$ represents good agreement between the simplest model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
+The peak at $(0, 0)$ represents good agreement between the simple model and comparison model, whereas the over-densities elsewhere are a consequence of testing the model on stars very different to the training set (e.g., dwarf stars tested on a model trained with only giant stars).\label{fig:joint-model-differences}}
 \end{figure}

### 2016-09-09 article/figures/plot_joint_model_metrics.py

@@ -13,9 +13,9 @@ import AnniesLasso as tc
 RESULTS_PATH = "/data/gaia-eso/arc/rave/results/"
 RESULTS_PATH = "../../"
 
-ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v37.fits.gz"))
+ms_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v43.fits.gz"))
 giant_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v42.fits.gz"))
-joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v41.fits.gz"))
+joint_results = Table.read(os.path.join(RESULTS_PATH, "rave-tgas-v46.fits.gz"))
 
 for t in (ms_results, giant_results, joint_results):
     if "Name" not in t.dtype.names:
@@ -64,7 +64,7 @@ fig.subplots_adjust(
 extent = (3000, 8000, 0.5, 5.5)
 
 titles = (
-    r"${\rm Joint}$ ${\rm model}$",
+    r"${\rm Simple}$ ${\rm model}$",
     r"${\rm Main{-}sequence}$ ${\rm model}$", 
     r"${\rm Giant}$ ${\rm branch}$ ${\rm model}$",
 )
@@ -141,8 +141,8 @@ axes[0].hexbin(x, y, **kwds)
 axes[0].set_xticks(ticks)
 axes[0].set_yticks(ticks)
 
-axes[0].set_xlabel(r"$(T_{{\rm eff},ms} - T_{{\rm eff},joint})/\delta_{T_{{\rm eff},ms}}$")
-axes[0].set_ylabel(r"$(\log{g}_{ms} - \log{g}_{joint})/\delta_{\log{g},ms}$")
+axes[0].set_xlabel(r"$(T_{{\rm eff},ms} - T_{{\rm eff},simple})/\delta_{T_{{\rm eff},ms}}$")
+axes[0].set_ylabel(r"$(\log{g}_{ms} - \log{g}_{simple})/\delta_{\log{g},ms}$")
 
 
 x_mu, y_mu = (0, 0)
@@ -157,8 +157,8 @@ hexbin = axes[1].hexbin(x2, y2, **kwds)
 axes[1].set_xticks(ticks)
 axes[1].set_yticks(ticks)
 
[…]

### 2016-09-09 article/figures/plot_pairwise.py

@@ -10,7 +10,7 @@ from matplotlib.ticker import MaxNLocator
 
 DATA_PATH = "../../"
 
-individual_results = Table.read(os.path.join(DATA_PATH, "unrave-v0.96.fits.gz"))
+individual_results = Table.read(os.path.join(DATA_PATH, "unrave-v0.97.fits.gz"))
 
 OK = individual_results["QC"]
 individual_results = individual_results[OK]
@@ -34,15 +34,15 @@ individual_results = individual_results.group_by("RAVEID")
 
 
 error_floor = {
-    "TEFF": 50, #100,
-    "LOGG": 0.10, #0.15,
-    "FE_H": 0.05, #0.08,
-    "MG_H": 0.06, #0.08,
-    "O_H": 0.06, #0.08,
-    "AL_H": 0.07, #0.08,
-    "CA_H": 0.06, #0.08,
-    "SI_H": 0.06, #0.08,
-    "NI_H": 0.05, #0.08,
+    "TEFF": 70, #100,
+    "LOGG": 0.12, #0.15,
+    "FE_H": 0.06, #0.08,
+    "MG_H": 0.07, #0.08,
+    "O_H": 0.07, #0.08,
+    "AL_H": 0.08, #0.08,
+    "CA_H": 0.07, #0.08,
+    "SI_H": 0.07, #0.08,
+    "NI_H": 0.06, #0.08,
 }
 
 pairwise_metrics = {}
@@ -62,7 +62,7 @@ for j, si in enumerate(indices[:-1]):
         error = np.repeat(individual_results["E_{}".format(label_name)].data[si:ei], N).reshape(-1, N)
 
         diff = value - value.T
-        summed_error = (error**2 + error.T**2 + 2*error_floor[label_name]**2 + 2*error**2 + 2*error.T**2)**0.5
+        summed_error = (error**2 + error.T**2 + 2*error_floor[label_name]**2)**0.5
 
         # Just take the upper diagonal of the matrix so we get all pair-wise
         # comparisons, just once.

### 2016-09-09 article/unrave.tex
Use correct numbers in abstract
update ges

@@
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-up to about 264,000 stars.  We present a data-driven re-analysis of \rave\ 
+up to 264,276 stars.  We present a data-driven re-analysis of \rave\ 
 spectra using \thecannon, and we provide more precise and accurate stellar
@@
 
-As an initial point of external comparison, we cross-matched our results against the 
-official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
+We cross-matched our results against the official fourth \rave\ data release 
+as an initial point of external comparison (Figure \ref{fig:rave-dr4-comparison}).
 In order to provide a fair comparison, we only show stars that meet a number
@@
 their analysis on high-resolution, high S/N spectra.  We found four notable
-studies with a sufficient level of overlap: the disk studies by \citet{Reddy_2003,
-Reddy_2006} and \citet{Bensby_2014}, as well as the \citet{Valenti_Fischer_2005} 
-work on exoplanet host star candidates.  These studies perform a careful (manual;
-expert) analysis using extremely high-resolution, high S/N spectra, and make use 
-of \project{Hipparcos} parallaxes where possible.  Most of the stars in these 
-samples are main-sequence or sub-giant stars.  Therefore, these works constitute 
-an excellent comparison to evaluate the accuracy of our results on the 
+studies with a sufficient level of overlap: the Milky Way disk studies by 
+\citet{Reddy_2003,Reddy_2006} and \citet{Bensby_2014}, as well as the 
+\citet{Valenti_Fischer_2005} work on exoplanet host star candidates.  These 
+studies perform a careful (manual; expert) analysis using extremely 
+high-resolution, high S/N spectra, and make use of \project{Hipparcos} 
+parallaxes where possible.  Most of the stars in these samples are 
+main-sequence or sub-giant stars.  Therefore, these works constitute an 
+excellent comparison to evaluate the accuracy of our results on the 
 main-sequence and sub-giant branch.
@@
 metallicity, near $[{\rm Fe/H}] \approx -1.5$.  The scatter increases for
-the four stars in the overlap sample with [Fe/H] $< -1$, however these
-spectra have low S/N ratios, which is reflected in the larger uncertainties
+the few stars in the overlap sample with $[{\rm Fe/H}] < -1$, in the regime
+where the influence from atomic transitions of these elements becomes very
+small in \rave\ spectra. Moreover, these particular stars have lower S/N
+ratios, which is reflected in the larger uncertainties
 reported for these metallicities.
@@
 factors.  The absolute bias for individual elements varies from as low as
-0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
+0.06~dex ([Al,Mg/H]) to as high as 0.26~dex ([Si/H]), where we over-estimate 
 [Si/H] abundances relative to the \ges\ survey.  The large bias in [Si/H]
@@
 \apogee, the source of our training set for giant star abundances.  
-The RMS deviation in each
-label is small for stars with [X/H] $> -0.5$, before increasing at lower
-metallicities.  If we consider all stars, the smallest abundance RMS we see
-with respect to \ges\ is 0.18~dex for [Ca/H].  The increasing RMS at low 
-metallicity is likely a consequence of multiple factors, namely: inaccurate
-abundance labels for metal-poor stars (Section~\ref{sec:the-training-set});
-only weak, blended lines being available in \rave, which cease to be visible
-in hot and/or metal-poor stars; and to a lesser extent, low S/N ratios for 
-those particular stars being compared.  Unfortunately not all of these
-factors are represented by the quoted errors in each label.  For these reasons, 
-although it affects only a small number of stars, we recommend caution when 
-using individual abundances for very metal-poor giant stars in our sample.
+The RMS deviation in each label is small for stars with [X/H] $> -0.5$, 
+before increasing at lower metallicities.  If we consider all stars, the 
+smallest abundance RMS we see with respect to \ges\ is 0.16~dex for [Ca/H] 
+and [Al/H].  The increasing RMS at low metallicity is likely a consequence
+of multiple factors, namely: inaccurate abundance labels for metal-poor stars
+(Section~\ref{sec:the-training-set}); only weak, blended lines being available
+in \rave, which cease to be visible in hot and/or metal-poor stars; and to a 
+lesser extent, low S/N ratios for those particular stars being compared.  
+Unfortunately not all of these factors are represented by the quoted errors 
+in each label.  For these reasons, although it affects only a small number of
+stars, we recommend caution when using individual abundances for very 
+metal-poor giant stars in our sample.
 

### 2016-09-09 article/unrave.tex
Include specifics of OC selection #32

@@
 The fourth \rave\ data release made use of a number of high-resolution studies
-to verify the accuracy of the reported stellar atmospheric parameters.  These
+to verify the accuracy of their derived stellar atmospheric parameters.  These
 samples include main-sequence stars, giant stars, with a particular focus to
 include metal-poor stars to identify (and correct) any deviations at low
-and high metallicities.  We refer the reader to \citet{Kordopatis_2013} for
-the full compilation of literature sources.  Although the stellar atmospheric 
-parameters in this compilation come from multiple (heterogeneous) sources, we 
-find generally good agreement with these works (Figure 
-\ref{fig:kordopatis-calibration}).  However, some reservation is warranted 
-when evaluating this comparison, as some of the metal-poor stars in this 
-calibration sample formed part of our training set.
+metallicities.  We refer the reader to \citet{Kordopatis_2013} for the full 
+compilation of literature sources.  Although the stellar atmospheric parameters
+in this compilation come from multiple (heterogeneous) sources, we find 
+generally good agreement with these works 
+(Figure \ref{fig:kordopatis-calibration}).  However, we note that some 
+reservation is warranted when evaluating this comparison, as some of the 
+metal-poor stars in this calibration sample formed part of our training set.
 
@@
 After verifying that our atmospheric parameters and abundances are comparable 
-with other high-resolution studies, here we verify that the abundances we find
-are consistent with expectations from astrophysics.  In the \rave\ survey, 
-\cite{Anguiano_2015} identified 70 stars with positions and radial velocities that are
-consistent with being members of globular clusters: 49 stars belonging to NGC~5139
+with high-resolution studies, here we verify that our results are consistent
+with expectations from astrophysics.  In the \rave\ survey, \cite{Anguiano_2015}
+identified 70 stars with positions and radial velocities that are consistent 
+with being members of globular clusters: 49 stars belonging to NGC~5139
 ($\omega$~Centauri), 11 members of the retrograde globular cluster NGC~3201, and 10 
@@
 are adopted from \citet{Kunder_2016,Marin-Franch_2009}, and the 
-\citet[][accessed 6 September 2016]{Harris_1996} online catalog of globular 
+\citet[][accessed 6 September 2016]{Harris_1996} catalog of globular 
 cluster properties.  The globular cluster with the most number of members is 
 NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
-consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,Carretta_2013}. 
-Based on the pre-defined membership criteria, we find the mean metallicity
-of $\omega$-Centauri to be $[{\rm Fe/H}] = -0.85$.  However, it is clear that the 
-membership criteria could be improved with our revised metallicities and 
-detailed chemical abundances.  Indeed, our individual abundance labels could 
-be used to identify globular cluster members that are now tidally disrupted 
+that is consistent with high-resolution studies \citep{Marino_2011,Carretta_2009,
+Carretta_2013}. Based on the pre-defined membership criteria, we find the mean 
+metallicity of $\omega$-Centauri to be $[{\rm Fe/H}] = -0.85$.  However, it 
+is clear that the membership criteria could be improved with our revised 
+metallicities and detailed chemical abundances.  Indeed, our individual abundance
+labels could be further used to identify globular cluster members --- at least, of
+relatively metal-rich clusters ---  that are now tidally disrupted 
 \citep{Anguiano_2016,Kuzma_2016,Navin_2016}, even for stars with low S/N ratios.
@@
 
-The \rave\ survey observed $\sim$160 highly probable members of several open clusters:
-78 stars in the Pleiades, 26 in Hyades, 30 stars in the solar-metallicity cluster
-M67, and 13 stars in IC4651. Details on the open cluster membership and selection can 
-be found in \cite{Kunder_2016}.  We show the effective temperature $\teff$ and surface
-gravity $\logg$ for these stars in Figure \ref{fig:open-cluster-HRD}.  The isochrones
-are sourced from \citet{Bressan_2012}, with cluster properties adopted from 
-\cite{Kunder_2016} and \cite{Kharchenko_2013}.
+Using positions, proper motions, and metallicities from the \rave\ official
+data release (such that they can be used as comparison), we identify $\sim$160 
+probable members of four open clusters that were observed by \rave.  
+Upon requiring $(\alpha,\delta) = (51.5^\circ$ to $57.5^\circ$, 21.5$^\circ$ to 28$^\circ)$,
+have error in metallicity \texttt{e\_M\_H\_K} $< 0.5$~dex, and radial
+velocities between $-10$~km~s$^{-1}$ and 20~km~s$^{-1}$, we identify 78 
+Pleiades members.  For the Hyades, by requiring stars within $3.5^\circ$ of the
+cluster centre to have: radial velocities between 36~km~s$^{-1}$ and 44~km~s$^{-1}$,
+$[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35, and \rave\ metallicity errors
+$<0.12$~dex, we identify 26 potential members.  Similarly for IC4561 we select
+13 stars by requiring: radial velocities within $-20$~km~s$^{-1}$ to $-40$~km~s$^{-1}$,
+an error in radial velocity $<3$~km~s$^{-1}$; $[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35;
+$\alpha$ between $261.1^\circ$ and $261.6^\circ$; and $\delta$ within $-49.7^\circ$
+and $-50.1^\circ$.  For the solar-metallicity open cluster member, we 30 select stars solely on
+position: $(\alpha,\delta) = (133.1^\circ$ to $132.5^\circ$, $11.6^\circ$ to $12^\circ)$.
+We show the effective temperature $\teff$ and surface gravity $\logg$ for these
+cluster candidates in Figure \ref{fig:open-cluster-HRD}.  The isochrones are 
+sourced from \citet{Bressan_2012}, with cluster properties adopted from 
+\cite{Kharchenko_2013}.
 

### 2016-09-09 article/unrave.tex
Leave the specifics out; we will want to refine this selection using our values at respond-to-referee

@@
 probable members of four open clusters that were observed by \rave.
-Upon requiring $(\alpha,\delta) = (51.5^\circ$ to $57.5^\circ$, 21.5$^\circ$ to 28$^\circ)$,
-have error in metallicity \texttt{e\_M\_H\_K} $< 0.5$~dex, and radial
-velocities between $-10$~km~s$^{-1}$ and 20~km~s$^{-1}$, we identify 78 
-Pleiades members.  For the Hyades, by requiring stars within $3.5^\circ$ of the
-cluster centre to have: radial velocities between 36~km~s$^{-1}$ and 44~km~s$^{-1}$,
-$[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35, and \rave\ metallicity errors
-$<0.12$~dex, we identify 26 potential members.  Similarly for IC4561 we select
-13 stars by requiring: radial velocities within $-20$~km~s$^{-1}$ to $-40$~km~s$^{-1}$,
-an error in radial velocity $<3$~km~s$^{-1}$; $[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35;
-$\alpha$ between $261.1^\circ$ and $261.6^\circ$; and $\delta$ within $-49.7^\circ$
-and $-50.1^\circ$.  For the solar-metallicity open cluster member, we 30 select stars solely on
-position: $(\alpha,\delta) = (133.1^\circ$ to $132.5^\circ$, $11.6^\circ$ to $12^\circ)$.
+Specifically we identify 78 potential Pleiades members, 26 candidates in the
+Hyades, another 13 in IC4561, and 30 stars in the solar-metallicity open
+cluster M67.
+%Upon requiring $(\alpha,\delta) = (51.5^\circ$ to $57.5^\circ$, 21.5$^\circ$ to 28$^\circ)$,
+%have error in metallicity \texttt{e\_M\_H\_K} $< 0.5$~dex, and radial
+%velocities between $-10$~km~s$^{-1}$ and 20~km~s$^{-1}$, we identify 78 
+%Pleiades members.  For the Hyades, by requiring stars within $3.5^\circ$ of the
+%cluster centre to have: radial velocities between 36~km~s$^{-1}$ and 44~km~s$^{-1}$,
+%$[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35, and \rave\ metallicity errors
+%$<0.12$~dex, we identify 26 potential members.  Similarly for IC4561 we select
+%13 stars by requiring: radial velocities within $-20$~km~s$^{-1}$ to $-40$~km~s$^{-1}$,
+%an error in radial velocity $<3$~km~s$^{-1}$; $[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35;
+%$\alpha$ between $261.1^\circ$ and $261.6^\circ$; and $\delta$ within $-49.7^\circ$
+%and $-50.1^\circ$.  For the solar-metallicity open cluster member, we 30 select stars solely on
+%position: $(\alpha,\delta) = (133.1^\circ$ to $132.5^\circ$, $11.6^\circ$ to $12^\circ)$.
 We show the effective temperature $\teff$ and surface gravity $\logg$ for these

### 2016-09-09 article/unrave.tex
Update figures to 0.97 to incl error floors; fix minor things

@@
 Luca Casagrande\altaffilmark{7}, 
-Gal~Matijevic\altaffilmark{6}, 
+Gal~Matijevi\v{c}\altaffilmark{6}, 
 George Seabroke\altaffilmark{8}, 
@@
 Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-up to 264,276 stars.  We present a data-driven re-analysis of \rave\ 
+over 200,000 stars.  We present a data-driven re-analysis of \rave\ 
 spectra using \thecannon, and we provide more precise and accurate stellar
@@
 main-sequence/sub-giant star or not.  The results for the entire survey sample are shown
-in the centre panel of Figure \ref{fig:test-set-density}.  The increased density of 
+in the center panel of Figure \ref{fig:test-set-density}.  The increased density of 
 solar-type stars is consistent with \rave\ observing stars in the local neighbourhood, 
@@
 
-Using positions, proper motions, and metallicities from the \rave\ official
-data release (such that they can be used as comparison), we identify $\sim$160 
+Using positions, proper motions, and metallicities from the \rave\ survey
+(i.e., not ours, such that they can be used as comparison), we identify $\sim$160 
 probable members of four open clusters that were observed by \rave.
@@
 %Pleiades members.  For the Hyades, by requiring stars within $3.5^\circ$ of the
-%cluster centre to have: radial velocities between 36~km~s$^{-1}$ and 44~km~s$^{-1}$,
+%cluster center to have: radial velocities between 36~km~s$^{-1}$ and 44~km~s$^{-1}$,
 %$[{\rm M/H}]_{RAVE}$ within $-0.15$ and 0.35, and \rave\ metallicity errors
@@
 to consider this method is as \emph{ad hoc} as any other approach.  The relative 
-weighting has no warranty to be (formally) correct, and may introduce inconsistencies
-or systematic errors rather than minimizing them.
+weighting has no warranty to be (formally) correct, and therefore may introduce 
+inconsistencies or systematic errors rather than minimizing them.
 
@@
 \citep{Fulbright_2010,Ruchti_2011}, but it is not known if the stellar parameters are
-of high fidelity because we have limited quality metrics available.  Moreover, there
-is no guarantee that the stellar parameters \emph{or} abundances are on the same scale
-as \apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
+of high fidelity because we have a limited number of quality statistics available. 
+Moreover, there is no guarantee that the stellar parameters \emph{or} abundances are 
+on the same scale as \apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.
 
@@
 considered trustworthy in any sense.  For this reason we have updated the electronic
-catalog to discard these results, and included a quality control flag to indicate
-the result was discarded (see catalog \texttt{README} file for specifics related to
-the data structure and content).
+catalog to discard these results as erroneous.
 
@@
 calculate relative weights.  However, spectra that are unusual from the 
-training set used in the simple model still project as systematic artefacts.
+training set used in the simple model could still project as systematic 
+artefacts in label space.
 
@@
 
-We have included a boolean flag column in our catalog for quality control,
-to signify whether a result is thought to be reasonable.  Specifically, 
-stars where \texttt{SPARV} indicates a $\teff > 8000$~K do not meet our quality 
-control flag.  For all other stars, we further require $\chi_r^2 < 3$ and
-$S/N >10$~pixel$^{-1}$.  These quality constraints represent the minimum
-acceptable constraints and should not be taken verbatim: judicious use of
-the morphological classifications \citep{Matijevic_2012} or additional 
-metadata from the \rave\ pre-processing pipelines is strongly encouraged.
+This line of reasoning extends to spectra with other peculiar characteristics 
+(e.g., chromospheric emission), and for these reasons we recommend the use of
+additional metadata to investigate possible artefacts.  In our catalog we have
+included a column containing a boolean flag to indicate whether the labels pass
+very weak quality constraints.  Specifically, we flag results as failing our
+quality constraints if \texttt{SPARV} indicates a $\teff > 8000$~K, or if 
+$\chi_r^2 > 3$, or if $S/N < 10$~pixel$^{-1}$.  These quality constraints 
+represent the minimum acceptable conditions and should not be taken verbatim: 
+judicious use of the morphological classifications \citep{Matijevic_2012} or 
+additional metadata from the \rave\ pre-processing pipelines is strongly 
+encouraged.
 
@@
 \caption{Distribution of differences in label estimates from multiple visits, divided by the quadrature sum of their formal errors, and a minimum error value for each observation.
-If our measurements were unbiased and our errors were representative, no minimum error floor would be required and this distribution would by normally-distributed with zero mean and unity variance.
-We increased the error floor for a given label until the distribution variance approximately reached unity. The error floors are given for each label.\label{fig:pairwise-comparison}}
+If our measurements were unbiased and our errors were representative, no minimum error floor would be required and the distribution of $\eta$ would by normally-distributed with zero mean and unity variance.
+We increased the error floor for each label until the variance in the distribution of $\eta$ approximately reached unity. Derived error floors are shown for each label.\label{fig:pairwise-comparison}}
 \end{figure*}
@@
 Hot stars are not included in our training set, and appear above the turnoff in our labels.
-For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars with $\teff > 8000$~K (as reported by \texttt{SPARV}), and other results that are of poor quality.\label{fig:hot-stars}}
+For this reason, we supply a boolean quality control flag \texttt{QC} that flags stars (as \texttt{False}, for failing to meet our quality constraints) with $\teff > 8000$~K (as reported by \texttt{SPARV}), or results that do not meet other minimum constraints (see Section \ref{sec:discussion} for details).\label{fig:hot-stars}}
 \end{figure}

### 2016-09-09 article/figures/plot_gold_standard_comparisons.py

@@ -189,7 +189,7 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
         allx.extend(x)
         ally.extend(y)
         if ax.is_first_row():
-            #ax.errorbar(x, y - y, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
+            ax.errorbar(x, y - x, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
             _ = ax.scatter(x, y - x, c=c0, cmap="plasma", **kwds)
             diffs.extend(y-x)
 
@@ -197,7 +197,7 @@ for ax, label_name in zip(axes, list(label_names) + list(label_names)):
                 handles.append(ax.scatter([-10000], [-10000], facecolor="#cccccc", **kwds))
 
         else:
-            #ax.errorbar(x, y, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
+            ax.errorbar(x, y, yerr=yerr, fmt=None, ecolor="#666666", zorder=-1)
             scat = ax.scatter(x, y, c=c0, cmap="plasma", **kwds)
     print(label_name, np.nanmedian(diffs), np.nanstd(diffs), np.isfinite(diffs).sum())
 

### 2016-09-09 article/figures/plot_hrd_test.py

@@ -49,7 +49,7 @@ for i, snr in enumerate(snrs):
     # On top figure show density.
     # On bottom axes show metallicity.
     
-    ok = cannon_dr1["QC"]
+    ok = cannon_dr1["QC"] * (cannon_dr1["SNR"] > snr)
 
     ax = axes[0, i]
 

### 2016-09-09 article/figures/plot_hrd_train.py

@@ -20,17 +20,18 @@ else:
 
 """
 names = (
-    r"${\rm Joint}$ ${\rm model}$",
+    r"${\rm Simple}$ ${\rm model}$",
     r"${\rm Main{-}sequence}$ ${\rm model}$", 
     r"${\rm Giant}$ ${\rm model}$", 
 )
 
 import AnniesLasso as tc
 labelled_sets = (
-    tc.load_model("rave-tgas-v41.model").labelled_set, # joint
-    tc.load_model("rave-tgas-v16b.model").labelled_set, # ms result
+    tc.load_model("rave-tgas-v46.model").labelled_set, # joint
+    tc.load_model("rave-tgas-v43.model").labelled_set, # ms result
     tc.load_model("rave-tgas-v42.model").labelled_set, # giant result
 )
+
 labelled_sets[1]["TEFF"] = labelled_sets[1]["EPIC_TEFF"]
 labelled_sets[1]["LOGG"] = labelled_sets[1]["EPIC_LOGG"]
 labelled_sets[1]["FE_H"] = labelled_sets[1]["EPIC_FEH"]

### 2016-09-09 article/figures/plot_open_clusters.py

@@ -67,7 +67,7 @@ for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
 
     isochrone, isochrone_path = (None, "{}_padova_iso.dat".format(cluster_name))
     if os.path.exists(isochrone_path):
-        isochrone = Table.read(isochrone_path, format="ascii", header_start=12)
+        isochrone = Table.read(isochrone_path, format="ascii")
         isochrone[iso_teff] = 10**isochrone["logTe"]
 
     match = (dr4["Cluster"] == cluster_name) * dr4_ok

### 2016-09-09 article/figures/plot_v34_search.py

@@ -86,6 +86,8 @@ for i, (model, filenames, base_filename, color, title) in enumerate(comparisons)
         x.append(Lambda)
         y.append(np.nanstd(model.labels_array - test_set, axis=0))
 
+        if i == 0 and "-5.736" in filename:
+            print(i, filename, np.nanmean(model.labels_array - test_set, axis=0), np.nanstd(model.labels_array - test_set, axis=0))
 
     x = np.array(x)
     y = np.array(y)

### 2016-09-09 article/unrave.tex
Updated abstract, final paragraphs, etc

@@
-%% This file is part of the unRAVE Project 
+%% This file is part of the RAVE-on Project 
 %% Copyright 2016 the authors.  All rights reserved.
@@
 \newcommand{\tgas}{\project{\acronym{TGAS}}}
-\newcommand{\unrave}{\project{unRAVE}}
+\newcommand{\raveon}{\rave-on}
 
@@
 
-% TODO: Final checks based on released version.
-\newcommand{\Nstarsqc}{447,207} % v0.95 finite values
-\newcommand{\Nabundances}{1,692,825} %v0.95 finite values
+\newcommand{\ReportedStellarParameters}{441,397} % catalog v0.97
+\newcommand{\ReportedAbundances}{1,685,851} % catalog v0.97
 
@@
 \begin{document}
-\title{The \project{unRAVE} catalog of stellar atmospheric parameters and\\
+\title{The \raveon\ catalog of stellar atmospheric parameters and\\
 chemical abundances for chemo-dynamic studies in the \gaia\ era}
@@
 \author{
-Andrew R. Casey\altaffilmark{1}, 
-Keith Hawkins\altaffilmark{1}, 
-David W. Hogg\altaffilmark{2,3,4,5}, 
-Melissa Ness\altaffilmark{5}, 
-Hans Walter-Rix\altaffilmark{5}, 
-Georges Kordopatis\altaffilmark{6}, 
-Andrea Kunder\altaffilmark{6}, 
-Matthias Steinmetz\altaffilmark{6}, 
-Sergey Koposov\altaffilmark{1}, 
-Harry Enke\altaffilmark{6}, 
-Jason Sanders\altaffilmark{1}, 
-Gerry Gilmore\altaffilmark{1}, 
-Toma\v{z} Zwitter\altaffilmark{6}, 
-Kenneth C. Freeman\altaffilmark{7}, 
-Luca Casagrande\altaffilmark{7}, 
+Andrew~R.~Casey\altaffilmark{1},
+Keith~Hawkins\altaffilmark{1},
+David~W.~Hogg\altaffilmark{2,3,4,5},
+Melissa~Ness\altaffilmark{5},
+Hans~Walter-Rix\altaffilmark{5},
+Georges~Kordopatis\altaffilmark{6},
+Andrea~Kunder\altaffilmark{6},
+Matthias~Steinmetz\altaffilmark{6},
+Sergey~Koposov\altaffilmark{1},
+Harry~Enke\altaffilmark{6},
+Jason~Sanders\altaffilmark{1},
+Gerry~Gilmore\altaffilmark{1},
+Toma\v{z}~Zwitter\altaffilmark{6},
+Kenneth~C.~Freeman\altaffilmark{7},
+Luca~Casagrande\altaffilmark{7},
 Gal~Matijevi\v{c}\altaffilmark{6},
-George Seabroke\altaffilmark{8}, 
-Olivier Bienayme\altaffilmark{9}, 
-Joss Bland-Hawthorn\altaffilmark{10}, 
+George~Seabroke\altaffilmark{8},
+Olivier~Bienaym\'e\altaffilmark{9},
+Joss~Bland-Hawthorn\altaffilmark{10},
 Brad~K.~Gibson\altaffilmark{11},
 Eva~K.~Grebel\altaffilmark{12},
-Amina Helmi\altaffilmark{13}, 
-Ulisse Munari\altaffilmark{14}, 
-Julio F. Navarro\altaffilmark{15,16}, 
-Warren Reid\altaffilmark{17,18}, 
-Arnaud Siebert\altaffilmark{19},
-Rosemary Wyse\altaffilmark{20}\\
+Amina~Helmi\altaffilmark{13},
+Ulisse~Munari\altaffilmark{14},
+Julio~F.~Navarro\altaffilmark{15,16},
+Warren~Reid\altaffilmark{17,18},
+Arnaud~Siebert\altaffilmark{9},
+Rosemary~Wyse\altaffilmark{19}\\
 \vspace{2em}
@@
 \begin{abstract}
-The Milky Way is a powerful laboratory for understanding galaxy formation,
-as it can provide orbits, ages, atmospheric parameters, as well as chemical 
-abundances for vast sets of individual stars.  These inferences require 
-both astrometric and spectroscopic data.  Indeed, in order to be maximally
-useful for chemo-dynamic studies, the Tycho-Gaia Astrometric Solution 
-(\tgas) sample in the first \gaia\ data release requires a spectroscopic 
-complement that includes radial velocities, stellar atmospheric parameters, and 
-elemental abundances.  Among existing spectroscopic samples, the RAdial 
-Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
-over 200,000 stars.  We present a data-driven re-analysis of \rave\ 
-spectra using \thecannon, and we provide more precise and accurate stellar
-parameters and abundances than available in previous \rave\ data releases. 
-We constructed our model using high-fidelity \apogee\ stellar parameters 
-and abundances for red giant branch stars that overlap with \rave, and 
-stellar parameters from the \project{K2}/\project{Ecliptic Pole Input 
-Catalog} for main-sequence and sub-giant branch stars.  We derive, and 
-validate, improved effective temperature $\teff$, surface gravity $\logg$, 
-and chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
-for \Nstarsqc\ stars.  In total we report \Nabundances\ elemental 
-abundances, with a typical precision of 0.07~dex.  The \unrave\ catalog 
-presented here, in conjunction with \tgas, represents the most powerful 
-data set for chemo-dynamic analysis at the dawn of broadly available 
-\gaia\ data.
+The orbits, atmospheric parameters, chemical abundances, and ages of 
+individual stars in the Milky Way provide the most comprehensive 
+illustration of galaxy formation available.  The Tycho-\gaia\ 
+Astrometric Solution (\tgas) will deliver astrometric parameters for
+the largest ever sample of Milky Way stars, though its full potential
+cannot be realized without the addition of complementary spectroscopy.  
+Among existing spectroscopic surveys, the RAdial Velocity Experiment 
+(\rave) has the largest overlap with \tgas\ ($\gtrsim$200,000 stars).  
+We present a data-driven re-analysis of \Nspectra\ \rave\ spectra using 
+\thecannon.  For red giants, we build our model using high-fidelity 
+\apogee\ stellar parameters and abundances for stars that overlap 
+with \rave.  For main-sequence and sub-giant stars, our model uses 
+stellar parameters from the \epic\ catalog.  We derive and validate 
+effective temperature $\teff$, surface gravity $\logg$, and chemical 
+abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni).  We 
+report a total of \ReportedAbundances\ elemental abundances with a 
+typical precision of 0.07~dex, a substantial improvement over past 
+\rave\ data releases.  The synthesis of \raveon\ and \tgas\ is the 
+most powerful data set for chemo-dynamic analyses of the Milky Way ever produced.
 \end{abstract}
@@
 survey is expected to have the largest overlap with the first \gaia\ 
-% TODO: Sanders: 292,036; Kunder: 264,276
 data release: up to 264,276 stars.  We used the \gaia\ universe model snapshot 
@@
 to be very satisfactory.  Indeed, given the metallicity precision available
-in the \unrave\ catalog, these results will likely be useful for future 
+in the \raveon\ catalog, these results will likely be useful for future 
 studies based on exoplanet host star properties (e.g., 
@@
 low S/N ratios in \rave\ ($\approx 25$~pixel$^{-1}$), there is good agreement
-in with \ges\ and the \unrave\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
+in with \ges\ and the \raveon\ stellar parameters (Figure~\ref{fig:ges-stellar-parameters}).  
 The RMS in effective temperature, surface gravity and metallicity is 233~K, 
@@
 We have presented a comprehensive collection of precise stellar abundances 
-for stars in the first \gaia\ data release.  In total we derive reliable 
-stellar atmospheric parameters for \Nstarsqc, and report more than 1.69 
-million abundances.  Despite the caveats and limitations discussed here, 
-our validation experiments and comparisons with high-resolution 
-spectroscopic studies suggests that our results have sufficient accuracy 
-and precision to be useful for chemo-dynamic studies that become imminently 
-feasible only in the era of \gaia\ data.  Indeed, we are optimistic that 
-the \unrave\ catalog will advance understanding of the Milky Way's 
-formation and evolution.
+for stars in the first \gaia\ data release.  In total we derive stellar 
+atmospheric parameters for \ReportedStellarParameters\ stars, and report more 
+than 1.69 million abundances.  Despite the caveats and limitations discussed here, 
+our validation experiments and comparisons with high-resolution spectroscopic
+studies suggests that our results have sufficient accuracy and precision to
+be useful for chemo-dynamic studies that become imminently feasible only in
+the era of \gaia\ data.  We are optimistic that the \raveon\ catalog will 
+advance understanding of the Milky Way's formation and evolution.
 
@@
 
-% TODO: Replace ZENODO-URL with a data url
 \noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
 and this document was compiled from revision hash \texttt{\githash} in that repository.
-Inferred labels, associated errors, and relevant metadata are available electronically
-at \texttt{\url{ZENODO-URL}} \citep{ZENODO-CITATION}.  These results are also available
-through the \rave\ DR5 database \citep{Kunder_2016}. Please note that it is a condition
+Derived labels, associated errors, and relevant metadata are available electronically
+through the \rave\ from 19 September 2016 onwards. Please note that it is a condition
 of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must 
@@
 
-\bibitem[Wojno et al.(2016)]{Wojno_2016} Wojno, J., et al.\ 2016, submitted
+\bibitem[Wojno et al.(2016)]{Wojno_2016} Wojno, J., et al.\ 2016, in preparation
 
@@
 Stars are colored by the metallicity of each study.
-Circles indicate literature markers in the first three panels, and the linked triangles indicate \project{unRAVE} parameters for the same object.
+Circles indicate literature markers in the first three panels, and the linked triangles indicate \raveon\ parameters for the same object.
 This figure illustrates the good qualitative agreement in the shape of the turnoff and sub-giant branch.\label{fig:gold-standard-hrd}}
@@
 \includegraphics[width=0.65\textwidth]{figures/gce.pdf}
-\caption{Detailed chemical abundances ([X/Fe]) for giant stars in \unrave\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
+\caption{Detailed chemical abundances ([X/Fe]) for giant stars in \raveon\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
 Bin densities are scaled logarithmically.
@@
 
-\altaffiltext{1}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
-\altaffiltext{2}{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
-\altaffiltext{3}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
-\altaffiltext{4}{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
-\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-\altaffiltext{6}{University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia}
-\altaffiltext{7}{Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia}
-\altaffiltext{8}{Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK}
-\altaffiltext{9}{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France}
-\altaffiltext{10}{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia}
-\altaffiltext{11}{E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom}
-\altaffiltext{12}{Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany}
-\altaffiltext{13}{Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands}
-\altaffiltext{14}{INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy}
-\altaffiltext{15}{Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2}
-\altaffiltext{16}{Senior CIfAR Fellow}
-\altaffiltext{17}{Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia}
-\altaffiltext{18}{Western Sydney University, Penrith South DC, NSW 1797}
-\altaffiltext{19}{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l?Universit\'e, F-67000 Strasbourg, France}
-\altaffiltext{20}{Johns Hopkins University, Baltimore, MD, USA}
+% A magic, bad, hack.
+\noindent{}$^{1}${Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK} \\
+$^{2}${Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA} \\
+$^{3}${Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Pl., room 424, New York, NY 10003, USA} \\
+$^{4}${Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA} \\
+$^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
+$^{6}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
+$^{7}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
+$^{8}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
+$^{9}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
+$^{10}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
+$^{11}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
+$^{12}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
+$^{13}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
+$^{14}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
+$^{15}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
+$^{16}${Senior CIfAR Fellow} \\
+$^{17}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
+$^{18}${Western Sydney University, Penrith South DC, NSW 1797} \\
+$^{19}${Johns Hopkins University, Baltimore, MD, USA} \\
 

### 2016-09-09 article/unrave.tex
typos

@@
 Derived labels, associated errors, and relevant metadata are available electronically
-through the \rave\ from 19 September 2016 onwards. Please note that it is a condition
-of using these results that the fifth \rave\ data release by \citet{Kunder_2016} must 
-also be cited, as the work presented here would not have been possible without the 
+through the \rave\ database from 19 September 2016 onwards. Please note that it is a 
+condition of using these results that the \rave\ data release by \citet{Kunder_2016} 
+must also be cited, as the work presented here would not have been possible without the 
 tireless efforts of the entire \rave\ collaboration, past and present.

### 2016-09-09 article/unrave.tex
ATM machine

@@
 with \rave.  For main-sequence and sub-giant stars, our model uses 
-stellar parameters from the \epic\ catalog.  We derive and validate 
+stellar parameters from the \epic.  We derive and validate 
 effective temperature $\teff$, surface gravity $\logg$, and chemical 
@@
 \begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/hrd-train-set.pdf}
+\includegraphics[width=\textwidth]{hrd-train-set.pdf}
 \caption{Effective temperature $\teff$ and surface gravity $\logg$ for all stars in the training sets.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/test-set-density.pdf}
+\includegraphics[width=\textwidth]{test-set-density.pdf}
 \caption{The logarithmic density of effective temperature $\teff$ and surface gravity $\logg$ for all \Nspectra\ \rave\ spectra, as derived using the simple model (left panel; Section~\ref{sec:a-simple-model}), the main-sequence star model (center panel; Section~\ref{sec:unevolved-star-model}), and the giant star model (right panel; Section~\ref{sec:evolved-star-model}).
@@
 \center
-\includegraphics[width=0.45\textwidth]{figures/set-hyperparameters.pdf}
+\includegraphics[width=0.45\textwidth]{set-hyperparameters.pdf}
 \caption{The percentage change in RMS deviation between inferred and training labels at different regularization strengths.
@@
 \center
-\includegraphics[width=0.5\textwidth]{figures/joint-model-differences.pdf}
+\includegraphics[width=0.5\textwidth]{joint-model-differences.pdf}
 \caption{The normalized differences in effective temperature $\teff$ and surface gravity $\logg$ between the main-sequence model and the simple model (top panel), and the giant model and the simple model (bottom panel).
@@
 \center
-\includegraphics[width=0.5\textwidth]{figures/model-weights.pdf}
+\includegraphics[width=0.5\textwidth]{model-weights.pdf}
 \caption{The mean relative main-sequence model weight $w_{ms}/(w_{ms} + w_{giant})$ at each hexagonal bin of weighted effective temperature $\teff$ and surface gravity $\logg$.
@@
 \begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
+\includegraphics[width=\textwidth]{hrd-test-set.pdf}
 \caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/pairwise-metrics.pdf}
+\includegraphics[width=\textwidth]{pairwise-metrics.pdf}
 \caption{Distribution of differences in label estimates from multiple visits, divided by the quadrature sum of their formal errors, and a minimum error value for each observation.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/repeat-visits.pdf}
+\includegraphics[width=\textwidth]{repeat-visits.pdf}
 \caption{The RMS deviation of labels for repeated observations in the test set.  The RMS deviation is binned as a function of the S/N ratio of the individual visit spectra.  The
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/dr4-comparison.pdf}
+\includegraphics[width=\textwidth]{dr4-comparison.pdf}
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth \rave\ data release \citep{Kordopatis_2013} and this work.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-hrd.pdf}
+\includegraphics[width=\textwidth]{gold-standard-hrd.pdf}
 \caption{Hertsprung-Russell diagrams of stars in common between this work and that of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/gold-standard-comparison.pdf}
+\includegraphics[width=\textwidth]{gold-standard-comparison.pdf}
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparisons for stars in common between this work and `gold standard' studies that use high-resolution, high S/N spectra and \hipparcos\ parallaxes where available: \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-comparison.pdf}
+\includegraphics[width=\textwidth]{ges-comparison.pdf}
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison between the fourth internal data release from the \ges\ survey, and this work.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/ges-abundances.pdf}
+\includegraphics[width=\textwidth]{ges-abundances.pdf}
 \caption{Detailed chemical abundances in the fourth internal data release from the \ges\ survey compared to this work.
@@
 \begin{figure*}[p]
-\includegraphics[width=\textwidth]{figures/kordopatis-calibration.pdf}
+\includegraphics[width=\textwidth]{kordopatis-calibration.pdf}
 \caption{Stellar parameter ($\teff$, $\logg$, [Fe/H]) comparison with the literature calibration sources used by \citet{Kordopatis_2013} and \citet{Kunder_2016}.
@@
 \center
-\includegraphics[height=\textheight]{figures/globular-clusters.pdf}
+\includegraphics[height=\textheight]{globular-clusters.pdf}
 \caption{Effective temperature $\teff$ and surface gravity $\logg$ for globular cluster members identified by \citet{Kunder_2014,Anguiano_2015}.
@@
 \center
-\includegraphics[height=\textheight]{figures/open-clusters.pdf}
+\includegraphics[height=\textheight]{open-clusters.pdf}
 \caption{Effective temperature $\teff$ and surface gravity $\logg$ for open cluster members identified by \citep{Kunder_2016}.
@@
 \center
-\includegraphics[width=0.65\textwidth]{figures/gce.pdf}
+\includegraphics[width=0.65\textwidth]{gce.pdf}
 \caption{Detailed chemical abundances ([X/Fe]) for giant stars in \raveon\ with respect to [Fe/H], showing the Galactic chemical evolution derived for each element.
@@
 \center
-\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
+\includegraphics[width=0.65\textwidth]{hot-stars.pdf}
 \caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV} \citep{Steinmetz_2006,Zwitter_2008}.

### 2016-09-09 article/unrave.tex
Correct figures, add Schlaufman in ack

--- 2016-09-09 article/unrave.tex
@@
 	Sven Buder (MPIA), 
+	Sofia Randich (INAF),
 and
-	Sofia Randich (INAF).
+	Kevin Schlaufman (Carnegie Observatories).
 This research made use of: 
@@
 
-%\bibitem[Norris \& Da Costa(1995)]{Norris_Da_Costa_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apj, 447, 680
-
 \bibitem[Pasquini et al.(2004)]{Pasquini_2004} Pasquini, L., Randich, S., Zoccali, M., et al.\ 2004, \aap, 424, 951

### 2016-09-09 article/unrave.tex
Move figures for arXiv; include githash from master version

--- 2016-09-09 article/unrave.tex
@@
 
-\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}\newcommand{\giturl}{UNKNOWN}}
+\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{db0ab84}\newcommand{\giturl}{https://www.github.com/AnnieJumpCannon/rave}}
 

### 2016-09-09 article/unrave.tex
A word
A word; synced hash

@@
 
-\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{db0ab84}\newcommand{\giturl}{https://www.github.com/AnnieJumpCannon/rave}}
+\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{794a221}\newcommand{\giturl}{https://www.github.com/AnnieJumpCannon/rave}}
 
@@
 report a total of \ReportedAbundances\ elemental abundances with a 
-typical precision of 0.07~dex, a substantial improvement over past 
+typical precision of 0.07~dex, a substantial improvement over previous 
 \rave\ data releases.  The synthesis of \raveon\ and \tgas\ is the 

### 2017-02-26 referee-response-scientific-v1.txt
Complete referee response

I have carefully reviewed the manuscript by Casey et al. "The RAVE-on
catalog of stellar atmospheric chemical abundances..." and have the
following recommendations.

This paper presents the results of a valuable and comprehensive effort
to obtain greatly improved estimates of stellar parameters and elemental
abundances for the over 500K stars from RAVE DR5. It is clearly worthy
of publication in the ApJ, and I urge rapid acceptance.

The paper is extremely well written, and describes the procedures and
cautions in interpretation of the results very well indeed. In fact, I
have only one primary recommendation for the authors to consider.

That is, the figure I was hoping to see is one that provides the reader
a quick way to assess the distributions of the derived params and
abundances (including [Fe/H]) in the form of panels of histograms.
Basically, what I found myself wanting to see was a single set of panels
which show the full set of derived Teff, logg, [Fe/H], and [X/Fe] for
the > 500K stars that the authors have made available electronically.
This would be quite useful for other workers considering these data, and
are of course trivial to produce.

My apologies for the slight delay in responding to this manuscript, but
I was on travel, and only now carved out the time to take care of this
review. This is an extremely important paper, and one that sets a high
standard for those to follow, based on the ever-expanding data now
coming available from other surveys.

-- Author's response:

    We kindly thank the referee for their positive review of our manuscript.

    We have included the figure requested by the referee, and added a paragraph
    to the text that references this figure in the appropriate section of the
    manuscript:

        The weighted labels for all stars are shown as normalized historgrams in
        Figure \ref{fig:label-histograms}. The normalization in each axis is arbitrary, 
        since main-sequence stars do not have detailed abundances here, there are fewer
        stars with -- for example, [Ni/H] labels -- than there are stars with $\teff$
        labels. The [X/H] abundance distributions peak near Solar values, and the peaks
        in the $\teff$ and $\logg$ histograms are consistent with our expectations from
        astrophysics: the build of clump stars is visible for cool giant stars 
        ($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars
        at $\teff \approx 6000$~K is expected given the longer timescales for the turnoff.

    Unfortunately the earliest that we could reasonably include this Figure is 
    in Section 3.6 (making this new figure, Figure 6) without disturbing the
    explanatory flow of the paper.

    As a side-note: there is no need to apologise for the delay, as the lead
    author was moving internationally during this period and would have been
    unable to immediately attend to this referee response anyway.

### 2017-02-26 referee-response-statistical-v1.txt
Complete referee response

Statistical Reviewer's comments (Reviewer #2):

The following comments concern the statistical content of the article. They are all relatively minor, and as whole the article is well written and was enjoyable to review.

1. The authors use the term "labels" throughout to refer to e.g. stellar parameters and chemical abundances, and refer to "misclassification" (e.g. the second-to-last bullet point prior to Section 3.1). However, the quantities referred to are continuous parameters, not discrete/categorical, and so I find the label/misclassification terminology to be a bit confusing. What is meant by a "misclassification" when considering a continuous parameter?

-- Author's response:

    This is a very valid point, one which we failed to clarify.

    In this paper we have opted to follow the terminology introduced by Ness et al. (2015)
    where stellar parameters and chemical abundances are collectively described as stellar
    *labels*. There are similar examples in the astrophysics literature where continuous
    values have been described as labels, but this no doubt differs from the machine
    learning community where the term *label* is used to describe a discrete or catagorical
    value.

    Because this point raises valid confusion, we have added a footnote to the manuscript
    which explains the origin of the terminology, and gives a better description of what
    a label represents in the context of this paper. The following footnote has been
    added at the start of Section 3 (Method), in the first instance where *label* is 
    introduced in the paper:

        In order to dispel any (well founded) terminology confusion, we note that throughout
        this \article\ we follow the terminology introduced by \citet{Ness_2015}, where 
        \emph{continuous} values of stellar parameters and chemical abundances are 
        collectively described as stellar \emph{labels}. This varies from the machine 
        learning literature where the term `label' is more frequently used to describe
        discrete or categorical values. Here we have opted to follow the \citet{Ness_2015} 
        terminology in order to be consistent with other papers making use of \thecannon.
        Similarly, when we refer to a \emph{label misclassification}, we qualitatively
        refer to a scenario when a `significantly' incorrect label (from The Truth) has
        been assigned. In other words, we will use the term \emph{label misclassification}
        as a qualitative description, not as a quantitative definition.

    We thank the referee for raising this point, and ask their forgiveness for this
    initial confusion in terminology.

2. In Section 3.1 the authors state that Lambda is "a regularization parameter which [they] will heuristically set in later sections." I suggest that it be stated here that leave-one-out cross-validation is used (and which I encourage!), even if a full explanation is not given until Section 3.4; I found myself skipping to the later sections to learn what was meant by "heuristically." The authors may also want to mention, perhaps in Section 3.4, that leave-one-out cross-validation cannot be
used to determine the "best" value of Lambda in a global sense, but rather that it is used to select between the 30 pre-specified values (which is known to those familiar with LOOCV but is nonetheless a caveat).

-- Author's response:

    We have removed the "heuristic" sentence and replaced it with: "where $\Lambda$ is a regularization parameter which we will set by leave-one-out cross-validation in later sections,"

    We have also added the following paragraph to Section 3.4, which describes the
    performance of leave-one-out cross-validation, and comments on the qualitatively
    similar approach of a Bayesian hierarchical model (thereby addressing #4 below):

        We note that while leave-one-out cross-validation has been used to choose a justified
        regularization strength, it cannot be used to find the \emph{optimal} regularization
        strength across all pixels. In the same sense, the regularization strength $\Lambda$
        could -- in principle -- differ for each pixel \emph{and} each label. A penalized
        likelihood function of that description could qualitatively be similar to a
        Bayesian hierarchical model with strong priors on the spectral derivatives $\theta$ being
        zero. Rather, in this case we have performed 30 iterations of leave-one-out
        cross-validation and we have fixed one global $\Lambda$ value based on the RMS 
        improvement in our 30 pre-selected regularization strengths.

3. Given the statistical methods adopted, as described in Section 3.1 and elsewhere, the authors should cite the wealth of statistics literature on fitting penalized likelihood functions. For example, their methodology is closely related to the well-known LASSO (Tibshirani 1996: http://www.jstor.org/stable/2346178?seq=1#page_scan_tab_contents).

-- Author's response:

    Good catch! We have added in the citation to Tibshirani (1996), which we should have included earlier.

4. I would welcome in Section 3.1 some comments on other potential modeling approaches and their strengths/limitations. For example, a Bayesian hierarchical model would naturally provide regularization and the "borrowing of strength" across the population, but I would guess that the size of the data leads to computational challenges that (currently) prohibit fitting such a model. If the authors see fit, such comments may also lead into a brief discussion of the connection between penalized
likelihood approaches and Bayesian methods (e.g. one can consider the penalization term to be a prior distribution on the parameters, and the penalized-likelihood estimates to be the maximum a posteriori, MAP, estimates); this may helpful to readers familiar with Bayesian methods (as their use in astrophysics is increasing) but not with penalized likelihood approaches.

-- Author's response:

    We have addressed this in our answer to item #2 above. If requested we can expand on this discussion, but we thought it fit more appropriately in Section 3.4 rather than Section 3.1, because our penalized likelihood (and decisions based on different regularization strengths) is not fully introduced until Section 3.4

5. In Section 3.6, the authors introduce scaling factors that are chosen empirically to derive weighted "labels." Some discussion of the sensitivity to the choice of factors would be welcome. Presumably other values of the factors than those ultimately adopted were examined during the empirical fitting, so some discussion of the sensitivity should be feasible.

-- Author's response:

    We have added a paragraph of discussion on this topic. In practice, the normalization factors were not sensitive past the ~30% level. The paragraph we have added reads:

        It is important to note that these $\delta$ values do not represent any kind of 
        intrinsic uncertainty or precision: they are merely normalization factors.  
        Empirically, we found that adopting substantially different scaling factors (e.g.,
        a relative factor of two change) would produce clear inconsistencies in our results (e.g.,
        sub-giants being misclassified as giants). Thus while the normalization factors
        are likely sensitive at the 100\% level, our tests suggested they were not
        sensitive within the $\approx$30\% level (e.g., 120~K and 65~K, respectively). Therefore we 
        chose these factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
        approximately isotropic, and to some extent, comparable. We also note that
        these normalization factors are comparable to the RMS scatter in the training sets
        of the giant model and the main-sequence model, which qualitatively describes why 
        differences with respect to the simple model become approximately isotropic when
        scaled with these normalization factors.

6. In the first paragraph of Section 5 the authors state: "However, our analysis has caveats. Inferences based on these results should recognize those caveats, and acknowledge that these result are subject to our explicit assumptions, some of which are provably incorrect." I would go a step further and claim that all statistical inferences are based on assumptions and decisions that can be called into question (e.g. what data to use and any preprocessing of the data that is performed, how to
treat missing data and/or outliers, etc.). That the authors shine a light on their decisions/assumptions and acknowledge potential shortcomings that result is commendable.

-- Author's response:

    We agree! We have amended the paragraph in question to read:

        However, \emph{all} statistical inferences -- in any study --
        are crucially reliant on assumptions, and any number of decisions that can be
        called into question. Our analysis, and our results, are no different. Inferences based on these 
        results should recognize those caveats, and acknowledge that these results are 
        subject to our explicit assumptions, some of which are provably incorrect.

7. The authors discuss the weighting of the giant-branch and main-sequence models in the second paragraph of Section 5, noting that the "relative weighting has no warranty to be (formally) correct, and therefore may introduce inconsistencies or systematic errors rather than minimizing them." A more formal method for determining the weighting is Bayesian Model Averaging (e.g. Hoeting et al. 1999: http://www.stat.colostate.edu/~jah/papers/statsci.pdf), among others. While such approaches
may not be computationally feasible, or not preferable for other reasons, some discussion of their unsuitability for this particular analysis would be beneficial. Were any other approaches considered?

-- Author's response:

    The referee is correct. We did not consider Bayesian Model Averaging here in
    part due to computational cost. We have added some text to Section 5 which
    remarks on this:

        Because the relative weights have no formal interpretation, it is reasonable
        to consider this method is as \emph{ad hoc} as any other approach.  The relative 
        weighting has no warranty to be (formally) correct, and therefore may introduce 
        inconsistencies or systematic errors rather than minimizing them.  While Bayesian
        model averaging \citep[e.g., ][and similar methods]{Hoeting_1999} repreresent more
        formal and considered approaches to weighting -- or deciding between -- multiple 
        models, these approaches were not considered here due in part to their higher 
        computational cost.

### 2017-02-26 article/unrave.tex
Complete referee response

--- 2016-09-09 article/unrave.tex
@@
 \newcommand{\acronym}[1]{{\small{#1}}}
-
+\newcommand{\article}{\emph{Article}}
 \newcommand{\project}[1]{\textsl{#1}}
@@
 Gerry~Gilmore\altaffilmark{1},
-Toma\v{z}~Zwitter\altaffilmark{6},
-Kenneth~C.~Freeman\altaffilmark{7},
-Luca~Casagrande\altaffilmark{7},
-Gal~Matijevi\v{c}\altaffilmark{6},
-George~Seabroke\altaffilmark{8},
-Olivier~Bienaym\'e\altaffilmark{9},
-Joss~Bland-Hawthorn\altaffilmark{10},
-Brad~K.~Gibson\altaffilmark{11},
-Eva~K.~Grebel\altaffilmark{12},
-Amina~Helmi\altaffilmark{13},
-Ulisse~Munari\altaffilmark{14},
-Julio~F.~Navarro\altaffilmark{15,16},
-Warren~Reid\altaffilmark{17,18},
-Arnaud~Siebert\altaffilmark{9},
-Rosemary~Wyse\altaffilmark{19}\\
+Toma\v{z}~Zwitter\altaffilmark{7},
+Kenneth~C.~Freeman\altaffilmark{8},
+Luca~Casagrande\altaffilmark{8},
+Gal~Matijevi\v{c}\altaffilmark{7},
+George~Seabroke\altaffilmark{9},
+Olivier~Bienaym\'e\altaffilmark{10},
+Joss~Bland-Hawthorn\altaffilmark{11},
+Brad~K.~Gibson\altaffilmark{12},
+Eva~K.~Grebel\altaffilmark{13},
+Amina~Helmi\altaffilmark{14},
+Ulisse~Munari\altaffilmark{15},
+Julio~F.~Navarro\altaffilmark{16,17},
+Warren~Reid\altaffilmark{18,19},
+Arnaud~Siebert\altaffilmark{10},
+Rosemary~Wyse\altaffilmark{20}\\
 \vspace{2em}
@@
 report a total of \ReportedAbundances\ elemental abundances with a 
-typical precision of 0.07~dex, a substantial improvement over past 
+typical precision of 0.07~dex, a substantial improvement over previous 
 \rave\ data releases.  The synthesis of \raveon\ and \tgas\ is the 
-most powerful data set for chemo-dynamic analyses of the Milky Way ever produced.
+most powerful data set for chemo-dynamic analyses of the Milky Way 
+ever produced.
 \end{abstract}
@@
 required to fully characterize the formation and evolution of the Milky Way. 
-\gaia\ will also provide radial velocities, stellar parameters and chemical 
+\gaia\ will also provide radial velocities, stellar parameters, and chemical 
 abundances for a subset of brighter stars, but these measurements will not be 
@@
 wavelength coverage of \rave\ is comparable to the Radial Velocity Spectrometer on
-board the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011,Recio-Blanco_2016}, 
+board the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011},
 and the wavelength range overlaps with one of the key setups used for the ground-based 
@@
 begin to dominate.  In other words, in the low S/N regime, a well-constructed 
-data-driven model will yield more precise \emph{labels} (e.g., stellar parameters 
-and chemical abundances) than most physics-driven models\footnote{However, see 
-\citet{Casey_2016a}.}.  This is particularly relevant for the low-resolution 
-\rave\ data analysed here, because about half of the spectra have S/N 
-$\lesssim 50$~pixel$^{-1}$.
+data-driven model will yield more precise \emph{labels}\footnote{In order to 
+dispel any (well founded) terminology confusion, we note that throughout this
+\article\ we follow the terminology introduced by \citet{Ness_2015}, where 
+\emph{continuous} values of stellar parameters and chemical abundances are 
+collectively described as stellar \emph{labels}. This varies from the machine 
+learning literature where the term `label' is more frequently used to describe
+discrete or categorical values. Here we have opted to follow the \citet{Ness_2015} 
+terminology in order to be consistent with other papers making use of \thecannon.
+Similarly, when we refer to a \emph{label misclassification}, we qualitatively
+refer to a scenario when a `significantly' incorrect label (from The Truth) has
+been assigned. In other words, we will use the term \emph{label misclassification}
+as a qualitative description, not as a quantitative definition.}
+(stellar parameters and chemical abundances) than most physics-driven 
+models\footnote{However, see \citet{Casey_2016a}.}.  This is particularly relevant 
+for the low-resolution \rave\ data analysed here, because about half of the 
+spectra have S/N $\lesssim 50$~pixel$^{-1}$.
 
@@
 
-\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
-in later sections, and $Q(\vectheta)$ is a L1 regularizing function that encourages 
+\noindent{}where $\Lambda$ is a regularization parameter which we will set by leave-one-out 
+cross-validation in later sections, and $Q(\vectheta)$ is a L1 regularizing function 
+\citep[e.g., see \emph{LASSO} in][and related literature]{Tibshirani_1996} that encourages 
 $\vectheta$ values to take on zero values without breaking convexity \citep{Casey_2016b}:
@@
 
-We inferred labels for all \Nspectra\ \rave\ spectra using this main-sequence/sub-giant
+We note that while leave-one-out cross-validation has been used to choose a justified
+regularization strength, it cannot be used to find the \emph{optimal} regularization
+strength across all pixels. In the same sense, the regularization strength $\Lambda$
+could -- in principle -- differ for each pixel \emph{and} each label. A penalized
+likelihood function of that description could qualitatively be similar to a
+Bayesian hierarchical model with strong priors on the spectral derivatives $\vectheta$ being
+zero. Rather, in this case we have performed 30 iterations of leave-one-out
+cross-validation and we have fixed one global $\Lambda$ value based on the RMS 
+improvement in our 30 pre-selected regularization strengths.
+
+We inferred labels for all \Nspectra\ \rave\ spectra using this regularized main-sequence/sub-giant
 star model; we made no \emph{a priori} decisions as to whether a star was likely a
@@
 It is important to note that these $\delta$ values do not represent any kind of 
-intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
-substantially different scaling factors would produce in clear inconsistencies in our
-results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
-factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
-approximately isotropic, and to some extent, comparable.  In Figure 
+intrinsic uncertainty or precision: they are merely normalization factors.  
+Empirically, we found that adopting substantially different scaling factors (e.g.,
+a relative factor of two change) would produce clear inconsistencies in our results (e.g.,
+sub-giants being misclassified as giants). Thus while the normalization factors
+are likely sensitive at the 100\% level, our tests suggested they were not
+sensitive within the $\approx$30\% level (e.g., 120~K and 65~K, respectively). Therefore we 
+chose these factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
+approximately isotropic, and to some extent, comparable. We also note that
+these normalization factors are comparable to the RMS scatter in the training sets
+of the giant model and the main-sequence model, which qualitatively describes why 
+differences with respect to the simple model become approximately isotropic when
+scaled with these normalization factors.  In Figure 
 \ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
@@
 $w_{ms} < 0.05$.
+
+The weighted labels for all stars are shown as normalized historgrams in
+Figure \ref{fig:label-histograms}. The normalization in each axis is arbitrary, 
+since main-sequence stars do not have detailed abundances here, there are fewer
+stars with -- for example, [Ni/H] labels -- than there are stars with $\teff$
+labels. The [X/H] abundance distributions peak near Solar values, and the peaks
+in the $\teff$ and $\logg$ histograms are consistent with our expectations from
+astrophysics: the build of clump stars is visible for cool giant stars 
+($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars
+at $\teff \approx 6000$~K is expected given the longer timescales for the turnoff.
 
@@
 combined with the \tgas\ sample, these results amount to a powerful compendium 
-for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
+for chemo-dynamic studies of the Milky Way.  However, \emph{all} statistical 
+inferences -- in any study -- are crucially reliant on assumptions, and any 
+number of decisions that can be called into question. Our analysis, and our
+results, are no different. 
 Inferences based on these results should recognize those caveats, and acknowledge
-that these results are subject to our explicit assumptions, some of which are provably
-incorrect.
+that these results are subject to our explicit assumptions, some of which are 
+provably incorrect.
 
@@
 weighting has no warranty to be (formally) correct, and therefore may introduce 
-inconsistencies or systematic errors rather than minimizing them.
+inconsistencies or systematic errors rather than minimizing them.  While Bayesian
+model averaging \citep[e.g., ][and similar methods]{Hoeting_1999} repreresent more
+formal and considered approaches to weighting -- or deciding between -- multiple 
+models, these approaches were not considered here due in part to their higher 
+computational cost.
 
@@
 Derived labels, associated errors, and relevant metadata are available electronically
-through the \rave\ database from 19 September 2016 onwards. Please note that it is a 
+from Zenodo \citep{Zenodo} at \url{http://doi.org/10.5281/zenodo.154381}, or 
+from the \rave\ online database. Please note that it is a 
 condition of using these results that the \rave\ data release by \citet{Kunder_2016} 
@@
 
-\bibitem[Casey(2016)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
-
-\bibitem[Casey et al.(2016)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
+\bibitem[Casey(2016a)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
+
+\bibitem[Casey et al.(2016b)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
+
+\bibitem[Casey et al. (2016c)]{Zenodo} Casey, A.~R., et al.\ 2016, Zenodo, 10.5281/zenodo.154381
 
@@
 
+\bibitem[Hoeting et al.(1999)]{Hoeting_1999} Hoeting, J.~A., et al.\ 1999, Statistical Science, 14, 4, 382--417
+
 \bibitem[H{\o}g et al.(2000)]{Hog_2000} H{\o}g, E., Fabricius, C., Makarov, V.~V., et al.\ 2000, \aap, 355, L27
@@
 
-\bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
-
 \bibitem[Reddy et al.(2003)]{Reddy_2003} Reddy, B.~E., Tomkin, J., Lambert, D.~L., \& Allende Prieto, C.\ 2003, \mnras, 340, 304
@@
 \bibitem[Taylor(2005)]{Taylor_2005} Taylor, M.~B.\ 2005, Astronomical Data Analysis Software and Systems XIV, 347, 29
+
+\bibitem[Tibshirani(1996)]{Tibshirani_1996} Tibshirani, R.\ 1996, Journal of the Royal Statistical Society. Series B (Methodological), 58, 267-288
 
@@
 \begin{figure}[p]
+\center
+\includegraphics[width=0.5\textwidth]{figures/label-histograms.pdf}
+\caption{Normalized histograms showing the derived labels for all \rave\ stars after combining labels from the main-sequence and giant star models.
+Only results meeting our quality constraints are shown (see Section~\ref{sec:discussion}).
+The abundance labels peak near Solar values, and the overdensities in $\teff$ and $\logg$ labels is expected given the longer phases of the giant branch clump and turn-off.\label{fig:label-histograms}}
+\end{figure}
+
+\begin{figure}[p]
 \includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
@@
 $^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
-$^{6}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
-$^{7}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
-$^{8}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
-$^{9}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
-$^{10}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
-$^{11}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
-$^{12}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
-$^{13}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
-$^{14}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
-$^{15}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
-$^{16}${Senior CIfAR Fellow} \\
-$^{17}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
-$^{18}${Western Sydney University, Penrith South DC, NSW 1797} \\
-$^{19}${Johns Hopkins University, Baltimore, MD, USA} \\
+$^{6}$Leibniz-Institut für Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany \\
+$^{7}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
+$^{8}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
+
+$^{9}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
+$^{10}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
+$^{11}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
+$^{12}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
+$^{13}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
+$^{14}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
+$^{15}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
+$^{16}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
+$^{17}${Senior CIfAR Fellow} \\
+$^{18}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
+$^{19}${Western Sydney University, Penrith South DC, NSW 1797} \\
+$^{20}${Johns Hopkins University, Baltimore, MD, USA} \\
 

### 2017-02-26 article/figures/plot_label_histograms.py

@@ -1,0 +1,68 @@
+
+"""
+Plot histograms of parameters (as requested by referee #1).
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy.table import Table
+from matplotlib.ticker import MaxNLocator
+
+
+catalog = Table.read("RAVE-on-v1.0.fits.gz")
+
+passed_qc = catalog["QC"]
+catalog = catalog[passed_qc]
+
+
+label_names = ('TEFF', 'LOGG', 'FE_H', 'O_H', 'MG_H', 'AL_H', 'SI_H', 'CA_H', 'NI_H')
+
+latex_labels = { 
+    "TEFF": r"$T_{\rm eff}$",
+    "LOGG": r"$\log{g}$",
+    "FE_H": r"$[{\rm Fe/H}]$",
+    "MG_H": r"$[{\rm Mg/H}]$",
+    "O_H": r"$[{\rm O/H}]$",
+    "AL_H": r"$[{\rm Al/H}]$",
+    "CA_H": r"$[{\rm Ca/H}]$",
+    "SI_H": r"$[{\rm Si/H}]$",
+    "NI_H": r"$[{\rm Ni/H}]$",
+}
+
+fig, axes = plt.subplots(3, 3, figsize=(7.66, 7.06))
+
+B = 30
+
+for ax, label_name in zip(axes.flatten(), label_names):
+
+    finite = np.isfinite(catalog[label_name])
+
+    if label_name == "TEFF":
+        bins = np.linspace(3500, 7000, B)
+
+    elif label_name == "LOGG":
+        bins = np.linspace(0, 5, B)
+
+    else:
+        bins = np.linspace(-1.2, 0.6, B)
+
+    ax.hist(catalog[label_name][finite], bins=bins, facecolor="#CCCCCC",
+        edgecolor="#666666", normed=True)
+    
+    #ax.text(0.10, 0.80, latex_labels[label_name], fontsize=14,
+    #    transform=ax.transAxes)
+
+    ax.xaxis.set_major_locator(MaxNLocator(4))
+
+    ax.set_xlabel(latex_labels[label_name])
+
+    ax.set_yticks([])    
+    ax.set_yticklabels([])
+
+    if ax.is_first_col():
+        ax.set_ylabel(r"$N$ $({\rm normalized})$")
+
+fig.tight_layout()
+
+fig.savefig("label-histograms.pdf", dpi=300)
+fig.savefig("label-histograms.png", dpi=300)

### 2017-03-14 referee-response-scientific-v1.txt
Address pixel correlations. Closes #35

@@
 
+    We have also included an additional paragraph that outlines how our assumption of 
+    pixel independence is violated, and the (negligible) impact that has on our results.
+    The added text reads:
+
+	One of these assumptions is that the noise in individual pixels is independent
+	(between adjacent pixels). The description of the data reduction in the fifth 
+	\rave\ data release paper \citep{Kunder_2016} shows that this assumption is
+	incorrect.  The noise in neighbouring pixels is correlated in two ways: first
+	due to the oversampling by the CCD pixels in the spectrograph's point spread
+	function, and due to a 3-pixel boxcar smoothing that was later applied
+	to the data.  While it is unlikely that our assumption of pixel independence
+	violates any of our results, it explains in part why our original errors were
+	underestimated.  However, there are other assumptions made that potentially
+	have more serious consequences on the validity or utility of our results.
+
     As a side-note: there is no need to apologise for the delay, as the lead

### 2017-03-14 referee-response-statistical-v1.txt
Address pixel correlations. Closes #35

@@
         computational cost.
+
+-- Author's note:
+
+    We note that we have included an additional paragraph that outlines how our assumption of pixel independence is violated, and the (negligible) impact that has on our results. The added text reads:
+
+	One of these assumptions is that the noise in individual pixels is independent
+	(between adjacent pixels). The description of the data reduction in the fifth 
+	\rave\ data release paper \citep{Kunder_2016} shows that this assumption is
+	incorrect.  The noise in neighbouring pixels is correlated in two ways: first
+	due to the oversampling by the CCD pixels in the spectrograph's point spread
+	function, and due to a 3-pixel boxcar smoothing that was later applied
+	to the data.  While it is unlikely that our assumption of pixel independence
+	violates any of our results, it explains in part why our original errors were
+	underestimated.  However, there are other assumptions made that potentially
+	have more serious consequences on the validity or utility of our results.

### 2017-03-14 article/unrave.tex
Address pixel correlations. Closes #35

@@
 
+One of these assumptions is that the noise in individual pixels is independent
+(between adjacent pixels). The description of the data reduction in the fifth 
+\rave\ data release paper \citep{Kunder_2016} shows that this assumption is
+incorrect.  The noise in neighbouring pixels is correlated in two ways: first
+due to the oversampling by the CCD pixels in the spectrograph's point spread
+function, and due to a 3-pixel boxcar smoothing that was later applied
+to the data.  While it is unlikely that our assumption of pixel independence
+violates any of our results, it explains in part why our original errors were
+underestimated.  However, there are other assumptions made that potentially
+have more serious consequences on the validity or utility of our results.
+
 For practical purposes we adopted separate models: one for the giant branch and one
@@
 
+\clearpage
+
 \begin{figure}[p]

### 2017-03-14 article/figures/plot_dr4_comparison.py
Update figures. Closes #37

@@ -77,7 +77,7 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_ylabel(" ".join([latex_label, r"$({\rm unRAVE})$"]))
+    ax.set_ylabel(" ".join([latex_label, r"$(RAVE{\rm -on})$"]))
     ax.set_xlabel(" ".join([latex_label, r"$({\rm RAVE}$ ${\rm DR4})$"]))
 
     diff = y - x

### 2017-03-14 article/figures/plot_ges_abundances.py

@@ -104,7 +104,7 @@ for i, (ax, species) in enumerate(zip(axes.flatten(), ges_species)):
     if not ax.is_first_col():
         ax.set_yticklabels([])
     else:
-        ax.set_ylabel(r"$[{\rm X/H}]$ $({\rm unRAVE})$")
+        ax.set_ylabel(r"$[{\rm X/H}]$ $(RAVE{\rm -on})$")
 
     if not ax.is_last_row():
         ax.set_xticklabels([])

### 2017-03-14 article/figures/plot_ges_comparison.py

@@ -108,7 +108,7 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm unRAVE})$"]))
+    ax.set_ylabel(" ".join([latex_labels[label_name], r"$(RAVE{\rm -on})$"]))
     ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm GES})$"]))
 
 

### 2017-03-14 article/figures/plot_globular_clusters.py

@@ -14,7 +14,7 @@ try:
 except NameError:
     from rave_io import get_cannon_dr1, get_rave_kordopatis_dr4
 
-    clusters = Table.read("RAVEDR4_GC_v3.fits")
+    clusters = Table.read("../../RAVEDR4_GC_v3.fits")
 
     unrave = join(get_cannon_dr1(), clusters, keys=("Name",))
     dr4 = join(get_rave_kordopatis_dr4(), clusters, keys=("Name",))
@@ -94,7 +94,7 @@ for i, (row, cluster_name) in enumerate(zip(axes, cluster_names)):
     match = (unrave["Cluster"] == cluster_name) * unrave_ok
     assert sum(match) > 0
 
-    # Show RHS unRAVE
+    # Show RHS RAVE-on 
     ax = row[1]
     ax.text(0.15, 0.82, r"${{\rm {0}}}$".format(cluster_name.strip().replace("NGC", "NGC\,")),
         horizontalalignment="left", transform=ax.transAxes, fontsize=16)
@@ -141,7 +141,7 @@ for ax in axes.flatten():
         if ax.is_first_col():
             ax.set_title(r"${\rm RAVE}$ ${\rm DR4}$")
         else:
-            ax.set_title(r"${\rm unRAVE}$")
+            ax.set_title(r"$RAVE{\rm -on}$")
 
 fig.subplots_adjust(top=0.9)
 cbar = plt.colorbar(scat, 

### 2017-03-14 article/figures/plot_gold_standard_comparisons.py

@@ -148,7 +148,7 @@ def latexify(label_name, axis):
         "FE_H": r"$[{\rm Fe/H}]$"
     }[label_name]
 
-    suffix = r"$({\rm Literature})$" if axis == "x" else r"$({\rm unRAVE})$"
+    suffix = r"$({\rm Literature})$" if axis == "x" else r"$(RAVE{\rm -on})$"
 
     return " ".join([latex, suffix])
 

### 2017-03-14 article/figures/plot_gold_standard_hrd.py

@@ -136,7 +136,7 @@ def latexify(label_name, axis):
         "FE_H": r"$[{\rm Fe/H}]$"
     }[label_name]
 
-    suffix = r"$({\rm Literature})$" if axis == "x" else r"$({\rm unRAVE})$"
+    suffix = r"$({\rm Literature})$" if axis == "x" else r"$(RAVE{\rm -on})$"
 
     return " ".join([latex, suffix])
 
@@ -147,7 +147,7 @@ def latexify(label_name, axis):
 
 
 # Make another 4-panel plot showing the HRD for:
-# Bensby, Reddy, Valenti & Fishcer, unRAVE.
+# Bensby, Reddy, Valenti & Fishcer, .
 
 
 comparison_labels = (
@@ -155,7 +155,7 @@ comparison_labels = (
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2003,2006)$", 
     r"${\rm Valenti}$ ${\rm +}$ ${\rm Fischer}$ $(2005)$",
     #r"${\rm Kordopatis}$ ${\rm et}$ ${\rm al.}$ $(2013{\rm ;}$ ${\rm RAVE}$ ${\rm DR4})$",
-    r"${\rm unRAVE}$"
+    r"$RAVE{\rm -on}$"
     )
 M = len(comparison_labels)
 

### 2017-03-14 article/figures/plot_kordopatis_calibration_sample.py

@@ -34,10 +34,10 @@ ok = data_table["QC"]# * (data_table["R"] > 10)
 
 latex_labels = {
     "TEFF_2": r"$T_{\rm eff}$ $[{\rm K}]$ $({\rm Literature})$",
-    "TEFF_1": r"$T_{\rm eff}$ $[{\rm K}]$ $({\rm unRAVE})$",
-    "LOGG_1": r"$\log{g}$ $({\rm unRAVE})$",
+    "TEFF_1": r"$T_{\rm eff}$ $[{\rm K}]$ $(RAVE{\rm -on})$",
+    "LOGG_1": r"$\log{g}$ $(RAVE{\rm -on})$",
     "LOGG_2": r"$\log{g}$ $({\rm Literature})$",
-    "FE_H": r"$[{\rm Fe/H}]$ $({\rm unRAVE})$",
+    "FE_H": r"$[{\rm Fe/H}]$ $(RAVE{\rm -on})$",
     "FEH": r"$[{\rm Fe/H}]$ $({\rm Literature})$"
 }
 

### 2017-03-14 article/figures/plot_open_clusters.py

@@ -14,7 +14,7 @@ try:
 except NameError:
     from rave_io import get_cannon_dr1, get_rave_kordopatis_dr4
 
-    clusters = Table.read("RAVEDR4_OC.fits")
+    clusters = Table.read("../../RAVEDR4_OC.fits")
 
     unrave = join(get_cannon_dr1(), clusters, keys=("Name",))
     dr4 = join(get_rave_kordopatis_dr4(), clusters, keys=("Name",))
@@ -141,7 +141,7 @@ for ax in axes.flatten():
         if ax.is_first_col():
             ax.set_title(r"${\rm RAVE}$ ${\rm DR4}$")
         else:
-            ax.set_title(r"${\rm unRAVE}$")
+            ax.set_title(r"$RAVE{\rm -on}$")
 
 fig.subplots_adjust(top=0.9)
 cbar = plt.colorbar(scat, 

### 2017-03-14 article/figures/rave_io.py

@@ -5,7 +5,7 @@ from astropy.table import Table
 from scipy.io import readsav
 
 DATA_PATH = "/data/gaia-eso/arc/rave-data-files/"
-DATA_PATH = ""
+DATA_PATH = "/Users/arc/research/rave/"
 
 def get_rave_kordopatis_dr4():
     return Table.read(os.path.join(DATA_PATH, "RAVE-DR4.fits"))
@@ -26,7 +26,7 @@ def get_kordopatis_comparisons():
 
 def get_cannon_dr1(filename=None):
 
-    filename = filename or os.path.join(DATA_PATH, "../../unrave-v0.97.fits.gz")
+    filename = filename or os.path.join(DATA_PATH, "RAVE-on-v1.0.fits")
 
     rave_cannon_dr1 = Table.read(os.path.join(DATA_PATH, filename))
 

### 2017-03-14 article/unrave.tex
Fix kerning of RAVE-on in figures

@@
 $^{8}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
-
 $^{9}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\

### 2017-03-14 article/figures/plot_dr4_comparison.py

@@ -77,7 +77,7 @@ for i, (ax, columns, latex_label) in enumerate(zip(axes, all_columns, latex_labe
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_ylabel(" ".join([latex_label, r"$(RAVE{\rm -on})$"]))
+    ax.set_ylabel(" ".join([latex_label, r"$({\rm \it{RAVE}}{\rm -on})$"]))
     ax.set_xlabel(" ".join([latex_label, r"$({\rm RAVE}$ ${\rm DR4})$"]))
 
     diff = y - x

### 2017-03-14 article/figures/plot_ges_abundances.py

@@ -104,7 +104,7 @@ for i, (ax, species) in enumerate(zip(axes.flatten(), ges_species)):
     if not ax.is_first_col():
         ax.set_yticklabels([])
     else:
-        ax.set_ylabel(r"$[{\rm X/H}]$ $(RAVE{\rm -on})$")
+        ax.set_ylabel(r"$[{\rm X/H}]$ $({\rm \it{RAVE}}{\rm -on})$")
 
     if not ax.is_last_row():
         ax.set_xticklabels([])

### 2017-03-14 article/figures/plot_ges_comparison.py

@@ -108,7 +108,7 @@ for i, (ax, label_name) in enumerate(zip(axes, label_names)):
     ax.set_ylim(limits)
     ax.xaxis.set_major_locator(MaxNLocator(6))
     ax.yaxis.set_major_locator(MaxNLocator(6))
-    ax.set_ylabel(" ".join([latex_labels[label_name], r"$(RAVE{\rm -on})$"]))
+    ax.set_ylabel(" ".join([latex_labels[label_name], r"$({\rm \it{RAVE}}{\rm -on})$"]))
     ax.set_xlabel(" ".join([latex_labels[label_name], r"$({\rm GES})$"]))
 
 

### 2017-03-14 article/figures/plot_globular_clusters.py

@@ -141,7 +141,7 @@ for ax in axes.flatten():
         if ax.is_first_col():
             ax.set_title(r"${\rm RAVE}$ ${\rm DR4}$")
         else:
-            ax.set_title(r"$RAVE{\rm -on}$")
+            ax.set_title(r"${\rm \it{RAVE}}{\rm -on}$")
 
 fig.subplots_adjust(top=0.9)
 cbar = plt.colorbar(scat, 

### 2017-03-14 article/figures/plot_gold_standard_comparisons.py

@@ -148,7 +148,7 @@ def latexify(label_name, axis):
         "FE_H": r"$[{\rm Fe/H}]$"
     }[label_name]
 
-    suffix = r"$({\rm Literature})$" if axis == "x" else r"$(RAVE{\rm -on})$"
+    suffix = r"$({\rm Literature})$" if axis == "x" else r"$({\rm \it{RAVE}}{\rm -on})$"
 
     return " ".join([latex, suffix])
 

### 2017-03-14 article/figures/plot_gold_standard_hrd.py

@@ -136,7 +136,7 @@ def latexify(label_name, axis):
         "FE_H": r"$[{\rm Fe/H}]$"
     }[label_name]
 
-    suffix = r"$({\rm Literature})$" if axis == "x" else r"$(RAVE{\rm -on})$"
+    suffix = r"$({\rm Literature})$" if axis == "x" else r"$({\rm \it{RAVE}}{\rm -on})$"
 
     return " ".join([latex, suffix])
 
@@ -155,7 +155,7 @@ comparison_labels = (
     r"${\rm Reddy}$ ${\rm et}$ ${\rm al.}$ $(2003,2006)$", 
     r"${\rm Valenti}$ ${\rm +}$ ${\rm Fischer}$ $(2005)$",
     #r"${\rm Kordopatis}$ ${\rm et}$ ${\rm al.}$ $(2013{\rm ;}$ ${\rm RAVE}$ ${\rm DR4})$",
-    r"$RAVE{\rm -on}$"
+    r"${\rm \it{RAVE}}{\rm -on}$"
     )
 M = len(comparison_labels)
 

### 2017-03-14 article/figures/plot_kordopatis_calibration_sample.py

@@ -34,10 +34,10 @@ ok = data_table["QC"]# * (data_table["R"] > 10)
 
 latex_labels = {
     "TEFF_2": r"$T_{\rm eff}$ $[{\rm K}]$ $({\rm Literature})$",
-    "TEFF_1": r"$T_{\rm eff}$ $[{\rm K}]$ $(RAVE{\rm -on})$",
-    "LOGG_1": r"$\log{g}$ $(RAVE{\rm -on})$",
+    "TEFF_1": r"$T_{\rm eff}$ $[{\rm K}]$ $({\rm \it{RAVE}}{\rm -on})$",
+    "LOGG_1": r"$\log{g}$ $({\rm \it{RAVE}}{\rm -on})$",
     "LOGG_2": r"$\log{g}$ $({\rm Literature})$",
-    "FE_H": r"$[{\rm Fe/H}]$ $(RAVE{\rm -on})$",
+    "FE_H": r"$[{\rm Fe/H}]$ $({\rm \it{RAVE}}{\rm -on})$",
     "FEH": r"$[{\rm Fe/H}]$ $({\rm Literature})$"
 }
 

### 2017-03-14 article/unrave.tex
typo

--- 2017-03-14 article/unrave.tex
@@
 $^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
-$^{6}$Leibniz-Institut für Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany \\
+$^{6}${Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany} \\
 $^{7}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
 $^{8}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
-
 $^{9}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\

### 2017-03-20 referee-response-scientific-v1.txt
Update referee response based on coauthor comments

@@
         in the $\teff$ and $\logg$ histograms are consistent with our expectations from
-        astrophysics: the build of clump stars is visible for cool giant stars 
+        astrophysics: the bulk of clump stars is visible for cool giant stars 
         ($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars

### 2017-03-20 article/unrave.tex
Update referee response based on coauthor comments

--- 2017-03-14 article/unrave.tex
@@
 Georges~Kordopatis\altaffilmark{6},
-Andrea~Kunder\altaffilmark{6},
+Andrea~Kunder\altaffilmark{6,7},
 Matthias~Steinmetz\altaffilmark{6},
@@
 Gerry~Gilmore\altaffilmark{1},
-Toma\v{z}~Zwitter\altaffilmark{7},
-Kenneth~C.~Freeman\altaffilmark{8},
-Luca~Casagrande\altaffilmark{8},
-Gal~Matijevi\v{c}\altaffilmark{7},
-George~Seabroke\altaffilmark{9},
-Olivier~Bienaym\'e\altaffilmark{10},
-Joss~Bland-Hawthorn\altaffilmark{11},
-Brad~K.~Gibson\altaffilmark{12},
-Eva~K.~Grebel\altaffilmark{13},
-Amina~Helmi\altaffilmark{14},
-Ulisse~Munari\altaffilmark{15},
-Julio~F.~Navarro\altaffilmark{16,17},
-Warren~Reid\altaffilmark{18,19},
-Arnaud~Siebert\altaffilmark{10},
-Rosemary~Wyse\altaffilmark{20}\\
+Toma\v{z}~Zwitter\altaffilmark{8},
+Kenneth~C.~Freeman\altaffilmark{9},
+Luca~Casagrande\altaffilmark{9},
+Gal~Matijevi\v{c}\altaffilmark{8},
+George~Seabroke\altaffilmark{10},
+Olivier~Bienaym\'e\altaffilmark{11},
+Joss~Bland-Hawthorn\altaffilmark{12},
+Brad~K.~Gibson\altaffilmark{13},
+Eva~K.~Grebel\altaffilmark{14},
+Amina~Helmi\altaffilmark{15},
+Ulisse~Munari\altaffilmark{16},
+Julio~F.~Navarro\altaffilmark{17,18},
+Warren~Reid\altaffilmark{19,20},
+Arnaud~Siebert\altaffilmark{11},
+Rosemary~Wyse\altaffilmark{21}\\
 \vspace{2em}
@@
 \citep{Gilmore_2012,Randich_2013}, \galah\ internal DR1 \citep{DeSilva_2015}, 
-\lamost\ DR1 \citep{Cui_2012}, and \rave\ DR4 \citep{Steinmetz_2006} catalogs.}, 
+\lamost\ DR1 \citep{Cui_2012}, and \rave\ DR4 \citep{Kordopatis_2013} catalogs.}, 
 we found that the RAdial Velocity Experiment \citep[\rave;][]{Steinmetz_2006} 
@@
 \item We assume that the training data are similar (in spectra) to the test data 
-where they overlap in label space, and we assume that the training data spans enough
+where they overlap in label space, and we assume that the training data span enough
 of the label space to capture the variation in the test-set spectra.
@@
 in the $\teff$ and $\logg$ histograms are consistent with our expectations from
-astrophysics: the build of clump stars is visible for cool giant stars 
+astrophysics: the bulk of clump stars is visible for cool giant stars 
 ($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars
@@
 We cross-matched our results against the official fourth \rave\ data release 
+\citep{Kordopatis_2013}
 as an initial point of external comparison (Figure \ref{fig:rave-dr4-comparison}).
@@
 in [Fe/H] (see Figure~\ref{fig:gold-standard-comparison}).  The RMS deviation
-in labels is 237~K, 0.30~dex, and 0.15~dex, respectively.  When considering 
+in labels are 237~K, 0.30~dex, and 0.15~dex, respectively.  When considering 
 the relative information content available in \rave\ (945~pixels in the near 
@@
 studies based on exoplanet host star properties (e.g., 
-\project{TESS}\footnote{At present, however, there are just $\approx$30 
+the Transiting Exoplanet Survey Satellite, \project{TESS}\footnote{At present, however, there are just $\approx$30 
 stars in \rave\ that overlap with the compilations of exoplanet host star 
@@
 This is a crucial observation, as the metallicities for stars in our 
-main-sequence training set have a principally different source to those on 
+main-sequence training set have a principally different source than those on 
 the giant branch.  We cannot make these same inferences based on other 
@@
 to verify the accuracy of their derived stellar atmospheric parameters.  These
-samples include main-sequence stars, giant stars, with a particular focus to
+samples include main-sequence stars and giant stars, with a particular focus to
 include metal-poor stars to identify (and correct) any deviations at low
@@
 \citet[][accessed 6 September 2016]{Harris_1996} catalog of globular 
-cluster properties.  The globular cluster with the most number of members is 
+cluster properties.  The globular cluster with the highest number of members is 
 NGC~5139 ($\omega$-Centauri), where we find a significant metallicity spread
@@
 inconsistencies or systematic errors rather than minimizing them.  While Bayesian
-model averaging \citep[e.g., ][and similar methods]{Hoeting_1999} repreresent more
+model averaging \citep[e.g., ][and similar methods]{Hoeting_1999} represent more
 formal and considered approaches to weighting -- or deciding between -- multiple 
@@
 are not on the same scale as \apogee, because we were forced to adopt abundances for 
-specific elements when they were unavailable.  Although we sought to adopt mean
+specific elements when they were unavailable.  Although we sought to adopt the mean
 level of Galactic chemical enrichment at a given overall metallicity, this is not a
@@
 We have shown some potential outcomes when the test set spectra differ 
-significantly from the spectra in the training set.  Test set spectra that is
+significantly from the spectra in the training set.  Test set spectra that are
 `unusual' from the training set can be projected as peculiar artefacts in 
@@
 Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
-Research Council of Canada; the Slovenian Research Agency; the Swiss National 
-Science Foundation; the Science \& Technology Facilities Council of the UK; 
-Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
-and Sydney. The RAVE web site is https://www.rave-survey.org.
+Research Council of Canada; the Slovenian Research Agency (research core funding 
+No. P1-0188); the Swiss National Science Foundation; the Science \& Technology 
+Facilities Council of the UK; Opticon; Strasbourg Observatory; and the 
+Universities of Groningen, Heidelberg and Sydney. The RAVE web site is 
+https://www.rave-survey.org.
 
@@
 
-\bibitem[Hawkins et al.(2016)]{Hawkins_2016} Hawkins, K., Masseron, T., Jofre, P., et al.\ 2016, arXiv:1604.08800
+\bibitem[Hawkins et al.(2016)]{Hawkins_2016} Hawkins, K., Masseron, T., Jofr{\'e}, P., et al.\ 2016, \aap, 594, A43
 
@@
 
-\bibitem[Ho et al.(2016)]{Ho_2016} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., et al.\ 2016, arXiv:1602.00303
+\bibitem[Ho et al.(2017)]{Ho_2016} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., et al.\ 2017, \apj, 836, 5
 
@@
 
-\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, submitted
+\bibitem[Kunder et al.(2017)]{Kunder_2016} Kunder, A., Kordopatis, G., Steinmetz, M., et al.\ 2017, \aj, 153, 75
 
@@
 
-\bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, arXiv:1606.03015
+\bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, \aap, 595, A59
 
@@
 $^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
-$^{6}$Leibniz-Institut für Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany \\
-$^{7}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
-$^{8}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
-
-$^{9}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
-$^{10}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
-$^{11}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
-$^{12}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
-$^{13}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
-$^{14}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
-$^{15}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
-$^{16}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
-$^{17}${Senior CIfAR Fellow} \\
-$^{18}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
-$^{19}${Western Sydney University, Penrith South DC, NSW 1797} \\
-$^{20}${Johns Hopkins University, Baltimore, MD, USA} \\
+$^{6}${Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany} \\
+$^{7}${Saint Martin's University, Old Main, 5000 Abbey Way SE, Lacey, WA 98503, USA} \\
+$^{8}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
+$^{9}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
+$^{10}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
+$^{11}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
+$^{12}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
+$^{13}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
+$^{14}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
+$^{15}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
+$^{16}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
+$^{17}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
+$^{18}${Senior CIfAR Fellow} \\
+$^{19}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
+$^{20}${Western Sydney University, Penrith South DC, NSW 1797} \\
+$^{21}${Johns Hopkins University, Baltimore, MD, USA} \\
 

### 2017-03-23 article/unrave.tex
I'm a failure

--- 2017-03-14 article/unrave.tex
@@
 
-\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}\newcommand{\giturl}{UNKNOWN}}
+\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{794a221}\newcommand{\giturl}{https://www.github.com/AnnieJumpCannon/rave}}
 
 \newcommand{\acronym}[1]{{\small{#1}}}
-\newcommand{\article}{\emph{Article}}
+
 \newcommand{\project}[1]{\textsl{#1}}
@@
 Melissa~Ness\altaffilmark{5},
-Hans~Walter-Rix\altaffilmark{5},
+Hans-Walter~Rix\altaffilmark{5},
 Georges~Kordopatis\altaffilmark{6},
-Andrea~Kunder\altaffilmark{6},
+Andrea~Kunder\altaffilmark{6,7},
 Matthias~Steinmetz\altaffilmark{6},
@@
 Gerry~Gilmore\altaffilmark{1},
-Toma\v{z}~Zwitter\altaffilmark{7},
+Toma\v{z}~Zwitter\altaffilmark{6},
 Kenneth~C.~Freeman\altaffilmark{8},
 Luca~Casagrande\altaffilmark{8},
-Gal~Matijevi\v{c}\altaffilmark{7},
+Gal~Matijevi\v{c}\altaffilmark{6},
 George~Seabroke\altaffilmark{9},
@@
 \rave\ data releases.  The synthesis of \raveon\ and \tgas\ is the 
-most powerful data set for chemo-dynamic analyses of the Milky Way 
-ever produced.
+most powerful data set for chemo-dynamic analyses of the Milky Way ever produced.
 \end{abstract}
@@
 required to fully characterize the formation and evolution of the Milky Way. 
-\gaia\ will also provide radial velocities, stellar parameters, and chemical 
+\gaia\ will also provide radial velocities, stellar parameters and chemical 
 abundances for a subset of brighter stars, but these measurements will not be 
@@
 wavelength coverage of \rave\ is comparable to the Radial Velocity Spectrometer on
-board the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011},
+board the \gaia\ space telescope \citep{Munari_2005,Kordopatis_2011,Recio-Blanco_2016}, 
 and the wavelength range overlaps with one of the key setups used for the ground-based 
@@
 begin to dominate.  In other words, in the low S/N regime, a well-constructed 
-data-driven model will yield more precise \emph{labels}\footnote{In order to 
-dispel any (well founded) terminology confusion, we note that throughout this
-\article\ we follow the terminology introduced by \citet{Ness_2015}, where 
-\emph{continuous} values of stellar parameters and chemical abundances are 
-collectively described as stellar \emph{labels}. This varies from the machine 
-learning literature where the term `label' is more frequently used to describe
-discrete or categorical values. Here we have opted to follow the \citet{Ness_2015} 
-terminology in order to be consistent with other papers making use of \thecannon.
-Similarly, when we refer to a \emph{label misclassification}, we qualitatively
-refer to a scenario when a `significantly' incorrect label (from The Truth) has
-been assigned. In other words, we will use the term \emph{label misclassification}
-as a qualitative description, not as a quantitative definition.}
-(stellar parameters and chemical abundances) than most physics-driven 
-models\footnote{However, see \citet{Casey_2016a}.}.  This is particularly relevant 
-for the low-resolution \rave\ data analysed here, because about half of the 
-spectra have S/N $\lesssim 50$~pixel$^{-1}$.
+data-driven model will yield more precise \emph{labels} (e.g., stellar parameters 
+and chemical abundances) than most physics-driven models\footnote{However, see 
+\citet{Casey_2016a}.}.  This is particularly relevant for the low-resolution 
+\rave\ data analysed here, because about half of the spectra have S/N 
+$\lesssim 50$~pixel$^{-1}$.
 
@@
 
-\noindent{}where $\Lambda$ is a regularization parameter which we will set by leave-one-out 
-cross-validation in later sections, and $Q(\vectheta)$ is a L1 regularizing function 
-\citep[e.g., see \emph{LASSO} in][and related literature]{Tibshirani_1996} that encourages 
+\noindent{}where $\Lambda$ is a regularization parameter which we will heuristically set
+in later sections, and $Q(\vectheta)$ is a L1 regularizing function that encourages 
 $\vectheta$ values to take on zero values without breaking convexity \citep{Casey_2016b}:
@@
 
-We note that while leave-one-out cross-validation has been used to choose a justified
-regularization strength, it cannot be used to find the \emph{optimal} regularization
-strength across all pixels. In the same sense, the regularization strength $\Lambda$
-could -- in principle -- differ for each pixel \emph{and} each label. A penalized
-likelihood function of that description could qualitatively be similar to a
-Bayesian hierarchical model with strong priors on the spectral derivatives $\vectheta$ being
-zero. Rather, in this case we have performed 30 iterations of leave-one-out
-cross-validation and we have fixed one global $\Lambda$ value based on the RMS 
-improvement in our 30 pre-selected regularization strengths.
-
-We inferred labels for all \Nspectra\ \rave\ spectra using this regularized main-sequence/sub-giant
+We inferred labels for all \Nspectra\ \rave\ spectra using this main-sequence/sub-giant
 star model; we made no \emph{a priori} decisions as to whether a star was likely a
@@
 It is important to note that these $\delta$ values do not represent any kind of 
-intrinsic uncertainty or precision: they are merely normalization factors.  
-Empirically, we found that adopting substantially different scaling factors (e.g.,
-a relative factor of two change) would produce clear inconsistencies in our results (e.g.,
-sub-giants being misclassified as giants). Thus while the normalization factors
-are likely sensitive at the 100\% level, our tests suggested they were not
-sensitive within the $\approx$30\% level (e.g., 120~K and 65~K, respectively). Therefore we 
-chose these factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
-approximately isotropic, and to some extent, comparable. We also note that
-these normalization factors are comparable to the RMS scatter in the training sets
-of the giant model and the main-sequence model, which qualitatively describes why 
-differences with respect to the simple model become approximately isotropic when
-scaled with these normalization factors.  In Figure 
+intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
+substantially different scaling factors would produce in clear inconsistencies in our
+results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
+factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
+approximately isotropic, and to some extent, comparable.  In Figure 
 \ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
@@
 $w_{ms} < 0.05$.
-
-The weighted labels for all stars are shown as normalized historgrams in
-Figure \ref{fig:label-histograms}. The normalization in each axis is arbitrary, 
-since main-sequence stars do not have detailed abundances here, there are fewer
-stars with -- for example, [Ni/H] labels -- than there are stars with $\teff$
-labels. The [X/H] abundance distributions peak near Solar values, and the peaks
-in the $\teff$ and $\logg$ histograms are consistent with our expectations from
-astrophysics: the build of clump stars is visible for cool giant stars 
-($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars
-at $\teff \approx 6000$~K is expected given the longer timescales for the turnoff.
 
@@
 combined with the \tgas\ sample, these results amount to a powerful compendium 
-for chemo-dynamic studies of the Milky Way.  However, \emph{all} statistical 
-inferences -- in any study -- are crucially reliant on assumptions, and any 
-number of decisions that can be called into question. Our analysis, and our
-results, are no different. 
+for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
 Inferences based on these results should recognize those caveats, and acknowledge 
-that these results are subject to our explicit assumptions, some of which are 
-provably incorrect.
-
-One of these assumptions is that the noise in individual pixels is independent
-(between adjacent pixels). The description of the data reduction in the fifth 
-\rave\ data release paper \citep{Kunder_2016} shows that this assumption is
-incorrect.  The noise in neighbouring pixels is correlated in two ways: first
-due to the oversampling by the CCD pixels in the spectrograph's point spread
-function, and due to a 3-pixel boxcar smoothing that was later applied
-to the data.  While it is unlikely that our assumption of pixel independence
-violates any of our results, it explains in part why our original errors were
-underestimated.  However, there are other assumptions made that potentially
-have more serious consequences on the validity or utility of our results.
+that these results are subject to our explicit assumptions, some of which are provably
+incorrect.
 
@@
 weighting has no warranty to be (formally) correct, and therefore may introduce 
-inconsistencies or systematic errors rather than minimizing them.  While Bayesian
-model averaging \citep[e.g., ][and similar methods]{Hoeting_1999} repreresent more
-formal and considered approaches to weighting -- or deciding between -- multiple 
-models, these approaches were not considered here due in part to their higher 
-computational cost.
+inconsistencies or systematic errors rather than minimizing them.
 
@@
 
-\noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
+\noindent{}Source code for this project is available at \texttt{\giturl},
 and this document was compiled from revision hash \texttt{\githash} in that repository.
 Derived labels, associated errors, and relevant metadata are available electronically
-from Zenodo \citep{Zenodo} at \url{http://doi.org/10.5281/zenodo.154381}, or 
-from the \rave\ online database. Please note that it is a 
+through the \rave\ database from 19 September 2016 onwards. Please note that it is a 
 condition of using these results that the \rave\ data release by \citet{Kunder_2016} 
@@
 
-\bibitem[Casey(2016a)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
-
-\bibitem[Casey et al.(2016b)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
-
-\bibitem[Casey et al. (2016c)]{Zenodo} Casey, A.~R., et al.\ 2016, Zenodo, 10.5281/zenodo.154381
+\bibitem[Casey(2016)]{Casey_2016a} Casey, A.~R.\ 2016, \apjs, 223, 8
+
+\bibitem[Casey et al.(2016)]{Casey_2016b} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040
 
@@
 
-\bibitem[Hoeting et al.(1999)]{Hoeting_1999} Hoeting, J.~A., et al.\ 1999, Statistical Science, 14, 4, 382--417
-
 \bibitem[H{\o}g et al.(2000)]{Hog_2000} H{\o}g, E., Fabricius, C., Makarov, V.~V., et al.\ 2000, \aap, 355, L27
@@
 
+\bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
+
 \bibitem[Reddy et al.(2003)]{Reddy_2003} Reddy, B.~E., Tomkin, J., Lambert, D.~L., \& Allende Prieto, C.\ 2003, \mnras, 340, 304
@@
 
-\bibitem[Tibshirani(1996)]{Tibshirani_1996} Tibshirani, R.\ 1996, Journal of the Royal Statistical Society. Series B (Methodological), 58, 267-288
-
 \bibitem[Valenti \& Fischer(2005)]{Valenti_Fischer_2005} Valenti, J.~A., \& Fischer, D.~A.\ 2005, \apjs, 159, 141
@@
 \begin{figure}[p]
-\center
-\includegraphics[width=0.5\textwidth]{figures/label-histograms.pdf}
-\caption{Normalized histograms showing the derived labels for all \rave\ stars after combining labels from the main-sequence and giant star models.
-Only results meeting our quality constraints are shown (see Section~\ref{sec:discussion}).
-The abundance labels peak near Solar values, and the overdensities in $\teff$ and $\logg$ labels is expected given the longer phases of the giant branch clump and turn-off.\label{fig:label-histograms}}
-\end{figure}
-
-\begin{figure}[p]
-\includegraphics[width=\textwidth]{figures/hrd-test-set.pdf}
+\includegraphics[width=\textwidth]{hrd-test-set.pdf}
 \caption{The effective temperature $\teff$ and surface gravity $\logg$ for \rave\ stars after combining labels from the main-sequence and giant star models.
@@
 
-\clearpage
-
 \begin{figure}[p]
 \center
-\includegraphics[width=0.65\textwidth]{figures/hot-stars.pdf}
+\includegraphics[width=0.65\textwidth]{hot-stars.pdf}
 \caption{Effective temperatures $\teff$ and surface gravities $\logg$ from this work, where each hexagonal bin is colored by the \emph{maximum} temperature for any star in that bin, as reported by the \rave\ pre-processing pipeline \texttt{SPARV} \citep{Steinmetz_2006,Zwitter_2008}.
@@
 $^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
-$^{6}$Leibniz-Institut für Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany \\
-$^{7}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
+$^{6}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
+$^{7}${Saint Martin's University, Old Main, 5000 Abbey Way SE, Lacey, WA 98503, USA} \\
 $^{8}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\

### 2017-03-23 article/unrave.tex
'Upgrade' to aastex v6.1

--- 2017-03-20 article/unrave.tex
@@
 
-\documentclass[preprint]{aastex}
+\documentclass[preprint]{aastex61}
 
@@
 
-\author{
-Andrew~R.~Casey\altaffilmark{1},
-Keith~Hawkins\altaffilmark{1},
-David~W.~Hogg\altaffilmark{2,3,4,5},
-Melissa~Ness\altaffilmark{5},
-Hans~Walter-Rix\altaffilmark{5},
-Georges~Kordopatis\altaffilmark{6},
-Andrea~Kunder\altaffilmark{6,7},
-Matthias~Steinmetz\altaffilmark{6},
-Sergey~Koposov\altaffilmark{1},
-Harry~Enke\altaffilmark{6},
-Jason~Sanders\altaffilmark{1},
-Gerry~Gilmore\altaffilmark{1},
-Toma\v{z}~Zwitter\altaffilmark{8},
-Kenneth~C.~Freeman\altaffilmark{9},
-Luca~Casagrande\altaffilmark{9},
-Gal~Matijevi\v{c}\altaffilmark{8},
-George~Seabroke\altaffilmark{10},
-Olivier~Bienaym\'e\altaffilmark{11},
-Joss~Bland-Hawthorn\altaffilmark{12},
-Brad~K.~Gibson\altaffilmark{13},
-Eva~K.~Grebel\altaffilmark{14},
-Amina~Helmi\altaffilmark{15},
-Ulisse~Munari\altaffilmark{16},
-Julio~F.~Navarro\altaffilmark{17,18},
-Warren~Reid\altaffilmark{19,20},
-Arnaud~Siebert\altaffilmark{11},
-Rosemary~Wyse\altaffilmark{21}\\
-\vspace{2em}
-\emph{(Affiliations can be found at the end of the article)}
-}
+\author{Andrew~R.~Casey}
+\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+
+\author{Keith~Hawkins}
+\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+
+\author{David~W.~Hogg}
+\affiliation{Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Pl., room 424, New York, NY 10003, USA}
+\affiliation{Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+
+\author{Melissa~Ness}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+
+\author{Hans-Walter~Rix}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+
+\author{Georges~Kordopatis}
+\affiliation{Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany}
+
+\author{Andrea~Kunder}
+\affiliation{Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany}
+\affiliation{Saint Martin's University, Old Main, 5000 Abbey Way SE, Lacey, WA 98503, USA}
+
+\author{Matthias~Steinmetz}
+\affiliation{Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany}
+
+\author{Sergey~Koposov}
+\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+
+\author{Harry~Enke}
+\affiliation{Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany}
+
+\author{Jason~Sanders}
+\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+
+\author{Gerry~Gilmore}
+\affiliation{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK}
+
+\author{Toma\v{z}~Zwitter}
+\affiliation{University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia}
+
+\author{Kenneth~C.~Freeman}
+\affiliation{Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia}
+
+\author{Luca~Casagrande}
+\affiliation{Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia}
+
+\author{Gal~Matijevi\v{c}}
+\affiliation{University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia}
+
+\author{George~Seabroke}
+\affiliation{Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK}
+
+\author{Olivier~Bienaym\'e}
+\affiliation{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France}
+
+\author{Joss~Bland-Hawthorn}
+\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia}
+
+\author{Brad~K.~Gibson}
+\affiliation{E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom}
+
+\author{Eva~K.~Grebel}
+\affiliation{Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany}
+
+\author{Amina~Helmi}
+\affiliation{Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands}
+
+\author{Ulisse~Munari}
+\affiliation{INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy}
+
+\author{Julio~F.~Navarro}
+\affiliation{Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2}
+\affiliation{Senior CIfAR Fellow}
+
+\author{Warren~Reid}
+\affiliation{Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia}
+\affiliation{Western Sydney University, Penrith South DC, NSW 1797}
+
+\author{Arnaud~Siebert}
+\affiliation{Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France}
+
+\author{Rosemary~Wyse}
+\affiliation{Johns Hopkins University, Baltimore, MD, USA}
 
@@
 
-\keywords{stars: fundamental parameters --- stars: abundances
-% Clear rest of page
-\clearpage}
+\keywords{stars: fundamental parameters --- stars: abundances}
 
@@
 	Kevin Schlaufman (Carnegie Observatories).
-This research made use of: 
-  	NASA's Astrophysics Data System Bibliographic Services;
-  	Astropy, a community-developed core Python package for Astronomy \citep{astropy};
-and 
-  	\project{TOPCAT} \citep{Taylor_2005}.
+
+\software{Astropy, \citep{astropy}, \project{TOPCAT} \citep{Taylor_2005}}
+
+This research made use of NASA's Astrophysics Data System Bibliographic Services.
 This work was partly supported by the European Union FP7 programme through ERC 
@@
 
-\clearpage
-
-% A magic, bad, hack.
-\noindent{}$^{1}${Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK} \\
-$^{2}${Simons Center for Data Analysis, 160 Fifth Avenue, 7th floor, New York, NY 10010, USA} \\
-$^{3}${Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Pl., room 424, New York, NY 10003, USA} \\
-$^{4}${Center for Data Science, New York University, 726 Broadway, 7th floor, New York, NY 10003, USA} \\
-$^{5}${Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany} \\
-$^{6}${Leibniz-Institut f\"ur Astrophysik Potsdam (AIP), An der Sternwarte 16, D-14482 Potsdam, Germany} \\
-$^{7}${Saint Martin's University, Old Main, 5000 Abbey Way SE, Lacey, WA 98503, USA} \\
-$^{8}${University of Ljubljana, Faculty of Mathematics and Physics, Jadranska 19, 1000 Ljubljana, Slovenia} \\
-$^{9}${Research School of Astronomy and Astrophysics, Mount Stromlo Observatory, The Australian National University, ACT 2611, Australia} \\
-$^{10}${Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, RH5 6NT, UK} \\
-$^{11}${Observatoire astronomique de Strasbourg, Universit\'e de Strasbourg, CNRS, UMR 7550, 11 rue de l'Universit\'e, F-67000 Strasbourg, France} \\
-$^{12}${Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW 2006, Australia} \\
-$^{13}${E.A. Milne Centre for Astrophysics, University of Hull, Hull, HU6 7RX, United Kingdom} \\
-$^{14}${Astronomisches Rechen-Institut, Zentrum f\"ur Astronomie der Universit\"at Heidelberg, M\"onchhofstr.\ 12--14, 69120 Heidelberg, Germany} \\
-$^{15}${Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands} \\
-$^{16}${INAF Astronomical Observatory of Padova, 36012 Asiago (VI), Italy} \\
-$^{17}${Department of Physics and Astronomy, University of Victoria, Victoria, BC, Canada V8P 5C2} \\
-$^{18}${Senior CIfAR Fellow} \\
-$^{19}${Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia} \\
-$^{20}${Western Sydney University, Penrith South DC, NSW 1797} \\
-$^{21}${Johns Hopkins University, Baltimore, MD, USA} \\
-
 \end{document}
